### 2003-07-15 docs/MSTK.tex
MSTK Mesh Toolkit by Rao V. Garimella (rao@lanl.gov), T-7, Los Alamos National Laboratory, Los Alamos, NM 87545

\documentclass[12pt]{article}

% TeX calculates margins from a point 1 inch from the top and 1 inch
% from the left of the page. So the total left margin for instance is
% the 'oddsidemargin' + 1 inch.
%
%\setlength{\textwidth}{4.5in}
%\setlength{\textheight}{7.0in}
%\setlength{\evensidemargin}{0.0in}
%\setlength{\oddsidemargin}{0.0in}
%\setlength{\topmargin}{0.5in}

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{algorithmic}
\usepackage[boxed]{algorithm}
\usepackage{rotating}
\usepackage{multicol}
%\usepackage{times}

%\displaybreak[0]       %% Break permitted but not encouraged

\begin{document}

\title{MSTK: Mesh Toolkit, v 1.0}

\author{Rao V. Garimella \\
  Mathematical Modeling and Analysis (T-7), Theoretical Division, \\
  Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
  rao@lanl.gov}

\maketitle

\thispagestyle{empty}

\newpage

\section{MSTK IO File Format}

\subsection{ASCII File Format}

\begin{verse}
\# \textit{This is a comment} \\
\# \textit{The string ``MSTK'' and File version number (1.0)} \\
\vspace{1ex}

MSTK Ver %
\vspace{1ex}

\# \textit{char *Mfilename - Geometric model file related to mesh} 
\vspace{1ex}

Mfilename
\vspace{1ex}

\# \textit{char *reptype - Type of representation} \\  
\# \textit{int NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
\vspace{1ex}

RepType \hspace{0.5ex} NV \hspace{0.5ex} NE \hspace{0.5ex} NF \hspace{0.5ex} NR \\
\hspace{2ex}

\# \textit{\textbf{VERTEX INFO}} \\
\# \textit{Each record has} \\
\# \textit{double X,Y,Z Coordinates} \\
\# \textit{int Mdim - Topological type or dimension of model entity that} \\
\# \textit{\hspace{4.7em} the vertex is on} \\
\# \textit{int Mid - ID of model entity the vertex is on} \\
\# \textit{int Mdim and Mid can be -1 if model information is absent} 
\vspace{1ex}

X\_Coord \hspace{0.5ex} Y\_Coord \hspace{0.5ex} Z\_Coord \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
X\_Coord \hspace{0.5ex} Y\_Coord \hspace{0.5ex} Z\_Coord \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NV times}
\vspace{2ex}

\# \textit{\textbf{EDGE INFO} - present only if NE $\ne$ 0} \\
\# \textit{Each edge record has} \\
\# \textit{int Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
\# \textit{int Mdim, Mid}
\vspace{1ex}

Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NEdges times}
\vspace{2ex}

\newpage
\# \textit{\textbf{FACE INFO} - present only if NF $\ne$ 0} \\
\# \textit{char *FLtype - Dimension of lower order entity describing faces} \\
\# \textit{Values: Vertex, Edge (case insensitive), e.g. VeRteX or EDGE} \\
\vspace{1ex}

FLtype 
\vspace{1ex}

\# \textit{If face described by vertices, then each face record has} \\
\# \textit{int NFV - Number of face vertices} \\
\# \textit{int Vid\_1 - ID of first vertex of face} \\
\# \textit{int Vid\_2 - ID of second vertex of face} \\
\# . . . \\
\# \textit{int Vid\_1 - ID of NFV'th vertex of face} \\
\# \textit{int Mdim, Mid} 
\vspace{1ex}

NFV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NFV \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
NFV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NFV \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NFaces times}
\vspace{2ex}\vspace{1ex}

\# \textit{If face described by edges, then each face record has} \\
\# \textit{int NFE - Number of face edges} \\
\# \textit{int $\pm$Eid\_1 - signed ID of first edge of face} \\
\# \textit{int $\pm$Eid\_2 - signed ID of second edge of face} \\
\# . . . \\
\# \textit{int $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
\# \textit{int Mdim, Mid} \\
\# \\
\# \textit{if sign of edge is +, face uses edge in direction it was defined} \\
\# \textit{if sign of edge is -, face uses edge in opposite direction} \\
\vspace{1ex}

NFE \hspace{0.5ex} $\pm$Eid\_1 \hspace{0.5ex} $\pm$Eid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Eid\_NFE \hspace{0.5ex} Mdim Mid \\
NFE \hspace{0.5ex} $\pm$Eid\_1 \hspace{0.5ex} $\pm$Eid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Eid\_NFE \hspace{0.5ex} Mdim Mid \\
. . . \\
\# \textit{Repeated NFaces times}
\vspace{2ex}\vspace{1ex}

\# \textit{\textbf{REGION INFO} - present only if NR $\ne$ 0} \\
\# \textit{char *RLtype - Dimension of Lower order entity describing region} \\
\# \textit{Values: Vertex, Face (case insensitive), e.g. VERtex or faCE}
\vspace{1ex}

RLtype
\vspace{1ex}

\# \textit{if region described by vertices, then each region record has} \\
\# \textit{int NRV - Number of region vertices} \\
\# \textit{int Vid\_1 - ID of first vertex of region} \\
\# \textit{int Vid\_2 - ID of second vertex of region} \\
\# . . . \\
\# \textit{int Vid\_NFE - ID of NRV'th vertex of region} \\
\# \textit{int Mid, (NOTE: Mdim is not specified, since it has to be 3)}
\vspace{1ex}

NRV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NRV \hspace{0.5ex} Mid \\
NRV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NRV \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeat NR times} 
\vspace{3ex}

\# \textit{if region described by faces, then each region record has} \\
\# \textit{int NRF - Number of region faces} \\
\# \textit{int Fid\_1 - signed ID of first face of region} \\
\# \textit{int Fid\_2 - signed ID of second face of region} \\
\# . . . \\
\# \textit{int Fid\_NRF - signed ID of NRF'th face of region} \\
\# \textit{int Mdim, Mid} \\
\#  \\
\# \textit{if sign of face is +, face normal points out of region} \\
\# \textit{if sign of edge is -, face normal points into region}
\vspace{1ex}

NRF \hspace{0.5ex} $\pm$Fid\_1 \hspace{0.5ex} $\pm$Fid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Fid\_NRF \hspace{0.5ex} Mid \\
NRF \hspace{0.5ex} $\pm$Fid\_1 \hspace{0.5ex} $\pm$Fid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Fid\_NRF \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NR times} 
\vspace{2ex}

\# \textit{\textbf{VERTEX ATTRIBUTES}} \\
\# \textit{int NVA - Number of Vertex attributes} \\
\# \\
\# \textit{char *VA\_name\_1 - Name of first vertex attribute} \\
\# \textit{int VA\_type\_1 - Type of first vertex attribute} \\
\# \textit{int VA\_dim\_1 - Dimension of first vertex attribute} \\
\# \\
\# \textit{char *VA\_name\_2 - Name of second vertex attribute} \\
\# \textit{int VA\_type\_2 - Type of second vertex attribute} \\
\# \textit{int VA\_dim\_2 - Dimension of first vertex attribute} \\
\# \\
\# . . .\\
\# \\
\# \textit{char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
\# \textit{int VA\_type\_NVA - Type of NVA'th vertex attribute} \\
\# \textit{int VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
\# \\
\# \textit{VA\_type can be 1 (int), 2 (double), 3 (string)} \\
\# \textit{VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
\# \textit{VA\_dim can only be 1 when VA\_type is string}
\vspace{1ex}

NVA \hspace{0.5ex} \\
VA\_name\_1 \hspace{0.5ex} VA\_type\_1 \hspace{0.5ex} VA\_dim\_1 \\ 
VA\_name\_2 \hspace{0.5ex} VA\_type\_2 \hspace{0.5ex} VA\_dim\_2 \\
. \\
VA\_dim\_NVA \hspace{0.5ex} VA\_type\_NVA \hspace{0.5ex} VA\_dim\_2  
\vspace{1ex}

\# \textit{For each vertex attribute record, list of attribute values} \\
\# \textit{E.G., there are 3 attributes for each vertex:} \\
\# \textit{a scalar int, a vector of 3 doubles and a string}

VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
. \\
. \\
. \\
\# \textit{Repeated NV times}
\vspace{2ex}

\newpage
\# \textit{\textbf{EDGE ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NEA \hspace{0.5ex} \\
EA\_name\_1 \hspace{0.5ex} EA\_type\_1 \hspace{0.5ex} EA\_dim\_1 \\ 
EA\_name\_2 \hspace{0.5ex} EA\_type\_2 \hspace{0.5ex} EA\_dim\_2 \\
. \\
EA\_dim\_NEA \hspace{0.5ex} EA\_type\_NEA \hspace{0.5ex} EA\_dim\_2  
\vspace{1ex}

\# \textit{For each edge attribute record, list of attribute values} \\
\# \textit{E.G., a scalar int, a scalar double and a string}
\vspace{1ex}

EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
. . . \\
\# \textit{Repeated NE times}
\vspace{2ex}

\# \textit{\textbf{FACE ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NFA \hspace{0.5ex} \\
FA\_name\_1 \hspace{0.5ex} FA\_type\_1 \hspace{0.5ex} FA\_dim\_1 \\ 
FA\_name\_2 \hspace{0.5ex} FA\_type\_2 \hspace{0.5ex} FA\_dim\_2 \\
. \\
FA\_dim\_NFA \hspace{0.5ex} FA\_type\_NEA \hspace{0.5ex} FA\_dim\_2  
\vspace{1ex}

\# \textit{For each face attribute record, list of attribute values} \\
\# \textit{E.G., a vector of 2 doubles and a string}
\vspace{1ex}

FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
. . . \\
\# \textit{Repeated NF times}
\vspace{2ex}

\newpage
\# \textit{\textbf{REGION ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NRA \hspace{0.5ex} \\
RA\_name\_1 \hspace{0.5ex} RA\_type\_1 \hspace{0.5ex} RA\_dim\_1 \\ 
RA\_name\_2 \hspace{0.5ex} RA\_type\_2 \hspace{0.5ex} RA\_dim\_2 \\
. \\
RA\_dim\_NRA \hspace{0.5ex} RA\_type\_NRA \hspace{0.5ex} RA\_dim\_NRA 
\vspace{1ex}

\# \textit{For each Region attribute record, list of attribute values} \\
\# \textit{E.G., a vector of 3 ints}
\vspace{1ex}

RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
. . .\\
\# \textit{Repeated NR times}
\vspace{2ex}

\end{verse}

\end{document}

### 2003-07-15 include/MRegion.h

@@ -1,0 +1,108 @@
+#ifndef _H_MRegion
+#define _H_MRegion
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "MSTK_types.h"
+#include "List.h"
+
+ extern int mrtype_nv[7], mrtype_ne[7], mrtype_nf[7];
+
+#ifdef _H_MRegion_Private
+  typedef struct MRegion {
+    int id;
+    int marker;
+    Mesh_ptr mesh;
+    char dim;
+    char gdim;
+    int gid;
+    GEntity_ptr gent;
+    RepType repType;
+    void *sameadj;
+    void *downadj;
+  } MRegion, *MRegion_ptr;
+
+  /*----- Upward adjacency definitions --------*/
+
+  /*      NONE        */
+
+  /*----- Same Level adjacency definitions ------*/
+
+  typedef struct MRegion_SameAdj_R2 {
+    int nradj;
+    MRegion_ptr *aregions;
+  } MRegion_SameAdj_R2;
+
+  /*----- Downward adjacency definitions --------*/
+
+  typedef struct MRegion_DownAdj_FN {
+    unsigned char nf;
+    unsigned char fdirs;
+    MFace_ptr *rfaces;
+  } MRegion_DownAdj_FN;
+
+  typedef struct MRegion_DownAdj_R1R2 {
+    unsigned char nv;
+    MVertex_ptr *rvertices;
+  } MRegion_DownAdj_R1R2;
+
+  typedef struct MRegion_DownAdj_R3R4 {
+    unsigned char nf;
+    unsigned char fdirs;
+    MFace_ptr *rfaces;
+  } MRegion_DownAdj_R3R4;
+
+#else
+  typedef void *MRegion_ptr;
+#endif
+
+
+  MRegion_ptr MR_New(Mesh_ptr mesh);
+  void MR_Set_GEntity(MRegion_ptr r, GEntity_ptr gent);
+  void MR_Set_GEntID(MRegion_ptr r, int gid);
+  void MR_Set_ID(MRegion_ptr r, int id);
+
+  /* Can be called by user/application after creating region by MR_New(); */
+  void MR_Set_Faces(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+  void MR_Set_Vertices(MRegion_ptr r, int nv, MVertex_ptr *mvertices);
+
+  /* Can be called by mesh modification routines */
[…]

### 2003-07-15 include/MRegion_jmp.h

@@ -1,0 +1,367 @@
+#define _H_MRegion_Private
+
+#include "MSTK.h"
+#include "MRegion.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+#ifdef DEBUG
+
+void MR_Set_RepType_F1(MRegion_ptr r);
+void MR_Set_RepType_F4(MRegion_ptr r);
+void MR_Set_RepType_R1(MRegion_ptr r);
+void MR_Set_RepType_R2(MRegion_ptr r);
+void MR_Set_RepType_R4(MRegion_ptr r);
+void MR_Set_RepType_FNR3R4(MRegion_ptr r);
+static void (*MR_Set_RepType_jmp[MSTK_MAXREP])(MRegion_ptr r) =
+{MR_Set_RepType_F1, MR_Set_RepType_F4, MR_Set_RepType_R1, 
+ MR_Set_RepType_R2, MR_Set_RepType_R4};
+
+void MR_Delete_F1(MRegion_ptr r);
+void MR_Delete_F4(MRegion_ptr r);
+void MR_Delete_R1(MRegion_ptr r);
+void MR_Delete_R2(MRegion_ptr r);
+void MR_Delete_R4(MRegion_ptr r);
+void MR_Delete_FNR3R4(MRegion_ptr r);
+static void (*MR_Delete_jmp[MSTK_MAXREP])(MRegion_ptr r) =
+{MR_Delete_F1, MR_Delete_F4, MR_Delete_R1, MR_Delete_R2, MR_Delete_R4};
+
+void MR_Set_Faces_F1(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+void MR_Set_Faces_F4(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+void MR_Set_Faces_R1(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+void MR_Set_Faces_R2(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+void MR_Set_Faces_R4(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+void MR_Set_Faces_FNR3R4(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs);
+static void 
[…]

### 2003-07-15 include/MSTK.h

@@ -1,0 +1,269 @@
+#ifndef _H_MSTK
+#define _H_MSTK
+
+#ifdef   __cplusplus
+extern "C" {
+#endif
+
+  /* #include "gmtk.h" */
+#include "MSTK_types.h"
+#include "MSTK_util.h"
+#include "List.h"
+#include "MSTK_malloc.h"
+
+
+/*******************************************************************/
+
+void        MSTK_Init(void);
+
+/********************************************************************/
+/*        MESH OBJECT OPERATORS                                     */
+/********************************************************************/  
+
+  Mesh_ptr   MESH_New(RepType type);
+  void       MESH_Delete(Mesh_ptr mesh);
+  int        MESH_InitFromFile(Mesh_ptr mesh, char *filename);
+  void       MESH_WriteToFile(Mesh_ptr mesh, char *filename);
+
+  GModel_ptr MESH_GModel(Mesh_ptr mesh);
+  RepType    MESH_RepType(Mesh_ptr mesh);
+
+  int        MESH_Num_Vertices(Mesh_ptr mesh);
+  int        MESH_Num_Edges(Mesh_ptr mesh);
+  int        MESH_Num_Faces(Mesh_ptr mesh);
+  int        MESH_Num_Regions(Mesh_ptr mesh);
+
+  MVertex_ptr  MESH_Vertex(Mesh_ptr mesh, int i);
+  MEdge_ptr    MESH_Edge(Mesh_ptr mesh, int i);
+  MFace_ptr    MESH_Face(Mesh_ptr mesh, int i);
+  MRegion_ptr  MESH_Region(Mesh_ptr mesh, int i);
+
+  MVertex_ptr  MESH_Next_Vertex(Mesh_ptr mesh, int *index);
+  MEdge_ptr    MESH_Next_Edge(Mesh_ptr mesh, int *index);
+  MFace_ptr    MESH_Next_Face(Mesh_ptr mesh, int *index);
+  MRegion_ptr  MESH_Next_Region(Mesh_ptr mesh, int *index);
+
+
+  void       MESH_Set_GModel(Mesh_ptr mesh, GModel_ptr geom);
+  int        MESH_Change_RepType(Mesh_ptr mesh, int nurep);
+
+/********************************************************************/
+/*        MESH VERTEX OPERATORS                                     */
+/********************************************************************/
+
+  MVertex_ptr MV_New(Mesh_ptr mesh);
[…]

### 2003-07-15 include/MSTK_private.h

@@ -1,0 +1,86 @@
+#ifndef _H_MSTK_PRIVATE
+#define _H_MSTK_PRIVATE
+
+#ifdef   __cplusplus
+extern "C" {
+#endif
+
+  /* #include "gmtk.h" */
+#include "MSTK_types.h"
+#include "MSTK_util.h"
+#include "List.h"
+#include "MSTK_malloc.h"
+#include "MSTK.h"
+
+  /*
+  void MSTK_Init();
+  */
+
+/* THIS FILE HAS ADDITIONAL FUNCTIONS THAT THE NORMAL USER NEED NOT SEE */
+
+
+  void       MESH_Add_Vertex(Mesh_ptr mesh, MVertex_ptr v);
+  void       MESH_Add_Edge(Mesh_ptr mesh, MEdge_ptr e);
+  void       MESH_Add_Face(Mesh_ptr mesh, MFace_ptr f);
+  void       MESH_Add_Region(Mesh_ptr mesh, MRegion_ptr r);
+
+  void       MESH_Rem_Vertex(Mesh_ptr mesh, MVertex_ptr v);
+  void       MESH_Rem_Edge(Mesh_ptr mesh, MEdge_ptr e);
+  void       MESH_Rem_Face(Mesh_ptr mesh, MFace_ptr f);
+  void       MESH_Rem_Region(Mesh_ptr mesh, MRegion_ptr r);
+
+
+/*
+  void MV_Set_RepType(MVertex_ptr v, RepType rtype);
+*/
+  void MV_Add_Edge(MVertex_ptr mvertex, MEdge_ptr medge);
+  void MV_Add_Face(MVertex_ptr mvertex, MFace_ptr mface);
+  void MV_Add_Region(MVertex_ptr mvertex, MRegion_ptr mregion);
+  void MV_Rem_Edge(MVertex_ptr mvertex, MEdge_ptr medge);
+  void MV_Rem_Face(MVertex_ptr mvertex, MFace_ptr mface);
+  void MV_Rem_Region(MVertex_ptr mvertex, MRegion_ptr mregion);
+
+/*
+  void ME_Set_RepType(MEdge_ptr medge, RepType rtype*/
+
+  void ME_Add_Face(MEdge_ptr medge, MFace_ptr mface);
+  void ME_Add_Region(MEdge_ptr medge, MRegion_ptr mregion);
+  void ME_Rem_Face(MEdge_ptr medge, MFace_ptr mface);
[…]

### 2003-07-15 include/MSTK_types.h

@@ -1,0 +1,62 @@
+#ifndef _H_TYPESMSTK
+#define _H_TYPESMSTK
+
+#define MSTK_UNKNOWN -1
+
+#ifdef   __cplusplus
+extern "C" {
+#endif
+#ifndef _H_Mesh_Private
+  typedef void     *Mesh_ptr;
+#endif
+#ifndef _H_MVertex_Private
+  typedef void     *MVertex_ptr;
+#endif
+#ifndef _H_MEdge_Private
+  typedef void     *MEdge_ptr;
+#endif
+#ifndef _H_MFace_Private
+  typedef void     *MFace_ptr;
+#endif
+#ifndef _H_MRegion_Private
+  typedef void     *MRegion_ptr;
+#endif
+#ifndef _H_MEntity_Private
+  typedef void     *MEntity_ptr;
+#endif
+#ifndef _H_GModel_Private
+  typedef void     *GModel_ptr;
+#endif
+#ifndef _H_GEntity_Private
+  typedef void     *GEntity_ptr;
+#endif
+
+#define MSTK_VER    1.0
+
+#define MSTK_MAXREP 10
+typedef enum RepType {UNKNOWN_REP=-1, F1=0, F2, F3, F4, F5, F6, R1, R2, R3, R4} RepType;
+
+
+typedef enum MFType {FDELETED=-1, FUNKNOWN=0, TRI=3, QUAD, POLYGON} MFType;
+typedef enum MRType {RDELETED=-1, RUNKNOWN=0, TET, PYRAMID, PRISM, HEX, POLYHED} MRType;
+
+typedef enum MType {MVERTEX=0, MEDGE, MFACE, MREGION} MType;
+typedef enum VType {VIGNORE=21, VPARENT=41, VDELETED=61} VType;
+
+
+/* typedefs needed for attributes */
+typedef enum AttType {INT=0, VINT, VINT2, DOUBLE, VDOUBLE, VDOUBLE2, CHAR, VCHAR} AttType;
+typedef enum Rank {SCALAR=4001, VECTOR, TENSOR} Rank;
+typedef enum Interp {CONSTANT=5001, COPY, SEQUENCE, LINEAR, LOG, ASINH, MAX,
+		     MIN, USER, AND, OR, INCMAX} Interp;
+
+/********
[…]

### 2003-07-15 src/base/ME_Face_Funcs_R1R2.c

@@ -1,0 +1,25 @@
+#define _H_MEdge_Private
+
+#include "MSTK_malloc.h"
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int ME_Num_Faces_R1R2(MEdge_ptr e) {
+    MSTK_Report("ME_Num_Faces_R1R2","Not Implemented",ERROR);
+    return 0;
+  }
+
+  List_ptr ME_Faces_R1R2(MEdge_ptr e) {
+    MSTK_Report("ME_Faces_R1R2","Not Implemented",ERROR);
+    return 0;
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/ME_Geometry.c

@@ -1,0 +1,43 @@
+#define _H_MEdge_Private
+
+#include <math.h>
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  double ME_LenSqr(MEdge_ptr e) {
+    double xyz0[3],xyz1[3],len2;
+
+    MV_Coords(e->vertex[0],xyz0);
+    MV_Coords(e->vertex[1],xyz1);
+
+    len2 = ((xyz1[0]-xyz0[0])*(xyz1[0]-xyz0[0])+
+	    (xyz1[0]-xyz0[0])*(xyz1[0]-xyz0[0])+
+	    (xyz1[0]-xyz0[0])*(xyz1[0]-xyz0[0]));
+    
+    return len2;
+  }
+
+  double ME_Len(MEdge_ptr e) {
+    return sqrt(ME_LenSqr(e));
+  }
+
+  void ME_Vec(MEdge_ptr e, double *evec) {
+    double xyz0[3],xyz1[3];
+
+    MV_Coords(e->vertex[0],xyz0);
+    MV_Coords(e->vertex[1],xyz1);
+
+    VDiff3(xyz1,xyz0,evec);
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/ME_Region_Funcs_R3R4.c

@@ -1,0 +1,56 @@
+#define _H_MEdge_Private
+
+#include "MSTK_malloc.h"
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int ME_Num_Regions_R3R4(MEdge_ptr e) {
+    MSTK_Report("ME_Num_Regions_R3R4","Not implemented",ERROR);
+    return 0;
+  }
+
+  List_ptr ME_Regions_R3R4(MEdge_ptr e) {
+    int i, k, mkr, ner, nef;
+    List_ptr eregions, efaces;
+    MFace_ptr eface;
+    MRegion_ptr freg;
+
+    ner = 0;
+    eregions = List_New(10);
+    mkr = MSTK_GetMarker();
+
+    efaces = ME_Faces(e);
+    nef = List_Num_Entries(efaces);
+
+    for (i = 0; i < nef; i++) {
+      eface = List_Entry(efaces,i);
+      for (k = 0; k < 2; k++) {
+	freg = MF_Region(eface,k);
+	if (freg && !MEnt_IsMarked(freg,mkr)) {
+	  MEnt_Mark(freg,mkr);
+	  List_Add(eregions,freg);
+	  ner++;
+	}
+      }
+    }
+    List_Delete(efaces);
+    List_Unmark(eregions,mkr);
+    MSTK_FreeMarker(mkr);
+
+    if (ner)
+      return eregions;
+    else {
+      List_Delete(eregions);
+      return 0;
+    }    
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/MEdge.c

@@ -1,0 +1,181 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  MEdge_ptr ME_New(Mesh_ptr mesh) {
+    MEdge_ptr e;
+    RepType rtype;
+
+    e = (MEdge_ptr) MSTK_malloc(sizeof(MEdge));
+
+    e->id = 0;
+    e->marker = 0;
+    e->mesh = mesh;
+    e->dim = 1;
+    e->gdim = 4; /* Nonsensical value since we don't know what it is */
+    e->gid = 0;
+    e->gent = (GEntity_ptr) NULL;
+    e->upadj = (void *) NULL;
+    e->vertex[0] = e->vertex[1] = (MVertex_ptr) NULL;
+
+    rtype = mesh ? MESH_RepType(mesh) : F1;
+    ME_Set_RepType(e,rtype);
+
+    if (mesh) MESH_Add_Edge(mesh,e);
+
+    return e;
+  }
+
+  void ME_Delete(MEdge_ptr e) {
+    MESH_Rem_Edge(e->mesh,e);
+    (*ME_Delete_jmp[e->repType])(e);
+  }
+
+  void ME_Set_RepType(MEdge_ptr e, RepType rtype) {
+    e->repType = rtype;
+    (*ME_Set_RepType_jmp[e->repType])(e);
+  }
+
+  void ME_Set_GEntity(MEdge_ptr e, GEntity_ptr gent) {
+    e->gent = gent;
+  }
+
+  void ME_Set_GEntDim(MEdge_ptr e, int gdim) {
+    e->gdim = gdim;
+  }
+
+  void ME_Set_GEntID(MEdge_ptr e, int gid) {
+    e->gid = gid;
+  }
+
+  void ME_Set_ID(MEdge_ptr e, int id) {
+    e->id = id;
+  }
+
+  void ME_Set_Vertex(MEdge_ptr e, int i, MVertex_ptr v) {
+    e->vertex[i] = v;
+    MV_Add_Edge(v,e);
+  }
+
+  Mesh_ptr ME_Mesh(MEdge_ptr e) {
+    return e->mesh;
+  }
+
+  int ME_ID(MEdge_ptr e) {
+    return e->id;
+  }
+
[…]

### 2003-07-15 src/base/MEdge_F4.c

@@ -1,0 +1,202 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void ME_Set_RepType_F4(MEdge_ptr e) {
+    MEdge_UpAdj_F4 *upadj;
+
+    upadj = e->upadj = (MEdge_UpAdj_F4 *) MSTK_malloc(sizeof(MEdge_UpAdj_F4));
+    upadj->nel = 0;
+    upadj->elements = List_New(10);
+  }
+
+  void ME_Delete_F4(MEdge_ptr e) {
+    MEdge_UpAdj_F4 *upadj;
+
+    upadj = (MEdge_UpAdj_F4 *) e->upadj;
+    List_Delete(upadj->elements);
+    MSTK_free(upadj);
+  }
+
+  int ME_Num_Faces_F4(MEdge_ptr e) {
+    List_ptr efaces;
+    int nf;
+
+#ifdef DEBUG
+    MSTK_Report("ME_Num_Faces",
+       	"May be inefficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    efaces = ME_Faces_F4(e);
+    nf = efaces ? List_Num_Entries(efaces) : 0;
+    if (efaces) List_Delete(efaces);
+
+    return nf;
+  }
+
+
+  /* Edge could be connected up to regions or isolated faces - so we
+     can't just return e->upadj->nel */
+  int ME_Num_Regions_F4(MEdge_ptr e) {
+    int nr;
+    List_ptr eregions;
+    
+    eregions = ME_Regions_F4(e);
+    nr = eregions ? List_Num_Entries(eregions) : 0;
+    if (eregions) List_Delete(eregions);
+    
+    return nr;
+  }
+
+
+  List_ptr ME_Faces_F4(MEdge_ptr e) {
+    MEdge_UpAdj_F4 *upadj;
+    List_ptr efaces, rfaces;
+    int nf, nrf, nel, mkr, i, j, dim;
+    MFace_ptr rface;
+    MEntity_ptr ent;
+
+    upadj = (MEdge_UpAdj_F4 *) e->upadj;
+
+    efaces = List_New(10);
+    nf = 0;
+    mkr = MSTK_GetMarker();
+
[…]

### 2003-07-15 src/base/MEdge_R1.c

@@ -1,0 +1,63 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void ME_Set_RepType_R1(MEdge_ptr e) {
+    MSTK_Report("ME_Set_RepType_R1","Not implemented",ERROR);
+  }
+
+  void ME_Delete_R1(MEdge_ptr e) {
+    MSTK_Report("ME_Delete_R1","Not implemented",ERROR);
+  }
+
+  int ME_Num_Faces_R1(MEdge_ptr e) {
+    return ME_Num_Faces_R1R2(e);
+  }
+
+  List_ptr ME_Faces_R1(MEdge_ptr e) {
+    return ME_Faces_R1R2(e);
+  }
+
+  int ME_Num_Regions_R1(MEdge_ptr e) {
+    return ME_Num_Regions_R1R2(e);
+  }
+
+  List_ptr ME_Regions_R1(MEdge_ptr e) {
+    return ME_Regions_R1R2(e);
+  }
+
+  void ME_Add_Face_R1(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+  
+  void ME_Rem_Face_R1(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Add_Region_R1(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Rem_Region_R1(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/MEdge_R4.c

@@ -1,0 +1,63 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void ME_Set_RepType_R4(MEdge_ptr e) {
+    MSTK_Report("ME_Set_RepType_R4","Not implemented",ERROR);
+  }
+
+  void ME_Delete_R4(MEdge_ptr e) {
+    MSTK_Report("ME_Delete_R4","Not implemented",ERROR);
+ }
+
+  int ME_Num_Faces_R4(MEdge_ptr e) {
+    return ME_Num_Faces_R3R4(e);
+  }
+
+  int ME_Num_Regions_R4(MEdge_ptr e) {
+    return ME_Num_Regions_R3R4(e);
+  }
+
+  List_ptr ME_Faces_R4(MEdge_ptr e) {
+    return ME_Faces_R3R4(e);
+  }
+
+  List_ptr ME_Regions_R4(MEdge_ptr e) {
+    return ME_Regions_R3R4(e);
+  }
+
+  void ME_Add_Face_R4(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+  
+  void ME_Rem_Face_R4(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Add_Region_R4(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Rem_Region_R4(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/MF_Region_Funcs_F1F3.c

@@ -1,0 +1,63 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  List_ptr MF_Regions_F1F3(MFace_ptr f) {
+    List_ptr fregs;
+    MFace_UpAdj_F1F3 *upadj;
+
+    upadj = (MFace_UpAdj_F1F3 *) f->upadj;
+    fregs = List_New(2);
+    if (upadj->fregions[0])
+      List_Add(fregs,upadj->fregions[0]);
+    if (upadj->fregions[1])
+      List_Add(fregs,upadj->fregions[1]);
+
+    return fregs;
+  }
+
+  MRegion_ptr MF_Region_F1F3(MFace_ptr f, int i) {
+    return ((MFace_UpAdj_F1F3 *) f->upadj)->fregions[i];
+  }
+
+  void MF_Add_Region_F1F3(MFace_ptr f, MRegion_ptr r, int side) {
+    MFace_UpAdj_F1F3 *upadj;
+
+
+    upadj = (MFace_UpAdj_F1F3 *) f->upadj;
+
+    if (side == MSTK_UNKNOWN) {
+      if (MR_FaceDir(r,f))
+	upadj->fregions[0] = r;
+      else
+	upadj->fregions[1] = r;
+    }
+    else
+      upadj->fregions[side] = r;
+  }
+
+  void MF_Rem_Region_F1F3(MFace_ptr f, MRegion_ptr r) {
+    MFace_UpAdj_F1F3 *upadj;
+
+
+    upadj = (MFace_UpAdj_F1F3 *) f->upadj;
+
+    if (upadj->fregions[0] == r)
+      upadj->fregions[0] = NULL;
+    else if (upadj->fregions[1] == r)
+      upadj->fregions[1] = NULL;
+    else
+      MSTK_Report("MF_Rem_Region_F1F3","Region not connected to face",ERROR);
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Vertex_Funcs_FN.c

@@ -1,0 +1,55 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  List_ptr MF_Vertices_FN(MFace_ptr f, int dir) {
+    int i, k, ne, edir;
+    List_ptr fverts;
+    MEdge_ptr e;
+    MVertex_ptr v;
+    MFace_DownAdj_FN *downadj;
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+    ne = downadj->ne;
+    fverts = List_New(ne);
+    for (i = 0; i < ne; i++) {
+      k = dir ? i : ne-1-i;
+      e = List_Entry(downadj->fedges,k);
+      edir = ((downadj->edirs)>>k) & 1;
+      v = ME_Vertex(e,edir^dir);
+      List_Add(fverts,v);
+    }
+
+    return fverts;
+  }
+	
+
+  int MF_UsesVertex_FN(MFace_ptr f, MVertex_ptr v) {
+    int ne, i;
+    MEdge_ptr e;
+    MFace_DownAdj_FN *downadj;
+
+    /* We have to check only ne-1 edges since they form a loop */
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+    ne = downadj->ne;
+    for (i = 0; i < ne-1; i++) {
+      e = List_Entry(downadj->fedges,i);
+      if (ME_UsesEntity(e,v,0))
+	return 1;
+    }
+
+    return 0;
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Vertex_Funcs_R1R2.c

@@ -1,0 +1,30 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_malloc.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int MF_Num_Vertices_R1R2(MFace_ptr f) {
+    MSTK_Report("MF_Num_Vertices_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+  List_ptr MF_Vertices_R1R2(MFace_ptr f, int dir) {
+    MSTK_Report("MF_Vertices_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+	
+
+  int MF_UsesVertex_R1R2(MFace_ptr f, MVertex_ptr v) {
+    MSTK_Report("MF_UsesVertex_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MFace_F4.c

@@ -1,0 +1,215 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  void MF_Set_RepType_F4(MFace_ptr f) {
+    MFace_DownAdj_FN *downadj;
+
+    downadj = f->downadj = (MFace_DownAdj_FN *) MSTK_malloc(sizeof(MFace_DownAdj_FN));
+    downadj->ne = 0;
+    downadj->edirs = 0;
+    downadj->fedges = NULL;
+  }
+
+  void MF_Delete_F4(MFace_ptr f) {
+    MFace_DownAdj_FN *downadj;
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+    List_Delete(downadj->fedges);
+    MSTK_free(downadj);
+  }
+
+  void MF_Set_Edges_F4(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+    MF_Set_Edges_FN(f,n,e,dir);
+  }
+
+  void MF_Replace_Edge_F4(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int dir) {
+    MF_Replace_Edge_FN(f,e,nue,dir);
+  }
+
+  void MF_Replace_Edge_i_F4(MFace_ptr f, int i, MEdge_ptr nue, int dir) {
+    MF_Replace_Edge_i_FN(f,i,nue,dir);
+  }
+
+  void MF_Set_Vertices_F4(MFace_ptr f, int n, MVertex_ptr *v) {
+#ifdef DEBUG
+    MSTK_Report("MF_Set_Vertices","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Vertex_i_F4(MFace_ptr f, int i, MVertex_ptr v) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Vertex","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Vertex_F4(MFace_ptr f, MVertex_ptr v, MVertex_ptr nuv) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Vertex","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  int MF_Num_Vertices_F4(MFace_ptr f) {
[…]

### 2003-07-15 src/base/MFace_R1.c

@@ -1,0 +1,128 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MF_Set_RepType_R1(MFace_ptr f) {
+    /* What info should be created for virtual entities */
+  }
+
+
+  void MF_Delete_R1(MFace_ptr f) {
+    MSTK_Report("MF_Delete_R1","Not Implemented",ERROR);
+  }
+
+  void MF_Set_Edges_R1(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Set_Edges","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_R1(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_i_R1(MFace_ptr f, int i, MEdge_ptr nue, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Set_Vertices_R1(MFace_ptr f, int n, MVertex_ptr *v) {
+    MSTK_Report("MF_Set_Vertices","Not implemented for this representation",WARN);
+  }
+
+  void MF_Replace_Vertex_R1(MFace_ptr f, MVertex_ptr v, MVertex_ptr nuv) {
+    MSTK_Report("MF_Replace_Vertex","Not implemented for this representation",WARN);
+  }
+
+  void MF_Replace_Vertex_i_R1(MFace_ptr f, int i, MVertex_ptr nuv) {
+    MSTK_Report("MF_Replace_Vertex","Not implemented for this representation",WARN);
+  }
+
+  void MF_Add_Region_R1(MFace_ptr f, MRegion_ptr r, int side) {
+#ifdef DEBUG
[…]

### 2003-07-15 src/base/MRegion_F1.c

@@ -1,0 +1,100 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_F1(MRegion_ptr r) {
+    MR_Set_RepType_FNR3R4(r);
+  }
+
+  void MR_Delete_F1(MRegion_ptr r) {
+    MR_Delete_FNR3R4(r);
+  }
+
+  void MR_Set_Faces_F1(MRegion_ptr r, int nf, MFace_ptr *rfaces, int *dirs) {
+    MR_Set_Faces_FNR3R4(r,nf,rfaces,dirs);
+  }
+
+  void MR_Set_Vertices_F1(MRegion_ptr r, int nv, MFace_ptr *mvertices) {
+    MR_Set_Vertices_FNR3R4(r,nv,mvertices);
+  }
+
+  int MR_Num_Faces_F1(MRegion_ptr r) {
+    return MR_Num_Faces_FNR3R4(r);
+  }
+
+  /*
+  int MR_Num_AdjRegions_F1(MRegion_ptr r) {
+    return MR_Num_AdjRegions_FNR3R4(r);
+  }
+  */
+
+  List_ptr MR_Vertices_F1(MRegion_ptr r) {
+    return MR_Vertices_FNR3R4(r);
+  }
+
+  List_ptr MR_Edges_F1(MRegion_ptr r) {
+    return MR_Edges_FNR3R4(r);
+  }
+
+  List_ptr MR_Faces_F1(MRegion_ptr r) {
+    return MR_Faces_FNR3R4(r);
+  }
+
+  List_ptr MR_AdjRegions_F1(MRegion_ptr r) {
+    return MR_AdjRegions_FNR3R4(r);
+  }
+
+  int MR_FaceDir_F1(MRegion_ptr r, MFace_ptr f) {
+    return MR_FaceDir_FNR3R4(r,f);
+  }
+
+  int MR_FaceDir_i_F1(MRegion_ptr r, int i) {
+    return MR_FaceDir_i_FNR3R4(r,i);
+  }
+
+  void MR_Replace_Face_F1(MRegion_ptr r, MFace_ptr f, MFace_ptr nuf,int nudir){
[…]

### 2003-07-15 src/base/MRegion_F4.c

@@ -1,0 +1,101 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_F4(MRegion_ptr r) {
+    MR_Set_RepType_FNR3R4(r);
+  }
+
+  void MR_Delete_F4(MRegion_ptr r) {
+    MR_Delete_FNR3R4(r);
+  }
+
+  void MR_Set_Faces_F4(MRegion_ptr r, int nf, MFace_ptr *rfaces, int *dirs) {
+    MR_Set_Faces_FNR3R4(r,nf,rfaces,dirs);
+  }
+
+  void MR_Set_Vertices_F4(MRegion_ptr r, int nv, MVertex_ptr *mvertices) {
+    MR_Set_Vertices_FNR3R4(r,nv,mvertices);
+  }
+
+  int MR_Num_Faces_F4(MRegion_ptr r) {
+    return MR_Num_Faces_FNR3R4(r);
+  }
+
+  /*
+  int MR_Num_AdjRegions_F4(MRegion_ptr r) {
+    return MR_Num_AdjRegions_FNR3R4(r);
+  }
+  */
+
+  List_ptr MR_Vertices_F4(MRegion_ptr r) {
+    return MR_Vertices_FNR3R4(r);
+  }
+
+  List_ptr MR_Edges_F4(MRegion_ptr r) {
+    return MR_Edges_FNR3R4(r);
+  }
+
+  List_ptr MR_Faces_F4(MRegion_ptr r) {
+    return MR_Faces_FNR3R4(r);
+  }
+
+  List_ptr MR_AdjRegions_F4(MRegion_ptr r) {
+    return MR_AdjRegions_FNR3R4(r);
+  }
+
+  int MR_FaceDir_F4(MRegion_ptr r, MFace_ptr f) {
+    return MR_FaceDir_FNR3R4(r,f);
+  }
+
+  int MR_FaceDir_i_F4(MRegion_ptr r, int i) {
+    return MR_FaceDir_i_FNR3R4(r,i);
+  }
+
+  void MR_Replace_Face_F4(MRegion_ptr r, MFace_ptr f, MFace_ptr nuf,int nudir){
[…]

### 2003-07-15 src/base/MRegion_FNR3R4.c

@@ -1,0 +1,466 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_FNR3R4(MRegion_ptr r) {
+    MRegion_DownAdj_FN *downadj;
+
+    r->downadj = (MRegion_DownAdj_FN *) MSTK_malloc(sizeof(MRegion_DownAdj_FN));
+    downadj = (MRegion_DownAdj_FN *) r->downadj;
+
+    downadj->nf = (unsigned char) 0;
+    downadj->fdirs = 0;
+    downadj->rfaces = NULL;
+  }
+
+  void MR_Delete_FNR3R4(MRegion_ptr r) {
+    MRegion_DownAdj_FN *downadj;
+    MFace_ptr f;
+    int i, nf;
+
+    downadj = (MRegion_DownAdj_FN *) r->downadj;
+
+    nf = List_Num_Entries(downadj->rfaces);
+    for (i = 0; i < nf; i++) {
+      f = List_Entry(downadj->rfaces,i);
+      MF_Rem_Region(f,r);
+    }
+    List_Delete(downadj->rfaces);
+    MSTK_free(downadj);
+
+    MSTK_free(r);
+  }
+
+  void MR_Set_Faces_FNR3R4(MRegion_ptr r, int nf, MFace_ptr *rfaces,int *dirs){
+    int i;
+    MRegion_DownAdj_FN *downadj;
+
+    downadj = (MRegion_DownAdj_FN *) r->downadj;
+    downadj->nf = nf;
+    downadj->fdirs = 0;
+    downadj->rfaces = List_New(nf);
+
+    for (i = 0; i < nf; i++) {
+      downadj->fdirs = downadj->fdirs | (dirs[i] << i);
+      List_Add(downadj->rfaces,rfaces[i]);
+      MF_Add_Region(rfaces[i],r,!dirs[i]);
+    }
+  }
+
+  void MR_Set_Vertices_FNR3R4(MRegion_ptr r, int nv, MFace_ptr *mvertices) {
+#ifdef DEBUG
+    MSTK_Report("MF_Rem_AdjRegion",
+		"Function call not suitable for this representation",WARN);
+#endif    
+  }
+
+  int MR_Num_Faces_FNR3R4(MRegion_ptr r) {
[…]

### 2003-07-15 src/base/MRegion_R2.c

@@ -1,0 +1,165 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_R2(MRegion_ptr r) {
+    MRegion_DownAdj_R1R2 *downadj;
+    MRegion_SameAdj_R2 *sameadj;
+
+    r->downadj = (MRegion_DownAdj_R1R2 *) MSTK_malloc(sizeof(MRegion_DownAdj_R1R2));
+    downadj = (MRegion_DownAdj_R1R2 *) r->downadj;
+    downadj->nv = (unsigned char) 0;
+    downadj->rvertices = NULL;
+
+    r->sameadj = (MRegion_SameAdj_R2 *) MSTK_malloc(sizeof(MRegion_SameAdj_R2));
+    sameadj = (MRegion_SameAdj_R2 *) r->sameadj;
+    sameadj->nradj = (unsigned char) 0;
+    sameadj->aregions = NULL;
+  }
+
+  void MR_Delete_R2(MRegion_ptr r) {
+    MRegion_DownAdj_R1R2 *downadj;
+    MRegion_SameAdj_R2 *sameadj;
+
+    downadj = (MRegion_DownAdj_R1R2 *) r->downadj;    
+    List_Delete(downadj->rvertices);
+    MSTK_free(downadj);
+
+    sameadj = (MRegion_SameAdj_R2 *) r->sameadj;
+    List_Delete(sameadj->aregions);
+    MSTK_free(sameadj);
+  }
+
+  void MR_Set_Vertices_R2(MRegion_ptr r, int nv, MVertex_ptr *rvertices) {
+    int i;
+    MRegion_DownAdj_R1R2 *downadj;
+
+    downadj = (MRegion_DownAdj_R1R2 *) r->downadj;
+    downadj->nv = nv;
+    downadj->rvertices = List_New(nv);
+    for (i = 0; i < nv; i++)
+      List_Add(downadj->rvertices,rvertices[i]);
+  }
+
+  void MR_Add_AdjRegion_R2(MRegion_ptr r, int facenum, MRegion_ptr aregion) {
+    MRegion_SameAdj_R2 *sameadj;
+
+    /* Is r->sameadj allocated ? */
+    /* We should make use of the facenum info */
[…]

### 2003-07-15 src/base/MRegion_R4.c

@@ -1,0 +1,103 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_R4(MRegion_ptr r) {
+    MR_Set_RepType_FNR3R4(r);
+  }
+
+  void MR_Delete_R4(MRegion_ptr r) {
+    MR_Delete_FNR3R4(r);
+  }
+
+  void MR_Set_Faces_R4(MRegion_ptr r, int nf, MFace_ptr *rfaces, int *dirs) {
+    MR_Set_Faces_FNR3R4(r,nf,rfaces,dirs);
+  }
+
+  void MR_Set_Vertices_R4(MRegion_ptr r, int nv, MFace_ptr *mvertices) {
+    MR_Set_Vertices_FNR3R4(r,nv,mvertices);
+  }
+
+  int MR_Num_Faces_R4(MRegion_ptr r) {
+    return MR_Num_Faces_FNR3R4(r);
+  }
+
+  /*
+  int MR_Num_AdjRegions_R4(MRegion_ptr r) {
+    return MR_Num_AdjRegions_FNR3R4(r);
+  }
+  */
+
+  List_ptr MR_Vertices_R4(MRegion_ptr r) {
+    return MR_Vertices_FNR3R4(r);
+  }
+
+  List_ptr MR_Edges_R4(MRegion_ptr r) {
+    return MR_Edges_FNR3R4(r);
+  }
+
+  List_ptr MR_Faces_R4(MRegion_ptr r) {
+    return MR_Faces_FNR3R4(r);
+  }
+
+  List_ptr MR_AdjRegions_R4(MRegion_ptr r) {
+    return MR_AdjRegions_FNR3R4(r);
+  }
+
+  int MR_FaceDir_R4(MRegion_ptr r, MFace_ptr f) {
+    return MR_FaceDir_FNR3R4(r,f);
+  }
+
+  int MR_FaceDir_i_R4(MRegion_ptr r, int i) {
+    return MR_FaceDir_i_FNR3R4(r,i);
+  }
+
+  int MR_UsesFace_R4(MRegion_ptr r, MFace_ptr f) {
+    return MR_UsesFace_FNR3R4(r, f);
+  }
+
+  int MR_UsesEdge_R4(MRegion_ptr r, MEdge_ptr e) {
[…]

### 2003-07-15 src/base/MSTK.c

@@ -1,0 +1,75 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "MSTK_private.h"
+#include "List.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  unsigned int MSTK_marker = 0;
+  int MSTK_MAXBITS = 8*sizeof(unsigned int);
+  int MSTK_lastbit = -1;
+
+  void MSTK_Init() {
+#ifdef DEBUG
+    List_ptr l = List_New(10);
+    List_PrintID(l);
+    List_Delete(l);
+#endif
+
+
+  }
+
+  int MSTK_GetMarker() {
+    int i;
+
+    /*** CHECK CAREFULLY ***/
+
+    if (MSTK_lastbit < MSTK_MAXBITS-1) {
+      MSTK_lastbit++;
+      MSTK_marker = MSTK_marker | 1<<MSTK_lastbit;
+      return (MSTK_lastbit+1); /* lastbit goes from 0 to N-1, markers from 1 to N */
+    }
+    else {
+      /* Search if any other bit is free */
+      i = 0;
+      while (i <= MSTK_MAXBITS-1) {
+	if (MSTK_marker & 1<<i)
+	  i++;
+	else {
+	  MSTK_marker = MSTK_marker | 1<<i;
+	  return i+1;
+	}
+      }
+
+#ifdef DEBUG
+      MSTK_Report("MSTK_GetMarker","Out of Markers",ERROR);
+#endif
+      return -1;
+    }
+
+#ifdef DEBUG
+    MSTK_Report("MSTK_GetMarker","Cannot allocate Marker",ERROR);
+#endif
+    return -1;
+  }
+
+  void MSTK_FreeMarker(int mkr) {
+
+#ifdef DEBUG
+    if (mkr < 1 || mkr > MSTK_MAXBITS)
+      MSTK_Report("MSTK_FreeMarker","Invalid Marker",ERROR);
+#endif
+
+    MSTK_marker = MSTK_marker & ~(1<<(mkr-1));
+    if (mkr-1 == MSTK_lastbit)
+      MSTK_lastbit--;
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MV_Face_Funcs_R1R2.c

@@ -1,0 +1,42 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Functions */
+
+  int MV_Num_Faces_R1R2(MVertex_ptr v) {
+    int i, nf;
+    List_ptr vfaces;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Faces_R1R2",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    vfaces = MV_Faces_R1R2(v);
+    nf = List_Num_Entries(vfaces);
+    for (i = 0; i < nf; i++) {
+      /* Must destroy the temporary faces */
+    }
+    List_Delete(vfaces);
+
+    return nf;
+  }
+
+
+  List_ptr MV_Faces_R1R2(MVertex_ptr v) {
+    MSTK_Report("MV_Faces","Not yet implemented for this representation",
+		MESG);
+    return NULL;
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MV_Region_Funcs_R3R4.c

@@ -1,0 +1,68 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Functions */
+
+  int MV_Num_Regions_R3R4(MVertex_ptr v) {
+    List_ptr vregions;
+    int nr;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Regions_R3R4",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    vregions = MV_Regions_R3R4(v);
+    nr = List_Num_Entries(vregions);
+    List_Delete(vregions);
+
+    return nr;
+  }
+
+
+  List_ptr MV_Regions_R3R4(MVertex_ptr v) {
+    int i, j, nr, nf, mkr;
+    MFace_ptr face;
+    MRegion_ptr reg;
+    List_ptr vregions, vfaces;
+ 
+    vregions = List_New(10);
+    nr = 0;
+    mkr = MSTK_GetMarker();
+
+    vfaces = MV_Faces_R4(v);
+    nf = List_Num_Entries(vfaces);
+    for (i = 0; i < nf; i++) {
+      face = List_Entry(vfaces,i);
+      for (j = 0; j < 2; j++) {
+	reg = MF_Region(face,j);
+	if (reg && !MEnt_IsMarked(reg,mkr)) {
+	  MEnt_Mark(reg,mkr);
+	  List_Add(vregions,reg);
+	  nr++;
+	}
+      }
+    }
+    List_Delete(vfaces);
+    List_Unmark(vregions,mkr);
+    MSTK_FreeMarker(mkr);
+
+    if (nr > 0)
+      return vregions;
+    else {
+      List_Delete(vregions);
+      return 0;
+    }
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MVertex_F4.c

@@ -1,0 +1,257 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  void MV_Set_RepType_F4(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+
+    upadj = v->upadj = (MVertex_UpAdj_F1F4 *) MSTK_malloc(sizeof(MVertex_UpAdj_F1F4));
+    upadj->ne = (unsigned int) 0;
+    upadj->vedges = List_New(10);
+  }
+
+  void MV_Delete_F4(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+
+    upadj = (MVertex_UpAdj_F1F4 *) v->upadj;
+    List_Delete(upadj->vedges);
+    MSTK_free(upadj);
+  }
+
+  int MV_Num_AdjVertices_F4(MVertex_ptr v) {
+    return ((MVertex_UpAdj_F1F4 *) v->upadj)->ne;
+  }
+
+  int MV_Num_Edges_F4(MVertex_ptr v) {
+    return ((MVertex_UpAdj_F1F4 *) v->upadj)->ne;
+  }
+
+  int MV_Num_Faces_F4(MVertex_ptr v) {
+    List_ptr vfaces;
+    int nf;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Faces",
+		"Ineficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    vfaces = MV_Faces_F4(v);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
+    return nf;
+  }
+
+  int MV_Num_Regions_F4(MVertex_ptr v) {
+    List_ptr vregions;
+    int nr;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Regions",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+	
+    vregions = MV_Regions_F4(v);
+    nr = List_Num_Entries(vregions);
+    List_Delete(vregions);
+    return nr;
+  }
+
+  List_ptr MV_AdjVertices_F4(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+    List_ptr vedges, adjv;
+    int ne, i;
+    MEdge_ptr vedge;
[…]

### 2003-07-15 src/base/MVertex_R1.c

@@ -1,0 +1,193 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MV_Set_RepType_R1(MVertex_ptr v) {
+    MVertex_UpAdj_R1R2 *upadj;
+
+    upadj = v->upadj = (MVertex_UpAdj_R1R2 *) MSTK_malloc(sizeof(MVertex_UpAdj_R1R2));
+    upadj->nel = (unsigned int) 0;
+    upadj->velements = List_New(10);
+  }
+
+  void MV_Delete_R1(MVertex_ptr v) {
+    MVertex_UpAdj_R1R2 *upadj;
+
+    upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
+    List_Delete(upadj->velements);
+    MSTK_free(upadj);
+  }
+
+  int MV_Num_AdjVertices_R1(MVertex_ptr v) {
+#ifdef DEBUG
+    MSTK_Report("MV_Num_AdjVertices",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    return MV_Num_Edges_R1(v);
+  }
+
+  int MV_Num_Edges_R1(MVertex_ptr v) {
+    int ne;
+    List_ptr vedges;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Edges",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    vedges = MV_Edges_R1(v);
+    ne = List_Num_Entries(vedges);
+    List_Delete(vedges);
+    return ne;
+  }
+
+  int MV_Num_Faces_R1(MVertex_ptr v) {
+    int nf;
+    List_ptr vfaces;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Faces",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    /* Have to account for the fact that this may be a surface mesh */
+    /* Then v->upadj->velements contains faces of vertex which can be
+       retrieved fairly efficiently */
+
+    vfaces = MV_Faces_R1(v);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
[…]

### 2003-07-15 src/base/MVertex_R2.c

@@ -1,0 +1,198 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MV_Set_RepType_R2(MVertex_ptr v) {
+    MVertex_UpAdj_R1R2 *upadj;
+    MVertex_SameAdj_R2R4 *sameadj;
+
+    upadj = v->upadj = (MVertex_UpAdj_R1R2 *) MSTK_malloc(sizeof(MVertex_UpAdj_R1R2));
+    upadj->nel = (unsigned int) 0;
+    upadj->velements = List_New(10);
+    sameadj = v->sameadj = (MVertex_SameAdj_R2R4 *) MSTK_malloc(sizeof(MVertex_SameAdj_R2R4));
+    sameadj->nvadj = (unsigned char) 0;
+    sameadj->adjverts = List_New(10);
+  }
+
+  void MV_Delete_R2(MVertex_ptr v) {
+    MVertex_UpAdj_R1R2 *upadj;
+    MVertex_SameAdj_R2R4 *sameadj;
+
+    upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
+    List_Delete(upadj->velements);
+    MSTK_free(upadj);
+
+    sameadj = (MVertex_SameAdj_R2R4 *) v->sameadj;
+    List_Delete(sameadj->adjverts);
+    MSTK_free(sameadj);
+  }
+    
+
+  int MV_Num_AdjVertices_R2(MVertex_ptr v) {
+    return ((MVertex_SameAdj_R2R4 *) v->sameadj)->nvadj;
+  }
+
+  int MV_Num_Edges_R2(MVertex_ptr v) {
+    return ((MVertex_SameAdj_R2R4 *)v->sameadj)->nvadj;
+  }
+
+  int MV_Num_Faces_R2(MVertex_ptr v) {
+    int nf;
+    List_ptr vfaces;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Faces",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    /* Have to account for the fact that this may be a surface mesh */
+    /* Then v->upadj->velements contains faces of vertex which can be
+       retrieved fairly efficiently */
+
[…]

### 2003-07-15 src/base/MVertex_R4.c

@@ -1,0 +1,137 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Functions */
+
+  void MV_Set_RepType_R4(MVertex_ptr v) {
+    MVertex_UpAdj_R3R4 *upadj;
+    MVertex_SameAdj_R2R4 *sameadj;
+
+    upadj = v->upadj = (MVertex_UpAdj_R3R4 *) MSTK_malloc(sizeof(MVertex_UpAdj_R3R4));
+    upadj->nf = (unsigned int) 0;
+    upadj->vfaces = List_New(10);
+    sameadj = v->sameadj = (MVertex_SameAdj_R2R4 *) MSTK_malloc(sizeof(MVertex_SameAdj_R2R4));
+    sameadj->nvadj = (unsigned char) 0;
+    sameadj->adjverts = List_New(10);
+  }
+
+  void MV_Delete_R4(MVertex_ptr v) {
+    MVertex_UpAdj_R3R4 *upadj;
+    MVertex_SameAdj_R2R4 *sameadj;
+   
+    upadj = (MVertex_UpAdj_R3R4 *) v->upadj;
+    List_Delete(upadj->vfaces);
+    MSTK_free(upadj);
+
+    sameadj = (MVertex_SameAdj_R2R4 *) v->sameadj;
+    List_Delete(sameadj->adjverts);
+    MSTK_free(sameadj);
+  }
+
+  int MV_Num_AdjVertices_R4(MVertex_ptr v) {
+    return ((MVertex_SameAdj_R2R4 *)v->sameadj)->nvadj;
+  }
+
+  int MV_Num_Edges_R4(MVertex_ptr v) {
+    return ((MVertex_SameAdj_R2R4 *)v->sameadj)->nvadj;
+  }
+
+  int MV_Num_Faces_R4(MVertex_ptr v) {
+    return ((MVertex_UpAdj_R3R4 *)v->upadj)->nf;
+  }
+  
+  int MV_Num_Regions_R4(MVertex_ptr v) {
+    return MV_Num_Regions_R3R4(v);
+  }
+
+  List_ptr MV_AdjVertices_R4(MVertex_ptr v) {
+    MVertex_SameAdj_R2R4 *sameadj;
+
[…]

### 2003-07-15 src/hilev/ME_Swap2D.c

@@ -1,0 +1,109 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "MSTK.h"
+
+
+int ME_Swap2D(MEdge_ptr eswp, MEdge_ptr *enew, MFace_ptr fnew[2]) {
+  Mesh_ptr mesh;
+  MEdge_ptr fe[4], fenew[6];
+  MVertex_ptr v[4];
+  MFace_ptr f[2];
+  int dir[3], gid, k, nf;
+  List_ptr fedges, efaces;
+
+  if (ME_GEntDim(eswp) != 2) {
+    MSTK_Report("ME_Swap2D","Can swap edges on model faces only",ERROR);
+    return 0;
+  } 
+  /*  gid = ME_GEntID(eswp); */
+  gid = 0;
+ 
+  /* Collect information */
+
+  mesh = ME_Mesh(eswp);
+  v[0] = ME_Vertex(eswp,0);
+  v[1] = ME_Vertex(eswp,1);
+
+  efaces = ME_Faces(eswp);
+  nf = List_Num_Entries(efaces);
+
+  if (nf != 2) {
+    MSTK_Report("ME_Swap2D","Edge must be connected to exactly two faces for 2D swap",ERROR);
+    List_Delete(efaces);
+    return 0;
+  }
+
+
+  k = 0;
+  f[0] = List_Entry(efaces,0);
+  if (!MF_EdgeDir(f[0],eswp)) {
+    k = 1;
+    f[0] = List_Entry(efaces,1);
+  }
+
+  fedges = MF_Edges(f[0],1,v[0]);
+  if (List_Num_Entries(fedges) != 3) {
+    MSTK_Report("ME_Swap2D","Cannot swap edge for non-triangular meshes",ERROR);
+    List_Delete(efaces);
+    List_Delete(fedges);
+    return 0;
+  }
+  fe[0] = List_Entry(fedges,1);
+  fe[1] = List_Entry(fedges,2);
+  List_Delete(fedges);
+
+  v[3] = ME_Vertex(fe[0],1);
+  if (v[3] == v[1])
+    v[3] = ME_Vertex(fe[0],0);
+
+  f[1] = List_Entry(efaces,!k);
+  fedges = MF_Edges(f[1],1,v[1]);
+  if (List_Num_Entries(fedges) != 3) {
[…]

### 2003-07-15 src/hilev/MF_Coords.c

@@ -1,0 +1,26 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "MSTK.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+void MF_Coords(MFace_ptr f, int *n, double (*xyz)[3]) {
+  int i;
+  MVertex_ptr fv;
+  List_ptr fverts;
+
+  fverts = MF_Vertices(f,1);
+  *n = List_Num_Entries(fverts);
+
+  for (i = 0; i < *n; i++) {
+    fv = List_Entry(fverts,i);
+    MV_Coords(fv,xyz[i]);
+  }
+  List_Delete(fverts);
+}
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/hilev/MFs_Join.c

@@ -1,0 +1,74 @@
+#include <stdlib.h>
+#include <stdio.h>
+#include "MSTK.h"
+
+
+MFace_ptr MFs_Join(MFace_ptr f1, MFace_ptr f2, MEdge_ptr e) {
+  int i, k1, k2, nfe1, nfe2, *fedir1, *fedir2, *fedir3;
+  MEdge_ptr *fe1, *fe2, *fe3;
+  MFace_ptr  nuface;
+  Mesh_ptr   mesh;
+  List_ptr   fedges;
+
+  mesh = MF_Mesh(f1);
+  
+
+  nfe1 = MF_Num_Edges(f1);
+  nfe2 = MF_Num_Edges(f2);
+
+  fe1 = (MEdge_ptr *) malloc(nfe1*sizeof(MEdge_ptr));
+  fedir1 = (int *) malloc(nfe1*sizeof(int));
+  fe2 = (MEdge_ptr *) malloc(nfe2*sizeof(MEdge_ptr));
+  fedir2 = (int *) malloc(nfe2*sizeof(int));
+  fe3 = (MEdge_ptr *) malloc((nfe1+nfe2-2)*sizeof(MEdge_ptr));
+  fedir3 = (int *) malloc((nfe1+nfe2-2)*sizeof(int));
+
+  fedges = MF_Edges(f1,1,0);
+  for (i = 0, k1 = -1; i < nfe1; i++) {
+    fe1[i] = List_Entry(fedges,i);
+    fedir1[i] = MF_EdgeDir_i(f1,i);
+    if (fe1[i] == e)
+      k1 = i;
+  }
+  List_Delete(fedges);
+
+  fedges = MF_Edges(f2,1,0);
+  for (i = 0, k2 = -1; i < nfe2; i++) {
+    fe2[i] = List_Entry(fedges,i);
+    fedir2[i] = MF_EdgeDir_i(f2,i);
+    if (fe2[i] == e)
+      k2 = i;
+  }
+  List_Delete(fedges);
+
+  if (k1 == -1 || k2 == -1) {
+    MSTK_Report("MFs_Join","Cannot find edge in face",ERROR);
+    return 0;
+  }
+
+  for (i = 0; i < nfe1-1; i++) {
+    fe3[i] = fe1[(k1+1+i)%nfe1];
+    fedir3[i] = fedir1[(k1+1+i)%nfe1];
+  }
+
[…]

### 2003-07-15 include/MEdge.h
Initial revision

@@ -1,0 +1,74 @@
+#ifndef _H_MEdge
+#define _H_MEdge
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "MSTK_types.h"
+#include "List.h"
+
+#ifdef _H_MEdge_Private
+  typedef struct MEdge {
+    int id;
+    int marker;
+    Mesh_ptr mesh;
+    char dim;
+    char gdim;
+    int gid;
+    GEntity_ptr gent;
+    RepType repType;
+    void *upadj;
+    MVertex_ptr vertex[2];
+  } MEdge, *MEdge_ptr;
+
+  /*----- Upward adjacency definitions --------*/
+
+  typedef struct MEdge_UpAdj_F1 {
+    unsigned int nf;
+    List_ptr efaces;
+  } MEdge_UpAdj_F1;
+
+  typedef struct MEdge_UpAdj_F4 {
+    unsigned int nel;
+    List_ptr elements;
+  } MEdge_UpAdj_F4;
+#else
+  typedef void *MEdge_ptr;
+#endif
+
+  MEdge_ptr ME_New(Mesh_ptr mesh);
+  void ME_Set_GEntity(MEdge_ptr medge, GEntity_ptr gent);
+  void ME_Set_GEntDim(MEdge_ptr medge, int gdim);
+  void ME_Set_GEntID(MEdge_ptr medge, int gid);
+  void ME_Set_ID(MEdge_ptr medge, int id);
+
+  void ME_Set_Vertex(MEdge_ptr medge, int i, MVertex_ptr vertex);
+  
+  int ME_ID(MEdge_ptr medge);
+  int ME_GEntDim(MEdge_ptr medge);
+  int ME_GEntID(MEdge_ptr medge);
+  GEntity_ptr ME_GEntity(MEdge_ptr medge);
+  MVertex_ptr ME_Vertex(MEdge_ptr medge, int i);
+  MVertex_ptr ME_OppVertex(MEdge_ptr e, MVertex_ptr v);
+
+  int ME_Num_Faces(MEdge_ptr medge);
+  int ME_Num_Regions(MEdge_ptr medge);
+  List_ptr ME_Faces(MEdge_ptr medge);
+  List_ptr ME_Regions(MEdge_ptr medge);
+  int ME_UsesEntity(MEdge_ptr medge, MEntity_ptr mentity, int etype);
+
+  /* Calling applications can only set the representation type for the
[…]

### 2003-07-15 include/MEntity.h

@@ -1,0 +1,46 @@
+#ifndef _H_MEntity
+#define _H_MEntity
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "MSTK_types.h"
+#include "List.h"
+
+#ifdef _H_MEntity_Private
+  typedef struct MEntity {
+    int id;
+    int marker;
+    Mesh_ptr mesh;
+    char dim;
+    char gdim;
+    int gid;
+    GEntity_ptr gent;
+  } MEntity, *MEntity_ptr;
+#else
+  typedef void *MEntity_ptr;
+#endif
+
+  void MEnt_Set_GEntity(MEntity_ptr mentity, GEntity_ptr gent);
+  void MEnt_Set_GEntDim(MEntity_ptr mentity, int gdim);
+  void MEnt_Set_ID(MEntity_ptr mentity, int id);
+
+  int MEnt_ID(MEntity_ptr mentity);
+  int MEnt_Dim(MEntity_ptr mentity);
+  Mesh_ptr MEnt_Mesh(MEntity_ptr mentity);
+  int MEnt_GEntDim(MEntity_ptr mentity);
+  GEntity_ptr MEnt_GEntity(MEntity_ptr mentity);
+
+  void MEnt_Mark(MEntity_ptr ent, int mkr);
+  int  MEnt_IsMarked(MEntity_ptr ent, int mkr);
+  void MEnt_UnMark(MEntity_ptr ent, int mkr);
+  void List_Mark(List_ptr list, int mkr);
+  void List_Unmark(List_ptr list, int mkr);
+
+  
+#ifdef __cplusplus
+}
+#endif
+
+#endif

### 2003-07-15 include/MEntity_data.h

@@ -1,0 +1,17 @@
+#ifndef _H_MEntityData
+#define _H_MEntityData
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int mentity_id;
+  Mesh_ptr mentity_mesh;
+  int mentity_gdim;
+  GEntity_ptr mentity_gent;
+
+#ifdef __cplusplus
+}
+#endif
+
+#endif

### 2003-07-15 include/MFace_jmp.h

@@ -1,0 +1,403 @@
+#define _H_MFace_Private
+
+#include "MSTK.h"
+#include "MFace.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* The following functions go through the appropriate F1 or F4
+     function to the common FN function. So in the non-Debug version
+     call the common function directly */
+
+#ifdef DEBUG
+
+void MF_Set_Edges_FN(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+void MF_Set_Edges_F1(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+void MF_Set_Edges_F4(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+void MF_Set_Edges_R1(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+void MF_Set_Edges_R2(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+void MF_Set_Edges_R4(MFace_ptr f, int n, MEdge_ptr *e, int *dir);
+static void 
+(*MF_Set_Edges_jmp[MSTK_MAXREP])(MFace_ptr f, int n, MEdge_ptr *e, int *dir) = 
+  {MF_Set_Edges_F1, MF_Set_Edges_F4, MF_Set_Edges_R1, MF_Set_Edges_R2, 
+   MF_Set_Edges_R4};
+
+void MF_Replace_Edge_FN(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
+void MF_Replace_Edge_F1(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
+void MF_Replace_Edge_F4(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
+void MF_Replace_Edge_R1(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
+void MF_Replace_Edge_R2(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
+void MF_Replace_Edge_R4(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir);
[…]

### 2003-07-15 include/MVertex.h

@@ -1,0 +1,129 @@
+#ifndef _H_MVertex
+#define _H_MVertex
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "MSTK_types.h"
+#include "List.h"
+
+#ifdef _H_MVertex_Private
+  typedef struct MVertex {
+    int id;
+    int marker;
+    Mesh_ptr mesh;
+    char dim;
+    char gdim;
+    int gid;
+    GEntity_ptr gent;
+    RepType repType;
+    void *upadj;
+    void *sameadj;
+    double xyz[3];
+  } MVertex, *MVertex_ptr;
+
+    /*----- Upward adjacency definitions --------*/
+
+  typedef struct MVertex_UpAdj_F1F4 {
+    unsigned int ne;
+    List_ptr vedges;
+  } MVertex_UpAdj_F1F4;
+
+  typedef struct MVertex_UpAdj_F2 {
+    unsigned int nr;
+    List_ptr vregions;
+  } MVertex_UpAdj_F2;
+
+  typedef struct MVertex_UpAdj_F3 {
+    unsigned int nf;
+    List_ptr vfaces;
+  } MVertex_UpAdj_F3;
+
+  typedef struct MVertex_UpAdj_F5 {
+    unsigned int ne;
+    unsigned int nf;
+    unsigned int nr;
+    List_ptr vedges;
+    List_ptr vfaces;
+    List_ptr vregions;
+  } MVertex_UpAdj_F5;
+
+  typedef struct MVertex_UpAdj_F6 {
+    unsigned int nf;
+    unsigned int nr;
+    List_ptr vfaces;
+    List_ptr vregions;
+  } MVertex_UpAdj_F6;
+
+  typedef struct MVertex_UpAdj_R1R2 {
+    unsigned int nel;
+    List_ptr velements;
+  } MVertex_UpAdj_R1R2;
+
+  typedef struct MVertex_UpAdj_R3R4 {
+    unsigned int nf;
+    List_ptr vfaces;
+  } MVertex_UpAdj_R3R4;
+
+  /*-------  Same Level adjacency definitions --------*/
+
+  typedef struct MVertex_SameAdj_R2R4 {
+    unsigned int nvadj;
+    List_ptr adjverts;
+  } MVertex_SameAdj_R2R4;
+
+  /*-------- Downward adjacency definitions ----------*/
+
+  /*    NONE     */
[…]

### 2003-07-15 include/MVertex_jmp.h

@@ -1,0 +1,213 @@
+#define _H_MVertex_Private
+
+#include "MSTK_types.h"
+#include "MVertex.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Jump tables */
+
+  void MV_Set_RepType_F1(MVertex_ptr v);
+  void MV_Set_RepType_F4(MVertex_ptr v);
+  void MV_Set_RepType_R1(MVertex_ptr v);
+  void MV_Set_RepType_R2(MVertex_ptr v);
+  void MV_Set_RepType_R4(MVertex_ptr v);
+  static void (*MV_Set_RepType_jmp[MSTK_MAXREP]) (MVertex_ptr v) = 
+  {MV_Set_RepType_F1, MV_Set_RepType_F4, MV_Set_RepType_R1, 
+   MV_Set_RepType_R2, MV_Set_RepType_R4};
+
+  void MV_Delete_F1(MVertex_ptr v);
+  void MV_Delete_F4(MVertex_ptr v);
+  void MV_Delete_R1(MVertex_ptr v);
+  void MV_Delete_R2(MVertex_ptr v);
+  void MV_Delete_R4(MVertex_ptr v);
+  static void (*MV_Delete_jmp[MSTK_MAXREP])(MVertex_ptr v) = 
+  {MV_Delete_F1, MV_Delete_F4, MV_Delete_R1, MV_Delete_R2, MV_Delete_R4};
+
+  int MV_Num_AdjVertices_F1(MVertex_ptr v);
+  int MV_Num_AdjVertices_F4(MVertex_ptr v);
+  int MV_Num_AdjVertices_R1(MVertex_ptr v);
+  int MV_Num_AdjVertices_R2(MVertex_ptr v);
+  int MV_Num_AdjVertices_R4(MVertex_ptr v);
+  static int (*MV_Num_AdjVertices_jmp[MSTK_MAXREP])(MVertex_ptr v) =
+  {MV_Num_AdjVertices_F1, MV_Num_AdjVertices_F4, MV_Num_AdjVertices_R1,
+   MV_Num_AdjVertices_R2, MV_Num_AdjVertices_R4};
+
+  int MV_Num_Edges_F1(MVertex_ptr v);
+  int MV_Num_Edges_F4(MVertex_ptr v);
[…]

### 2003-07-15 include/Mesh.h

@@ -1,0 +1,61 @@
+#ifndef _H_Mesh
+#define _H_Mesh
+
+#include "MSTK_types.h"
+#include "List.h"
+
+#ifdef _cplusplus
+extern "C" {
+#endif
+
+#ifdef _H_Mesh_Private
+  typedef struct Mesh {
+    RepType reptype;
+    int nv, ne, nf, nr;
+    List_ptr mvertex, medge, mface, mregion;
+    GModel_ptr geom;
+    int max_vid, max_eid, max_fid, max_rid;
+  } Mesh, *Mesh_ptr;
+#else
+  typedef void *Mesh;
+#endif
+
+
+
+  Mesh_ptr   MESH_New(RepType type);
+  int        MESH_InitFromFile(Mesh_ptr mesh, char *filename);
+  void       MESH_WriteToFile(Mesh_ptr mesh, char *filename);
+  
+  GModel_ptr MESH_GModel(Mesh_ptr mesh);
+  RepType    MESH_RepType(Mesh_ptr mesh);
+
+  int        MESH_Num_Vertices(Mesh_ptr mesh);
+  int        MESH_Num_Edges(Mesh_ptr mesh);
+  int        MESH_Num_Faces(Mesh_ptr mesh);
+  int        MESH_Num_Regions(Mesh_ptr mesh);
+  
+  MVertex_ptr  MESH_Vertex(Mesh_ptr mesh, int i);
+  MEdge_ptr    MESH_Edge(Mesh_ptr mesh, int i);
+  MFace_ptr    MESH_Face(Mesh_ptr mesh, int i);
+  MRegion_ptr  MESH_Region(Mesh_ptr mesh, int i);
+  
+  void       MESH_Add_Vertex(Mesh_ptr mesh, MVertex_ptr v);
+  void       MESH_Add_Edge(Mesh_ptr mesh, MEdge_ptr e);
+  void       MESH_Add_Face(Mesh_ptr mesh, MFace_ptr f);
+  void       MESH_Add_Region(Mesh_ptr mesh, MRegion_ptr r);
+
+  void       MESH_Rem_Vertex(Mesh_ptr mesh, MVertex_ptr v);
+  void       MESH_Rem_Edge(Mesh_ptr mesh, MEdge_ptr e);
+  void       MESH_Rem_Face(Mesh_ptr mesh, MFace_ptr f);
[…]

### 2003-07-15 src/base/ME_Face_Funcs_R3R4.c

@@ -1,0 +1,87 @@
+#define _H_MEdge_Private
+
+#include "MSTK_malloc.h"
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int ME_Num_Faces_R3R4(MEdge_ptr e) {
+    List_ptr efaces;
+    int nf;
+
+#ifdef DEBUG
+    MSTK_Report("ME_Num_Faces",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+
+    efaces = ME_Faces(e);
+    nf = efaces ? List_Num_Entries(efaces) : 0;
+    if (efaces) List_Delete(efaces);
+
+    return nf;
+  }
+
+  List_ptr ME_Faces_R3R4(MEdge_ptr e) {
+    int i, j, nef, nvf, nv, found;
+    List_ptr efaces, vfaces, fverts;
+    MFace_ptr vface;
+    MVertex_ptr ev[2];
+
+    nef = 0;
+    efaces = List_New(10);
+
+    for (i = 0; i < 2; i++)
+      ev[i] = ME_Vertex(e,i);
+
+    vfaces = MV_Faces(ev[0]);
+    nvf = List_Num_Entries(vfaces);
+
+    /* Go through all faces connected to ev[0] and pick out ones which
+       use this edge. Do this by checking if ev[0] and ev[1] are
+       consecutive vertices in the face */
+    for (i = 0; i < nvf; i++) {
+      vface = List_Entry(vfaces,i);
+      
+      /* Get vertices of i'th face connected to ev[0] */
+      fverts = MF_Vertices(vface,1);
+      nv = MF_Num_Vertices(vface);
+
+      /* If the j'th vertex is ev[0], check if the next or the
+         previous vertex is ev[1]; If it is then the edge is used by
+         the face */
+
+      found = 0;
+      for (j = 0; j < nv; j++) {
+	if (ev[0] == List_Entry(fverts,j) &&
+	    (ev[1] == List_Entry(fverts,(j+1)%nv) ||
+	     ev[1] == List_Entry(fverts,(j+nv-1)%nv))) {
+	  found = 1;
+	  break;
+	}
+      }
+      List_Delete(fverts);
+
+      if (found) {
+	List_Add(efaces,vface);
+	nef++;
+      }
+    }
+    List_Delete(vfaces);
+
[…]

### 2003-07-15 src/base/ME_Region_Funcs_R1R2.c

@@ -1,0 +1,25 @@
+#define _H_MEdge_Private
+
+#include "MSTK_malloc.h"
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int ME_Num_Regions_R1R2(MEdge_ptr e) {
+    MSTK_Report("ME_Num_Regions_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+  List_ptr ME_Regions_R1R2(MEdge_ptr e) {
+    MSTK_Report("ME_Regions_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/MEdge_F1.c

@@ -1,0 +1,129 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void ME_Set_RepType_F1(MEdge_ptr e) {
+    MEdge_UpAdj_F1 *upadj;
+
+    upadj = e->upadj = (MEdge_UpAdj_F1 *) MSTK_malloc(sizeof(MEdge_UpAdj_F1));
+    upadj->nf = 0;
+    upadj->efaces = List_New(10);
+  }
+
+  void ME_Delete_F1(MEdge_ptr e) {
+    MEdge_UpAdj_F1 *upadj;
+
+    upadj = (MEdge_UpAdj_F1 *) e->upadj;
+    List_Delete(upadj->efaces);
+    MSTK_free(upadj);
+
+    MV_Rem_Edge(e->vertex[0],e);
+    MV_Rem_Edge(e->vertex[1],e);
+
+    MSTK_free(e);
+  }
+
+  int ME_Num_Faces_F1(MEdge_ptr e) {
+    return ((MEdge_UpAdj_F1 *)e->upadj)->nf;
+  }
+
+  int ME_Num_Regions_F1(MEdge_ptr e) {
+    List_ptr eregs;
+    int nr;
+
+#ifdef DEBUG
+    MSTK_Report("ME_Num_Regions",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    eregs = ME_Regions_F1(e);
+    nr = eregs ? List_Num_Entries(eregs) : 0;
+    if (eregs) List_Delete(eregs);
+
+    return nr;
+  }
+
+  List_ptr ME_Faces_F1(MEdge_ptr e) {
+    return List_Copy(((MEdge_UpAdj_F1 *)e->upadj)->efaces);
+  }
+
+  List_ptr ME_Regions_F1(MEdge_ptr e) {
+    MEdge_UpAdj_F1 *upadj; 
+    int i, j, nr, mkr;
+    List_ptr eregs;
+    MFace_ptr eface;
+    MRegion_ptr freg;
+ 
+    upadj = (MEdge_UpAdj_F1 *) e->upadj;
+    nr = 0;
+    eregs = List_New(10);
+    mkr = MSTK_GetMarker();
+
+    for (i = 0; i < upadj->nf; i++) {
[…]

### 2003-07-15 src/base/MEdge_R2.c

@@ -1,0 +1,63 @@
+#define _H_MEdge_Private
+
+#include "MEdge.h"
+#include "MEdge_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void ME_Set_RepType_R2(MEdge_ptr e) {
+    MSTK_Report("ME_Set_RepType_R2","Not implemented",ERROR);
+  }
+
+  void ME_Delete_R2(MEdge_ptr e) {
+    MSTK_Report("ME_Set_RepType_R2","Not implemented",ERROR);
+  }
+
+  int ME_Num_Faces_R2(MEdge_ptr e) {
+    return ME_Num_Faces_R1R2(e);
+  }
+
+  int ME_Num_Regions_R2(MEdge_ptr e) {
+    return ME_Num_Regions_R1R2(e);
+  }
+
+  List_ptr ME_Faces_R2(MEdge_ptr e) {
+    return ME_Faces_R1R2(e);
+  }
+
+  List_ptr ME_Regions_R2(MEdge_ptr e) {
+    return ME_Regions_R1R2(e);
+  }
+
+  void ME_Add_Face_R2(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+  
+  void ME_Rem_Face_R2(MEdge_ptr e, MFace_ptr f) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Face","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Add_Region_R2(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Add_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void ME_Rem_Region_R2(MEdge_ptr e, MRegion_ptr r) {
+#ifdef DEBUG
+    MSTK_Report("ME_Rem_Region","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+#ifdef __cplusplus
+}
+#endif
+

### 2003-07-15 src/base/MEntity.c

@@ -1,0 +1,58 @@
+#define _H_MEntity_Private
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "MEntity.h"
+#include "List.h"
+#include "MSTK_malloc.h"
+#include "MSTK.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int MEnt_Dim(MEntity_ptr ent) {
+    return (ent->dim);
+  }
+
+  int MEnt_ID(MEntity_ptr ent) {
+    return (ent->id);
+  }
+
+  int MEnt_IsMarked(MEntity_ptr ent, int markerID) {
+    return (ent->marker & 1<<(markerID-1));
+  }
+
+  void MEnt_Mark(MEntity_ptr ent, int markerID) {
+    ent->marker = ent->marker | 1<<(markerID-1);
+  }
+
+  void MEnt_Unmark(MEntity_ptr ent, int markerID) {
+    ent->marker = ent->marker & ~(1<<(markerID-1));
+  }
+
+  void List_Mark(List_ptr list, int markerID) {
+    MEntity_ptr ent;
+    int i, n = List_Num_Entries(list);
+    
+    for (i = 0; i < n; i++) {
+      ent = (MEntity_ptr) List_Entry(list,i);
+      MEnt_Mark(ent,markerID);
+    }
+  }
+
+  void List_Unmark(List_ptr list, int markerID) {
+    MEntity_ptr ent;
+    int i, n = List_Num_Entries(list);
+    
+    for (i = 0; i < n; i++) {
+      ent = (MEntity_ptr) List_Entry(list,i);
+      MEnt_Unmark(ent,markerID);
+    }
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Edge_Funcs_FN.c

@@ -1,0 +1,164 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MF_Set_Edges_FN(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+    int i;
+    MFace_DownAdj_FN *downadj;
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+
+    downadj->ne = n;
+    downadj->edirs = 0;
+    downadj->fedges = List_New(n);
+    
+    for (i = 0; i < n; i++) {
+      downadj->edirs = downadj->edirs | (dir[i] << i);
+      List_Add(downadj->fedges,e[i]);
+      ME_Add_Face(e[i],f);
+    }
+  }
+
+  void MF_Replace_Edge_i_FN(MFace_ptr f, int i, MEdge_ptr e, int dir) {
+    MFace_DownAdj_FN *downadj;
+    MEdge_ptr olde;
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+
+    if (downadj->ne == 0)
+      MSTK_Report("MF_Replace_Edge_i","No initial set of edges for face",
+		  ERROR);
+
+    olde = List_Entry(downadj->fedges,i);
+      
+    downadj->edirs = (downadj->edirs & ~(1<<i)); /* set bit i to 0 */
+    downadj->edirs = (downadj->edirs | (dir<<i)); /* set to dir */
+    List_Replacei(downadj->fedges,i,e);
+
+    ME_Rem_Face(olde,f);
+    ME_Add_Face(e,f);
+  }
+
+  void MF_Replace_Edge_FN(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int dir) {
+    int i, found = 0;
+    MFace_DownAdj_FN *downadj;
+    MEdge_ptr olde;
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+
+
+    if (downadj->ne == 0)
+      MSTK_Report("MF_Replace_Edge_F1","No initial set of edges for face",
+		  ERROR);
+      
+    for (i = 0; i < downadj->ne; i++)
+      if ((olde = List_Entry(downadj->fedges,i)) == e) {
+	found = 1;
+	break;
+      }
[…]

### 2003-07-15 src/base/MF_Edge_Funcs_R1R2.c

@@ -1,0 +1,38 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int MF_Num_Edges_R1R2(MFace_ptr f) {
+    MSTK_Report("MF_Num_Edges_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+  List_ptr MF_Edges_R1R2(MFace_ptr f, int dir, MVertex_ptr v0) {
+    MSTK_Report("MF_Edges_R1R2","Not implemented",ERROR);
+    return 0;    
+  }
+
+  int MF_EdgeDir_R1R2(MFace_ptr f, MEdge_ptr e) {
+    MSTK_Report("MF_EdgeDir_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+  int MF_EdgeDir_i_R1R2(MFace_ptr f, int i) {
+    MSTK_Report("MF_EdgeDir_i_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+  int MF_UsesEdge_R1R2(MFace_ptr f, MEdge_ptr e) {
+    MSTK_Report("MF_UsesEdge_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Edge_Funcs_R3R4.c

@@ -1,0 +1,39 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  int MF_Num_Edges_R3R4(MFace_ptr f) {
+    return ((MFace_DownAdj_R3R4 *)f->downadj)->nv;
+  }
+
+  List_ptr MF_Edges_R3R4(MFace_ptr f, int dir, MVertex_ptr v0) {
+    /* Are we supposed to create edges and send them back ? */
+    MSTK_Report("MF_Edges_R3R4","Not implemented",ERROR);
+    return NULL;
+  }
+
+  int MF_EdgeDir_R3R4(MFace_ptr f, MEdge_ptr e) {
+    MSTK_Report("MF_EdgeDir_R3R4","Not Implemented",ERROR);
+    return 0;
+  }
+
+  int MF_EdgeDir_i_R3R4(MFace_ptr f, int i) {
+    MSTK_Report("MF_EdgeDir_i_R3R4","Not Implemented",ERROR);
+    return 0;
+  }
+
+  int MF_UsesEdge_R3R4(MFace_ptr f, MEdge_ptr e) {
+    MSTK_Report("MF_UsesEdge_R3R4","Not Implemented",ERROR);
+    return 0;
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Region_Funcs_R1R2.c

@@ -1,0 +1,24 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  List_ptr MF_Regions_R1R2(MFace_ptr f) {
+    MSTK_Report("MF_Regions_R1R2","Not implemented",ERROR);
+    return 0;    
+  }
+
+  MRegion_ptr MF_Region_R1R2(MFace_ptr f, int dir) {
+    MSTK_Report("MF_Region_R1R2","Not implemented",ERROR);
+    return 0;
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Region_Funcs_R3R4.c

@@ -1,0 +1,62 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  List_ptr MF_Regions_R3R4(MFace_ptr f) {
+    List_ptr fregs;
+    MFace_UpAdj_R3R4 *upadj;
+
+    upadj = (MFace_UpAdj_R3R4 *) f->upadj;
+
+    fregs = List_New(2);
+    if (upadj->fregions[0])
+      List_Add(fregs,upadj->fregions[0]);
+    if (upadj->fregions[1])
+      List_Add(fregs,upadj->fregions[1]);
+
+    return fregs;
+  }
+
+  MRegion_ptr MF_Region_R3R4(MFace_ptr f, int i) {
+    return ((MFace_UpAdj_R3R4 *) f->upadj)->fregions[i];
+  }
+
+  void MF_Add_Region_R3R4(MFace_ptr f, MRegion_ptr r, int side) {
+    MFace_UpAdj_R3R4 *upadj;
+
+    upadj = (MFace_UpAdj_R3R4 *) f->upadj;
+
+    if (side == MSTK_UNKNOWN) {
+      if (MR_FaceDir(r,f))
+	upadj->fregions[0] = r;
+      else
+	upadj->fregions[1] = r;
+    }
+    else
+      upadj->fregions[side] = r;
+  }
+
+  void MF_Rem_Region_R3R4(MFace_ptr f, MRegion_ptr r) {
+    MFace_UpAdj_R3R4 *upadj;
+
+    upadj = (MFace_UpAdj_R3R4 *) f->upadj;
+
+    if (upadj->fregions[0] == r)
+      upadj->fregions[0] = NULL;
+    else if (upadj->fregions[1] == r)
+      upadj->fregions[1] = NULL;
+    else
+      MSTK_Report("MF_Rem_Region_R3R4","Region not connected to face",ERROR);
+  }
+
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MF_Vertex_Funcs_R3R4.c

@@ -1,0 +1,64 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MF_Set_Vertices_R3R4(MFace_ptr f, int n, MVertex_ptr *v) {
+    MFace_DownAdj_R3R4 *downadj;
+    int i;
+
+    downadj = f->downadj;
+    downadj->nv =n;
+    downadj->fvertices = List_New(n);
+
+    for (i = 0; i < n; i++)
+      List_Add(downadj->fvertices,v[i]);
+  }
+
+  void MF_Replace_Vertex_i_R3R4(MFace_ptr f, int i, MVertex_ptr v) {
+    MFace_DownAdj_R3R4 *downadj;
+
+    downadj = f->downadj;
+    if (downadj->nv == 0)
+      MSTK_Report("MF_Replace_Vertex_R3R4","No initial set of vertices for face",ERROR);
+
+    List_Replacei(downadj->fvertices,i,v);
+  }
+
+  void MF_Replace_Vertex_R3R4(MFace_ptr f, MVertex_ptr v, MVertex_ptr nuv) {
+    MFace_DownAdj_R3R4 *downadj;
+
+    downadj = f->downadj;
+    if (downadj->nv == 0)
+      MSTK_Report("MF_Replace_Vertex_R3R4","No initial set of vertices for face",ERROR);
+
+    List_Replace(downadj->fvertices,v,nuv);
+  }
+
+  int MF_Num_Vertices_R3R4(MFace_ptr f) {
+    MFace_DownAdj_R3R4 *downadj;
+    downadj = (MFace_DownAdj_R3R4 *) f->downadj;
+    return downadj->nv;
+  }
+
+  List_ptr MF_Vertices_R3R4(MFace_ptr f, int dir) {
+    MFace_DownAdj_R3R4 *downadj;
+    downadj = (MFace_DownAdj_R3R4 *) f->downadj;
+    return List_Copy(downadj->fvertices);
+  }
+	
+  int MF_UsesVertex_R3R4(MFace_ptr f, MVertex_ptr v) {
+    MFace_DownAdj_R3R4 *downadj;
+    downadj = (MFace_DownAdj_R3R4 *) f->downadj;
+    return List_Contains(downadj->fvertices,v);
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MFace.c

@@ -1,0 +1,188 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  MFace_ptr MF_New(Mesh_ptr mesh) {
+    MFace_ptr f;
+    RepType rtype;
+
+    f = (MFace_ptr) MSTK_malloc(sizeof(MFace));
+
+    f->id = 0;
+    f->marker = 0;
+    f->mesh = mesh;
+    f->dim = 2;
+    f->gdim = 4; /* nonsensical value since we don;t know what it is */
+    f->gid = 0;
+    f->gent = (GEntity_ptr) NULL;
+    f->upadj = (void *) NULL;
+    f->sameadj = (void *) NULL;
+    f->downadj = (void *) NULL;
+
+    rtype = mesh ? MESH_RepType(mesh) : F1;
+    MF_Set_RepType(f,rtype);
+    
+    if (mesh) MESH_Add_Face(mesh,f);
+
+    return f;
+  }
+
+  void MF_Delete(MFace_ptr f) {
+    MESH_Rem_Face(f->mesh,f);
+    (*MF_Delete_jmp[f->repType])(f);
+  }
+
+  void MF_Set_RepType(MFace_ptr f, RepType rtype) {
+    f->repType = rtype;
+    (*MF_Set_RepType_jmp[f->repType])(f);
+  }
+
+  void MF_Set_GEntity(MFace_ptr f, GEntity_ptr gent) {
+    f->gent = gent;
+  }
+
+  void MF_Set_GEntDim(MFace_ptr f, int gdim) {
+    f->gdim = gdim;
+  }
+
+  void MF_Set_GEntID(MFace_ptr f, int gid) {
+    f->gid = gid;
+  }
+
+  void MF_Set_ID(MFace_ptr f, int id) {
+    f->id = id;
+  }
+
+  void MF_Set_Edges(MFace_ptr f, int n, MEdge_ptr *edges, int *dir) {
+    (*MF_Set_Edges_jmp[f->repType])(f,n,edges,dir);
+  }
+
+  void MF_Replace_Edge(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir) {
[…]

### 2003-07-15 src/base/MFace_F1.c

@@ -1,0 +1,166 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Most of these functions are identical to those of F4 */
+
+  void MF_Set_RepType_F1(MFace_ptr f) {
+    MFace_UpAdj_F1F3 *upadj;
+    MFace_DownAdj_FN *downadj;
+
+    upadj = f->upadj = (MFace_UpAdj_F1F3 *) MSTK_malloc(sizeof(MFace_UpAdj_F1F3));
+    upadj->fregions[0] = (MRegion_ptr) NULL;
+    upadj->fregions[1] = (MRegion_ptr) NULL;
+    downadj = f->downadj = (MFace_DownAdj_FN *) MSTK_malloc(sizeof(MFace_DownAdj_FN));
+    downadj->ne = 0;
+    downadj->edirs = 0;
+    downadj->fedges = NULL;
+  }
+
+  void MF_Delete_F1(MFace_ptr f) {
+    MFace_UpAdj_F1F3 *upadj;
+    MFace_DownAdj_FN *downadj;
+    MEdge_ptr e;
+    int i, ne;
+
+    upadj = (MFace_UpAdj_F1F3 *) f->upadj;
+    MSTK_free(upadj);
+
+    downadj = (MFace_DownAdj_FN *) f->downadj;
+
+    ne = List_Num_Entries(downadj->fedges);
+    for (i = 0; i < ne; i++) {
+      e = List_Entry(downadj->fedges,i);
+      ME_Rem_Face(e,f);
+    }
+    List_Delete(downadj->fedges);
+    MSTK_free(downadj);
+
+    MSTK_free(f);
+  }
+
+  void MF_Set_Edges_F1(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+    MF_Set_Edges_FN(f,n,e,dir);
+  }
+
+  void MF_Replace_Edge_F1(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir) {
+    MF_Replace_Edge_FN(f,e,nue,nudir);
+  }
+
+  void MF_Replace_Edge_i_F1(MFace_ptr f, int i, MEdge_ptr nue, int nudir) {
+    MF_Replace_Edge_i_FN(f,i,nue,nudir);
+  }
+
[…]

### 2003-07-15 src/base/MFace_R2.c

@@ -1,0 +1,125 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_malloc.h"
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MF_Set_RepType_R2(MFace_ptr f) {
+    /* What info should be created for virtual entities */
+    MSTK_Report("MF_Set_RepType_R2","Not Implemented",ERROR);
+  }
+
+  void MF_Delete_R2(MFace_ptr f) {
+    MSTK_Report("MF_Delete_R2","Not Implemented",ERROR);
+  }
+
+  void MF_Set_Edges_R2(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Set_Edges","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_R2(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_i_R2(MFace_ptr f, int i, MEdge_ptr e, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Set_Vertices_R2(MFace_ptr f, int n, MVertex_ptr *v) {
+    MSTK_Report("MF_Set_Vertices","Not implemented for this representation",WARN);
+  }
+
+  void MF_Replace_Vertex_R2(MFace_ptr f, MVertex_ptr v, MVertex_ptr nuv) {
+    MSTK_Report("MF_Replace_Vertex","Not implemented for this representation",WARN);
+  }
+
+  void MF_Replace_Vertex_i_R2(MFace_ptr f, int i, MVertex_ptr v) {
+    MSTK_Report("MF_Replace_Vertex","Not implemented for this representation",WARN);
+  }
+
+  void MF_Add_Region_R2(MFace_ptr f, MRegion_ptr r, int side) {
[…]

### 2003-07-15 src/base/MFace_R4.c

@@ -1,0 +1,144 @@
+#define _H_MFace_Private
+
+#include "MFace.h"
+#include "MFace_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+ 
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  void MF_Set_RepType_R4(MFace_ptr f) {
+    MFace_UpAdj_R3R4 *upadj;
+    MFace_DownAdj_R3R4 *downadj;
+
+    upadj = f->upadj =(MFace_UpAdj_R3R4 *) MSTK_malloc(sizeof(MFace_UpAdj_R3R4));
+    upadj->fregions[0] = upadj->fregions[1] = (MRegion_ptr) NULL;
+    downadj = (MFace_DownAdj_R3R4 *) MSTK_malloc(sizeof(MFace_DownAdj_R3R4));
+    downadj->nv = 0;
+    downadj->fvertices = NULL;
+  }
+
+  void MF_Delete_R4(MFace_ptr f) {
+    MFace_UpAdj_R3R4 *upadj;
+    MFace_DownAdj_R3R4 *downadj;
+
+    upadj  = (MFace_UpAdj_R3R4 *) f->upadj;
+    MSTK_free(upadj);
+
+    downadj = (MFace_DownAdj_R3R4 *) f->downadj;
+    List_Delete(downadj->fvertices);
+    MSTK_free(downadj);
+  }
+  
+
+  void MF_Set_Edges_R4(MFace_ptr f, int n, MEdge_ptr *e, int *dir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Set_Edges",
+		"Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_i_R4(MFace_ptr f, int i, MEdge_ptr nue, int nudir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge",
+		"Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MF_Replace_Edge_R4(MFace_ptr f, MEdge_ptr e, MEdge_ptr nue, int nudir) {
+#ifdef DEBUG
+    MSTK_Report("MF_Replace_Edge",
+		"Function call not suitable for this representation",WARN);
+#endif
+  }
+
[…]

### 2003-07-15 src/base/MRegion.c

@@ -1,0 +1,207 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+ int mrtype_nv[7]={0,0,4,5,6,8,-1};
+ int mrtype_ne[7]={0,0,6,8,9,12,-1};
+ int mrtype_nf[7]={0,0,4,5,5,6,-1};
+
+  MRegion_ptr MR_New(Mesh_ptr mesh) {
+    MRegion_ptr r;
+    RepType rtype;
+
+    r = (MRegion_ptr) MSTK_malloc(sizeof(MRegion));
+
+    r->id = 0;
+    r->marker = 0;
+    r->mesh = mesh;
+    r->dim = 3;
+    r->gdim = 3;
+    r->gid = 0;
+    r->gent = (GEntity_ptr) NULL;
+    r->sameadj = (void *) NULL;
+    r->downadj = (void *) NULL;
+
+    rtype = mesh ? MESH_RepType(mesh) : F1;
+    MR_Set_RepType(r,rtype);
+
+    if (mesh) MESH_Add_Region(mesh,r);
+
+    return r;
+  }
+
+  void MR_Delete(MRegion_ptr r) {
+    MESH_Rem_Region(r->mesh,r);
+    (*MR_Delete_jmp[r->repType])(r);
+  }
+
+  void MR_Set_RepType(MRegion_ptr r, RepType rtype) {
+    r->repType = rtype;
+    (*MR_Set_RepType_jmp[r->repType])(r);
+  }
+
+  void MR_Set_GEntity(MRegion_ptr r, GEntity_ptr gent) {
+    r->gent = gent;
+  }
+
+  void MR_Set_GEntDim(MRegion_ptr r, int gdim) {
+    r->gdim = gdim;
+  }
+
+  void MR_Set_GEntID(MRegion_ptr r, int gid) {
+    r->gid = gid;
+  }
+
+  void MR_Set_ID(MRegion_ptr r, int id) {
+    r->id = id;
+  }
+
+  void MR_Set_Faces(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs) {
+    (*MR_Set_Faces_jmp[r->repType])(r,nf,mfaces,dirs);
+  }
+
[…]

### 2003-07-15 src/base/MRegion_R1.c

@@ -1,0 +1,164 @@
+#define _H_MRegion_Private
+
+#include "MRegion.h"
+#include "MRegion_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  void MR_Set_RepType_R1(MRegion_ptr r) {
+    MRegion_DownAdj_R1R2 *downadj;
+
+    r->downadj = (MRegion_DownAdj_R1R2 *) MSTK_malloc(sizeof(MRegion_DownAdj_R1R2));
+    downadj = r->downadj;
+    downadj->nv = (unsigned char) 0;
+    downadj->rvertices = NULL;
+  }
+
+  void MR_Delete_R1(MRegion_ptr r) {
+    MRegion_DownAdj_R1R2 *downadj;
+
+    downadj = (MRegion_DownAdj_R1R2 *) r->downadj;
+    List_Delete(downadj->rvertices);
+    MSTK_free(downadj);
+  }
+
+  void MR_Set_Vertices_R1(MRegion_ptr r, int nv, MVertex_ptr *rvertices) {
+    int i;
+    MRegion_DownAdj_R1R2 *downadj;
+
+    downadj = (MRegion_DownAdj_R1R2 *) r->downadj;
+    downadj->nv = nv;
+    downadj->rvertices = List_New(nv);
+    for (i = 0; i < nv; i++)
+      List_Add(downadj->rvertices,rvertices[i]);
+  }
+
+  void MR_Set_Faces_R1(MRegion_ptr r, int nf, MFace_ptr *mfaces, int *dirs) {
+#ifdef DEBUG
+    MSTK_Report("MR_Set_Faces_R1","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MR_Replace_Face_R1(MRegion_ptr r, MFace_ptr f, MFace_ptr nuf, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MR_Replace_Face_R1","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  void MR_Replace_Face_i_R1(MRegion_ptr r, int i, MFace_ptr nuf, int dir) {
+#ifdef DEBUG
+    MSTK_Report("MR_Replace_Face_i_R1","Function call not suitable for this representation",WARN);
+#endif
+  }
+
+  int MR_Num_Faces_R1(MRegion_ptr r) {
+    return mrtype_nf[MR_ElementType(r)];
+  }
+
[…]

### 2003-07-15 src/base/MV_Edge_Funcs_R2R4.c

@@ -1,0 +1,33 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  /* Functions */
+
+
+  List_ptr MV_Edges_R2R4(MVertex_ptr v) {
+    int i;
+    MVertex_ptr adjv;
+    List_ptr vedges;
+    MVertex_SameAdj_R2R4 *sameadj;
+
+    /* Have to create volatile edges */
+
+    vedges = List_New(sameadj->nvadj);
+    for (i = 0; i < sameadj->nvadj; i++) {
+      /* create a volatile edge from v and the adjacent vertex */
+      adjv = List_Entry(sameadj->adjverts,i);
+    }
+
+    return vedges;
+  }
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/base/MVertex.c

@@ -1,0 +1,156 @@
+#define _H_MVertex_Private
+
+#include <string.h>
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  MVertex_ptr MV_New(Mesh_ptr mesh) {
+    MVertex_ptr v;
+    RepType rtype;
+    
+    v = (MVertex_ptr) MSTK_malloc(sizeof(MVertex));
+
+    v->id = 0;
+    v->marker = 0;
+    v->mesh = mesh;
+    v->dim  = 0;
+    v->gdim = 4; /* Nonsensical value since we don't know what it is */
+    v->gid  = 0;
+    v->gent = (GEntity_ptr) NULL;
+    v->xyz[0] = v->xyz[1] = v->xyz[2] = 0.0;
+    v->upadj = NULL;
+    v->sameadj = NULL;
+
+    rtype = mesh ? MESH_RepType(mesh) : F1;
+    MV_Set_RepType(v,rtype);
+
+    if (mesh) MESH_Add_Vertex(mesh,v);
+
+    return v;
+  } 
+
+  void MV_Delete(MVertex_ptr v) {
+    MESH_Rem_Vertex(v->mesh,v);
+    (*MV_Delete_jmp[v->repType])(v);
+  }
+
+  void MV_Set_RepType(MVertex_ptr v, RepType rtype) {
+    v->repType = rtype;
+    (*MV_Set_RepType_jmp[rtype])(v);
+  }
+
+  void MV_Set_Coords(MVertex_ptr v, double *xyz) {
+    v->xyz[0] = xyz[0];
+    v->xyz[1] = xyz[1];
+    v->xyz[2] = xyz[2];
+  }
+
+  void MV_Set_GEntity(MVertex_ptr v, GEntity_ptr gent) {
+    v->gent = gent;
+  }
+
+  void MV_Set_GEntDim(MVertex_ptr v, int gdim) {
+    v->gdim = gdim;
+  }
+
+  void MV_Set_GEntID(MVertex_ptr v, int gid) {
+    v->gid = gid;
+  }
+
+  void MV_Set_ID(MVertex_ptr v, int id) {
+    v->id = id;
+  }
+
[…]

### 2003-07-15 src/base/MVertex_F1.c

@@ -1,0 +1,250 @@
+#define _H_MVertex_Private
+
+#include "MVertex.h"
+#include "MVertex_jmp.h"
+#include "MSTK_private.h"
+#include "MSTK_malloc.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
+  void MV_Set_RepType_F1(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+
+    upadj = v->upadj = (MVertex_UpAdj_F1F4 *) MSTK_malloc(sizeof(MVertex_UpAdj_F1F4));
+    upadj->ne = (unsigned int) 0;
+    upadj->vedges = List_New(10);
+  }
+
+  void MV_Delete_F1(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+
+    upadj = (MVertex_UpAdj_F1F4 *) v->upadj;
+    List_Delete(upadj->vedges);
+    MSTK_free(upadj);
+
+    MSTK_free(v);
+  }
+
+  int MV_Num_AdjVertices_F1(MVertex_ptr v) {
+    return ((MVertex_UpAdj_F1F4 *) v->upadj)->ne; 
+  }
+
+  int MV_Num_Edges_F1(MVertex_ptr v) {
+    return ((MVertex_UpAdj_F1F4 *) v->upadj)->ne;
+  }
+
+  int MV_Num_Faces_F1(MVertex_ptr v) {
+    List_ptr vfaces;
+    int nf;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Faces",
+		"Ineficient to call this routine with this representation",
+		MESG);
+#endif
+    
+    vfaces = MV_Faces_F1(v);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
+    return nf;
+  }
+
+  int MV_Num_Regions_F1(MVertex_ptr v) {
+    int nr;
+    List_ptr vregions;
+
+#ifdef DEBUG
+    MSTK_Report("MV_Num_Regions",
+		"Inefficient to call this routine with this representation",
+		MESG);
+#endif
+	
+    vregions = MV_Regions_F1(v);
+    nr = List_Num_Entries(vregions);
+    List_Delete(vregions);
+
+    return nr;
+  }
+
+  List_ptr MV_AdjVertices_F1(MVertex_ptr v) {
+    MVertex_UpAdj_F1F4 *upadj;
+    List_ptr vedges, adjv;
[…]

### 2003-07-15 src/base/Mesh.c

@@ -1,0 +1,957 @@
+#define _H_Mesh_Private
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "Mesh.h"
+#include "MSTK.h"
+#include "List.h"
+#include "MSTK_malloc.h"
+
+#include "MSTK_private.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+  int MESH_rtype[MSTK_MAXREP] = {F1, F2, F3, F4, F5, F6, R1, R2, R3, R4};
+  char MESH_rtype_str[10][3] = {"F1\0","F2\0","F3\0","F4\0","F5\0","F6\0","R1\0","R2\0","R3\0","R4\0"};
+
+Mesh_ptr MESH_New(RepType type) {
+  Mesh_ptr newmesh;
+
+  newmesh = (Mesh_ptr) MSTK_malloc(sizeof(Mesh));
+
+  newmesh->reptype = type;
+  newmesh->nv = newmesh->ne = newmesh->nf = newmesh->nr = 0;
+  newmesh->mvertex = (List_ptr) NULL;
+  newmesh->medge = (List_ptr) NULL;
+  newmesh->mface = (List_ptr) NULL;
+  newmesh->mregion = (List_ptr) NULL;
+  newmesh->geom = (GModel_ptr) NULL;
+
+  newmesh->max_vid = newmesh->max_eid = newmesh->max_fid = newmesh->max_rid = 0;
+
+  return newmesh;
+}
+
+
+void MESH_Delete(Mesh_ptr mesh) {
+  int i, nv, ne, nf, nr;
+  MVertex_ptr mv;
+  MEdge_ptr me;
+  MFace_ptr mf;
+  MRegion_ptr mr;
+
+  if (mesh->mregion) {
+    nr = mesh->nr;
+    i = 0;
+    while (mr = List_Next_Entry(mesh->mregion,&i)) {
+      MR_Delete(mr);
+    }
+    List_Delete(mesh->mregion);
+  }
+  if (mesh->mface) {
+    nf = mesh->nf;
+    i = 0;
+    while (mf = List_Next_Entry(mesh->mface,&i)) {
+      MF_Delete(mf);
+    }
+    List_Delete(mesh->mface);
+  }
+  if (mesh->medge) {
+    ne = mesh->ne;
+    i = 0;
+    while (me = List_Next_Entry(mesh->medge,&i)) {
+      ME_Delete(me);
+    }
+    List_Delete(mesh->medge);
[…]

### 2003-07-15 src/hilev/MR_Coords.c

@@ -1,0 +1,26 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "MSTK.h"
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+void MR_Coords(MRegion_ptr r, int *n, double (*xyz)[3]) {
+  int i;
+  MVertex_ptr rv;
+  List_ptr rverts;
+
+  rverts = MR_Vertices(r);
+  *n = List_Num_Entries(rverts);
+
+  for (i = 0; i < *n; i++) {
+    rv = List_Entry(rverts,i);
+    MV_Coords(rv,xyz[i]);
+  }
+  List_Delete(rverts);
+}
+
+#ifdef __cplusplus
+}
+#endif

### 2003-07-15 src/misc/List_PrintID.c

@@ -1,0 +1,25 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "MSTK.h"
+
+#ifdef DEBUG
+
+  void List_PrintID(List_ptr l) {
+    int i, k, num;
+    MEntity_ptr entry;
+
+    num = List_Num_Entries(l);
+
+    fprintf(stderr,"List 0x%x:\n",l);
+    for (i = 0, k = 0; i < num; i++) {
+      entry = List_Entry(l,i);
+
+      fprintf(stderr,"%d ",MEnt_ID(entry));
+      k++;
+      if (k%10 == 0)
+	fprintf(stderr,"\n");
+    }
+    fprintf(stderr,"\n");
+  }
+
+#endif

### 2003-11-17 docs/MSTK.tex
Added PDF and PS files to the repository Removed the .aux and .log files

@@
 
-\title{MSTK: Mesh Toolkit, v 1.0}
+\title{MSTK: Mesh Toolkit, v 1.1 - DRAFT}
 
@@
 \thispagestyle{empty}
+\setlength{\parindent}{0.0in}
+\setlength{\parskip}{0.5em}
 
 \newpage
+\section{Introduction}
 
-\section{MSTK IO File Format}
+MSTK or Mesh Toolkit is a mesh framework that allows users to
+represent, manipulate and query unstructured 3D arbitrary topology
+meshes in a general manner without the need to code their own data
+structures. MSTK is a flexible framework in that it allows (or will
+eventually allow) a wide variety of underlying representations for the
+mesh while maintaining a common interface. It will allow users to
+choose from different mesh representations either at initialization or
+during the program execution so that the optimal data structures are
+used for the particular algorithm. The interaction of users and
+applications with MSTK is through a functional interface that acts as
+though the mesh always contains vertices, edges, faces and regions and
+maintains connectivity between all these entities.
 
-\subsection{ASCII File Format}
+\par MSTK allows for the simultaneous existence of an arbitrary number of
+meshes. However, any entity in MSTK can belong to only one mesh at a
+time.
+
+\par MSTK will eventually support distributed meshes for parallel
+computing. However, this is still not in place.
+
+\par MSTK will soon allow applications to attach application or field data
+to entities. This data may be integers, reals (doubles), integer
+vectors, real (double) vectors, integer tensors, real (double tensors)
+and pointers.
+
+\par The basis for development of MSTK is laid out in the following paper:
+
+Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
+FEA Applications,'' \textit{International Journal of Numerical Methods
+  in Engineering}, v55 n4, pp. 441-478, 2002.
+
+In the following sections, the data types of MSTK will be described
+followed by a description of the functional interface. The MSTK file
+format will be described in the last section.
+
+\newpage
+\section{MSTK Data Types}
+
+\begin{description}
+\item[]\textbf{\textit{Set\_ptr}}: Handle to a Set object
+
+\item[]\textbf{\textit{Mesh\_ptr}}: Handle to a Mesh object.
+
+\item[]\textbf{\textit{MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
+Dimension 0)
+
+\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1)
+
+\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2)
+
+\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3)
+
+\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity object. Any of
+the above types of entities can be cast as \textit{MEntity\_ptr}
+
+\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object
+\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object
+
+\item[]\textbf{\textit{RepType}}: Enumerated type describing the type
+  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
+  See Appendix~\ref{app:reptypes} for schematics of these
+  representations \textit{Currently only representation types F1 and
+    F4 are supported}
+
+\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type
+Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON
+
+\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type
+Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED
+\end{description}
+
+\newpage
+\section{MSTK Functional Interface}
+\subsection{Sets}
+
+Sets of entities in MSTK are returned as sets of type
+\textit{Set\_ptr}. The following are the set operations available in
+MSTK:
+
+\begin{description}
+\item[]\textbf{\textit{Set\_ptr} Set\_New(\textit{int} inisize):} Create a
+new set with an initial size, \textit{inisize}. If \textit{inisize} is
+0, the initial size is set to be 10.
+
+\item[]\textbf{\textit{void} Set\_Delete(\textit{Set\_ptr} l):} Delete a set.
+
+\item[]\textbf{\textit{Set\_ptr} Set\_Compress(\textit{Set\_ptr} l):} Compress
+a set.  Doing this while an algorithm is iterating through the set can
+currently cause problems!! Calling \textbf{Set\_Compress} could change the
+pointer for the set due to reallocation.
+
+\item[]\textbf{\textit{Set\_ptr} Set\_Copy(\textit{Set\_ptr} l):} Return a
+copy of a set.
+
+\item[]\textbf{\textit{Set\_ptr} Set\_Add(\textit{Set\_ptr} l, \textit{void
+  *}entry):} Add an entry to the set. The entry is appended to the end
+of the set.
+
+\item[]\textbf{\textit{Set\_ptr} Set\_ChknAdd(\textit{Set\_ptr} l,
+\textit{void *}entry):} Add an entry to a set only if it is not already
+in the set.
+
+\item[]\textbf{\textit{int} Set\_Rem(\textit{Set\_ptr} l, \textit{void
+  *}entry):} Remove an entry from the set. Returns 1 if successful, 0
+otherwise.
+
+\item[]\textbf{\textit{int} Set\_Remi(\textit{Set\_ptr} l, \textit{int} i):}
+Remove the i'th valid entry in the set. Returns 1 if successful, 0
+otherwise.
+
+\item[]\textbf{\textit{int} Set\_Replace(\textit{Set\_ptr} l, \textit{void
+  *}entry, \textit{void *}nuentry):} Replace 'entry' with 'nuentry' in
+set. Returns 1 if successful, 0 otherwise.
+
+\item[]\textbf{\textit{int} Set\_Replacei(\textit{Set\_ptr} l, \textit{int}
+i, void *nuentry):} Replace the i'th valid entry in the set with
+'nuentry'. Returns 1 if successful, 0 otherwise.
+
+\item[]\textbf{\textit{int} Set\_Contains(\textit{Set\_ptr} l, void *entry):}
+Returns 1 if set contains the entry, 0 otherwise.
+
+\item[]\textbf{\textit{int} Set\_Locate(\textit{Set\_ptr} l, void *entry):}
+Returns the positional index of the entry in the set. Returns -1 if
+the set does not contain the entry.
+
+\item[]\textbf{\textit{void *}Set\_Entry(\textit{Set\_ptr} l, \textit{int}
+i):} Return the i'th valid entry in the set.  Returns a NULL pointer if
+the i'th valid entry could not be found.
+
+\item[]\textbf{\textit{void *}Set\_Next\_Entry(\textit{Set\_ptr} l,
+\textit{int} *i):} Return the next valid entry in the set. This routine
+works like an iterator. To start iterating through the set, set the
+iteration index i=0 and call the routine to get the first entry in the
+set. Subsequent calls to the routine will iterate through the entries
+in the set. The routine will return a NULL to indicate that the end of
+the set is reached.
+
+The value of the iteration index i will be modified by the routine on
+each call to indicate where in the set it is. This value should not be
+modified externally while iterating through the set. Also, no specific
+meaning be derived from from the iteration index by other applications
+since the internal implementation and interpretation of the index may
+change at any time.
+
+\item[]\textbf{\textit{int} Set\_Num\_Entries(\textit{Set\_ptr} l):} Return
+the number of entries in a set
+\end{description}
+
+\newpage
+\subsection{Mesh Object}
+
+A mesh object is a set of vertices (nodes) possibly connected by other
+entities such as edges, faces, regions. Depending on the
+representation chosen and type of mesh, some or all of the entities
+may be explicitly stored.  Full representations contain all types of
+entities up to the highest dimension of the mesh. For example, a full
+representation of a tetrahedral mesh contains vertices, edges, faces
+and regions. However, one type of reduced representation of this mesh
+may contain only vertices and regions. For a surface mesh, a full
+representation includes vertices, edges and faces while a reduced
+representation only has vertices and faces. Also, depending on the
+type of representation, some adjacencies (information about which
+entities are connected to which other entities) are stored and others
+are derived.
+
+\begin{description}
+\item[]\textbf{\textit{Mesh\_ptr} MESH\_New(\textit{RepType} type):}
+Initialize a new mesh object with the given representation type which
+can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Not all of these types
+are implemented. If the representation type is not known at the
+present time (e.g. before reading the mesh from a file), the
+representation type of UNKNOWN\_REP can be specified.  Note that this
+only initializes a mesh object, it does not create or generate a mesh
+which is the work of high level mesh generation routines.
+
+\item[]\textbf{\textit{int} MESH\_InitFromFile(\textit{Mesh\_ptr} mesh,
+const char *filename):} Initialize or read a mesh from file into the
+given mesh object. Returns 1 if successful, 0 otherwise.
+
+\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr} mesh,
+const char *filename):} Save a mesh to a filename. The file is created
+if it does not exists. It is recommended that the .mstk extension be
+used for MSTK mesh files.  However, there is no such requirement.
+
+\item[]\textbf{\textit{GModel\_ptr} MESH\_GModel(\textit{Mesh\_ptr} mesh):}
+Return a handle to the underlying geometric model. If there is no
+geometric model associated with the mesh, NULL pointer is returned.
+
+\item[]\textbf{\textit{RepType} MESH\_RepType(\textit{Mesh\_ptr} mesh):}
+Representation type currently being used by the mesh.
+
+\item[]\textbf{\textit{int} MESH\_Num\_Vertices(\textit{Mesh\_ptr} mesh):}
+Number of vertices in the mesh.
+
+\item[]\textbf{\textit{int} MESH\_Num\_Edges(\textit{Mesh\_ptr} mesh):}
+Number of edges in the mesh. For reduced representations, this routine
+returns 0 since it is impractically expensive to count the number of
+edges when they do not explicitly exist. Applications must find a way
+to avoid using this routine for reduced representations.
+
+\item[]\textbf{\textit{int} MESH\_Num\_Faces(\textit{Mesh\_ptr} mesh):}
+Number of faces in the mesh. For reduced representations R1 or R2,
+this routine counts only the faces that are explicitly represented
+i.e. faces not connected to any mesh region. Therefore, a value of 0
+will be returned for the number of faces of a tetrahedral mesh with
+representation R1 or R2 but the correct number will be reported for a
+tetrahedral mesh in other representations. Also, the correct number
+will be reported for the number of faces in a surface mesh in
+representation R1 or R2.  Therefore, this routine must be used
+carefully.
+
+\item[]\textbf{\textit{int} MESH\_Num\_Regions(\textit{Mesh\_ptr} mesh):}
+Number of regions in the mesh.
+
+\item[]\textbf{\textit{MVertex\_ptr} MESH\_Vertex(\textit{Mesh\_ptr} mesh,
+\textit{int} i):} Return the i'th vertex in the mesh. Returns NULL if i
+$<$ 0 or i $>$ number of mesh vertices.
+
+\item[]\textbf{\textit{MEdge\_ptr} MESH\_Edge(\textit{Mesh\_ptr} mesh,
+\textit{int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
+0 or i $>$ number of mesh edges. Returns NULL for reduced
+representations.
+
+\item[]\textbf{\textit{MFace\_ptr} MESH\_Face(\textit{Mesh\_ptr} mesh,
+\textit{int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
+0 or i $>$ number of mesh faces. Only faces explicitly represented in
+the mesh are returned for reduced representation (See explanation for
+MESH\_Num\_Faces).
+
+\item[]\textbf{\textit{MRegion\_ptr} MESH\_Region(\textit{Mesh\_ptr} mesh,
+\textit{int} i):} Return the i'th region in the mesh. Returns NULL if i
+$<$ 0 or i $>$ number of mesh region.
+
+\item[]\textbf{\textit{MVertex\_ptr} MESH\_Next\_Vertex(\textit{Mesh\_ptr}
+mesh, \textit{int} *idx):} Returns the next vertex while iterating
+through the vertices of the mesh. See the routine \textbf{Set\_Next\_Entry}
+above for an explanation of how the iteration works.
+
+\item[]\textbf{\textit{MEdge\_ptr} MESH\_Next\_Edge(\textit{Mesh\_ptr} mesh,
+\textit{int} *idx):} Returns the next edge while iterating through the
+edges of the mesh. See the routine Set\_Next\_Entry above for an
+explanation of how the iteration works.  The routine always returns
+NULL for reduced representations.
+
+\item[]\textbf{\textit{MFace\_ptr} MESH\_Next\_Face(\textit{Mesh\_ptr} mesh,
+\textit{int} *idx):} Returns the next face while iterating through the
+faces of the mesh. See the routine \textbf{Set\_Next\_Entry} above for
+an explanation of how the iteration works.  Only faces explicitly
+represented in the mesh are returned for reduced representation (See
+explanation for \textbf{MESH\_Num\_Faces}).
+
+\item[]\textbf{\textit{void} MESH\_Add\_Vertex(\textit{Mesh\_ptr} mesh,
+\textit{MVertex\_ptr} v):} Add a vertex to the mesh. It is assumed that
+the vertex and its coordinates set are properly defined.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Edge(\textit{Mesh\_ptr} mesh,
+\textit{MEdge\_ptr} e):} Add an edge to the mesh. It is assumed that
+the edge is and its topology is defined.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Face(\textit{Mesh\_ptr} mesh,
+\textit{MFace\_ptr} f):} Add a face to the mesh. It is assumed that the
+face and its topology is properly defined.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Region(\textit{Mesh\_ptr} mesh,
+\textit{MRegion\_ptr} r):} Add a region to the mesh. It is assumed that
+the region and its topology is properly defined.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Vertex(\textit{Mesh\_ptr} mesh,
+\textit{MVertex\_ptr} v):} Remove vertex from mesh. Vertex is not
+deleted and must be deleted afterward separately.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Edge(\textit{Mesh\_ptr} mesh,
+\textit{MEdge\_ptr} e):} Remove edge from mesh. Edge is not deleted and
+must be deleted afterward separately.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Face(\textit{Mesh\_ptr} mesh,
+\textit{MFace\_ptr} f):} Remove face from mesh. Face is not deleted and
+must be deleted afterward separately.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Region(\textit{Mesh\_ptr} mesh,
+\textit{MRegion\_ptr} r):} Remove region from mesh. Region is not deleted and must be deleted afterward separately.
+
+\item[]\textbf{\textit{void} MESH\_Set\_GModel(\textit{Mesh\_ptr} mesh,
+GModel\_ptr geom):} Assign a geometric model handle to the mesh.
+
+\item[]\textbf{\textit{int} MESH\_Change\_RepType(\textit{Mesh\_ptr} mesh,
+\textit{int} nurep):} Change the representation type of the mesh. This
+routine can be used to modify the representation type dynamically to
+suit different algorithms.  However, the cost of making the change and
+reordering all adjacencies and creating or deleting entities has to be
+considered while invoking this routine. Also, once a conversion is
+made from a full representation to a reduced representation, not all
+information may be retrievable when switching back to a full
+representation.  (particularly classification information, i.e.,
+relationship of mesh entities to the geometric model).
+
+\end{description}
+
+\newpage
+\subsection{Mesh Vertex Object}
+
+\begin{description}
+\item[]\textbf{\textit{MVertex\_ptr} MV\_New(\textit{Mesh\_ptr} mesh):}
+Create a new vertex object. No geometric or topological information is
+embedded in the vertex when it is created. The vertex only knows which
+mesh it belongs to. The ID of the vertex is set by this function.
+
+\item[]\textbf{\textit{void} MV\_Delete(\textit{MVertex\_ptr} mvertex):}
+Delete the vertex. Deletes all topological and geometric
+information embedded in the vertex.
+
+\item[]\textbf{\textit{void} MV\_Set\_Coords(\textit{MVertex\_ptr} mvertex,
+double *xyz):} Set the coordinates of the vertex.
+
+\item[]\textbf{\textit{void} MV\_Set\_GEntity(\textit{MVertex\_ptr} mvertex,
+GEntity\_ptr gent):} Set the geometric model entity on which vertex is
+classified.
+
+\item[]\textbf{\textit{void} MV\_Set\_GEntDim(\textit{MVertex\_ptr} mvertex,
+\textit{int} gdim):} Set topological dimension of model entity on which
+vertex is classified.
+
+\item[]\textbf{\textit{void} MV\_Set\_GEntID(\textit{MVertex\_ptr} mvertex,
+\textit{int} gid):} Set ID of model entity on which vertex is
+classified.
+
+\item[]\textbf{\textit{void} MV\_Add\_AdjVertex(\textit{MVertex\_ptr}
+mvertex, \textit{MVertex\_ptr} adjvertex):} Add neighboring vertex,
+adjvertex, to ajdacent vertex list of vertex, mvertex.
+
+\item[]\textbf{\textit{void} MV\_Rem\_AdjVertex(\textit{MVertex\_ptr}
+mvertex, \textit{MVertex\_ptr} adjvertex):} Delete neighboring vertex
+of given vertex.
+
+\item[]\textbf{\textit{void} MV\_Set\_ID(\textit{MVertex\_ptr} mvertex,
+\textit{int} id):} Explicitly set ID of a vertex and overwrite the ID
+set by the MV\_New operator. Does not check for duplication of edge
+IDs.
+
+\item[]\textbf{\textit{Mesh\_ptr} MV\_Mesh(\textit{MVertex\_ptr} mv):}
+Returns the mesh that this vertex belongs to.
+
+\item[]\textbf{\textit{int} MV\_ID(\textit{MVertex\_ptr} mvertex):} Returns
+the ID of the vertex.
+
+\item[]\textbf{\textit{int} MV\_GEntDim(\textit{MVertex\_ptr} mvertex):}
+Returns the dimension of the geometric model entity that the vertex is
+classified on. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} MV\_GEntID(\textit{MVertex\_ptr} mvertex):}
+Returns the ID of the geometric model entity that the vertex is
+classified on. Returns 0 if this information is not known.
+
+\item[]\textbf{\textit{GEntity\_ptr} MV\_GEntity(\textit{MVertex\_ptr}
+  mvertex):} Returns a pointer or handle to the geometric model entity
+  that the vertex is classified on. Returns NULL if this information
+  is not known.
+
+\item[]\textbf{\textit{void} MV\_Coords(\textit{MVertex\_ptr} mvertex,
+double *xyz):} Returns the coordinates of the vertex.
+
+\item[]\textbf{\textit{int} MV\_Num\_AdjVertices(\textit{MVertex\_ptr}
+mvertex):} Returns the number of edge connected neighboring vertices of
+vertex. \textit{Not efficient for all representations}.
+
+\item[]\textbf{\textit{int} MV\_Num\_Edges(\textit{MVertex\_ptr} mvertex):}
+Returns the number of edges connected to the vertex.
+
+\item[]\textbf{\textit{int} MV\_Num\_Faces(\textit{MVertex\_ptr} mvertex):}
+Returns the number of faces connected to the vertex.
+
+\item[]\textbf{\textit{int} MV\_Num\_Regions(\textit{MVertex\_ptr} mvertex):}
+Returns the number of regions connected to the vertex
+
+\item[]\textbf{\textit{Set\_ptr} MV\_AdjVertices(\textit{MVertex\_ptr}
+mvertex):} Set of adjacent or edge connected neighboring vertices of vertex.
+
+\item[]\textbf{\textit{Set\_ptr} MV\_Edges(\textit{MVertex\_ptr} mvertex):}
+Set of edges connected to the vertex.
+
+\item[]\textbf{\textit{Set\_ptr} MV\_Faces(\textit{MVertex\_ptr} mvertex):}
+Set of faces connected to the vertex.
+
+\item[]\textbf{\textit{Set\_ptr} MV\_Regions(\textit{MVertex\_ptr} mvertex):}
+Set of regions connected to the vertex.
+\end{description}
+
+\newpage
+\subsection{Mesh Edge Object}
+
+\begin{description}
+\item[]\textbf{\textit{MEdge\_ptr} ME\_New(\textit{Mesh\_ptr} mesh):} Create
+a new edge object. No topological information is embedded in the edge
+when it is created. The edge only knows which mesh it belongs
+to. The ID of the edge is set by this function.
+
+\item[]\textbf{\textit{void} ME\_Delete(\textit{MEdge\_ptr} medge):} Delete
+the edge. Deletes all topological information embedded in the edge.
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntity(\textit{MEdge\_ptr} medge,
+GEntity\_ptr gent):} Set the geometric model entity on which the edge
+is classified.
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntDim(\textit{MEdge\_ptr} medge,
+\textit{int} gdim):} Set the topological dimension of model entity on
+which edge is classified.
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntID(\textit{MEdge\_ptr} medge,
+\textit{int} gid):} Set ID of model entity on which edge is classified.
+
+\item[]\textbf{\textit{void} ME\_Set\_ID(\textit{MEdge\_ptr} medge,
+\textit{int} id):} Explicitly set ID of an edge and overwrite the ID set
+by the ME\_New function. Does not check for duplication of edge IDs.
+
+\item[]\textbf{\textit{void} ME\_Set\_Vertex(\textit{MEdge\_ptr} medge,
+\textit{int} i, \textit{MVertex\_ptr} vertex):} Set the i'th vertex of
+the edge. i can be 0 or 1.
+
+\item[]\textbf{\textit{void} ME\_Replace\_Vertex(\textit{MEdge\_ptr} medge,
+\textit{MVertex\_ptr} vert, \textit{MVertex\_ptr} nuvert):} Replace
+i'th vertex by new vertex.
+
+\item[]\textbf{\textit{Mesh\_ptr} ME\_Mesh(\textit{MEdge\_ptr} medge):}
+Returns the mesh that this edge belongs to.
+
+\item[]\textbf{\textit{int} ME\_ID(\textit{MEdge\_ptr} medge):} Returns the
+ID of the vertex. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} ME\_GEntDim(\textit{MEdge\_ptr} medge):} Returns
+the dimension of the geometric model entity that the vertex is
+classified on. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} ME\_GEntID(\textit{MEdge\_ptr} medge):} Returns
+the ID of the geometric model entity that the vertex is classified on.
+Returns 0 if this information is not known.
+
+\item[]\textbf{\textit{GEntity\_ptr} ME\_GEntity(\textit{MEdge\_ptr} medge):} Returns
+a pointer or handle to the geometric model entity that the vertex is
+classified on. Returns NULL if this information is not known.
+
+\item[]\textbf{\textit{int} ME\_Num\_Faces(\textit{MEdge\_ptr} medge):}
+Returns the number of faces connected to the edge.
+
+\item[]\textbf{\textit{int} ME\_Num\_Regions(\textit{MEdge\_ptr} medge):}
+Returns the number of regions connected to the edge.
+
+\item[]\textbf{\textit{MVertex\_ptr} ME\_Vertex(\textit{MEdge\_ptr} medge,
+\textit{int} i):} Returns the i'th vertex of the edge. i=0 returns the
+first vertex and i=1 returns the second vertex.
+
+\item[]\textbf{\textit{MVertex\_ptr} ME\_OppVertex(\textit{MEdge\_ptr}
+medge, \textit{MVertex\_ptr} ov):} Return the vertex opposite to given
+vertex in edge.
+
+\item[]\textbf{\textit{int} ME\_UsesEntity(\textit{MEdge\_ptr} medge,
+\textit{MEntity\_ptr} mentity, \textit{int} etype):} Check if edge uses
+given lower dimension entity, \textit{mentity}. The dimension of the
+entity is specified by the \textit{etype} variable. For an edge, the
+only lower dimensional entity is a vertex. If the edge uses the
+vertex, the function returns 1; otherwise it returns 0. If any
+other type of entity is specified, the function returns 0.
+
+\item[]\textbf{\textit{Set\_ptr} ME\_Faces(\textit{MEdge\_ptr} medge):}
+Returns the set of faces using this edge.
+
+\item[]\textbf{\textit{Set\_ptr} ME\_Regions(\textit{MEdge\_ptr} medge):}
+Returns the set of regions using this edge.
+
+\item[]\textbf{\textit{MEdge\_ptr} MVs\_CommonEdge(\textit{MVertex\_ptr} v1,
+\textit{MVertex\_ptr} v2):} Return the edge connecting vertices v1 and
+v2, if it exists. If such an edge does not exist, the function returns
+0.
+
+\item[]\textbf{\textit{double} ME\_Len(\textit{MEdge\_ptr} e):} Return
+  the length of the straight line connecting the two vertices of the
+  edge.
+
+\item[]\textbf{\textit{double} ME\_LenSqr(\textit{MEdge\_ptr} e):}
+  Return the square of the length of the straight line connecting the
+  two vertices of the edge.
+
+\item[]\textbf{\textit{void} ME\_Vec(\textit{MEdge\_ptr} e, double *evec):}
+Return the vector going from the first vertex of the edge to the
+second vertex of the edge.
+\end{description}
+
+\newpage
+\subsection{Mesh Face Object}
+
+\begin{description}
+\item[]\textbf{\textit{MFace\_ptr} MF\_New(\textit{Mesh\_ptr} mesh):} Create
+a new face object. No topological information is embedded in the face
+when it is created. The face only knows which mesh it belongs
+to. The ID of the face is set by this function.
+
+\item[]\textbf{\textit{void} MF\_Delete(\textit{MFace\_ptr} mface):} Delete
+the face. Delete all topological information embedded in the face.
+
+\item[]\textbf{\textit{void} MF\_Set\_GEntity(\textit{MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the edge is classified.
+
+\item[]\textbf{\textit{void} MF\_Set\_GEntDim(\textit{MFace\_ptr} mface,
+\textit{int} gdim):} Set the dimension of the geometric model entity on
+which the face is classified.
+
+\item[]\textbf{\textit{void} MF\_Set\_GEntID(\textit{MFace\_ptr} mface,
+\textit{int} gid):} Set the ID of the geometric model entity on which
+the face is classified.
+
+\item[]\textbf{\textit{void} MF\_Set\_ID(\textit{MFace\_ptr} mface,
+\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
+set by the MF\_New operator. Does not check for duplication of face
+IDs.
+
+\item[]\textbf{\textit{void} MF\_Set\_Edges(\textit{MFace\_ptr} mface,
+\textit{int} n, \textit{MEdge\_ptr} *edges, \textit{int} *dirs):} Set
+the edges of the face along with their directions. The ordered set of
+edge pointers and their directions are passed in through arrays along
+with the number of edges. The edges are assumed to be ordered
+clockwise around the face. If an edge direction is along the clockwise
+direction of the face then the entry in the 'dirs' array must be 1;
+otherwise it must be 0. This function is relevant only for full
+representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Set\_Vertices(\textit{MFace\_ptr} mface,
+\textit{int} n, \textit{MVertex\_ptr} *verts):} Set the vertices of the
+face. The ordered set of vertices is passed in through an array along
+with the number of vertices. The vertices are assumed to be ordered
+clockwise around the face. This function is relevant only for reduced
+representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edge(\textit{MFace\_ptr} mface,
+\textit{MEdge\_ptr} edge, \textit{MEdge\_ptr} nuedge, \textit{int}
+dir):} Replace an edge in the face with another edge. The direction
+in which the new edge is used in the face must also be supplied. This
+function is relevant only for full representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Vertex(\textit{MFace\_ptr} mface,
+\textit{MVertex\_ptr} mvertex, \textit{MVertex\_ptr} nuvertex):}
+Replace a vertex in the face with another vertex. This function is
+relevant only for reduced representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edge\_i(\textit{MFace\_ptr} mface,
+\textit{int} i, \textit{MEdge\_ptr} nuedge, \textit{int} dir):} Replace
+the i'th edge in the face with another edge. The direction in which the
+new edge is used in the face must also be supplied. This function is
+relevant only for full representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
+mface, \textit{int} i, \textit{MVertex\_ptr} nuvertex):} Replace the
+i'th vertex in the face with a new vertex. This function is relevant
+only for reduced representations in MSTK.
+
+\item[]\textbf{\textit{Mesh\_ptr} MF\_Mesh(\textit{MFace\_ptr} mf):} Returns the mesh that this mesh belongs to.
+
+\item[]\textbf{\textit{int} MF\_ID(\textit{MFace\_ptr} mface):} Returns the ID of the face. Returns 0 if not known.
+
+\item[]\textbf{\textit{int} MF\_GEntDim(\textit{MFace\_ptr} mface):} Returns the dimensions of the geometric model entity that the vertex is classified on. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} MF\_GEntID(\textit{MFace\_ptr} mface):} Returns the ID of the geometric model entity that the vertex is classified on. Returns 0 if this information is not known.
+
+\item[]\textbf{\textit{GEntity\_ptr} MF\_GEntity(\textit{MFace\_ptr} mface):} Returns a pointer or handle to the geometric model entity that the vertex is classified on.
+Returns NULL if this information is not known.
+
+\item[]\textbf{\textit{int} MF\_Num\_Vertices(\textit{MFace\_ptr} mface):} Returns the number of vertices of the face.
+
+\item[]\textbf{\textit{int} MF\_Num\_Edges(\textit{MFace\_ptr} mface):} Returns the number of edges of the face.
+
+\item[]\textbf{\textit{int} MF\_Num\_AdjFaces(\textit{MFace\_ptr}
+    mface):} Returns the number of adjacent faces of a face. This
+  operator is relevant only in planar or surface meshes, i.e., for
+  boundary faces not connected to any regions.
+
+\item[]\textbf{\textit{Set\_ptr} MF\_Vertices(\textit{MFace\_ptr}
+    mface, \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
+  ordered set of the vertices of the face. The vertices are ordered in
+  ccw direction while looking down the face 'normal', if 'dir' is 1
+  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
+  vertex set is reordered so that it is the first vertex (This
+  argument will be added soon to the function. For now, omit this
+  argument). The behavior of this function can be illustrated using
+  Figure~\ref{fig:face_def}. For the face shown in the figure, a
+  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
+  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
+  A vertex set with ccw ordering starting with vertex $V_2$ is
+  ${V_2,V_3,V_0,V_1}$.
+
+\begin{figure}[!h]
+  \begin{center}
+  \includegraphics[scale=1.0]{figures/face_def} 
+  \caption{Face definition}
+  \label{fig:face_def}
+  \end{center}
+\end{figure}
+
+\item[]\textbf{\textit{Set\_ptr} MF\_Edges(\textit{MFace\_ptr} mface,
+    \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
+  ordered set of edges of the face. The edges are ordered in the ccw
+  while looking down the face 'normal' if dir is 1 and in the cw if
+  dir is 0. If 'mvert' is specified, the edge set is reordered so that
+  the first edge in the set contains this vertex. More precisely, if
+  'dir' is 1, and the first edge is 'e' used in the face in direction
+  'd', then \textbf{ME\_Vertex(e,!d)} = mvert. With reference to
+  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
+  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
+  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
+  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
+  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
+  ${E_1,E_0,E_3,E_2}$.
+
+\item[]\textbf{\textit{Set\_ptr} MF\_AdjFaces(\textit{MFace\_ptr} mface):}
+Set of adacent faces of a face. This operator is relevant only in
+planar or surface surface meshes, i.e., for boundary faces not
+connected to any regions.
+
+\item[]\textbf{\textit{int} MF\_EdgeDir(\textit{MFace\_ptr} mface,
+\textit{MEdge\_ptr} medge):} Returns the direction in which the face
+uses the given edge. If the faces use the edge in the positive direction,
+the function returns 1; otherwise it returns 0.
+
+\item[]\textbf{\textit{int} MF\_EdgeDir\_i(\textit{MFace\_ptr} mface,
+\textit{int} i):} Returns the direction in which the face uses its i'th
+edge. If the face uses the edge in the positive direction the function
+returns 1; otherwise it returns 0;
+
+\item[]\textbf{\textit{int} MF\_UsesEntity(\textit{MFace\_ptr} mface,
+\textit{MEntity\_ptr} mentity, \textit{int} type):} Check if the face
+uses the given lower dimension entity, 'mentity'. The type of the
+entity is specified by the 'etype' variable. For a face, a lower
+dimensional entity can be a vertex or an edge. If the face uses the
+vertex or the edge, the function returns 1; otherwise it returns 0. If
+any other type of entity is specified, the function returns 0.
+
+\item[]\textbf{\textit{Set\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
+Return the set of regions connected to the face. If the face is not
+used by any regions, the function returns NULL to indicate that the
+set is empty. If not the set may contain one or two regions.
+
+\item[]\textbf{\textit{MRegion\_ptr} MF\_Region(\textit{MFace\_ptr} mface,
+\textit{int} side):} Returns the region on the specified side of the
+face. The positive side of the face (side = 1) is the side towards
+which the face normal points. The negative side of the face (side = 0)
+is the opposite side.
+
+\item[]\textbf{\textit{void} MF\_Coords(\textit{MFace\_ptr} mface,
+\textit{int} *n, double (*xyz)[3], \textit{int} dir):} Returns the
+coordinates of the face vertices in an array along with the number of
+vertices. If 'dir' is 1, the coordinates are returned with a  ccw
+ordering while looking down the face normal; if 'dir' is 0, they are
+returned with a cw ordering (See Figure~\ref{fig:face_def}).
+\end{description}
+
+\newpage
+\subsection{Mesh Region Object}
+
+\begin{description}
+\item[]\textbf{\textit{MRegion\_ptr} MR\_New(\textit{Mesh\_ptr} mesh):}
+Create a new region object. No topological information is embedded in
+the region when it is created. The region only knows which mesh it
+belongs to. The ID of the region is set by this function.
+
+\item[]\textbf{\textit{void} MR\_Delete(\textit{MRegion\_ptr} mregion):}
+Delete the region. Deletes all topological information embedded in the
+edge.
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntity(\textit{MRegion\_ptr} mregion,
+GEntity\_ptr gent):} Set the geometric model entity on which the edge
+is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntDim(\textit{MRegion\_ptr} mregion,
+\textit{int} gdim):} Set the dimension of the geometric model entity on
+which the edge is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntID(\textit{MRegion\_ptr} mregion,
+\textit{int} gid):} Set the ID of the geometric model entity on which
+the edge is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_ID(\textit{MRegion\_ptr} mregion,
+\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
+set by the \textbf{MR\_New} function. Does not check for duplication
+of region IDs.
+
+\item[]\textbf{\textit{void} MR\_Set\_Faces(\textit{MRegion\_ptr} mregion,
+\textit{int} nf, \textit{MFace\_ptr} *mfaces, \textit{int} *dirs):} Set
+the faces of the region along with their directions. The
+\textit{unordered} set of faces and their directions are passed in
+through arrays along with the number of faces. If the normal of the
+face points out of the region, the associated direction to be passed
+in is 1; otherwise it is 0. This function is only relevant for full
+representations in MSTK.
+
+\item[]\textbf{\textit{void} MR\_Set\_Vertices(\textit{MRegion\_ptr}
+mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set the
+vertices of the region. This function is relevant for reduced
+representations only. For standard elements, the vertices must be
+ordered as indicated in Appendix~\ref{app:reg_conventions}. This
+function does not apply for defining general regions since there
+cannot be any implicit ordering.
+
+\item[]\textbf{\textit{void} MR\_Replace\_Face(\textit{MRegion\_ptr}
+mregion, \textit{MFace\_ptr} mface, \textit{MFace\_ptr} nuface,
+\textit{int} dir):} Replace a face of the region with another edge. The
+direction in which the new face is used in the region must also be
+supplied. This function is only relevant for full representations in
+MSTK.
+
+\item[]\textbf{\textit{void} MR\_Replace\_Vertex(\textit{MRegion\_ptr}
+mregion, \textit{MVertex\_ptr} mvertex, \textit{MVertex\_ptr}
+nuvertex):} Replace a vertex of a region with another vertex. This
+function is relevant only for reduced representations in MSTK.
+
+\item[]\textbf{\textit{void} MR\_Replace\_Face\_i(\textit{MRegion\_ptr}
+mregion, \textit{int} i, \textit{MFace\_ptr} mface, \textit{int} dir):}
+Replace the i'th face in the region with another face.The direction in
+which the new face is used in the region must also be supplied. This
+function is only relevant for full representations in MSTK.
+
+\item[]\textbf{\textit{void} MR\_Replace\_Vertex\_i(\textit{MRegion\_ptr}
+mregion, \textit{int} i, \textit{MVertex\_ptr} mvertex):} Replace the
+i'th vertex of the region with another vertex. This function is only
+relevant for reduced representations in MSTK.
+
+\item[]\textbf{\textit{Mesh\_ptr} MR\_Mesh(\textit{MRegion\_ptr} mregion):}
+Returns the mesh that the region belongs to.
+
+\item[]\textbf{\textit{int} MR\_ID(\textit{MRegion\_ptr} mregion):} Returns
+the ID of the region. Returns 0 if not known.
+
+\item[]\textbf{\textit{int} MR\_GEntDim(\textit{MRegion\_ptr} mregion):}
+Returns the dimension of the geometric model entity the region is
+classified on. Always returns 3 since a mesh region can be classified
+only on a model region.
+
+\item[]\textbf{\textit{int} MR\_GEntID(\textit{MRegion\_ptr} mregion):}
+Returns the ID of the geometric model entity that the region is
+classified on. Returns 0 if not known.
+
+\item[]\textbf{\textit{GEntity\_ptr} MR\_GEntity(\textit{MRegion\_ptr}
+  mregion):} Returns a pointer or handle to the geometric model entity
+  that the vertex is classified on. Returns NULL if this information
+  is not known.
+
+\item[]\textbf{\textit{int} MR\_Num\_Vertices(\textit{MRegion\_ptr}
+mregion):} Returns the number of vertices of a region.
+
+\item[]\textbf{\textit{int} MR\_Num\_Edges(\textit{MRegion\_ptr} mregion):}
+Returns the number of edges of a region.
+
+\item[]\textbf{\textit{int} MR\_Num\_Faces(\textit{MRegion\_ptr} mregion):}
+Returns the number of faces of a region.
+
+\item[]\textbf{\textit{int} MR\_Num\_AdjRegions(\textit{MRegion\_ptr}
+mregion):} Returns the number of adjacent regions of a region, i.e.,
+regions sharing a face with this region.
+
+\item[]\textbf{\textit{Set\_ptr} MR\_Vertices(\textit{MRegion\_ptr}
+mregion):} Returns the set of vertices of a region. For standard
+elements the vertices are ordered as indicated in
+Appendix~\ref{app:reg_conventions}. For non-standard elements the set
+is unordered.
+
+\item[]\textbf{\textit{Set\_ptr} MR\_Edges(\textit{MRegion\_ptr} mregion):}
+Return the \underline{unordered} set of edges of a region.
+
+\item[]\textbf{\textit{Set\_ptr} MR\_Faces(\textit{MRegion\_ptr} mregion):}
+Returns the set of faces of a region.
+
+\item[]\textbf{\textit{Set\_ptr} MR\_AdjRegions(\textit{MRegion\_ptr}
+mregion):} Returns the set of adjacent regions of a region, i.e.,
+regions sharing a face with this region. The set is not ordered.
+
+\item[]\textbf{\textit{int} MR\_FaceDir(\textit{MRegion\_ptr} mregion,
+\textit{MFace\_ptr} mface):} Returns the direction in which the region
+uses the given face. Returns 1 if the face normal points out of the
+region and returns 0 if the face normal points into the region.
+
+\item[]\textbf{\textit{int} MR\_FaceDir\_i(\textit{MRegion\_ptr} mregion,
+\textit{int} i):} Returns the direction in which the region uses the
+i'th face. Returns 1 if the face normal points out of the region and
+returns 0 if the face normal points into the region.
+
+\item[]\textbf{\textit{int} MR\_UsesEntity(\textit{MRegion\_ptr} mregion,
+\textit{MEntity\_ptr} ment, \textit{int} type):} Check if the region
+uses the given lower dimension entity, 'mentity'. The type of the
+entity is
+
+\item[]\textbf{\textit{void} MR\_Coords(\textit{MRegion\_ptr} mregion,
+\textit{int} *n, double (*xyz)[3]):} Returns the coordinates of the
+region vertices in an array along with the number of vertices. For
+standard elements, the ordering is as given in
+Appendix~\ref{app:reg_conventions}. For non-standard elements, the
+ordering is arbitrary.
+\end{description}
+
+\newpage
+\subsection{Generic Entity Object}
+
+The following functions operate on generic mesh entities of type
+\textit{MEntity\_ptr}. This implies that variables of type
+\textit{MVertex\_ptr}, \textit{MEdge\_ptr}, \textit{MFace\_ptr},
+\textit{MRegion\_ptr} can all be passed in place of
+\textit{MEntity\_ptr} variables in the following functions.
+
+\begin{description}
+%%  \textit{void} MEnt\_Set\_GEntity(\textit{MEntity\_ptr} mentity, GEntity\_ptr gent);
+%%
+%%  \textit{void} MEnt\_Set\_GEntDim(\textit{MEntity\_ptr} mentity, \textit{int} gdim);
+%%
+%%  \textit{void} MEnt\_Set\_ID(\textit{MEntity\_ptr} mentity, \textit{int} id);
+
+\item[]\textbf{\textit{int} MEnt\_ID(\textit{MEntity\_ptr} mentity):} Returns
+the ID of a generic entity.
+
+\item[]\textbf{\textit{int} MEnt\_Dim(\textit{MEntity\_ptr} mentity):}
+Returns the topological dimension or type of generic entity.
+
+\item[]\textbf{\textit{Mesh\_ptr} MEnt\_Mesh(\textit{MEntity\_ptr} mentity):}
+Returns the mesh that the entity belongs to.
+
+\item[]\textbf{\textit{int} MEnt\_GEntDim(\textit{MEntity\_ptr} mentity):}
+Returns the dimension of the geometric model entity that the entity is
+classified on.
+
+\item[]\textbf{\textit{GEntity\_ptr} MEnt\_GEntity(\textit{MEntity\_ptr} mentity):}
+Returns a pointer or handle to geometric model entity that the entity
+is classified on.
+\end{description}
+
+\newpage
+\subsection{Entity Marks}
+
+Entity marks or markers are a way of tagging entities. Such
+functionality is useful in algorithms which must keep track of
+processed entities to avoid duplication of work. An example of such an
+operation is creating a union of entity sets while extracting upward
+adjacency information such as the regions connected to an edge.
+
+\begin{description}
+\item[]\textbf{\textit{int} MSTK\_GetMarker():} Returns a unique marker ID
+which may be used to tag entities.
+
+\item[]\textbf{\textit{void} MEnt\_Mark(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Mark an entity with the given marker 'mkr'.
+
+\item[]\textbf{\textit{int} MEnt\_IsMarked(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Check if an entity is marked with the given marker 'mkr'.
+
+\item[]\textbf{\textit{void} MEnt\_Unmark(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Unmark an entity with respect to the given marker 'mkr'
+
+\item[]\textbf{\textit{void} Set\_Mark(\textit{Set\_ptr} list, \textit{int} mkr):} Mark a set of entities with given marker.
+
+\item[]\textbf{\textit{void} Set\_Unmark(\textit{Set\_ptr} list, \textit{int} mkr):}
+Unmark a set of entities with respect to the given marker.
+
+\item[]\textbf{\textit{void} MSTK\_FreeMarker(\textit{int} mkr):} Release the
+marker ID given by \textbf{MSTK\_GetMarker()} so that it can be
+reused. Care must be taken to unmark all entities marked with this
+marker ID before releasing it. If not, subsequent operations with
+reassigned marker will find a tag on some entities and mistake them
+for being processed.
+\end{description}
+
+\newpage
+\subsection{Mesh Modification}
+
+\begin{description}
+
+\item[]\textbf{\textit{int} ME\_Swap2D(\textit{MEdge\_ptr} e,
+\textit{MEdge\_ptr} *enew, \textit{MFace\_ptr} fnew[2]):} Swap an edge
+in a triangular mesh. No checks are performed for topological or
+geometric validity.
+
+\item[]\textbf{\textit{MFace\_ptr} MFs\_Join(\textit{MFace\_ptr} f1,
+\textit{MFace\_ptr} f2, \textit{MEdge\_ptr} e):} Join two faces along
+common edge and create new face by eliminating the common edge as
+shown in Figure~\ref{fig:MFs_Join}. If 'f1' has 'n1' edges and 'f2'
+has 'n2' edges, then the new face has ('n1'+'n2'-2) edges.
+
+\end{description}
+
+\begin{figure}[!h]
+  \begin{center}
+    \includegraphics[scale=1.0]{figures/MFs_Join}
+
+    \caption{Joining two faces (a) Two faces $F_1$ and $F_2$ sharing a
+      common edge (b) New pentagonal face $F_{new}$ created by
+      eliminating the common edge.  }
+    \label{fig:MFs_Join}
+  \end{center}
+\end{figure}
+
+\newpage
+\subsection{Utilities}
+
+\begin{description}
+
+\item[]\textbf{\textit{void} MSTK\_Report(\textit{char} *module,
+  \textit{char} *message, \textit{ErrType} severity):} Error handler
+  for MSTK. 'module' is the name of the function in which the error
+  occurs. 'message' is the error message and is recommended to be less
+  than 1024 characters in length. 'severity' is an error code and can
+  be MESSG, WARN, ERROR or FATAL. If the error code is FATAL, the
+  program will quit after printing the error. If the same message is
+  repeated successively, then the message is printed only the first
+  time.
+
+\item[]\textbf{\textit{void} Set\_PrintID(\textit{Set\_ptr} l):} Debugging
+utility to print the IDs of the entities in a set.
+
+\item[]\textbf{\textit{void} MV\_Print(\textit{MVertex\_ptr} v, int
+    lev):} Debugging utility to print information about a mesh vertex,
+  \textbf{v}. The argument \textbf{lev} controls the level of detail
+  of the information printed. \textbf{lev} = 0 prints the minimum
+  information, i.e., vertex pointer, its ID and its coordinates. If
+  \textbf{lev} = 1, the function prints classification information for
+  the vertex (if available), i.e., ID and dimension of the model
+  entity that the vertex is on. If \textbf{lev} $>$ 1, then upward
+  detailed adjacency information is also printed for the vertex, i.e.,
+  information is printed about the edges, faces and regions connected
+  to the vertex.
+
+\item[]\textbf{\textit{void} ME\_Print(\textit{MEdge\_ptr} e, int
+    lev):} Debugging utility to print information about a mesh edge,
+  \textbf{e}. The argument \textbf{lev} controls the level of detail
+  of the information printed. \textbf{lev} = 0 prints the minimum
+  information, i.e., edge pointer, its ID and the IDs of its two
+  vertices. If \textbf{lev} = 1, the function prints classification
+  information for the edge (if available), i.e., ID and dimension of
+  the model entity that the edge is on. Also, more detailed vertex
+  information printed in this case. If \textbf{lev} $>$ 1, the
+  function prints detailed upward adjacency information for the edge,
+  i.e., information is printed about the faces and regions connected
+  to the edge.
+
+  \item[]\textbf{\textit{void} MF\_Print(\textit{MFace\_ptr} f, int
+      lev):} Debugging utility to print information about a mesh face,
+    \textbf{f}. The argument \textbf{lev} controls the level of detail
+    of the information printed. \textbf{lev} = 0 prints the minimum
+    information, i.e., the face pointer and its ID. If \textbf{lev} =
+    1, the function prints classification information for the edge (if
+    available), i.e., ID and dimension of the model entity that the
+    face is on. Also, a signed list of the edges of the face is
+    printed.  If \textbf{lev} $>$ 1, the function prints detailed
+    downward and upward adjacency information for the face, i.e.,
+    information is printed about the edges and vertices of the face,
+    and about the regions connected to the face.
+
+  \item[]\textbf{\textit{void} MR\_Print(\textit{MRegion\_ptr} r, int
+      lev):} Debugging utility to print information about a mesh
+    region, \textbf{r}. The argument \textbf{lev} controls the level
+    of detail of the information printed. \textbf{lev} = 0 prints the
+    minimum information, i.e., region pointer and its ID. If
+    \textbf{lev} = 1, the function prints classification information
+    for the region (if available), i.e., ID of the model entity that
+    the region is on. Also, a signed list of the faces of the region
+    is printed. If \textbf{lev} $>$ 1, the function prints detailed
+    downward adjacency information for the region, i.e., information
+    is printed about the faces, edges and vertices forming the region.
+
+\end{description}
+
+\newpage
+\appendix
+
+\section{Mesh Representation Types  in MSTK}
+\label{app:reptypes}
+
+\begin{figure}[!h]
+%  \begin{center}
+    \begin{minipage}{2.525in}
+      \begin{center}
+        \includegraphics[scale=0.9]{figures/repF1} \\
+        Representation F1
+      \end{center}
+    \end{minipage}\hfill%
+    \begin{minipage}{2.525in}
+      \begin{center}
+        \includegraphics[scale=0.9]{figures/repF4} \\
+        Representation F4
+      \end{center}
+    \end{minipage}
+    \vspace{5em}
+
+    \begin{minipage}{1.75in}
+      \begin{center}
+        \includegraphics[scale=0.9]{figures/repR1} \\
+        Representation R1
+      \end{center}
+    \end{minipage}
+    \begin{minipage}{1.75in}
+      \begin{center}
+        \includegraphics[scale=0.9]{figures/repR2} \\
+        Representation R2
+      \end{center}
+    \end{minipage}
+    \begin{minipage}{1.75in}
+      \begin{center}
+        \includegraphics[scale=0.9]{figures/repR4} \\
+        Representation R4
+      \end{center}
+    \end{minipage}
+%  \end{center}
+\end{figure}
+
+\newpage
+\section{Conventions for Vertex, Edge Numbering in Standard Region Types}
+\label{app:reg_conventions}
+
+\begin{figure}[!h]
+  \begin{center}
+    \includegraphics[scale=1.0]{figures/reg_conventions}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{MSTK File Format}
+\label{app:file_format}
+
+\subsection{MSTK ASCII File Format}
 
@@
 
-\# \textit{char *Mfilename - Geometric model file related to mesh} 
-\vspace{1ex}
-
-Mfilename
-\vspace{1ex}
-
 \# \textit{char *reptype - Type of representation} \\  
-\# \textit{int NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
+\# \textit{\textit{int} NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
 \vspace{1ex}
@@
 \# \textit{double X,Y,Z Coordinates} \\
-\# \textit{int Mdim - Topological type or dimension of model entity that} \\
+\# \textit{\textit{int} Mdim - Topological type or dimension of model entity that} \\
 \# \textit{\hspace{4.7em} the vertex is on} \\
-\# \textit{int Mid - ID of model entity the vertex is on} \\
-\# \textit{int Mdim and Mid can be -1 if model information is absent} 
+\# \textit{\textit{int} Mid - ID of model entity the vertex is on} \\
+\# \textit{Mdim and Mid can be -1 and 0 resp. if model info. is absent} 
 \vspace{1ex}
@@
 \# \textit{\textbf{EDGE INFO} - present only if NE $\ne$ 0} \\
+\# \textit{Keyword 'edges' followed by edge records}
 \# \textit{Each edge record has} \\
-\# \textit{int Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
-\# \textit{int Mdim, Mid}
+\# \textit{\textit{int} Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
+\# \textit{\textit{int} Mdim, Mid}
 \vspace{1ex}
 
+edges
 Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
@@
 \# \textit{\textbf{FACE INFO} - present only if NF $\ne$ 0} \\
-\# \textit{char *FLtype - Dimension of lower order entity describing faces} \\
+\# \textit{Keyword 'faces'}
+\# \textit{char *FLtype: Keyword for lower order entity describing faces} \\
 \# \textit{Values: Vertex, Edge (case insensitive), e.g. VeRteX or EDGE} \\
@@
 
-FLtype 
+faces FLtype 
 \vspace{1ex}
@@
 \# \textit{If face described by vertices, then each face record has} \\
-\# \textit{int NFV - Number of face vertices} \\
-\# \textit{int Vid\_1 - ID of first vertex of face} \\
-\# \textit{int Vid\_2 - ID of second vertex of face} \\
+\# \textit{\textit{int} NFV - Number of face vertices} \\
+\# \textit{\textit{int} Vid\_1 - ID of first vertex of face} \\
+\# \textit{\textit{int} Vid\_2 - ID of second vertex of face} \\
 \# . . . \\
-\# \textit{int Vid\_1 - ID of NFV'th vertex of face} \\
-\# \textit{int Mdim, Mid} 
+\# \textit{\textit{int} Vid\_1 - ID of NFV'th vertex of face} \\
+\# \textit{\textit{int} Mdim, Mid} 
 \vspace{1ex}
@@
 \# \textit{If face described by edges, then each face record has} \\
-\# \textit{int NFE - Number of face edges} \\
-\# \textit{int $\pm$Eid\_1 - signed ID of first edge of face} \\
-\# \textit{int $\pm$Eid\_2 - signed ID of second edge of face} \\
+\# \textit{\textit{int} NFE - Number of face edges} \\
+\# \textit{\textit{int} $\pm$Eid\_1 - signed ID of first edge of face} \\
+\# \textit{\textit{int} $\pm$Eid\_2 - signed ID of second edge of face} \\
 \# . . . \\
-\# \textit{int $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
-\# \textit{int Mdim, Mid} \\
+\# \textit{\textit{int} $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
+\# \textit{\textit{int} Mdim, Mid} \\
 \# \\
@@
 \# \textit{\textbf{REGION INFO} - present only if NR $\ne$ 0} \\
-\# \textit{char *RLtype - Dimension of Lower order entity describing region} \\
+\# \textit{Keyword 'regions'}
+\# \textit{char *RLtype - keyword for lower order entity describing region} \\
 \# \textit{Values: Vertex, Face (case insensitive), e.g. VERtex or faCE}
@@
 
-RLtype
+regions RLtype
 \vspace{1ex}
@@
 \# \textit{if region described by vertices, then each region record has} \\
-\# \textit{int NRV - Number of region vertices} \\
-\# \textit{int Vid\_1 - ID of first vertex of region} \\
-\# \textit{int Vid\_2 - ID of second vertex of region} \\
+\# \textit{\textit{int} NRV - Number of region vertices} \\
+\# \textit{\textit{int} Vid\_1 - ID of first vertex of region} \\
+\# \textit{\textit{int} Vid\_2 - ID of second vertex of region} \\
 \# . . . \\
-\# \textit{int Vid\_NFE - ID of NRV'th vertex of region} \\
-\# \textit{int Mid, (NOTE: Mdim is not specified, since it has to be 3)}
+\# \textit{\textit{int} Vid\_NFE - ID of NRV'th vertex of region} \\
+\# \textit{\textit{int} Mid, (NOTE: Mdim is not specified, since it has to be 3)}
 \vspace{1ex}
@@
 \# \textit{if region described by faces, then each region record has} \\
-\# \textit{int NRF - Number of region faces} \\
-\# \textit{int Fid\_1 - signed ID of first face of region} \\
-\# \textit{int Fid\_2 - signed ID of second face of region} \\
+\# \textit{\textit{int} NRF - Number of region faces} \\
+\# \textit{\textit{int} Fid\_1 - signed ID of first face of region} \\
+\# \textit{\textit{int} Fid\_2 - signed ID of second face of region} \\
 \# . . . \\
-\# \textit{int Fid\_NRF - signed ID of NRF'th face of region} \\
-\# \textit{int Mdim, Mid} \\
+\# \textit{\textit{int} Fid\_NRF - signed ID of NRF'th face of region} \\
+\# \textit{\textit{int} Mdim, Mid} \\
 \#  \\
@@
 
+\# \textit{\textbf{NOT IMPLEMENTED}}
+
 \# \textit{\textbf{VERTEX ATTRIBUTES}} \\
-\# \textit{int NVA - Number of Vertex attributes} \\
+\# \textit{\textit{int} NVA - Number of Vertex attributes} \\
 \# \\
 \# \textit{char *VA\_name\_1 - Name of first vertex attribute} \\
-\# \textit{int VA\_type\_1 - Type of first vertex attribute} \\
-\# \textit{int VA\_dim\_1 - Dimension of first vertex attribute} \\
+\# \textit{\textit{int} VA\_type\_1 - Type of first vertex attribute} \\
+\# \textit{\textit{int} VA\_dim\_1 - Dimension of first vertex attribute} \\
 \# \\
 \# \textit{char *VA\_name\_2 - Name of second vertex attribute} \\
-\# \textit{int VA\_type\_2 - Type of second vertex attribute} \\
-\# \textit{int VA\_dim\_2 - Dimension of first vertex attribute} \\
+\# \textit{\textit{int} VA\_type\_2 - Type of second vertex attribute} \\
+\# \textit{\textit{int} VA\_dim\_2 - Dimension of first vertex attribute} \\
 \# \\
@@
 \# \textit{char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
-\# \textit{int VA\_type\_NVA - Type of NVA'th vertex attribute} \\
-\# \textit{int VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
+\# \textit{\textit{int} VA\_type\_NVA - Type of NVA'th vertex attribute} \\
+\# \textit{\textit{int} VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
 \# \\
@@
 
-\# \textit{For each vertex attribute record, list of attribute values} \\
+\# \textit{For each vertex attribute record, set of attribute values} \\
 \# \textit{E.G., there are 3 attributes for each vertex:} \\
@@
 
-VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
-VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
+VA\_\textit{int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
+VA\_\textit{int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
 . \\
@@
 
-\# \textit{For each edge attribute record, list of attribute values} \\
+\# \textit{For each edge attribute record, set of attribute values} \\
 \# \textit{E.G., a scalar int, a scalar double and a string}
@@
 
-EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
-EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
+EA\_\textit{int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
+EA\_\textit{int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
 . . . \\
@@
 
-\# \textit{For each face attribute record, list of attribute values} \\
+\# \textit{For each face attribute record, set of attribute values} \\
 \# \textit{E.G., a vector of 2 doubles and a string}
@@
 
-\# \textit{For each Region attribute record, list of attribute values} \\
+\# \textit{For each Region attribute record, set of attribute values} \\
 \# \textit{E.G., a vector of 3 ints}
@@
 
+\newpage
+
 \end{document}

### 2004-01-15 docs/MSTK.tex
Switched back to using List_* instead of Set_* since the word Set can mean assign as well as a group

@@
 \begin{description}
-\item[]\textbf{\textit{Set\_ptr}}: Handle to a Set object
+\item[]\textbf{\textit{List\_ptr}}: Handle to a List object
 
@@
 \section{MSTK Functional Interface}
-\subsection{Sets}
-
-Sets of entities in MSTK are returned as sets of type
-\textit{Set\_ptr}. The following are the set operations available in
+\subsection{List}
+
+Lists of entities in MSTK are returned as sets of type
+\textit{List\_ptr}. The following are the set operations available in
 MSTK:
@@
 \begin{description}
-\item[]\textbf{\textit{Set\_ptr} Set\_New(\textit{int} inisize):} Create a
+\item[]\textbf{\textit{List\_ptr} List\_New(\textit{int} inisize):} Create a
 new set with an initial size, \textit{inisize}. If \textit{inisize} is
@@
 
-\item[]\textbf{\textit{void} Set\_Delete(\textit{Set\_ptr} l):} Delete a set.
-
-\item[]\textbf{\textit{Set\_ptr} Set\_Compress(\textit{Set\_ptr} l):} Compress
+\item[]\textbf{\textit{void} List\_Delete(\textit{List\_ptr} l):} Delete a set.
+
+\item[]\textbf{\textit{List\_ptr} List\_Compress(\textit{List\_ptr} l):} Compress
 a set.  Doing this while an algorithm is iterating through the set can
-currently cause problems!! Calling \textbf{Set\_Compress} could change the
+currently cause problems!! Calling \textbf{List\_Compress} could change the
 pointer for the set due to reallocation.
 
-\item[]\textbf{\textit{Set\_ptr} Set\_Copy(\textit{Set\_ptr} l):} Return a
+\item[]\textbf{\textit{List\_ptr} List\_Copy(\textit{List\_ptr} l):} Return a
 copy of a set.
 
-\item[]\textbf{\textit{Set\_ptr} Set\_Add(\textit{Set\_ptr} l, \textit{void
+\item[]\textbf{\textit{List\_ptr} List\_Add(\textit{List\_ptr} l, \textit{void
   *}entry):} Add an entry to the set. The entry is appended to the end
@@
 
-\item[]\textbf{\textit{Set\_ptr} Set\_ChknAdd(\textit{Set\_ptr} l,
+\item[]\textbf{\textit{List\_ptr} List\_ChknAdd(\textit{List\_ptr} l,
 \textit{void *}entry):} Add an entry to a set only if it is not already
@@
 
-\item[]\textbf{\textit{int} Set\_Rem(\textit{Set\_ptr} l, \textit{void
+\item[]\textbf{\textit{int} List\_Rem(\textit{List\_ptr} l, \textit{void
   *}entry):} Remove an entry from the set. Returns 1 if successful, 0
@@
 
-\item[]\textbf{\textit{int} Set\_Remi(\textit{Set\_ptr} l, \textit{int} i):}
+\item[]\textbf{\textit{int} List\_Remi(\textit{List\_ptr} l, \textit{int} i):}
 Remove the i'th valid entry in the set. Returns 1 if successful, 0
@@
 
-\item[]\textbf{\textit{int} Set\_Replace(\textit{Set\_ptr} l, \textit{void
+\item[]\textbf{\textit{int} List\_Replace(\textit{List\_ptr} l, \textit{void
   *}entry, \textit{void *}nuentry):} Replace 'entry' with 'nuentry' in
@@
 
-\item[]\textbf{\textit{int} Set\_Replacei(\textit{Set\_ptr} l, \textit{int}
+\item[]\textbf{\textit{int} List\_Replacei(\textit{List\_ptr} l, \textit{int}
 i, void *nuentry):} Replace the i'th valid entry in the set with
@@
 
-\item[]\textbf{\textit{int} Set\_Contains(\textit{Set\_ptr} l, void *entry):}
+\item[]\textbf{\textit{int} List\_Contains(\textit{List\_ptr} l, void *entry):}
 Returns 1 if set contains the entry, 0 otherwise.
 
-\item[]\textbf{\textit{int} Set\_Locate(\textit{Set\_ptr} l, void *entry):}
+\item[]\textbf{\textit{int} List\_Locate(\textit{List\_ptr} l, void *entry):}
 Returns the positional index of the entry in the set. Returns -1 if
@@
 
-\item[]\textbf{\textit{void *}Set\_Entry(\textit{Set\_ptr} l, \textit{int}
+\item[]\textbf{\textit{void *}List\_Entry(\textit{List\_ptr} l, \textit{int}
 i):} Return the i'th valid entry in the set.  Returns a NULL pointer if
@@
 
-\item[]\textbf{\textit{void *}Set\_Next\_Entry(\textit{Set\_ptr} l,
+\item[]\textbf{\textit{void *}List\_Next\_Entry(\textit{List\_ptr} l,
 \textit{int} *i):} Return the next valid entry in the set. This routine
@@
 
-\item[]\textbf{\textit{int} Set\_Num\_Entries(\textit{Set\_ptr} l):} Return
+\item[]\textbf{\textit{int} List\_Num\_Entries(\textit{List\_ptr} l):} Return
 the number of entries in a set
@@
 mesh, \textit{int} *idx):} Returns the next vertex while iterating
-through the vertices of the mesh. See the routine \textbf{Set\_Next\_Entry}
+through the vertices of the mesh. See the routine \textbf{List\_Next\_Entry}
 above for an explanation of how the iteration works.
@@
 \textit{int} *idx):} Returns the next edge while iterating through the
-edges of the mesh. See the routine Set\_Next\_Entry above for an
+edges of the mesh. See the routine List\_Next\_Entry above for an
 explanation of how the iteration works.  The routine always returns
@@
 \textit{int} *idx):} Returns the next face while iterating through the
-faces of the mesh. See the routine \textbf{Set\_Next\_Entry} above for
+faces of the mesh. See the routine \textbf{List\_Next\_Entry} above for
 an explanation of how the iteration works.  Only faces explicitly
@@
 
-\item[]\textbf{\textit{Set\_ptr} MV\_AdjVertices(\textit{MVertex\_ptr}
-mvertex):} Set of adjacent or edge connected neighboring vertices of vertex.
-
-\item[]\textbf{\textit{Set\_ptr} MV\_Edges(\textit{MVertex\_ptr} mvertex):}
-Set of edges connected to the vertex.
-
-\item[]\textbf{\textit{Set\_ptr} MV\_Faces(\textit{MVertex\_ptr} mvertex):}
-Set of faces connected to the vertex.
-
-\item[]\textbf{\textit{Set\_ptr} MV\_Regions(\textit{MVertex\_ptr} mvertex):}
-Set of regions connected to the vertex.
+\item[]\textbf{\textit{List\_ptr} MV\_AdjVertices(\textit{MVertex\_ptr}
+mvertex):} List of adjacent or edge connected neighboring vertices of vertex.
+
+\item[]\textbf{\textit{List\_ptr} MV\_Edges(\textit{MVertex\_ptr} mvertex):}
+List of edges connected to the vertex.
+
+\item[]\textbf{\textit{List\_ptr} MV\_Faces(\textit{MVertex\_ptr} mvertex):}
+List of faces connected to the vertex.
+
+\item[]\textbf{\textit{List\_ptr} MV\_Regions(\textit{MVertex\_ptr} mvertex):}
+List of regions connected to the vertex.
 \end{description}
@@
 
-\item[]\textbf{\textit{Set\_ptr} ME\_Faces(\textit{MEdge\_ptr} medge):}
+\item[]\textbf{\textit{List\_ptr} ME\_Faces(\textit{MEdge\_ptr} medge):}
 Returns the set of faces using this edge.
 
-\item[]\textbf{\textit{Set\_ptr} ME\_Regions(\textit{MEdge\_ptr} medge):}
+\item[]\textbf{\textit{List\_ptr} ME\_Regions(\textit{MEdge\_ptr} medge):}
 Returns the set of regions using this edge.
@@
 
-\item[]\textbf{\textit{Set\_ptr} MF\_Vertices(\textit{MFace\_ptr}
+\item[]\textbf{\textit{List\_ptr} MF\_Vertices(\textit{MFace\_ptr}
     mface, \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
@@
 
-\item[]\textbf{\textit{Set\_ptr} MF\_Edges(\textit{MFace\_ptr} mface,
+\item[]\textbf{\textit{List\_ptr} MF\_Edges(\textit{MFace\_ptr} mface,
     \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
@@
 
-\item[]\textbf{\textit{Set\_ptr} MF\_AdjFaces(\textit{MFace\_ptr} mface):}
-Set of adacent faces of a face. This operator is relevant only in
+\item[]\textbf{\textit{List\_ptr} MF\_AdjFaces(\textit{MFace\_ptr} mface):}
+List of adacent faces of a face. This operator is relevant only in
 planar or surface surface meshes, i.e., for boundary faces not
@@
 
-\item[]\textbf{\textit{Set\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
+\item[]\textbf{\textit{List\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
 Return the set of regions connected to the face. If the face is not
@@
 
-\item[]\textbf{\textit{Set\_ptr} MR\_Vertices(\textit{MRegion\_ptr}
+\item[]\textbf{\textit{List\_ptr} MR\_Vertices(\textit{MRegion\_ptr}
 mregion):} Returns the set of vertices of a region. For standard
@@
 
-\item[]\textbf{\textit{Set\_ptr} MR\_Edges(\textit{MRegion\_ptr} mregion):}
+\item[]\textbf{\textit{List\_ptr} MR\_Edges(\textit{MRegion\_ptr} mregion):}
 Return the \underline{unordered} set of edges of a region.
 
-\item[]\textbf{\textit{Set\_ptr} MR\_Faces(\textit{MRegion\_ptr} mregion):}
+\item[]\textbf{\textit{List\_ptr} MR\_Faces(\textit{MRegion\_ptr} mregion):}
 Returns the set of faces of a region.
 
-\item[]\textbf{\textit{Set\_ptr} MR\_AdjRegions(\textit{MRegion\_ptr}
+\item[]\textbf{\textit{List\_ptr} MR\_AdjRegions(\textit{MRegion\_ptr}
 mregion):} Returns the set of adjacent regions of a region, i.e.,
@@
 
-\item[]\textbf{\textit{void} Set\_Mark(\textit{Set\_ptr} list, \textit{int} mkr):} Mark a set of entities with given marker.
-
-\item[]\textbf{\textit{void} Set\_Unmark(\textit{Set\_ptr} list, \textit{int} mkr):}
+\item[]\textbf{\textit{void} List\_Mark(\textit{List\_ptr} list, \textit{int} mkr):} Mark a set of entities with given marker.
+
+\item[]\textbf{\textit{void} List\_Unmark(\textit{List\_ptr} list, \textit{int} mkr):}
 Unmark a set of entities with respect to the given marker.
@@
 
-\item[]\textbf{\textit{void} Set\_PrintID(\textit{Set\_ptr} l):} Debugging
+\item[]\textbf{\textit{void} List\_PrintID(\textit{List\_ptr} l):} Debugging
 utility to print the IDs of the entities in a set.

### 2004-01-15 example/src/main.c

@@ -14,7 +14,7 @@ int main(int argc, char *argv[]) {
   MEdge_ptr e;
   MFace_ptr f;
   GEntity_ptr gent;
-  Set_ptr fedges;
+  List_ptr fedges;
 
   
   fprintf(stderr,"\n");
@@ -93,11 +93,11 @@ int main(int argc, char *argv[]) {
 
     /* Edges of face */
     fedges = MF_Edges(f,1,0);
-    ne = Set_Num_Entries(fedges);
+    ne = List_Num_Entries(fedges);
     fprintf(stderr,"Edges: %-d\n",ne);
     fprintf(stderr,"Object        ID      GEntID   GEntDim    Vertex IDs\n");
     idx2 = 0; i = 0;
-    while (e = Set_Next_Entry(fedges,&idx2)) {
+    while (e = List_Next_Entry(fedges,&idx2)) {
       edir = MF_EdgeDir_i(f,i);
       if (edir) 
 	fprintf(stderr,"0x%-8x    %-8d %-8d     %-1d       %-d  %-d\n",
@@ -110,7 +110,7 @@ int main(int argc, char *argv[]) {
       i++;
     }
     fprintf(stderr,"\n");
-    Set_Delete(fedges);
+    List_Delete(fedges);
   }
 
 

### 2004-01-15 src/misc/MF_CondNums.c
Switched back to using List_* instead of Set_* since the word Set can mean assign as well as a group

@@ -14,15 +14,15 @@ void MF_CondNums(MFace_ptr f, double *conds) {
   double xyz1[3][3], eps=1.0E-14;
   int i, k, nfv;
   MVertex_ptr vtx[MAXPV2];
-  Set_ptr fverts;
+  List_ptr fverts;
 
   fverts = MF_Vertices(f,1,0);
-  nfv = Set_Num_Entries(fverts);
+  nfv = List_Num_Entries(fverts);
   for (i = 0; i < nfv; i++) {
-    vtx[i] = Set_Entry(fverts,i);
+    vtx[i] = List_Entry(fverts,i);
     MV_Coords(vtx[i],xyz[i]);
   }
-  Set_Delete(fverts);
+  List_Delete(fverts);
     
 
   for (i = 0; i < nfv; i++) {

### 2004-01-15 src/misc/MSTK_Print.c

@@ -14,7 +14,7 @@ extern "C" {
     MEdge_ptr e;
     MFace_ptr f;
     MRegion_ptr r;
-    Set_ptr vedges, vfaces, vregions;
+    List_ptr vedges, vfaces, vregions;
     int idx, ne, nf, nr;
     GEntity_ptr gent;
     
@@ -48,51 +48,51 @@ extern "C" {
 	
 	
 	vedges = MV_Edges(v);
-	ne = Set_Num_Entries(vedges);
+	ne = List_Num_Entries(vedges);
 	
 	if (ne) {
 	  fprintf(stderr,"Edges:%-d\n",ne);
 	  
 	  fprintf(stderr,"Object        ID      GEntID   GEntDim    Vertex IDs\n");
 	  idx = 0;
-	  while ((e = Set_Next_Entry(vedges,&idx))) {
+	  while ((e = List_Next_Entry(vedges,&idx))) {
 	    fprintf(stderr,"0x%-8x    %-8d %-8d     %-1d    %-d  %-d\n",
 		    e,ME_ID(e),ME_GEntID(e),ME_GEntDim(e),
 		    MV_ID(ME_Vertex(e,0)),MV_ID(ME_Vertex(e,1)));
 	  }
 	  fprintf(stderr,"\n");
-	  Set_Delete(vedges);
+	  List_Delete(vedges);
 	}
 	
 	
 	vfaces = MV_Faces(v);
 	if (vfaces) {
-	  nf = Set_Num_Entries(vfaces);
+	  nf = List_Num_Entries(vfaces);
 	  fprintf(stderr,"Faces: %-d\n",nf);
 	  
 	  fprintf(stderr,"Object        ID      GEntID   GEntDim\n");
 	  idx = 0;
-	  while ((f = Set_Next_Entry(vfaces,&idx)))
+	  while ((f = List_Next_Entry(vfaces,&idx)))
 	    fprintf(stderr,"0x%-8x    %-8d %-8d     %-1d\n",f,MF_ID(f),
 		    MF_GEntID(f),ME_GEntDim(f));
 	  fprintf(stderr,"\n");
-	  Set_Delete(vfaces);
+	  List_Delete(vfaces);
 	}
 	
 	
 	vregions= MV_Regions(v);
 	if (vregions) {
-	  nr = Set_Num_Entries(vregions);
+	  nr = List_Num_Entries(vregions);
 	  fprintf(stderr,"Regions: %-d",nr);
 	  
[…]

### 2004-01-15 include/MFace.h
Switched back to using List_* instead of Set_* since the word Set can mean assign as well as a group

@@ -6,7 +6,7 @@ extern "C" {
 #endif
 
 #include "MSTK_types.h"
-#include "Set.h"
+#include "List.h"
 
 #ifdef _H_MFace_Private
   typedef struct MFace {
@@ -37,7 +37,7 @@ extern "C" {
 
   typedef struct MFace_SameAdj_R2R4 {
     unsigned char nadj;
-    Set_ptr adjfaces;
+    List_ptr adjfaces;
   } MFace_SameAdj_R2R4;
 
   /*----- Downward adjacency definitions --------*/
@@ -45,12 +45,12 @@ extern "C" {
   typedef struct MFace_DownAdj_FN {
     unsigned char ne;
     int edirs;
-    Set_ptr fedges;
+    List_ptr fedges;
   } MFace_DownAdj_FN;
 
   typedef struct MFace_DownAdj_R3R4 {
     unsigned char nv;
-    Set_ptr *fvertices;
+    List_ptr *fvertices;
   } MFace_DownAdj_R3R4;
   
 #else
@@ -83,13 +83,13 @@ extern "C" {
   int MF_Num_Vertices(MFace_ptr f);
   int MF_Num_Edges(MFace_ptr f);
   int MF_Num_AdjFaces(MFace_ptr f);
-  Set_ptr MF_Vertices(MFace_ptr f, int dir, MVertex_ptr v);
-  Set_ptr MF_Edges(MFace_ptr f, int dir, MVertex_ptr v);
-  Set_ptr MF_AdjFaces(MFace_ptr f);
+  List_ptr MF_Vertices(MFace_ptr f, int dir, MVertex_ptr v);
+  List_ptr MF_Edges(MFace_ptr f, int dir, MVertex_ptr v);
+  List_ptr MF_AdjFaces(MFace_ptr f);
   int MF_EdgeDir(MFace_ptr f, MEdge_ptr e);
   int MF_EdgeDir_i(MFace_ptr f, int i);
   int MF_UsesEntity(MFace_ptr f, MEntity_ptr e, int type);
-  Set_ptr MF_Regions(MFace_ptr f);
+  List_ptr MF_Regions(MFace_ptr f);
   MRegion_ptr MF_Region(MFace_ptr f, int side);
 
   /* Calling applications can only set the representation type for the

### 2004-01-15 include/MRegion_jmp.h

@@ -51,41 +51,41 @@ static void
 {MR_Set_Vertices_F1, MR_Set_Vertices_F4, MR_Set_Vertices_R1, 
  MR_Set_Vertices_R2, MR_Set_Vertices_R4};
 
-Set_ptr MR_Vertices_F1(MRegion_ptr r);
-Set_ptr MR_Vertices_F4(MRegion_ptr r);
-Set_ptr MR_Vertices_R1(MRegion_ptr r);
-Set_ptr MR_Vertices_R2(MRegion_ptr r);
-Set_ptr MR_Vertices_R4(MRegion_ptr r);
-Set_ptr MR_Vertices_FNR3R4(MRegion_ptr r);
-static Set_ptr (*MR_Vertices_jmp[MSTK_MAXREP])(MRegion_ptr r) =
+List_ptr MR_Vertices_F1(MRegion_ptr r);
+List_ptr MR_Vertices_F4(MRegion_ptr r);
+List_ptr MR_Vertices_R1(MRegion_ptr r);
+List_ptr MR_Vertices_R2(MRegion_ptr r);
+List_ptr MR_Vertices_R4(MRegion_ptr r);
+List_ptr MR_Vertices_FNR3R4(MRegion_ptr r);
+static List_ptr (*MR_Vertices_jmp[MSTK_MAXREP])(MRegion_ptr r) =
 {MR_Vertices_F1, MR_Vertices_F4, MR_Vertices_R1, MR_Vertices_R2, 
 MR_Vertices_R4};
 
-Set_ptr MR_Edges_F1(MRegion_ptr r);
-Set_ptr MR_Edges_F4(MRegion_ptr r);
-Set_ptr MR_Edges_R1(MRegion_ptr r);
-Set_ptr MR_Edges_R2(MRegion_ptr r);
-Set_ptr MR_Edges_R4(MRegion_ptr r);
-Set_ptr MR_Edges_FNR3R4(MRegion_ptr r);
-static Set_ptr (*MR_Edges_jmp[MSTK_MAXREP])(MRegion_ptr r) =
+List_ptr MR_Edges_F1(MRegion_ptr r);
+List_ptr MR_Edges_F4(MRegion_ptr r);
+List_ptr MR_Edges_R1(MRegion_ptr r);
+List_ptr MR_Edges_R2(MRegion_ptr r);
+List_ptr MR_Edges_R4(MRegion_ptr r);
[…]

### 2004-01-15 include/MSTK.h

@@ -9,7 +9,7 @@ extern "C" {
 #include "MSTK_defines.h"
 #include "MSTK_types.h"
 #include "MSTK_util.h"
-#include "Set.h"
+#include "List.h"
 #include "MSTK_malloc.h"
 
 
@@ -75,10 +75,10 @@ void        MSTK_Init(void);
   int MV_Num_Edges(MVertex_ptr mvertex);
   int MV_Num_Faces(MVertex_ptr mvertex);
   int MV_Num_Regions(MVertex_ptr mvertex);
-  Set_ptr MV_AdjVertices(MVertex_ptr mvertex);
-  Set_ptr MV_Edges(MVertex_ptr mvertex);
-  Set_ptr MV_Faces(MVertex_ptr mvertex);
-  Set_ptr MV_Regions(MVertex_ptr mvertex);
+  List_ptr MV_AdjVertices(MVertex_ptr mvertex);
+  List_ptr MV_Edges(MVertex_ptr mvertex);
+  List_ptr MV_Faces(MVertex_ptr mvertex);
+  List_ptr MV_Regions(MVertex_ptr mvertex);
 
 
 /********************************************************************/
@@ -106,8 +106,8 @@ void        MSTK_Init(void);
   MVertex_ptr ME_Vertex(MEdge_ptr medge, int i);
   MVertex_ptr ME_OppVertex(MEdge_ptr medge, MVertex_ptr ov);
   int ME_UsesEntity(MEdge_ptr medge, MEntity_ptr mentity, int etype);
-  Set_ptr ME_Faces(MEdge_ptr medge);
-  Set_ptr ME_Regions(MEdge_ptr medge);
+  List_ptr ME_Faces(MEdge_ptr medge);
+  List_ptr ME_Regions(MEdge_ptr medge);
 
 
   MEdge_ptr MVs_CommonEdge(MVertex_ptr v1, MVertex_ptr v2);
@@ -145,13 +145,13 @@ void        MSTK_Init(void);
   int MF_Num_Vertices(MFace_ptr mface);
   int MF_Num_Edges(MFace_ptr mface);
   int MF_Num_AdjFaces(MFace_ptr mface);
-  Set_ptr MF_Vertices(MFace_ptr mface, int dir, MVertex_ptr mvert);
-  Set_ptr MF_Edges(MFace_ptr mface, int dir, MVertex_ptr mvert);
-  Set_ptr MF_AdjFaces(MFace_ptr mface);
[…]

### 2004-01-15 include/MSTK_private.h

@@ -8,7 +8,7 @@ extern "C" {
   /* #include "gmtk.h" */
 #include "MSTK_types.h"
 #include "MSTK_util.h"
-#include "Set.h"
+#include "List.h"
 #include "MSTK_malloc.h"
 #include "MSTK.h"
 

### 2004-01-15 include/MVertex_jmp.h

@@ -77,59 +77,59 @@ extern "C" {
    MV_Num_Regions_R2, MV_Num_Regions_R3R4};
 #endif
 
-  Set_ptr MV_AdjVertices_F1(MVertex_ptr v);
-  Set_ptr MV_AdjVertices_F4(MVertex_ptr v);
-  Set_ptr MV_AdjVertices_R1(MVertex_ptr v);
-  Set_ptr MV_AdjVertices_R2(MVertex_ptr v);
-  Set_ptr MV_AdjVertices_R4(MVertex_ptr v);
-  static Set_ptr (*MV_AdjVertices_jmp[MSTK_MAXREP])(MVertex_ptr v) =
+  List_ptr MV_AdjVertices_F1(MVertex_ptr v);
+  List_ptr MV_AdjVertices_F4(MVertex_ptr v);
+  List_ptr MV_AdjVertices_R1(MVertex_ptr v);
+  List_ptr MV_AdjVertices_R2(MVertex_ptr v);
+  List_ptr MV_AdjVertices_R4(MVertex_ptr v);
+  static List_ptr (*MV_AdjVertices_jmp[MSTK_MAXREP])(MVertex_ptr v) =
   {MV_AdjVertices_F1, MV_AdjVertices_F4, MV_AdjVertices_R1,
    MV_AdjVertices_R2, MV_AdjVertices_R4};
 
-  Set_ptr MV_Edges_F1(MVertex_ptr v);
-  Set_ptr MV_Edges_F4(MVertex_ptr v);
-  Set_ptr MV_Edges_R1(MVertex_ptr v);
-  Set_ptr MV_Edges_R2(MVertex_ptr v);
-  Set_ptr MV_Edges_R4(MVertex_ptr v);
-  Set_ptr MV_Edges_R2R4(MVertex_ptr v);
+  List_ptr MV_Edges_F1(MVertex_ptr v);
+  List_ptr MV_Edges_F4(MVertex_ptr v);
+  List_ptr MV_Edges_R1(MVertex_ptr v);
+  List_ptr MV_Edges_R2(MVertex_ptr v);
+  List_ptr MV_Edges_R4(MVertex_ptr v);
+  List_ptr MV_Edges_R2R4(MVertex_ptr v);
 #ifdef DEBUG
-  static Set_ptr (*MV_Edges_jmp[MSTK_MAXREP])(MVertex_ptr v) = 
[…]

### 2004-01-15 include/Mesh.h

@@ -2,7 +2,7 @@
 #define _H_Mesh
 
 #include "MSTK_types.h"
-#include "Set.h"
+#include "List.h"
 
 #ifdef _cplusplus
 extern "C" {
@@ -12,7 +12,7 @@ extern "C" {
   typedef struct Mesh {
     RepType reptype;
     int nv, ne, nf, nr;
-    Set_ptr mvertex, medge, mface, mregion;
+    List_ptr mvertex, medge, mface, mregion;
     GModel_ptr geom;
     int max_vid, max_eid, max_fid, max_rid;
   } Mesh, *Mesh_ptr;

### 2004-01-15 src/base/ME_Face_Funcs_R1R2.c

@@ -14,7 +14,7 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr ME_Faces_R1R2(MEdge_ptr e) {
+  List_ptr ME_Faces_R1R2(MEdge_ptr e) {
     MSTK_Report("ME_Faces_R1R2","Not Implemented",ERROR);
     return 0;
   }

### 2004-01-15 src/base/ME_Region_Funcs_R3R4.c

@@ -14,38 +14,38 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr ME_Regions_R3R4(MEdge_ptr e) {
+  List_ptr ME_Regions_R3R4(MEdge_ptr e) {
     int i, k, mkr, ner, nef;
-    Set_ptr eregions, efaces;
+    List_ptr eregions, efaces;
     MFace_ptr eface;
     MRegion_ptr freg;
 
     ner = 0;
-    eregions = Set_New(10);
+    eregions = List_New(10);
     mkr = MSTK_GetMarker();
 
     efaces = ME_Faces(e);
-    nef = Set_Num_Entries(efaces);
+    nef = List_Num_Entries(efaces);
 
     for (i = 0; i < nef; i++) {
-      eface = Set_Entry(efaces,i);
+      eface = List_Entry(efaces,i);
       for (k = 0; k < 2; k++) {
 	freg = MF_Region(eface,k);
 	if (freg && !MEnt_IsMarked(freg,mkr)) {
 	  MEnt_Mark(freg,mkr);
-	  Set_Add(eregions,freg);
+	  List_Add(eregions,freg);
 	  ner++;
 	}
       }
     }
-    Set_Delete(efaces);
-    Set_Unmark(eregions,mkr);
+    List_Delete(efaces);
+    List_Unmark(eregions,mkr);
     MSTK_FreeMarker(mkr);
 
     if (ner)
       return eregions;
     else {
-      Set_Delete(eregions);
+      List_Delete(eregions);
       return 0;
     }    
   }

### 2004-01-15 src/base/MEdge.c

@@ -100,11 +100,11 @@ extern "C" {
     return (*ME_Num_Regions_jmp[e->repType])(e);
   }
 
-  Set_ptr ME_Faces(MEdge_ptr e) {
+  List_ptr ME_Faces(MEdge_ptr e) {
     return (*ME_Faces_jmp[e->repType])(e);
   }
 
-  Set_ptr ME_Regions(MEdge_ptr e) {
+  List_ptr ME_Regions(MEdge_ptr e) {
     return (*ME_Regions_jmp[e->repType])(e);
   }
 
@@ -148,7 +148,7 @@ extern "C" {
 
 
   MEdge_ptr MVs_CommonEdge(MVertex_ptr v1, MVertex_ptr v2) {
-    Set_ptr vedges;
+    List_ptr vedges;
     MEdge_ptr edge=0;
     MVertex_ptr ev;
     int i, found, nve;
@@ -157,10 +157,10 @@ extern "C" {
     if (!vedges)
       return 0;
 
-    nve = Set_Num_Entries(vedges);
+    nve = List_Num_Entries(vedges);
 
     for (i = 0, found = 0; i < nve; i++) {
-      edge = Set_Entry(vedges,i);
+      edge = List_Entry(vedges,i);
       ev = ME_Vertex(edge,0);
       if (ev == v2) {
         found = 1;
@@ -174,7 +174,7 @@ extern "C" {
         }
       }
     }
-    Set_Delete(vedges);
+    List_Delete(vedges);
 
     return (found ? edge : 0);
   }

### 2004-01-15 src/base/MEdge_R1.c

@@ -21,7 +21,7 @@ extern "C" {
     return ME_Num_Faces_R1R2(e);
   }
 
-  Set_ptr ME_Faces_R1(MEdge_ptr e) {
+  List_ptr ME_Faces_R1(MEdge_ptr e) {
     return ME_Faces_R1R2(e);
   }
 
@@ -29,7 +29,7 @@ extern "C" {
     return ME_Num_Regions_R1R2(e);
   }
 
-  Set_ptr ME_Regions_R1(MEdge_ptr e) {
+  List_ptr ME_Regions_R1(MEdge_ptr e) {
     return ME_Regions_R1R2(e);
   }
 

### 2004-01-15 src/base/MEdge_R4.c

@@ -25,11 +25,11 @@ extern "C" {
     return ME_Num_Regions_R3R4(e);
   }
 
-  Set_ptr ME_Faces_R4(MEdge_ptr e) {
+  List_ptr ME_Faces_R4(MEdge_ptr e) {
     return ME_Faces_R3R4(e);
   }
 
-  Set_ptr ME_Regions_R4(MEdge_ptr e) {
+  List_ptr ME_Regions_R4(MEdge_ptr e) {
     return ME_Regions_R3R4(e);
   }
 

### 2004-01-15 src/base/MEntity.c

@@ -4,7 +4,7 @@
 #include <stdlib.h>
 #include <string.h>
 #include "MEntity.h"
-#include "Set.h"
+#include "List.h"
 #include "MSTK_malloc.h"
 #include "MSTK.h"
 
@@ -32,22 +32,22 @@ extern "C" {
     ent->marker = ent->marker & ~(1<<(markerID-1));
   }
 
-  void Set_Mark(Set_ptr list, int markerID) {
+  void List_Mark(List_ptr list, int markerID) {
     MEntity_ptr ent;
-    int i, n = Set_Num_Entries(list);
+    int i, n = List_Num_Entries(list);
     
     for (i = 0; i < n; i++) {
-      ent = (MEntity_ptr) Set_Entry(list,i);
+      ent = (MEntity_ptr) List_Entry(list,i);
       MEnt_Mark(ent,markerID);
     }
   }
 
-  void Set_Unmark(Set_ptr list, int markerID) {
+  void List_Unmark(List_ptr list, int markerID) {
     MEntity_ptr ent;
-    int i, n = Set_Num_Entries(list);
+    int i, n = List_Num_Entries(list);
     
     for (i = 0; i < n; i++) {
-      ent = (MEntity_ptr) Set_Entry(list,i);
+      ent = (MEntity_ptr) List_Entry(list,i);
       MEnt_Unmark(ent,markerID);
     }
   }

### 2004-01-15 src/base/MF_Edge_Funcs_FN.c

@@ -16,11 +16,11 @@ extern "C" {
 
     downadj->ne = n;
     downadj->edirs = 0;
-    downadj->fedges = Set_New(n);
+    downadj->fedges = List_New(n);
     
     for (i = 0; i < n; i++) {
       downadj->edirs = downadj->edirs | (dir[i] << i);
-      Set_Add(downadj->fedges,e[i]);
+      List_Add(downadj->fedges,e[i]);
       ME_Add_Face(e[i],f);
     }
   }
@@ -35,11 +35,11 @@ extern "C" {
       MSTK_Report("MF_Replace_Edge_i","No initial set of edges for face",
 		  ERROR);
 
-    olde = Set_Entry(downadj->fedges,i);
+    olde = List_Entry(downadj->fedges,i);
       
     downadj->edirs = (downadj->edirs & ~(1<<i)); /* set bit i to 0 */
     downadj->edirs = (downadj->edirs | (dir<<i)); /* set to dir */
-    Set_Replacei(downadj->fedges,i,e);
+    List_Replacei(downadj->fedges,i,e);
 
     ME_Rem_Face(olde,f);
     ME_Add_Face(e,f);
@@ -58,7 +58,7 @@ extern "C" {
 		  ERROR);
       
     for (i = 0; i < downadj->ne; i++)
-      if ((olde = Set_Entry(downadj->fedges,i)) == e) {
+      if ((olde = List_Entry(downadj->fedges,i)) == e) {
 	found = 1;
 	break;
       }
@@ -69,7 +69,7 @@ extern "C" {
 
     downadj->edirs = (downadj->edirs & ~(1<<i)); /* set bit i to 0 */
     downadj->edirs = (downadj->edirs | (dir<<i)); /* set to dir */
-    Set_Replacei(downadj->fedges,i,e);
+    List_Replacei(downadj->fedges,i,e);
 
     ME_Rem_Face(olde,f);
     ME_Add_Face(e,f);
@@ -79,9 +79,9 @@ extern "C" {
     return ((MFace_DownAdj_FN *) f->downadj)->ne;
   }	
 
-  Set_ptr MF_Edges_FN(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Edges_FN(MFace_ptr f, int dir, MVertex_ptr v0) {
     int i, k, n, fnd, edir;
-    Set_ptr fedges;
+    List_ptr fedges;
[…]

### 2004-01-15 src/base/MF_Edge_Funcs_R1R2.c

@@ -13,7 +13,7 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr MF_Edges_R1R2(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Edges_R1R2(MFace_ptr f, int dir, MVertex_ptr v0) {
     MSTK_Report("MF_Edges_R1R2","Not implemented",ERROR);
     return 0;    
   }

### 2004-01-15 src/base/MF_Edge_Funcs_R3R4.c

@@ -13,7 +13,7 @@ extern "C" {
     return ((MFace_DownAdj_R3R4 *)f->downadj)->nv;
   }
 
-  Set_ptr MF_Edges_R3R4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Edges_R3R4(MFace_ptr f, int dir, MVertex_ptr v0) {
     /* Are we supposed to create edges and send them back ? */
     MSTK_Report("MF_Edges_R3R4","Not implemented",ERROR);
     return NULL;

### 2004-01-15 src/base/MF_Region_Funcs_R1R2.c

@@ -8,7 +8,7 @@
 extern "C" {
 #endif
 
-  Set_ptr MF_Regions_R1R2(MFace_ptr f) {
+  List_ptr MF_Regions_R1R2(MFace_ptr f) {
     MSTK_Report("MF_Regions_R1R2","Not implemented",ERROR);
     return 0;    
   }

### 2004-01-15 src/base/MF_Region_Funcs_R3R4.c

@@ -9,17 +9,17 @@ extern "C" {
 #endif
 
 
-  Set_ptr MF_Regions_R3R4(MFace_ptr f) {
-    Set_ptr fregs;
+  List_ptr MF_Regions_R3R4(MFace_ptr f) {
+    List_ptr fregs;
     MFace_UpAdj_R3R4 *upadj;
 
     upadj = (MFace_UpAdj_R3R4 *) f->upadj;
 
-    fregs = Set_New(2);
+    fregs = List_New(2);
     if (upadj->fregions[0])
-      Set_Add(fregs,upadj->fregions[0]);
+      List_Add(fregs,upadj->fregions[0]);
     if (upadj->fregions[1])
-      Set_Add(fregs,upadj->fregions[1]);
+      List_Add(fregs,upadj->fregions[1]);
 
     return fregs;
   }

### 2004-01-15 src/base/MF_Vertex_Funcs_R1R2.c

@@ -13,7 +13,7 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr MF_Vertices_R1R2(MFace_ptr f, int dir) {
+  List_ptr MF_Vertices_R1R2(MFace_ptr f, int dir) {
     MSTK_Report("MF_Vertices_R1R2","Not implemented",ERROR);
     return 0;
   }

### 2004-01-15 src/base/MF_Vertex_Funcs_R3R4.c

@@ -15,10 +15,10 @@ extern "C" {
 
     downadj = f->downadj;
     downadj->nv =n;
-    downadj->fvertices = Set_New(n);
+    downadj->fvertices = List_New(n);
 
     for (i = 0; i < n; i++)
-      Set_Add(downadj->fvertices,v[i]);
+      List_Add(downadj->fvertices,v[i]);
   }
 
   void MF_Replace_Vertex_i_R3R4(MFace_ptr f, int i, MVertex_ptr v) {
@@ -28,7 +28,7 @@ extern "C" {
     if (downadj->nv == 0)
       MSTK_Report("MF_Replace_Vertex_R3R4","No initial set of vertices for face",ERROR);
 
-    Set_Replacei(downadj->fvertices,i,v);
+    List_Replacei(downadj->fvertices,i,v);
   }
 
   void MF_Replace_Vertex_R3R4(MFace_ptr f, MVertex_ptr v, MVertex_ptr nuv) {
@@ -38,7 +38,7 @@ extern "C" {
     if (downadj->nv == 0)
       MSTK_Report("MF_Replace_Vertex_R3R4","No initial set of vertices for face",ERROR);
 
-    Set_Replace(downadj->fvertices,v,nuv);
+    List_Replace(downadj->fvertices,v,nuv);
   }
 
   int MF_Num_Vertices_R3R4(MFace_ptr f) {
@@ -47,9 +47,9 @@ extern "C" {
     return downadj->nv;
   }
 
-  Set_ptr MF_Vertices_R3R4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Vertices_R3R4(MFace_ptr f, int dir, MVertex_ptr v0) {
     MFace_DownAdj_R3R4 *downadj;
-    Set_ptr fverts;
+    List_ptr fverts;
     int i, k, nv, fnd;
 
     downadj = (MFace_DownAdj_R3R4 *) f->downadj;
@@ -57,19 +57,19 @@ extern "C" {
 
     if (!v0) {
       if (dir) 
-	fverts = Set_Copy(downadj->fvertices);
+	fverts = List_Copy(downadj->fvertices);
       else {
-	fverts = Set_New(nv);
+	fverts = List_New(nv);
 
 	for (i = 0; i < nv; i++)
-	  Set_Add(fverts,Set_Entry(downadj->fvertices,i));
[…]

### 2004-01-15 src/base/MFace_F4.c

@@ -23,7 +23,7 @@ extern "C" {
     MFace_DownAdj_FN *downadj;
 
     downadj = (MFace_DownAdj_FN *) f->downadj;
-    Set_Delete(downadj->fedges);
+    List_Delete(downadj->fedges);
     MSTK_free(downadj);
   }
 
@@ -76,12 +76,12 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr MF_Vertices_F4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Vertices_F4(MFace_ptr f, int dir, MVertex_ptr v0) {
     return MF_Vertices_FN(f,dir,v0);
   }
 	
 
-  Set_ptr MF_Edges_F4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Edges_F4(MFace_ptr f, int dir, MVertex_ptr v0) {
     return MF_Edges_FN(f,dir,v0);
   }
 
@@ -101,7 +101,7 @@ extern "C" {
     return MF_UsesVertex_FN(f,v);
   }
 
-  Set_ptr MF_AdjFaces_F4(MFace_ptr f) {
+  List_ptr MF_AdjFaces_F4(MFace_ptr f) {
 #ifdef DEBUG
     MSTK_Report("MF_AdjFaces",
 		"Function call not suitable for this representation",WARN);
@@ -109,8 +109,8 @@ extern "C" {
     return NULL;
   }
   
-  Set_ptr MF_Regions_F4(MFace_ptr f) {
-    Set_ptr fregs, eregs;
+  List_ptr MF_Regions_F4(MFace_ptr f) {
+    List_ptr fregs, eregs;
     MRegion_ptr r;
     MEdge_ptr e;
     MFace_DownAdj_FN *downadj;
@@ -118,37 +118,37 @@ extern "C" {
     
     downadj = (MFace_DownAdj_FN *) f->downadj;
     
-    fregs = Set_New(2);
-    e = Set_Entry(downadj->fedges,0);
+    fregs = List_New(2);
+    e = List_Entry(downadj->fedges,0);
     
     eregs = ME_Regions(e);
     if (eregs) {
-      nr = Set_Num_Entries(eregs);
+      nr = List_Num_Entries(eregs);
       
       for (i = 0; i < nr; i++) {
-	r = Set_Entry(eregs,i);
+	r = List_Entry(eregs,i);
 	if (MR_UsesEntity(r,f,2)) {
[…]

### 2004-01-15 src/base/MFace_R4.c

@@ -29,7 +29,7 @@ extern "C" {
     MSTK_free(upadj);
 
     downadj = (MFace_DownAdj_R3R4 *) f->downadj;
-    Set_Delete(downadj->fvertices);
+    List_Delete(downadj->fvertices);
     MSTK_free(downadj);
   }
   
@@ -83,12 +83,12 @@ extern "C" {
     return MF_Num_Edges_R3R4(f);
   }
 
-  Set_ptr MF_Vertices_R4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Vertices_R4(MFace_ptr f, int dir, MVertex_ptr v0) {
     return MF_Vertices_R3R4(f,dir,v0);
   }
 	
 
-  Set_ptr MF_Edges_R4(MFace_ptr f, int dir, MVertex_ptr v0) {
+  List_ptr MF_Edges_R4(MFace_ptr f, int dir, MVertex_ptr v0) {
     return MF_Edges_R3R4(f,dir,v0); 
   }
 
@@ -108,7 +108,7 @@ extern "C" {
     return MF_UsesVertex_R3R4(f,v);
   }
 
-  Set_ptr MF_Regions_R4(MFace_ptr f) {
+  List_ptr MF_Regions_R4(MFace_ptr f) {
     return MF_Regions_R3R4(f);
   }
 
@@ -122,7 +122,7 @@ extern "C" {
     return 0;
   }
 
-  Set_ptr MF_AdjFaces_R4(MFace_ptr f) {
+  List_ptr MF_AdjFaces_R4(MFace_ptr f) {
     MSTK_Report("MF_AdjFaces_R1","Not yet implemented for this representation",WARN);
     return 0;
   }

### 2004-01-15 src/base/MRegion.c

@@ -104,7 +104,7 @@ extern "C" {
       /* it could be a pyramid or a triangular prism */
       nt = 0;
       for (i = 0; i < downadj->nf; i++) {
-	face = Set_Entry(downadj->rfaces,i);
+	face = List_Entry(downadj->rfaces,i);
 	if (MF_Num_Edges(face) == 3)
 	  nt++;
       }
@@ -140,19 +140,19 @@ extern "C" {
   }
   */
 
-  Set_ptr MR_Vertices(MRegion_ptr r) {
+  List_ptr MR_Vertices(MRegion_ptr r) {
     return (*MR_Vertices_jmp[r->repType])(r);
   }
 
-  Set_ptr MR_Edges(MRegion_ptr r) {
+  List_ptr MR_Edges(MRegion_ptr r) {
     return (*MR_Edges_jmp[r->repType])(r);
   }
 
-  Set_ptr MR_Faces(MRegion_ptr r) {
+  List_ptr MR_Faces(MRegion_ptr r) {
     return (*MR_Faces_jmp[r->repType])(r);
   }
 
-  Set_ptr MR_AdjRegions(MRegion_ptr r) {
+  List_ptr MR_AdjRegions(MRegion_ptr r) {
     return (*MR_AdjRegions_jmp[r->repType])(r);
   }
 

### 2004-01-15 src/base/MRegion_F1.c

@@ -35,19 +35,19 @@ extern "C" {
   }
   */
 
-  Set_ptr MR_Vertices_F1(MRegion_ptr r) {
+  List_ptr MR_Vertices_F1(MRegion_ptr r) {
     return MR_Vertices_FNR3R4(r);
   }
 
-  Set_ptr MR_Edges_F1(MRegion_ptr r) {
+  List_ptr MR_Edges_F1(MRegion_ptr r) {
     return MR_Edges_FNR3R4(r);
   }
 
-  Set_ptr MR_Faces_F1(MRegion_ptr r) {
+  List_ptr MR_Faces_F1(MRegion_ptr r) {
     return MR_Faces_FNR3R4(r);
   }
 
-  Set_ptr MR_AdjRegions_F1(MRegion_ptr r) {
+  List_ptr MR_AdjRegions_F1(MRegion_ptr r) {
     return MR_AdjRegions_FNR3R4(r);
   }
 

### 2004-01-15 src/base/MRegion_FNR3R4.c

@@ -27,12 +27,12 @@ extern "C" {
 
     downadj = (MRegion_DownAdj_FN *) r->downadj;
 
-    nf = Set_Num_Entries(downadj->rfaces);
+    nf = List_Num_Entries(downadj->rfaces);
     for (i = 0; i < nf; i++) {
-      f = Set_Entry(downadj->rfaces,i);
+      f = List_Entry(downadj->rfaces,i);
       MF_Rem_Region(f,r);
     }
-    Set_Delete(downadj->rfaces);
+    List_Delete(downadj->rfaces);
     MSTK_free(downadj);
 
     MSTK_free(r);
@@ -45,11 +45,11 @@ extern "C" {
     downadj = (MRegion_DownAdj_FN *) r->downadj;
     downadj->nf = nf;
     downadj->fdirs = 0;
-    downadj->rfaces = Set_New(nf);
+    downadj->rfaces = List_New(nf);
 
     for (i = 0; i < nf; i++) {
       downadj->fdirs = downadj->fdirs | (dirs[i] << i);
-      Set_Add(downadj->rfaces,rfaces[i]);
+      List_Add(downadj->rfaces,rfaces[i]);
       MF_Add_Region(rfaces[i],r,!dirs[i]);
     }
   }
@@ -66,7 +66,7 @@ extern "C" {
   }
 
   int MR_Num_AdjRegions_FNR3R4(MRegion_ptr r) {
-    Set_ptr adjr;
+    List_ptr adjr;
     int nr;
 
 #ifdef DEBUG
@@ -77,20 +77,20 @@ extern "C" {
 
     adjr = MR_AdjRegions(r);
     if (adjr) {
-      nr = Set_Num_Entries(adjr);
-      Set_Delete(adjr);
+      nr = List_Num_Entries(adjr);
+      List_Delete(adjr);
       return nr;
     }
     else
       return 0;
   }
 
-  Set_ptr MR_Vertices_FNR3R4(MRegion_ptr r) {
+  List_ptr MR_Vertices_FNR3R4(MRegion_ptr r) {
     int i, j, n, mkr, found, diradj0, diropp, edir, fdir;
     MFace_ptr face, fadj0, fopp;
     MEdge_ptr edge;
     MVertex_ptr vert, rv0, rvopp0;
-    Set_ptr rvertices, fverts, fedges;
+    List_ptr rvertices, fverts, fedges;
     MRegion_DownAdj_FN *downadj;
 
     downadj = (MRegion_DownAdj_FN *) r->downadj;
[…]

### 2004-01-15 src/base/MV_Edge_Funcs_R2R4.c

@@ -11,18 +11,18 @@ extern "C" {
   /* Functions */
 
 
-  Set_ptr MV_Edges_R2R4(MVertex_ptr v) {
+  List_ptr MV_Edges_R2R4(MVertex_ptr v) {
     int i;
     MVertex_ptr adjv;
-    Set_ptr vedges;
+    List_ptr vedges;
     MVertex_SameAdj_R2R4 *sameadj;
 
     /* Have to create volatile edges */
 
-    vedges = Set_New(sameadj->nvadj);
+    vedges = List_New(sameadj->nvadj);
     for (i = 0; i < sameadj->nvadj; i++) {
       /* create a volatile edge from v and the adjacent vertex */
-      adjv = Set_Entry(sameadj->adjverts,i);
+      adjv = List_Entry(sameadj->adjverts,i);
     }
 
     return vedges;

### 2004-01-15 src/base/MVertex_F1.c

@@ -15,14 +15,14 @@ extern "C" {
 
     upadj = v->upadj = (MVertex_UpAdj_F1F4 *) MSTK_malloc(sizeof(MVertex_UpAdj_F1F4));
     upadj->ne = (unsigned int) 0;
-    upadj->vedges = Set_New(10);
+    upadj->vedges = List_New(10);
   }
 
   void MV_Delete_F1(MVertex_ptr v) {
     MVertex_UpAdj_F1F4 *upadj;
 
     upadj = (MVertex_UpAdj_F1F4 *) v->upadj;
-    Set_Delete(upadj->vedges);
+    List_Delete(upadj->vedges);
     MSTK_free(upadj);
 
     MSTK_free(v);
@@ -37,7 +37,7 @@ extern "C" {
   }
 
   int MV_Num_Faces_F1(MVertex_ptr v) {
-    Set_ptr vfaces;
+    List_ptr vfaces;
     int nf;
 
 #ifdef DEBUG
@@ -47,14 +47,14 @@ extern "C" {
 #endif
     
     vfaces = MV_Faces_F1(v);
-    nf = Set_Num_Entries(vfaces);
-    Set_Delete(vfaces);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
     return nf;
   }
 
   int MV_Num_Regions_F1(MVertex_ptr v) {
     int nr;
-    Set_ptr vregions;
+    List_ptr vregions;
 
 #ifdef DEBUG
     MSTK_Report("MV_Num_Regions",
@@ -63,15 +63,15 @@ extern "C" {
 #endif
 	
     vregions = MV_Regions_F1(v);
-    nr = Set_Num_Entries(vregions);
-    Set_Delete(vregions);
+    nr = List_Num_Entries(vregions);
+    List_Delete(vregions);
 
     return nr;
   }
 
-  Set_ptr MV_AdjVertices_F1(MVertex_ptr v) {
+  List_ptr MV_AdjVertices_F1(MVertex_ptr v) {
     MVertex_UpAdj_F1F4 *upadj;
-    Set_ptr vedges, adjv;
+    List_ptr vedges, adjv;
     int ne, i;
     MEdge_ptr vedge;
     MVertex_ptr ov;
@@ -82,30 +82,30 @@ extern "C" {
       return 0;
 
     ne = upadj->ne;
-    adjv = Set_New(ne);
+    adjv = List_New(ne);
[…]

### 2004-01-15 src/base/MVertex_F4.c

@@ -15,14 +15,14 @@ extern "C" {
 
     upadj = v->upadj = (MVertex_UpAdj_F1F4 *) MSTK_malloc(sizeof(MVertex_UpAdj_F1F4));
     upadj->ne = (unsigned int) 0;
-    upadj->vedges = Set_New(10);
+    upadj->vedges = List_New(10);
   }
 
   void MV_Delete_F4(MVertex_ptr v) {
     MVertex_UpAdj_F1F4 *upadj;
 
     upadj = (MVertex_UpAdj_F1F4 *) v->upadj;
-    Set_Delete(upadj->vedges);
+    List_Delete(upadj->vedges);
     MSTK_free(upadj);
   }
 
@@ -35,7 +35,7 @@ extern "C" {
   }
 
   int MV_Num_Faces_F4(MVertex_ptr v) {
-    Set_ptr vfaces;
+    List_ptr vfaces;
     int nf;
 
 #ifdef DEBUG
@@ -45,13 +45,13 @@ extern "C" {
 #endif
     
     vfaces = MV_Faces_F4(v);
-    nf = Set_Num_Entries(vfaces);
-    Set_Delete(vfaces);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
     return nf;
   }
 
   int MV_Num_Regions_F4(MVertex_ptr v) {
-    Set_ptr vregions;
+    List_ptr vregions;
     int nr;
 
 #ifdef DEBUG
@@ -61,14 +61,14 @@ extern "C" {
 #endif
 	
     vregions = MV_Regions_F4(v);
-    nr = Set_Num_Entries(vregions);
-    Set_Delete(vregions);
+    nr = List_Num_Entries(vregions);
+    List_Delete(vregions);
     return nr;
   }
 
-  Set_ptr MV_AdjVertices_F4(MVertex_ptr v) {
+  List_ptr MV_AdjVertices_F4(MVertex_ptr v) {
     MVertex_UpAdj_F1F4 *upadj;
-    Set_ptr vedges, adjv;
+    List_ptr vedges, adjv;
     int ne, i;
     MEdge_ptr vedge;
     MVertex_ptr ov;
@@ -79,30 +79,30 @@ extern "C" {
       return 0;
 
     ne = upadj->ne;
-    adjv = Set_New(ne);
+    adjv = List_New(ne);
     for (i = 0; i < ne; i++) {
-      vedge = Set_Entry(vedges,i);
+      vedge = List_Entry(vedges,i);
[…]

### 2004-01-15 src/base/MVertex_R1.c

@@ -14,14 +14,14 @@ extern "C" {
 
     upadj = v->upadj = (MVertex_UpAdj_R1R2 *) MSTK_malloc(sizeof(MVertex_UpAdj_R1R2));
     upadj->nel = (unsigned int) 0;
-    upadj->velements = Set_New(10);
+    upadj->velements = List_New(10);
   }
 
   void MV_Delete_R1(MVertex_ptr v) {
     MVertex_UpAdj_R1R2 *upadj;
 
     upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
-    Set_Delete(upadj->velements);
+    List_Delete(upadj->velements);
     MSTK_free(upadj);
   }
 
@@ -36,7 +36,7 @@ extern "C" {
 
   int MV_Num_Edges_R1(MVertex_ptr v) {
     int ne;
-    Set_ptr vedges;
+    List_ptr vedges;
 
 #ifdef DEBUG
     MSTK_Report("MV_Num_Edges",
@@ -45,14 +45,14 @@ extern "C" {
 #endif
     
     vedges = MV_Edges_R1(v);
-    ne = Set_Num_Entries(vedges);
-    Set_Delete(vedges);
+    ne = List_Num_Entries(vedges);
+    List_Delete(vedges);
     return ne;
   }
 
   int MV_Num_Faces_R1(MVertex_ptr v) {
     int nf;
-    Set_ptr vfaces;
+    List_ptr vfaces;
 
 #ifdef DEBUG
     MSTK_Report("MV_Num_Faces",
@@ -65,8 +65,8 @@ extern "C" {
        retrieved fairly efficiently */
 
     vfaces = MV_Faces_R1(v);
-    nf = Set_Num_Entries(vfaces);
-    Set_Delete(vfaces);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
     return nf;
   }
   
@@ -77,49 +77,49 @@ extern "C" {
 
     upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
     for (i = 0; i < upadj->nel; i++) {
-      ent = (MEntity_ptr) Set_Entry(upadj->velements,i);
+      ent = (MEntity_ptr) List_Entry(upadj->velements,i);
       if (MEnt_Dim(ent) == 3)
 	nr++;
     }
     return nr;
   }
 
-  Set_ptr MV_AdjVertices_R1(MVertex_ptr v) {
+  List_ptr MV_AdjVertices_R1(MVertex_ptr v) {
[…]

### 2004-01-15 src/base/MVertex_R2.c

@@ -15,10 +15,10 @@ extern "C" {
 
     upadj = v->upadj = (MVertex_UpAdj_R1R2 *) MSTK_malloc(sizeof(MVertex_UpAdj_R1R2));
     upadj->nel = (unsigned int) 0;
-    upadj->velements = Set_New(10);
+    upadj->velements = List_New(10);
     sameadj = v->sameadj = (MVertex_SameAdj_R2R4 *) MSTK_malloc(sizeof(MVertex_SameAdj_R2R4));
     sameadj->nvadj = (unsigned char) 0;
-    sameadj->adjverts = Set_New(10);
+    sameadj->adjverts = List_New(10);
   }
 
   void MV_Delete_R2(MVertex_ptr v) {
@@ -26,11 +26,11 @@ extern "C" {
     MVertex_SameAdj_R2R4 *sameadj;
 
     upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
-    Set_Delete(upadj->velements);
+    List_Delete(upadj->velements);
     MSTK_free(upadj);
 
     sameadj = (MVertex_SameAdj_R2R4 *) v->sameadj;
-    Set_Delete(sameadj->adjverts);
+    List_Delete(sameadj->adjverts);
     MSTK_free(sameadj);
   }
     
@@ -45,7 +45,7 @@ extern "C" {
 
   int MV_Num_Faces_R2(MVertex_ptr v) {
     int nf;
-    Set_ptr vfaces;
+    List_ptr vfaces;
 
 #ifdef DEBUG
     MSTK_Report("MV_Num_Faces",
@@ -58,8 +58,8 @@ extern "C" {
        retrieved fairly efficiently */
 
     vfaces = MV_Faces_R1(v);
-    nf = Set_Num_Entries(vfaces);
-    Set_Delete(vfaces);
+    nf = List_Num_Entries(vfaces);
+    List_Delete(vfaces);
     return nf;
   }
   
@@ -70,49 +70,49 @@ extern "C" {
 
     upadj = (MVertex_UpAdj_R1R2 *) v->upadj;
     for (i = 0; i < upadj->nel; i++) {
-      ent = (MEntity_ptr) Set_Entry(upadj->velements,i);
+      ent = (MEntity_ptr) List_Entry(upadj->velements,i);
       if (MEnt_Dim(ent) == 3)
 	nr++;
     }
     return nr;
   }
 
[…]

### 2004-01-15 src/base/MVertex_R4.c

@@ -17,10 +17,10 @@ extern "C" {
 
     upadj = v->upadj = (MVertex_UpAdj_R3R4 *) MSTK_malloc(sizeof(MVertex_UpAdj_R3R4));
     upadj->nf = (unsigned int) 0;
-    upadj->vfaces = Set_New(10);
+    upadj->vfaces = List_New(10);
     sameadj = v->sameadj = (MVertex_SameAdj_R2R4 *) MSTK_malloc(sizeof(MVertex_SameAdj_R2R4));
     sameadj->nvadj = (unsigned char) 0;
-    sameadj->adjverts = Set_New(10);
+    sameadj->adjverts = List_New(10);
   }
 
   void MV_Delete_R4(MVertex_ptr v) {
@@ -28,11 +28,11 @@ extern "C" {
     MVertex_SameAdj_R2R4 *sameadj;
    
     upadj = (MVertex_UpAdj_R3R4 *) v->upadj;
-    Set_Delete(upadj->vfaces);
+    List_Delete(upadj->vfaces);
     MSTK_free(upadj);
 
     sameadj = (MVertex_SameAdj_R2R4 *) v->sameadj;
-    Set_Delete(sameadj->adjverts);
+    List_Delete(sameadj->adjverts);
     MSTK_free(sameadj);
   }
 
@@ -52,26 +52,26 @@ extern "C" {
     return MV_Num_Regions_R3R4(v);
   }
 
-  Set_ptr MV_AdjVertices_R4(MVertex_ptr v) {
+  List_ptr MV_AdjVertices_R4(MVertex_ptr v) {
     MVertex_SameAdj_R2R4 *sameadj;
 
     sameadj = (MVertex_SameAdj_R2R4 *)v->sameadj;
-    return Set_Copy(sameadj->adjverts);
+    return List_Copy(sameadj->adjverts);
   }
 
-  Set_ptr MV_Edges_R4(MVertex_ptr v) {
+  List_ptr MV_Edges_R4(MVertex_ptr v) {
     /* Have to create volatile edges */
     return MV_Edges_R2R4(v);
   }
 
-  Set_ptr MV_Faces_R4(MVertex_ptr v) {
+  List_ptr MV_Faces_R4(MVertex_ptr v) {
     MVertex_UpAdj_R3R4 *upadj;
     upadj = (MVertex_UpAdj_R3R4 *)v->upadj;
 
[…]

### 2004-01-15 src/hilev/MFs_Join.c

@@ -8,7 +8,7 @@ MFace_ptr MFs_Join(MFace_ptr f1, MFace_ptr f2, MEdge_ptr e) {
   MEdge_ptr *fe1, *fe2, *fe3;
   MFace_ptr  nuface;
   Mesh_ptr   mesh;
-  Set_ptr   fedges;
+  List_ptr   fedges;
 
   mesh = MF_Mesh(f1);
   gid = MF_GEntID(f1);
@@ -35,21 +35,21 @@ MFace_ptr MFs_Join(MFace_ptr f1, MFace_ptr f2, MEdge_ptr e) {
 
   fedges = MF_Edges(f1,1,0);
   for (i = 0, k1 = -1; i < nfe1; i++) {
-    fe1[i] = Set_Entry(fedges,i);
+    fe1[i] = List_Entry(fedges,i);
     fedir1[i] = MF_EdgeDir_i(f1,i);
     if (fe1[i] == e)
       k1 = i;
   }
-  Set_Delete(fedges);
+  List_Delete(fedges);
 
   fedges = MF_Edges(f2,1,0);
   for (i = 0, k2 = -1; i < nfe2; i++) {
-    fe2[i] = Set_Entry(fedges,i);
+    fe2[i] = List_Entry(fedges,i);
     fedir2[i] = MF_EdgeDir_i(f2,i);
     if (fe2[i] == e)
       k2 = i;
   }
-  Set_Delete(fedges);
+  List_Delete(fedges);
 
   if (k1 == -1 || k2 == -1) {
     MSTK_Report("MFs_Join","Cannot find edge in face",ERROR);

### 2004-01-27 docs/MSTK.tex
Updated version 1.2 manual with an example

@@
 %
-%\setlength{\textwidth}{4.5in}
+\setlength{\textwidth}{6.5in}
 %\setlength{\textheight}{7.0in}
-%\setlength{\evensidemargin}{0.0in}
-%\setlength{\oddsidemargin}{0.0in}
+\setlength{\evensidemargin}{0.0in}
+\setlength{\oddsidemargin}{0.0in}
 %\setlength{\topmargin}{0.5in}
@@
 
-\title{MSTK: Mesh Toolkit, v 1.1 - DRAFT}
+\title{MSTK: Mesh Toolkit, v 1.2 - DRAFT}
 
@@
 
-\begin{figure}[!h]
-%  \begin{center}
+\begin{figure}[!ht]
+  \begin{center}
     \begin{minipage}{2.525in}
@@
       \end{center}
-    \end{minipage}\hfill%
+    \end{minipage}
     \begin{minipage}{2.525in}
@@
     \end{minipage}
-%  \end{center}
+  \end{center}
 \end{figure}
@@
 
+\section{Example program}
+
+NOTE: This program is included in the distribution.
+
+\begin{verbatim}
+#include <stdio.h>
+#include <stdlib.h>
+#include "MSTK.h"
+#include "test.h"
+
+int main(int argc, char *argv[]) {
+ int i, idx, idx2, ok, edir, nv, ne;
+ double xyz[3];
+ char meshname[256];
+ Mesh_ptr mesh;
+ MVertex_ptr v;
+ MEdge_ptr e;
+ MFace_ptr f;
+ GEntity_ptr gent;
+ List_ptr fedges;
+
+ /* Initialize MSTK - Always do this even if it does
+    not seem to matter in this version of MSTK */
+
+ MSTK_Init();
+
+ /* Load the mesh */
+
+ strcpy(meshname,argv[1]);
+ strcat(meshname,".mstk");
+
+ mesh = MESH_New(UNKNOWN_REP);
+ ok = MESH_InitFromFile(mesh,meshname);
+ if (!ok) {
+   fprintf(stderr,"Cannot file input file %s\n\n\n",meshname);
+   exit(-1);
+ }
+
+ /* Print some info about the mesh */
+
+ nv = MESH_Num_Vertices(mesh);
+ for (i = 0; i < nv; i++) {
+  v = MESH_Vertex(mesh,i);
+
+  /* Basic info */
+  printf("\n");
+  printf("Vertex: 0x%-x   ID: %-d   ",v,MV_ID(v));
+
+  /* Classification w.r.t. geometric model */
+
+  if (MV_GEntDim(v) == -1)
+    fprintf(stderr,"Unknown Classification\n");
+  else {
+   printf("GEntID: %-d   GEntDim: %-d\n",MV_GEntID(v),MV_GEntDim(v));
+   if ((gent = MV_GEntity(v)))
+     printf("Model entity pointer: 0x%-x\n",gent);   
+  }
+
+  /* Coordinates */
+  MV_Coords(v,xyz);
+  printf("Coords: %16.8lf %16.8lf %16.8lf\n",xyz[0],xyz[1],xyz[2]);
+ }
+
+ idx = 0;
+ while (f = MESH_Next_Face(mesh,&idx)) {
+
+  /* Basic info */
+  printf("\n");
+  printf("Face: 0x%-x   ID: %-d   ",f,MF_ID(f));
+
+  /* Classification w.r.t. geometric model */
+
+  if (MF_GEntDim(f) == -1)
+   fprintf(stderr,"Unknown Classification\n");
+  else {
+   printf("GEntID: %-d   GEntDim: %-d\n",MF_GEntID(f),MF_GEntDim(f));
+   if ((gent = MF_GEntity(f)))
+    printf("Model entity pointer: 0x%-x\n",gent);
+  }
+  printf("\n");
+
+  /* Edges of face */
+  fedges = MF_Edges(f,1,0);
+  ne = List_Num_Entries(fedges);
+  printf("Edges: %-d\n",ne);
+  printf("Object       ID     GEntID  GEntDim   Vertex IDs\n");
+  idx2 = 0; i = 0;
+  while (e = List_Next_Entry(fedges,&idx2)) {
+   edir = MF_EdgeDir_i(f,i);
+   if (edir) 
+    printf("0x%-8x    %-8d %-8d     %-1d       %-d  %-d\n",
+           e,ME_ID(e),ME_GEntID(e),ME_GEntDim(e),
+           MV_ID(ME_Vertex(e,0)),MV_ID(ME_Vertex(e,1)));
+   else 
+    printf("0x%-8x    %-8d %-8d     %-1d       %-d  %-d\n",
+           e,-ME_ID(e),ME_GEntID(e),ME_GEntDim(e),
+           MV_ID(ME_Vertex(e,0)),MV_ID(ME_Vertex(e,1)));
+   i++;
+  }
+  printf("\n");
+  List_Delete(fedges);
+ }
+
+ /* Write out a copy of the mesh */
+
+ strcpy(meshname,argv[1]);
+ strcat(meshname,"-copy.mstk");
+ MESH_WriteToFile(mesh,meshname);
+
+ /* No need to delete a mesh if program ends right afterwards */
+
+ MESH_Delete(mesh);
+ return 1;
+}
+\end{verbatim}
+
 \end{document}

### 2004-07-09 docs/MSTK.tex
Updated docs to reflect new functions

@@
 
-\title{MSTK: Mesh Toolkit, v 1.2 - DRAFT}
+\title{MSTK: Mesh Toolkit, v 1.3 - DRAFT}
 
@@
 
-\par MSTK will soon allow applications to attach application or field data
-to entities. This data may be integers, reals (doubles), integer
-vectors, real (double) vectors, integer tensors, real (double tensors)
-and pointers.
+\par To support numerical analysis and other applications, MSTK allows
+applications to attach application or field data to entities. This
+data may be integers, reals (doubles), integer vectors, real (double)
+vectors, integer tensors, real (double tensors) and pointers.
 
@@
 \begin{description}
-\item[]\textbf{\textit{List\_ptr}}: Handle to a List object
+\item[]\textbf{\textit{List\_ptr}}: Handle to a List object.
 
@@
 \item[]\textbf{\textit{MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
-Dimension 0)
-
-\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1)
-
-\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2)
-
-\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3)
-
-\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity object. Any of
-the above types of entities can be cast as \textit{MEntity\_ptr}
-
-\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object
-\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object
+Dimension 0).
+
+\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1).
+
+\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2).
+
+\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3).
+
+\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity
+  object. Any of the above types of entities can be cast as
+  \textit{MEntity\_ptr}.
+
+\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object.
+\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object.
+
+\item[]\textbf{\textit{MAttrib\_ptr}}: Handle to a mesh attribute.
 
@@
   See Appendix~\ref{app:reptypes} for schematics of these
-  representations \textit{Currently only representation types F1 and
-    F4 are supported}
-
-\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type
-Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON
-
-\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type
-Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED
+  representations. \textit{Currently only representation types F1 and
+    F4 are supported}.
+
+\item[]\textbf{\textit{MType}}: Enumerated type for mesh entity type. Can be MDELETED, MVERTEX, MEDGE, MFACE, MREGION, MUNKNOWNTYPE, MALLTYPE.
+
+\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type.
+Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON.
+
+\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type.
+Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED.
+
+\item[]\textbf{\textit{AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
 \end{description}
@@
 
-\item[]\textbf{\textit{List\_ptr} List\_Compress(\textit{List\_ptr} l):} Compress
-a set.  Doing this while an algorithm is iterating through the set can
-currently cause problems!! Calling \textbf{List\_Compress} could change the
-pointer for the set due to reallocation.
+\item[]\textbf{\textit{List\_ptr} List\_Compress(\textit{List\_ptr}
+    l):} Compress a set.  Doing this while an algorithm is iterating
+  through the set can currently cause problems!! Calling
+  \textbf{List\_Compress} could change the pointer for the set due to
+  reallocation.
 
@@
 modified externally while iterating through the set. Also, no specific
-meaning be derived from from the iteration index by other applications
+meaning should be derived from the iteration index by other applications
 since the internal implementation and interpretation of the index may
@@
 
-\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr} mesh,
-const char *filename):} Save a mesh to a filename. The file is created
-if it does not exists. It is recommended that the .mstk extension be
-used for MSTK mesh files.  However, there is no such requirement.
+\item[]\textbf{\textit{int} MESH\_ImportFromFile(\textit{Mesh\_ptr}
+    mesh, const char *filename, const char *format):} Import a mesh
+  data from an external file format and construct MSTK mesh. Currently
+  the only format supported is the
+  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
+  file format. The routine imports as many attributes as it can from
+  the input file. It also uses special attributes or keywords as data
+  about element and node classification.  For GMV files, the routine
+  uses the ``material'' data to indicate region or face classification
+  depending on the dimensionality of the mesh. If ``material'' data is
+  not specified, it uses the ``itetclr'' attribute to assign region or
+  face classification. It also uses the ``icr'' keyword describing the
+  number of constraints on a node to interpret if the node is
+  classified on a model region, model face, model edge or a model
+  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
+  meshes generated by
+  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.
+
+\item[]\textbf{\textit{int} MESH\_BuildClassfn(\textit{Mesh\_ptr}
+    mesh):} Build classification information for mesh entities if only
+  partial information is present. In other words, if the only data
+  known is the IDs of geometric model entities (``material regions'')
+  of the highest level mesh entities (faces or regions), then this
+  procedure will build information about the type of geometric model
+  entities that the lower dimension entities are on. Faces will be
+  classified as being on a model face (external or interior) or inside
+  model region. Edges will be classified as being on a model edge,
+  model face or in a model region. Vertices will be classified as
+  being on a model vertex, model edge, model face or in a model
+  region. If no classification data is associated with even the
+  highest level entities, the procedure will assume that all the
+  highest level entities are classified on one model entitiy of that
+  dimension. If partial information is available for lower order
+  entities, this routine will not destroy that information. In the
+  future, the procedure will be more sophisticated, incorporating
+  features such as classifying mesh edges as being on model edges if
+  the dihedral angle between the boundary faces connected to it is
+  very large.
+
+\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr}
+    mesh, const char *filename):} Save a mesh to a filename. The file
+  is created if it does not exists. It is recommended that the .mstk
+  extension be used for MSTK mesh files.  However, there is no such
+  requirement.
+
+%% \item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
+%%     mesh, const char *filname, const char *format, int natt, char
+%%     **attnames):} Export a mesh to an external file format. Currently,
+%%   the formats supported are the GMV file (format string: ``gmv'') and
+%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
+%%   ``dx''). Only integer and double attributes of the mesh are exported
+%%   to the files.
+
+\item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
+    mesh, const char *filname, const char *format, int natt, char
+    **attnames):} Export a mesh to an external file format. Currently,
+  the only format supported is the GMV file format (format string:
+  ``gmv''). Only integer and double attributes of the mesh are
+  exported to the files.
+
+\item[]
 
@@
 Representation type currently being used by the mesh.
+
+\item[]
 
@@
 
+\item[]
+
 \item[]\textbf{\textit{MVertex\_ptr} MESH\_Vertex(\textit{Mesh\_ptr} mesh,
@@
 $<$ 0 or i $>$ number of mesh region.
+
+\item[]
 
@@
 
-\item[]\textbf{\textit{void} MESH\_Add\_Vertex(\textit{Mesh\_ptr} mesh,
-\textit{MVertex\_ptr} v):} Add a vertex to the mesh. It is assumed that
-the vertex and its coordinates set are properly defined.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Edge(\textit{Mesh\_ptr} mesh,
-\textit{MEdge\_ptr} e):} Add an edge to the mesh. It is assumed that
-the edge is and its topology is defined.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Face(\textit{Mesh\_ptr} mesh,
-\textit{MFace\_ptr} f):} Add a face to the mesh. It is assumed that the
-face and its topology is properly defined.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Region(\textit{Mesh\_ptr} mesh,
-\textit{MRegion\_ptr} r):} Add a region to the mesh. It is assumed that
-the region and its topology is properly defined.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Vertex(\textit{Mesh\_ptr} mesh,
-\textit{MVertex\_ptr} v):} Remove vertex from mesh. Vertex is not
-deleted and must be deleted afterward separately.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Edge(\textit{Mesh\_ptr} mesh,
-\textit{MEdge\_ptr} e):} Remove edge from mesh. Edge is not deleted and
-must be deleted afterward separately.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Face(\textit{Mesh\_ptr} mesh,
-\textit{MFace\_ptr} f):} Remove face from mesh. Face is not deleted and
-must be deleted afterward separately.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Region(\textit{Mesh\_ptr} mesh,
-\textit{MRegion\_ptr} r):} Remove region from mesh. Region is not deleted and must be deleted afterward separately.
+\item[]\textbf{\textit{MRegion\_ptr}
+    MESH\_Next\_Region(\textit{Mesh\_ptr} mesh, \textit{int} *idx):}
+  Returns the next region while iterating through the regions of the
+  mesh. See the routine \textbf{List\_Next\_Entry} above for an
+  explanation of how the iteration works.
+
+\item[]
+
+\item[]\textbf{\textit{MVertex\_ptr}
+    MESH\_VertexFromID(\textit{Mesh\_ptr} mesh, \textit{int} id):}
+  Return mesh vertex with given ID if it exists; return NULL
+  otherwise.
+
+\item[]\textbf{\textit{MEdge\_ptr} MESH\_EdgeFromID(\textit{Mesh\_ptr}
+    mesh, \textit{int} id):} Return mesh edge with given ID if it
+  exists; return NULL otherwise. This routine will return NULL for all
+  reduced representations.
+
+\item[]\textbf{\textit{MFace\_ptr} MESH\_FaceFromID(\textit{Mesh\_ptr}
+    mesh, \textit{int} i):} Return mesh face with given ID if it
+  exists; return NULL otherwise. If faces are not explicitly
+  represented in the mesh, it will return NULL.
+
+\item[]\textbf{\textit{MRegion\_ptr}
+    MESH\_RegionFromID(\textit{Mesh\_ptr} mesh, \textit{int} id):}
+  Return mesh region with given ID if it exists; return NULL
+  otherwise.
+
+\item[]
+
+\item[]\textbf{\textit{int} MESH\_Num\_Attribs(\textit{Mesh\_ptr}
+    mesh):} Number of attributes associated with the mesh.
+
+\item[]\textbf{\textit{MAttrib\_ptr} MESH\_Attrib(\textit{Mesh\_ptr}
+    mesh, \textit{int} i):} Return the i'th attribute in the mesh.
+  Returns NULL if i $<$ 0 or i $>$ number of mesh attributes.
+
+\item[]\textbf{\textit{MAttrib\_ptr}
+    MESH\_Next\_Attrib(\textit{Mesh\_ptr} mesh, \textit{int} *index):}
+  Returns the next attribute while iterating through the attributes of
+  the mesh. See the routine List\_Next\_Entry above for an explanation
+  of how the iteration works.
+
+\item[]\textbf{\textit{MAttrib\_ptr}
+    MESH\_AttribByName(\textit{Mesh\_ptr} mesh, \textit{const char}
+    *name):} Return a mesh attribute with given name if it exisists in
+  the mesh. Returns NULL if mesh has no such attribute.
+
+\item[]
 
@@
 
+\item[]
+
+\item[]\textbf{\textit{void} MESH\_Add\_Vertex(\textit{Mesh\_ptr}
+    mesh, \textit{MVertex\_ptr} v):} Add a vertex to the mesh. It is
+  assumed that the vertex and its coordinates set are properly
+  defined. Normally, one need not call this since
+  \textbf{\textit{MV\_New}} will add the vertex to the mesh. Use this
+  only if you know exactly what you are doing.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Edge(\textit{Mesh\_ptr} mesh,
+    \textit{MEdge\_ptr} e):} Add an edge to the mesh. It is assumed
+  that the edge is and its topology is defined. Normally, one need not
+  call this since \textbf{\textit{ME\_New}} will add the edge to the
+  mesh. Use this only if you know exactly what you are doing.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Face(\textit{Mesh\_ptr} mesh,
+    \textit{MFace\_ptr} f):} Add a face to the mesh. It is assumed
+  that the face and its topology is properly defined. Normally, one
+  need not call this since \textbf{\textit{MF\_New}} will add the face
+  to the mesh. Use this only if you know exactly what you are doing.
+
+\item[]\textbf{\textit{void} MESH\_Add\_Region(\textit{Mesh\_ptr}
+    mesh, \textit{MRegion\_ptr} r):} Add a region to the mesh. It is
+  assumed that the region and its topology is properly defined.
+  Normally, one need not call this since \textbf{\textit{MR\_New}}
+  will add the region to the mesh. Use this only if you know exactly
+  what you are doing.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Vertex(\textit{Mesh\_ptr}
+    mesh, \textit{MVertex\_ptr} v):} Remove vertex from mesh. Vertex
+  is not deleted and must be deleted afterward separately. Normally,
+  one need not call this since \textbf{\textit{MV\_Delete}} will
+  remove the vertex from the mesh. Use this only if you know exactly
+  what you are doing.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Edge(\textit{Mesh\_ptr} mesh,
+    \textit{MEdge\_ptr} e):} Remove edge from mesh. Edge is not
+  deleted and must be deleted afterward separately. Normally, one need
+  not call this since \textbf{\textit{ME\_Delete}} will remove the
+  edge from the mesh. Use this only if you know exactly what you are
+  doing.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Face(\textit{Mesh\_ptr} mesh,
+    \textit{MFace\_ptr} f):} Remove face from mesh. Face is not
+  deleted and must be deleted afterward separately. Normally, one need
+  not call this since \textbf{\textit{MF\_Delete}} will remove the
+  face from the mesh. Use this only if you know exactly what you are
+  doing.
+
+\item[]\textbf{\textit{void} MESH\_Rem\_Region(\textit{Mesh\_ptr}
+    mesh, \textit{MRegion\_ptr} r):} Remove region from mesh. Region
+  is not deleted and must be deleted afterward separately. Normally,
+  one need not call this since \textbf{\textit{MR\_Delete}} will
+  remove the region from the mesh. Use this only if you know exactly
+  what you are doing.
 \end{description}
@@
 \begin{description}
-\item[]\textbf{\textit{MVertex\_ptr} MV\_New(\textit{Mesh\_ptr} mesh):}
-Create a new vertex object. No geometric or topological information is
-embedded in the vertex when it is created. The vertex only knows which
-mesh it belongs to. The ID of the vertex is set by this function.
-
-\item[]\textbf{\textit{void} MV\_Delete(\textit{MVertex\_ptr} mvertex):}
-Delete the vertex. Deletes all topological and geometric
-information embedded in the vertex.
-
-\item[]\textbf{\textit{void} MV\_Set\_Coords(\textit{MVertex\_ptr} mvertex,
-double *xyz):} Set the coordinates of the vertex.
-
-\item[]\textbf{\textit{void} MV\_Set\_GEntity(\textit{MVertex\_ptr} mvertex,
-GEntity\_ptr gent):} Set the geometric model entity on which vertex is
-classified.
+\item[]\textbf{\textit{MVertex\_ptr} MV\_New(\textit{Mesh\_ptr}
+    mesh):} Create a new vertex object. No geometric or topological
+  information is embedded in the vertex when it is created. The vertex
+  only knows which mesh it belongs to. The ID of the vertex is set by
+  this function.
+
+\item[]\textbf{\textit{void} MV\_Delete(\textit{MVertex\_ptr} mvertex,
+    \textit{int} keep):} Delete the vertex. If \textbf{keep} is 0, the
+  vertex is removed from the mesh and all topological and geometric
+  information embedded in the vertex is destroyed. If \textbf{keep} is
+  1, the vertex is marked as type \textit{MDELVERTEX} and removed from
+  the mesh but the vertex is not destroyed.
+
+\item[]\textbf{\textit{void} MV\_Restore(\textit{MVertex\_ptr}
+    mvertex):} Restore a deleted vertex. The vertex type is restored
+  from \textit{MDELVERTEX} to \textit{MVERTEX} and the vertex is
+  added back to the mesh.
+
+\item[]
+
+\item[]\textbf{\textit{void} MV\_Set\_Coords(\textit{MVertex\_ptr}
+    mvertex, double *xyz):} Set the coordinates of the vertex.
+
+\item[]\textbf{\textit{void} V\_Set\_GEntity(\textit{MVertex\_ptr}
+    mvertex, GEntity\_ptr gent):} Set the geometric model entity on
+  which vertex is classified.
 
@@
 
+\item[]
+
 \item[]\textbf{\textit{Mesh\_ptr} MV\_Mesh(\textit{MVertex\_ptr} mv):}
@@
 
+\item[]
+
 \item[]\textbf{\textit{int} MV\_Num\_AdjVertices(\textit{MVertex\_ptr}
@@
 \begin{description}
-\item[]\textbf{\textit{MEdge\_ptr} ME\_New(\textit{Mesh\_ptr} mesh):} Create
-a new edge object. No topological information is embedded in the edge
-when it is created. The edge only knows which mesh it belongs
-to. The ID of the edge is set by this function.
-
-\item[]\textbf{\textit{void} ME\_Delete(\textit{MEdge\_ptr} medge):} Delete
-the edge. Deletes all topological information embedded in the edge.
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntity(\textit{MEdge\_ptr} medge,
-GEntity\_ptr gent):} Set the geometric model entity on which the edge
-is classified.
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntDim(\textit{MEdge\_ptr} medge,
-\textit{int} gdim):} Set the topological dimension of model entity on
-which edge is classified.
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntID(\textit{MEdge\_ptr} medge,
-\textit{int} gid):} Set ID of model entity on which edge is classified.
+\item[]\textbf{\textit{MEdge\_ptr} ME\_New(\textit{Mesh\_ptr} mesh):}
+  Create a new edge object. No topological information is embedded in
+  the edge when it is created. The edge only knows which mesh it
+  belongs to. The ID of the edge is set by this function.
+
+\item[]\textbf{\textit{void} ME\_Delete(\textit{MEdge\_ptr} medge,
+    \textit{int} keep):} Delete the edge. If \textbf{keep} is 0, the
+  edge is removed from the mesh and all topological and geometric
+  information embedded in the edge is destroyed. If \textbf{keep} is
+  1, the edge is marked as type \textit{MDELEDGE} and removed from the
+  mesh but the edge is not destroyed. Also, the vertices of this edge
+  no longer point to this edge.
+
+\item[]\textbf{\textit{void} ME\_Restore(\textit{MEdge\_ptr}
+    medge):} Restore a temporarily deleted edge. The edge type is
+  restored from \textit{MDELEDGE} to \textit{MEDGE} and the edge is
+  added back to the mesh. The vertices of the edge once again point
+  back to the edge.
+
+\item[]
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntity(\textit{MEdge\_ptr}
+    medge, GEntity\_ptr gent):} Set the geometric model entity on
+  which the edge is classified.
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntDim(\textit{MEdge\_ptr}
+    medge, \textit{int} gdim):} Set the topological dimension of model
+  entity on which edge is classified.
+
+\item[]\textbf{\textit{void} ME\_Set\_GEntID(\textit{MEdge\_ptr}
+    medge, \textit{int} gid):} Set ID of model entity on which edge is
+  classified.
 
 \item[]\textbf{\textit{void} ME\_Set\_ID(\textit{MEdge\_ptr} medge,
-\textit{int} id):} Explicitly set ID of an edge and overwrite the ID set
-by the ME\_New function. Does not check for duplication of edge IDs.
-
-\item[]\textbf{\textit{void} ME\_Set\_Vertex(\textit{MEdge\_ptr} medge,
-\textit{int} i, \textit{MVertex\_ptr} vertex):} Set the i'th vertex of
-the edge. i can be 0 or 1.
-
-\item[]\textbf{\textit{void} ME\_Replace\_Vertex(\textit{MEdge\_ptr} medge,
-\textit{MVertex\_ptr} vert, \textit{MVertex\_ptr} nuvert):} Replace
-i'th vertex by new vertex.
-
-\item[]\textbf{\textit{Mesh\_ptr} ME\_Mesh(\textit{MEdge\_ptr} medge):}
-Returns the mesh that this edge belongs to.
-
-\item[]\textbf{\textit{int} ME\_ID(\textit{MEdge\_ptr} medge):} Returns the
-ID of the vertex. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} ME\_GEntDim(\textit{MEdge\_ptr} medge):} Returns
-the dimension of the geometric model entity that the vertex is
-classified on. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} ME\_GEntID(\textit{MEdge\_ptr} medge):} Returns
-the ID of the geometric model entity that the vertex is classified on.
-Returns 0 if this information is not known.
-
-\item[]\textbf{\textit{GEntity\_ptr} ME\_GEntity(\textit{MEdge\_ptr} medge):} Returns
-a pointer or handle to the geometric model entity that the vertex is
-classified on. Returns NULL if this information is not known.
-
-\item[]\textbf{\textit{int} ME\_Num\_Faces(\textit{MEdge\_ptr} medge):}
-Returns the number of faces connected to the edge.
-
-\item[]\textbf{\textit{int} ME\_Num\_Regions(\textit{MEdge\_ptr} medge):}
-Returns the number of regions connected to the edge.
-
-\item[]\textbf{\textit{MVertex\_ptr} ME\_Vertex(\textit{MEdge\_ptr} medge,
-\textit{int} i):} Returns the i'th vertex of the edge. i=0 returns the
-first vertex and i=1 returns the second vertex.
+    \textit{int} id):} Explicitly set ID of an edge and overwrite the
+  ID set by the ME\_New function. Does not check for duplication of
+  edge IDs.
+
+\item[]\textbf{\textit{void} ME\_Set\_Vertex(\textit{MEdge\_ptr}
+    medge, \textit{int} i, \textit{MVertex\_ptr} vertex):} Set the
+  i'th vertex of the edge. i can be 0 or 1.
+
+\item[]\textbf{\textit{void} ME\_Replace\_Vertex(\textit{MEdge\_ptr}
+    medge, \textit{MVertex\_ptr} vert, \textit{MVertex\_ptr} nuvert):}
+  Replace i'th vertex by new vertex.
+
+\item[]
+
+\item[]\textbf{\textit{Mesh\_ptr} ME\_Mesh(\textit{MEdge\_ptr}
+    medge):} Returns the mesh that this edge belongs to.
+
+\item[]\textbf{\textit{int} ME\_ID(\textit{MEdge\_ptr} medge):}
+  Returns the ID of the vertex. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} ME\_GEntDim(\textit{MEdge\_ptr} medge):}
+  Returns the dimension of the geometric model entity that the vertex
+  is classified on. Returns -1 if not known.
+
+\item[]\textbf{\textit{int} ME\_GEntID(\textit{MEdge\_ptr} medge):}
+  Returns the ID of the geometric model entity that the vertex is
+  classified on.  Returns 0 if this information is not known.
+
+\item[]\textbf{\textit{GEntity\_ptr} ME\_GEntity(\textit{MEdge\_ptr}
+    medge):} Returns a pointer or handle to the geometric model entity
+  that the vertex is classified on. Returns NULL if this information
+  is not known.
+
+\item[]
+
+\item[]\textbf{\textit{int} ME\_Num\_Faces(\textit{MEdge\_ptr}
+    medge):} Returns the number of faces connected to the edge.
+
+\item[]\textbf{\textit{int} ME\_Num\_Regions(\textit{MEdge\_ptr}
+    medge):} Returns the number of regions connected to the edge.
+
+\item[]\textbf{\textit{MVertex\_ptr} ME\_Vertex(\textit{MEdge\_ptr}
+    medge, \textit{int} i):} Returns the i'th vertex of the edge. i=0
+  returns the first vertex and i=1 returns the second vertex.
 
 \item[]\textbf{\textit{MVertex\_ptr} ME\_OppVertex(\textit{MEdge\_ptr}
-medge, \textit{MVertex\_ptr} ov):} Return the vertex opposite to given
-vertex in edge.
+    medge, \textit{MVertex\_ptr} ov):} Return the vertex opposite to
+  given vertex in edge.
 
 \item[]\textbf{\textit{int} ME\_UsesEntity(\textit{MEdge\_ptr} medge,
-\textit{MEntity\_ptr} mentity, \textit{int} etype):} Check if edge uses
-given lower dimension entity, \textit{mentity}. The dimension of the
-entity is specified by the \textit{etype} variable. For an edge, the
-only lower dimensional entity is a vertex. If the edge uses the
-vertex, the function returns 1; otherwise it returns 0. If any
-other type of entity is specified, the function returns 0.
-
-\item[]\textbf{\textit{List\_ptr} ME\_Faces(\textit{MEdge\_ptr} medge):}
-Returns the set of faces using this edge.
-
-\item[]\textbf{\textit{List\_ptr} ME\_Regions(\textit{MEdge\_ptr} medge):}
-Returns the set of regions using this edge.
-
-\item[]\textbf{\textit{MEdge\_ptr} MVs\_CommonEdge(\textit{MVertex\_ptr} v1,
-\textit{MVertex\_ptr} v2):} Return the edge connecting vertices v1 and
-v2, if it exists. If such an edge does not exist, the function returns
-0.
+    \textit{MEntity\_ptr} mentity, \textit{int} etype):} Check if edge
+  uses given lower dimension entity, \textit{mentity}. The dimension
+  of the entity is specified by the \textit{etype} variable. For an
+  edge, the only lower dimensional entity is a vertex. If the edge
+  uses the vertex, the function returns 1; otherwise it returns 0. If
+  any other type of entity is specified, the function returns 0.
+
+\item[]\textbf{\textit{List\_ptr} ME\_Faces(\textit{MEdge\_ptr}
+    medge):} Returns the set of faces using this edge.
+
+\item[]\textbf{\textit{List\_ptr} ME\_Regions(\textit{MEdge\_ptr}
+    medge):} Returns the set of regions using this edge.
+
+\item[]
+
+\item[]\textbf{\textit{MEdge\_ptr}
+    MVs\_CommonEdge(\textit{MVertex\_ptr} v1, \textit{MVertex\_ptr}
+    v2):} Return the edge connecting vertices v1 and v2, if it exists.
+  If such an edge does not exist, the function returns 0.
 
@@
 
-\item[]\textbf{\textit{void} ME\_Vec(\textit{MEdge\_ptr} e, double *evec):}
-Return the vector going from the first vertex of the edge to the
-second vertex of the edge.
+\item[]\textbf{\textit{void} ME\_Vec(\textit{MEdge\_ptr} e, double
+    *evec):} Return the vector going from the first vertex of the edge
+  to the second vertex of the edge.
 \end{description}
@@
 
-\item[]\textbf{\textit{void} MF\_Delete(\textit{MFace\_ptr} mface):} Delete
-the face. Delete all topological information embedded in the face.
-
-\item[]\textbf{\textit{void} MF\_Set\_GEntity(\textit{MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the edge is classified.
+\item[]\textbf{\textit{void} MF\_Delete(\textit{MFace\_ptr} mface):}
+  Delete the face. If \textbf{keep} is 0, the face is removed from the
+  mesh and all topological and geometric information embedded in the
+  face is destroyed. If \textbf{keep} is 1, the face is marked as type
+  \textit{MDELFACE} and removed from the mesh but the face is not
+  destroyed. Also, the vertices/edges of this face no longer point up
+  to this face.
+
+\item[]\textbf{\textit{void} MF\_Restore(\textit{MFace\_ptr}
+    mface):} Restore a temporarily deleted face. The face type is
+  restored from \textit{MDELFACE} to \textit{MFACE} and the face is
+  added back to the mesh. The vertices/edges of the face once again point
+  back to the face.
+
+\item[]
+
+\item[]\textbf{\textit{void} MF\_Set\_GEntity(\textit{MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the face is classified.
 
@@
 \item[]\textbf{\textit{void} MF\_Set\_ID(\textit{MFace\_ptr} mface,
-\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
-set by the MF\_New operator. Does not check for duplication of face
+\textit{int} id):} Explicitly set ID of a face and overwrite the ID
+set by the \textbf{MF\_New} operator. Does not check for duplication of face
 IDs.
@@
 
-\item[]\textbf{\textit{void} MF\_Set\_Vertices(\textit{MFace\_ptr} mface,
-\textit{int} n, \textit{MVertex\_ptr} *verts):} Set the vertices of the
-face. The ordered set of vertices is passed in through an array along
-with the number of vertices. The vertices are assumed to be ordered
-clockwise around the face. This function is relevant only for reduced
-representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Edge(\textit{MFace\_ptr} mface,
-\textit{MEdge\_ptr} edge, \textit{MEdge\_ptr} nuedge, \textit{int}
-dir):} Replace an edge in the face with another edge. The direction
-in which the new edge is used in the face must also be supplied. This
-function is relevant only for full representations in MSTK.
+\item[]\textbf{\textit{void} MF\_Set\_Vertices(\textit{MFace\_ptr}
+    mface, \textit{int} n, \textit{MVertex\_ptr} *verts):} Set the
+  vertices of the face. The ordered set of vertices (ccw around the
+  face) is passed in through an array along with the number of
+  vertices. This routine will collect/build all lower order
+  topological information (edges, edge directions in the face) as needed by the face.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edge(\textit{MFace\_ptr}
+    mface, \textit{MEdge\_ptr} edge, \textit{MEdge\_ptr} nuedge,
+    \textit{int} dir):} \textit{THIS FUNCTION HAS BEEN SUPERSEDED BY
+    \textbf{MF\_Replace\_Edges}}.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edge\_i(\textit{MFace\_ptr}
+    mface, \textit{int} i, \textit{MEdge\_ptr} nuedge, \textit{int}
+    dir):} \textit{THIS FUNCTION HAS BEEN SUPERSEDED BY
+    \textbf{MF\_Replace\_Edges\_i}}.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edges(\textit{MFace\_ptr}
+    mface, \textit{int} nold, \textit{MEdge\_ptr} *oldedges,
+    \textit{int} nnu, \textit{MEdge\_ptr} *nuedges):} Replace a set of
+  edges in the face with another set of edges. The direction in which
+  the new edges are to be used in the face are automatically deduced.
+  This function is relevant only for full representations in MSTK.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Edges\_i(\textit{MFace\_ptr}
+    mface, \textit{int} nold, \textit{int} i, \textit{int} nnu,
+    \textit{MEdge\_ptr} *nuedges):} Replace the \textit{nold} edges in
+  the face starting with the i'th edge and going ccw around the face
+  with a new set of edges. The directions in which the new edges are
+  used in the face are automatically deduced. This function is
+  relevant only for full representations in MSTK.
 
@@
 
-\item[]\textbf{\textit{void} MF\_Replace\_Edge\_i(\textit{MFace\_ptr} mface,
-\textit{int} i, \textit{MEdge\_ptr} nuedge, \textit{int} dir):} Replace
-the i'th edge in the face with another edge. The direction in which the
-new edge is used in the face must also be supplied. This function is
-relevant only for full representations in MSTK.
-
 \item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
@@
 
+\item[]\textbf{\textit{void} MF\_Insert\_Vertex(\textit{MFace\_ptr} mface,
+\textit{MVertex\_ptr} nuvertex, \textit{MVertex\_ptr} b4vertex):}
+Insert a vertex in the face before \textbf{b4vertex} w.r.t. a ccw
+ordering of the vertex faces.
+
+\item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
+mface, \textit{MVertex\_ptr} nuvertex, \textit{int} i):} Insert a new
+vertex before vertex i in the face w.r.t. a ccw ordering of vertices.
+
+\item[]
+
 \item[]\textbf{\textit{Mesh\_ptr} MF\_Mesh(\textit{MFace\_ptr} mf):} Returns the mesh that this mesh belongs to.
@@
 Returns NULL if this information is not known.
+
+\item[]
 
@@
 
+\item[]\textbf{\textit{List\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
+Return the set of regions connected to the face. If the face is not
+used by any regions, the function returns NULL to indicate that the
+set is empty. If not the set may contain one or two regions.
+
+\item[]\textbf{\textit{MRegion\_ptr} MF\_Region(\textit{MFace\_ptr} mface,
+\textit{int} side):} Returns the region on the specified side of the
+face. The positive side of the face (side = 1) is the side towards
+which the face normal points. The negative side of the face (side = 0)
+is the opposite side.
+
+\item[]
+
 \item[]\textbf{\textit{int} MF\_UsesEntity(\textit{MFace\_ptr} mface,
@@
 
-\item[]\textbf{\textit{List\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
-Return the set of regions connected to the face. If the face is not
-used by any regions, the function returns NULL to indicate that the
-set is empty. If not the set may contain one or two regions.
-
-\item[]\textbf{\textit{MRegion\_ptr} MF\_Region(\textit{MFace\_ptr} mface,
-\textit{int} side):} Returns the region on the specified side of the
-face. The positive side of the face (side = 1) is the side towards
-which the face normal points. The negative side of the face (side = 0)
-is the opposite side.
+\item[]\textbf{\textit{MFace\_ptr} MVs\_CommonFace(\textit{int} nv,
+    \textit{MVertex\_ptr} *fverts):} Return a face connected to all
+  the given mesh vertices. Returns NULL if no such face exists.
+
+\item[]\textbf{\textit{MFace\_ptr} MEs\_CommonFace(\textit{int} ne,
+    \textit{MEdge\_ptr} *fedges):} Return a face connected to all
+  the given mesh edges. Returns NULL if no such face exists.
+
+\item[]
 
@@
 \begin{description}
-\item[]\textbf{\textit{MRegion\_ptr} MR\_New(\textit{Mesh\_ptr} mesh):}
-Create a new region object. No topological information is embedded in
-the region when it is created. The region only knows which mesh it
-belongs to. The ID of the region is set by this function.
-
-\item[]\textbf{\textit{void} MR\_Delete(\textit{MRegion\_ptr} mregion):}
-Delete the region. Deletes all topological information embedded in the
-edge.
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntity(\textit{MRegion\_ptr} mregion,
-GEntity\_ptr gent):} Set the geometric model entity on which the edge
-is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntDim(\textit{MRegion\_ptr} mregion,
-\textit{int} gdim):} Set the dimension of the geometric model entity on
-which the edge is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntID(\textit{MRegion\_ptr} mregion,
-\textit{int} gid):} Set the ID of the geometric model entity on which
-the edge is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_ID(\textit{MRegion\_ptr} mregion,
-\textit{int} id):} Explicitly set ID of an edge and overwrite the ID
-set by the \textbf{MR\_New} function. Does not check for duplication
-of region IDs.
-
-\item[]\textbf{\textit{void} MR\_Set\_Faces(\textit{MRegion\_ptr} mregion,
-\textit{int} nf, \textit{MFace\_ptr} *mfaces, \textit{int} *dirs):} Set
-the faces of the region along with their directions. The
-\textit{unordered} set of faces and their directions are passed in
-through arrays along with the number of faces. If the normal of the
-face points out of the region, the associated direction to be passed
-in is 1; otherwise it is 0. This function is only relevant for full
-representations in MSTK.
+\item[]\textbf{\textit{MRegion\_ptr} MR\_New(\textit{Mesh\_ptr}
+    mesh):} Create a new region object. No topological information is
+  embedded in the region when it is created. The region only knows
+  which mesh it belongs to. The ID of the region is set by this
+  function.
+
+\item[]\textbf{\textit{void} MR\_Delete(\textit{MRegion\_ptr} mregion,
+    \textit{int} keep):} Delete the region. Deletes all topological
+  information embedded in the region. If \textbf{keep} is 0, the
+  region is removed from the mesh and all topological and geometric
+  information embedded in the region is destroyed. If \textbf{keep} is
+  1, the region is marked as type \textit{MDELREGION} and removed from
+  the mesh but the region is not destroyed. Also, the vertices, edges
+  and faces of this region no longer point up to this region.
+
+\item[]\textbf{\textit{void} MR\_Restore(\textit{MRegion\_ptr}
+    mregion):} Restore a temporarily deleted region. The region type
+  is restored from \textit{MDELREGION} to \textit{MREGION} and the
+  region is added back to the mesh. The vertices, edges and faces of
+  the region once again point back to the region.
+
+\item[]
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntity(\textit{MRegion\_ptr}
+    mregion, GEntity\_ptr gent):} Set the geometric model entity on
+  which the region is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntDim(\textit{MRegion\_ptr}
+    mregion, \textit{int} gdim):} Set the dimension of the geometric
+  model entity on which the region is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_GEntID(\textit{MRegion\_ptr}
+    mregion, \textit{int} gid):} Set the ID of the geometric model
+  entity on which the region is classified.
+
+\item[]\textbf{\textit{void} MR\_Set\_ID(\textit{MRegion\_ptr}
+    mregion, \textit{int} id):} Explicitly set ID of a region and
+  overwrite the ID set by the \textbf{MR\_New} function. Does not
+  check for duplication of region IDs.
+
+\item[]\textbf{\textit{void} MR\_Set\_Faces(\textit{MRegion\_ptr}
+    mregion, \textit{int} nf, \textit{MFace\_ptr} *mfaces,
+    \textit{int} *dirs):} Set the faces of the region along with their
+  directions. The \textit{unordered} set of faces and their directions
+  are passed in through arrays along with the number of faces. If the
+  normal of the face points out of the region, the associated
+  direction to be passed in is 1; otherwise it is 0. This function is
+  only relevant for full representations in MSTK.
 
 \item[]\textbf{\textit{void} MR\_Set\_Vertices(\textit{MRegion\_ptr}
-mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set the
-vertices of the region. This function is relevant for reduced
-representations only. For standard elements, the vertices must be
-ordered as indicated in Appendix~\ref{app:reg_conventions}. This
-function does not apply for defining general regions since there
-cannot be any implicit ordering.
+    mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set
+  the vertices of the region. This routine will collect/build all
+  lower order topological information (edges, edge directions in the
+  face) as needed by the face. For standard elements, the vertices
+  must be ordered as indicated in Appendix~\ref{app:reg_conventions},
+  while \textbf{nf} can be 0 and \textbf{rfv\_template} can be NULL.
+  For non-standard elements, \textbf{nf} is the number of faces of the
+  polyhedron and \textbf{rfv\_template} gives the vertices used in each
+  face. More precisely, \textbf{rfv\_template} has \textbf{nfv}
+  pointers to integer array $i$. The first entry of each
+  \textbf{rfv\_template$[i]$} represents the number of vertices in that
+  face and the remaining entries represent the list of vertices of the
+  face listed so that the face defined by them points outward.
 
@@
     mregion, \textit{MFace\_ptr} mface, \textit{MFace\_ptr} nuface,
-\textit{int} dir):} Replace a face of the region with another edge. The
-direction in which the new face is used in the region must also be
-supplied. This function is only relevant for full representations in
-MSTK.
+    \textit{int} dir):} Replace a face of the region with another
+  face. The direction in which the new face is used in the region must
+  also be supplied. This function is only relevant for full
+  representations in MSTK.
 
@@
 
-\item[]\textbf{\textit{void} MR\_Replace\_Face\_i(\textit{MRegion\_ptr}
-mregion, \textit{int} i, \textit{MFace\_ptr} mface, \textit{int} dir):}
-Replace the i'th face in the region with another face.The direction in
-which the new face is used in the region must also be supplied. This
-function is only relevant for full representations in MSTK.
-
-\item[]\textbf{\textit{void} MR\_Replace\_Vertex\_i(\textit{MRegion\_ptr}
-mregion, \textit{int} i, \textit{MVertex\_ptr} mvertex):} Replace the
-i'th vertex of the region with another vertex. This function is only
-relevant for reduced representations in MSTK.
-
-\item[]\textbf{\textit{Mesh\_ptr} MR\_Mesh(\textit{MRegion\_ptr} mregion):}
-Returns the mesh that the region belongs to.
-
-\item[]\textbf{\textit{int} MR\_ID(\textit{MRegion\_ptr} mregion):} Returns
-the ID of the region. Returns 0 if not known.
-
-\item[]\textbf{\textit{int} MR\_GEntDim(\textit{MRegion\_ptr} mregion):}
-Returns the dimension of the geometric model entity the region is
-classified on. Always returns 3 since a mesh region can be classified
-only on a model region.
-
-\item[]\textbf{\textit{int} MR\_GEntID(\textit{MRegion\_ptr} mregion):}
-Returns the ID of the geometric model entity that the region is
-classified on. Returns 0 if not known.
+\item[]\textbf{\textit{void}
+    MR\_Replace\_Face\_i(\textit{MRegion\_ptr} mregion, \textit{int}
+    i, \textit{MFace\_ptr} mface, \textit{int} dir):} Replace the i'th
+  face in the region with another face.The direction in which the new
+  face is used in the region must also be supplied. This function is
+  only relevant for full representations in MSTK.
+
+\item[]\textbf{\textit{void}
+    MR\_Replace\_Vertex\_i(\textit{MRegion\_ptr} mregion, \textit{int}
+    i, \textit{MVertex\_ptr} mvertex):} Replace the i'th vertex of the
+  region with another vertex. This function is only relevant for
+  reduced representations in MSTK.
+
+\item[]
+
+\item[]\textbf{\textit{Mesh\_ptr} MR\_Mesh(\textit{MRegion\_ptr}
+    mregion):} Returns the mesh that the region belongs to.
+
+\item[]\textbf{\textit{int} MR\_ID(\textit{MRegion\_ptr} mregion):}
+  Returns the ID of the region. Returns 0 if not known.
+
+\item[]\textbf{\textit{int} MR\_GEntDim(\textit{MRegion\_ptr}
+    mregion):} Returns the dimension of the geometric model entity the
+  region is classified on. Always returns 3 since a mesh region can be
+  classified only on a model region.
+
+\item[]\textbf{\textit{int} MR\_GEntID(\textit{MRegion\_ptr}
+    mregion):} Returns the ID of the geometric model entity that the
+  region is classified on. Returns 0 if not known.
 
 \item[]\textbf{\textit{GEntity\_ptr} MR\_GEntity(\textit{MRegion\_ptr}
-  mregion):} Returns a pointer or handle to the geometric model entity
-  that the vertex is classified on. Returns NULL if this information
-  is not known.
+    mregion):} Returns a pointer or handle to the geometric model
+  entity that the vertex is classified on. Returns NULL if this
+  information is not known.
+
+\item[]
 
@@
 regions sharing a face with this region.
+
+\item[]
 
@@
 
+\item[]
+
 \item[]\textbf{\textit{int} MR\_UsesEntity(\textit{MRegion\_ptr} mregion,
@@
 entity is
+
+\item[]
 
@@
 
+\item[]\textbf{\textit{int} MEnt\_OrigDim(\textit{MEntity\_ptr} mentity):}
+Returns the original topological dimension or type of generic entity before it
+was temporarily deleted.
+
 \item[]\textbf{\textit{Mesh\_ptr} MEnt\_Mesh(\textit{MEntity\_ptr} mentity):}
@@
 
-\item[]\textbf{\textit{int} MEnt\_GEntDim(\textit{MEntity\_ptr} mentity):}
-Returns the dimension of the geometric model entity that the entity is
-classified on.
-
-\item[]\textbf{\textit{GEntity\_ptr} MEnt\_GEntity(\textit{MEntity\_ptr} mentity):}
-Returns a pointer or handle to geometric model entity that the entity
-is classified on.
+\item[]\textbf{\textit{int} MEnt\_GEntDim(\textit{MEntity\_ptr}
+    mentity):} Returns the dimension of the geometric model entity
+  that the entity is classified on.
+
+\item[]\textbf{\textit{GEntity\_ptr}
+    MEnt\_GEntity(\textit{MEntity\_ptr} mentity):} Returns a pointer
+  or handle to geometric model entity that the entity is classified
+  on.
+
+\item[]
+
+\item[]\textbf{\textit{void} MEnt\_Set\_AttVal(\textit{MEntity\_ptr}
+    ent, \textit{MAttrib\_ptr} attrib, \textit{int} ival,
+    \textit{double} lval, \textit{void} *pval):} Set the value of an
+  attribute for given mesh entity. Depending on whether the attribute
+  is integer, double or pointer, the appropriate entry in the argument
+  list is specified.
+
+\item[]\textbf{\textit{void} MEnt\_Rem\_AttVal(\textit{MEntity\_ptr}
+    ent, \textit{MAttrib\_ptr mattrib}:} Clear the value of the given
+  attribute from the entity.
+
+\item[]\textbf{\textit{int} MEnt\_Get\_AttVal(\textit{MEntity\_ptr}
+    ent, \textit{MAttrib\_ptr} int *ival, double *rval, double
+    **pval):} Get the value of an attribute from an entity.  Depending
+  on the attribute type, either \textbf{ival}, \textbf{rval} or
+  \textbf{pval} is returned.
+
+\end{description}
+
+\newpage
+\subsection{Mesh Attributes}
+
+\begin{description}
+
+\item[]\textbf{\textit{MAttrib\_ptr} MAttrib\_New(\textit{Mesh\_ptr}
+    mesh, \textit{const char} *att\_name, \textit{MAttType} att\_type,
+    \textit{MType} entdim):} Define a new mesh attribute. Along with
+  the mesh to which this attribute is assigned (\textbf{mesh}) and the
+  name of the attribute (\textbf{att\_name}), the type of the
+  attribute (\textbf{att\_type}) must be specified as \textbf{INT},
+  \textbf{DOUBLE} or \textbf{POINTER}. Also, the dimension of the entity
+    for which the attribute value can be set must be specified as
+    \textbf{MVERTEX}, \textbf{MEDGE}, \textbf{MFACE}, \textbf{MREGION}
+    or \textbf{MALLTYPE}.
+
+\item[]\textbf{\textit{char} *MAttrib\_Get\_Name(\textit{MAttrib\_ptr}
+    attrib, \textit{char} *att\_name):} Get the name of the given attribute.
+
+\item[]\textbf{\textit{MAttType}
+    MAttrib\_Get\_Type(\textit{MAttrib\_ptr} attrib):} Get the type of
+  the attribute (Can return \textbf{INT}, \textbf{DOUBLE},
+  \textbf{PVAL}).
+
+\item[]\textbf{\textit{MType}
+    MAttrib\_Get\_EntDim(\textit{MAttrib\_ptr} attrib):} Get the
+  dimension (or type) of entity attribute can be asssigned to. Can be
+  \textbf{MVERTEX}, \textbf{MEDGE}, \textbf{MFACE}, \textbf{MREGION}
+  or \textbf{MALLTYPE}.
+
+\item[]\textbf{\textit{void} MAttrib\_Delete(\textit{MAttrib\_ptr}
+    attrib):} Delete an attribute.
+
 \end{description}

### 2004-07-28 docs/MSTK.tex
Updated documentation for version 1.3

@@
   very large.
+
+\item[]\textbf{\textit{int} MESH\_DelInterior(\textit{Mesh\_ptr}
+    mesh):} Delete the interior of a mesh and retain only its
+    boundaries (including interior boundaries). For solid meshes, this
+    results in a surface mesh (i.e., all mesh regions, and  mesh
+    faces, edges and vertices classified on model regions are deleted). For
+    surface meshes, this results in curve mesh. For an edge mesh, only
+    the end vertices are retained (unlikely to be used this
+    way). Undefined for a vertex mesh.
 

### 2004-07-28 docs/MSTK.tex
*** empty log message ***

@@
 \item[]\textbf{\textit{void} MR\_Set\_Vertices(\textit{MRegion\_ptr}
-    mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set
+    mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices, 
+    \textit{int} nf, \textit{int} **rfv\_template):} Set
   the vertices of the region. This routine will collect/build all
@@
   face and the remaining entries represent the list of vertices of the
-  face listed so that the face defined by them points outward.
+  face listed so that the face defined by them points out of the region.
 

### 2010-02-25 docs/MSTK.tex
Fixed error in description of MF_Coords
Updated!

--- 2004-07-28 docs/MSTK.tex
@@
 
-\title{MSTK: Mesh Toolkit, v 1.3 - DRAFT}
+\title{MSTK: Mesh Toolkit, v 1.7 - DRAFT}
 
 \author{Rao V. Garimella \\
-  Mathematical Modeling and Analysis (T-7), Theoretical Division, \\
+  T-5, Theoretical Division, \\
   Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
@@
 
+\subsection{What is MSTK?}
 MSTK or Mesh Toolkit is a mesh framework that allows users to
@@
 meshes in a general manner without the need to code their own data
-structures. MSTK is a flexible framework in that it allows (or will
-eventually allow) a wide variety of underlying representations for the
-mesh while maintaining a common interface. It will allow users to
-choose from different mesh representations either at initialization or
-during the program execution so that the optimal data structures are
-used for the particular algorithm. The interaction of users and
-applications with MSTK is through a functional interface that acts as
-though the mesh always contains vertices, edges, faces and regions and
-maintains connectivity between all these entities.
+structures.
+
+\subsection{What MSTK is not?}
+
+MSTK is not a mesh generator - it can be used to write more easily
+than starting from scratch. Also, MSTK cannot answer computational
+questions related to the mesh (for example, is this point in the given
+mesh element?).
+
+\subsection{Salient Features of MSTK}
+
+MSTK is a flexible framework in that it allows a variety
+of underlying representations for the mesh while maintaining a common
+interface. It will allow users to choose from different mesh
+representations either at initialization (implemented) or during the
+program execution (not implemented) so that the optimal data
+structures are used for the particular algorithm. The interaction of
+users and applications with MSTK is through a functional interface
+that acts as though the mesh always contains vertices, edges, faces
+and regions and maintains connectivity between all these entities.
 
@@
 Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
-FEA Applications,'' \textit{International Journal of Numerical Methods
+FEA Applications,'' {\em International Journal of Numerical Methods
   in Engineering}, v55 n4, pp. 441-478, 2002.
 
-In the following sections, the data types of MSTK will be described
-followed by a description of the functional interface. The MSTK file
-format will be described in the last section.
-
-\newpage
-\section{MSTK Data Types}
-
-\begin{description}
-\item[]\textbf{\textit{List\_ptr}}: Handle to a List object.
-
-\item[]\textbf{\textit{Mesh\_ptr}}: Handle to a Mesh object.
-
-\item[]\textbf{\textit{MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
-Dimension 0).
-
-\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1).
-
-\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2).
-
-\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3).
-
-\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity
-  object. Any of the above types of entities can be cast as
-  \textit{MEntity\_ptr}.
-
-\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object.
-\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object.
-
-\item[]\textbf{\textit{MAttrib\_ptr}}: Handle to a mesh attribute.
-
-\item[]\textbf{\textit{RepType}}: Enumerated type describing the type
-  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
-  See Appendix~\ref{app:reptypes} for schematics of these
-  representations. \textit{Currently only representation types F1 and
-    F4 are supported}.
-
-\item[]\textbf{\textit{MType}}: Enumerated type for mesh entity type. Can be MDELETED, MVERTEX, MEDGE, MFACE, MREGION, MUNKNOWNTYPE, MALLTYPE.
-
-\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type.
-Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON.
-
-\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type.
-Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED.
-
-\item[]\textbf{\textit{AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
-\end{description}
-
-\newpage
-\section{MSTK Functional Interface}
-\subsection{List}
-
-Lists of entities in MSTK are returned as sets of type
-\textit{List\_ptr}. The following are the set operations available in
-MSTK:
-
-\begin{description}
-\item[]\textbf{\textit{List\_ptr} List\_New(\textit{int} inisize):} Create a
-new set with an initial size, \textit{inisize}. If \textit{inisize} is
-0, the initial size is set to be 10.
-
-\item[]\textbf{\textit{void} List\_Delete(\textit{List\_ptr} l):} Delete a set.
-
-\item[]\textbf{\textit{List\_ptr} List\_Compress(\textit{List\_ptr}
-    l):} Compress a set.  Doing this while an algorithm is iterating
-  through the set can currently cause problems!! Calling
-  \textbf{List\_Compress} could change the pointer for the set due to
-  reallocation.
-
-\item[]\textbf{\textit{List\_ptr} List\_Copy(\textit{List\_ptr} l):} Return a
-copy of a set.
-
-\item[]\textbf{\textit{List\_ptr} List\_Add(\textit{List\_ptr} l, \textit{void
-  *}entry):} Add an entry to the set. The entry is appended to the end
-of the set.
-
-\item[]\textbf{\textit{List\_ptr} List\_ChknAdd(\textit{List\_ptr} l,
-\textit{void *}entry):} Add an entry to a set only if it is not already
-in the set.
-
-\item[]\textbf{\textit{int} List\_Rem(\textit{List\_ptr} l, \textit{void
-  *}entry):} Remove an entry from the set. Returns 1 if successful, 0
-otherwise.
-
-\item[]\textbf{\textit{int} List\_Remi(\textit{List\_ptr} l, \textit{int} i):}
-Remove the i'th valid entry in the set. Returns 1 if successful, 0
-otherwise.
-
-\item[]\textbf{\textit{int} List\_Replace(\textit{List\_ptr} l, \textit{void
-  *}entry, \textit{void *}nuentry):} Replace 'entry' with 'nuentry' in
-set. Returns 1 if successful, 0 otherwise.
-
-\item[]\textbf{\textit{int} List\_Replacei(\textit{List\_ptr} l, \textit{int}
-i, void *nuentry):} Replace the i'th valid entry in the set with
-'nuentry'. Returns 1 if successful, 0 otherwise.
-
-\item[]\textbf{\textit{int} List\_Contains(\textit{List\_ptr} l, void *entry):}
-Returns 1 if set contains the entry, 0 otherwise.
-
-\item[]\textbf{\textit{int} List\_Locate(\textit{List\_ptr} l, void *entry):}
-Returns the positional index of the entry in the set. Returns -1 if
-the set does not contain the entry.
-
-\item[]\textbf{\textit{void *}List\_Entry(\textit{List\_ptr} l, \textit{int}
-i):} Return the i'th valid entry in the set.  Returns a NULL pointer if
-the i'th valid entry could not be found.
-
-\item[]\textbf{\textit{void *}List\_Next\_Entry(\textit{List\_ptr} l,
-\textit{int} *i):} Return the next valid entry in the set. This routine
-works like an iterator. To start iterating through the set, set the
-iteration index i=0 and call the routine to get the first entry in the
-set. Subsequent calls to the routine will iterate through the entries
-in the set. The routine will return a NULL to indicate that the end of
-the set is reached.
-
-The value of the iteration index i will be modified by the routine on
-each call to indicate where in the set it is. This value should not be
-modified externally while iterating through the set. Also, no specific
-meaning should be derived from the iteration index by other applications
-since the internal implementation and interpretation of the index may
-change at any time.
-
-\item[]\textbf{\textit{int} List\_Num\_Entries(\textit{List\_ptr} l):} Return
-the number of entries in a set
-\end{description}
-
-\newpage
-\subsection{Mesh Object}
-
-A mesh object is a set of vertices (nodes) possibly connected by other
-entities such as edges, faces, regions. Depending on the
-representation chosen and type of mesh, some or all of the entities
-may be explicitly stored.  Full representations contain all types of
-entities up to the highest dimension of the mesh. For example, a full
-representation of a tetrahedral mesh contains vertices, edges, faces
-and regions. However, one type of reduced representation of this mesh
-may contain only vertices and regions. For a surface mesh, a full
-representation includes vertices, edges and faces while a reduced
-representation only has vertices and faces. Also, depending on the
-type of representation, some adjacencies (information about which
-entities are connected to which other entities) are stored and others
-are derived.
-
-\begin{description}
-\item[]\textbf{\textit{Mesh\_ptr} MESH\_New(\textit{RepType} type):}
-Initialize a new mesh object with the given representation type which
-can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Not all of these types
-are implemented. If the representation type is not known at the
-present time (e.g. before reading the mesh from a file), the
-representation type of UNKNOWN\_REP can be specified.  Note that this
-only initializes a mesh object, it does not create or generate a mesh
-which is the work of high level mesh generation routines.
-
-\item[]\textbf{\textit{int} MESH\_InitFromFile(\textit{Mesh\_ptr} mesh,
-const char *filename):} Initialize or read a mesh from file into the
-given mesh object. Returns 1 if successful, 0 otherwise.
-
-\item[]\textbf{\textit{int} MESH\_ImportFromFile(\textit{Mesh\_ptr}
-    mesh, const char *filename, const char *format):} Import a mesh
-  data from an external file format and construct MSTK mesh. Currently
-  the only format supported is the
-  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
-  file format. The routine imports as many attributes as it can from
-  the input file. It also uses special attributes or keywords as data
-  about element and node classification.  For GMV files, the routine
-  uses the ``material'' data to indicate region or face classification
-  depending on the dimensionality of the mesh. If ``material'' data is
-  not specified, it uses the ``itetclr'' attribute to assign region or
-  face classification. It also uses the ``icr'' keyword describing the
-  number of constraints on a node to interpret if the node is
-  classified on a model region, model face, model edge or a model
-  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
-  meshes generated by
-  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.
-
-\item[]\textbf{\textit{int} MESH\_BuildClassfn(\textit{Mesh\_ptr}
-    mesh):} Build classification information for mesh entities if only
-  partial information is present. In other words, if the only data
-  known is the IDs of geometric model entities (``material regions'')
-  of the highest level mesh entities (faces or regions), then this
-  procedure will build information about the type of geometric model
-  entities that the lower dimension entities are on. Faces will be
-  classified as being on a model face (external or interior) or inside
-  model region. Edges will be classified as being on a model edge,
-  model face or in a model region. Vertices will be classified as
-  being on a model vertex, model edge, model face or in a model
-  region. If no classification data is associated with even the
-  highest level entities, the procedure will assume that all the
-  highest level entities are classified on one model entitiy of that
-  dimension. If partial information is available for lower order
-  entities, this routine will not destroy that information. In the
-  future, the procedure will be more sophisticated, incorporating
-  features such as classifying mesh edges as being on model edges if
-  the dihedral angle between the boundary faces connected to it is
-  very large.
-
-\item[]\textbf{\textit{int} MESH\_DelInterior(\textit{Mesh\_ptr}
-    mesh):} Delete the interior of a mesh and retain only its
-    boundaries (including interior boundaries). For solid meshes, this
-    results in a surface mesh (i.e., all mesh regions, and  mesh
-    faces, edges and vertices classified on model regions are deleted). For
-    surface meshes, this results in curve mesh. For an edge mesh, only
-    the end vertices are retained (unlikely to be used this
-    way). Undefined for a vertex mesh.
-
-\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr}
-    mesh, const char *filename):} Save a mesh to a filename. The file
-  is created if it does not exists. It is recommended that the .mstk
-  extension be used for MSTK mesh files.  However, there is no such
-  requirement.
-
-%% \item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
-%%     mesh, const char *filname, const char *format, int natt, char
-%%     **attnames):} Export a mesh to an external file format. Currently,
-%%   the formats supported are the GMV file (format string: ``gmv'') and
-%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
-%%   ``dx''). Only integer and double attributes of the mesh are exported
-%%   to the files.
-
-\item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
-    mesh, const char *filname, const char *format, int natt, char
-    **attnames):} Export a mesh to an external file format. Currently,
-  the only format supported is the GMV file format (format string:
-  ``gmv''). Only integer and double attributes of the mesh are
-  exported to the files.
-
-\item[]
-
-\item[]\textbf{\textit{GModel\_ptr} MESH\_GModel(\textit{Mesh\_ptr} mesh):}
-Return a handle to the underlying geometric model. If there is no
-geometric model associated with the mesh, NULL pointer is returned.
-
-\item[]\textbf{\textit{RepType} MESH\_RepType(\textit{Mesh\_ptr} mesh):}
-Representation type currently being used by the mesh.
-
-\item[]
-
-\item[]\textbf{\textit{int} MESH\_Num\_Vertices(\textit{Mesh\_ptr} mesh):}
-Number of vertices in the mesh.
-
-\item[]\textbf{\textit{int} MESH\_Num\_Edges(\textit{Mesh\_ptr} mesh):}
-Number of edges in the mesh. For reduced representations, this routine
-returns 0 since it is impractically expensive to count the number of
-edges when they do not explicitly exist. Applications must find a way
-to avoid using this routine for reduced representations.
-
-\item[]\textbf{\textit{int} MESH\_Num\_Faces(\textit{Mesh\_ptr} mesh):}
-Number of faces in the mesh. For reduced representations R1 or R2,
-this routine counts only the faces that are explicitly represented
-i.e. faces not connected to any mesh region. Therefore, a value of 0
-will be returned for the number of faces of a tetrahedral mesh with
-representation R1 or R2 but the correct number will be reported for a
-tetrahedral mesh in other representations. Also, the correct number
-will be reported for the number of faces in a surface mesh in
-representation R1 or R2.  Therefore, this routine must be used
-carefully.
-
-\item[]\textbf{\textit{int} MESH\_Num\_Regions(\textit{Mesh\_ptr} mesh):}
-Number of regions in the mesh.
-
-\item[]
-
-\item[]\textbf{\textit{MVertex\_ptr} MESH\_Vertex(\textit{Mesh\_ptr} mesh,
-\textit{int} i):} Return the i'th vertex in the mesh. Returns NULL if i
-$<$ 0 or i $>$ number of mesh vertices.
-
-\item[]\textbf{\textit{MEdge\_ptr} MESH\_Edge(\textit{Mesh\_ptr} mesh,
-\textit{int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
-0 or i $>$ number of mesh edges. Returns NULL for reduced
-representations.
-
-\item[]\textbf{\textit{MFace\_ptr} MESH\_Face(\textit{Mesh\_ptr} mesh,
-\textit{int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
-0 or i $>$ number of mesh faces. Only faces explicitly represented in
-the mesh are returned for reduced representation (See explanation for
-MESH\_Num\_Faces).
-
-\item[]\textbf{\textit{MRegion\_ptr} MESH\_Region(\textit{Mesh\_ptr} mesh,
-\textit{int} i):} Return the i'th region in the mesh. Returns NULL if i
-$<$ 0 or i $>$ number of mesh region.
-
-\item[]
-
-\item[]\textbf{\textit{MVertex\_ptr} MESH\_Next\_Vertex(\textit{Mesh\_ptr}
-mesh, \textit{int} *idx):} Returns the next vertex while iterating
-through the vertices of the mesh. See the routine \textbf{List\_Next\_Entry}
-above for an explanation of how the iteration works.
-
-\item[]\textbf{\textit{MEdge\_ptr} MESH\_Next\_Edge(\textit{Mesh\_ptr} mesh,
-\textit{int} *idx):} Returns the next edge while iterating through the
-edges of the mesh. See the routine List\_Next\_Entry above for an
-explanation of how the iteration works.  The routine always returns
-NULL for reduced representations.
-
-\item[]\textbf{\textit{MFace\_ptr} MESH\_Next\_Face(\textit{Mesh\_ptr} mesh,
-\textit{int} *idx):} Returns the next face while iterating through the
-faces of the mesh. See the routine \textbf{List\_Next\_Entry} above for
-an explanation of how the iteration works.  Only faces explicitly
-represented in the mesh are returned for reduced representation (See
-explanation for \textbf{MESH\_Num\_Faces}).
-
-\item[]\textbf{\textit{MRegion\_ptr}
-    MESH\_Next\_Region(\textit{Mesh\_ptr} mesh, \textit{int} *idx):}
-  Returns the next region while iterating through the regions of the
-  mesh. See the routine \textbf{List\_Next\_Entry} above for an
-  explanation of how the iteration works.
-
-\item[]
-
-\item[]\textbf{\textit{MVertex\_ptr}
-    MESH\_VertexFromID(\textit{Mesh\_ptr} mesh, \textit{int} id):}
-  Return mesh vertex with given ID if it exists; return NULL
-  otherwise.
-
-\item[]\textbf{\textit{MEdge\_ptr} MESH\_EdgeFromID(\textit{Mesh\_ptr}
-    mesh, \textit{int} id):} Return mesh edge with given ID if it
-  exists; return NULL otherwise. This routine will return NULL for all
-  reduced representations.
-
-\item[]\textbf{\textit{MFace\_ptr} MESH\_FaceFromID(\textit{Mesh\_ptr}
-    mesh, \textit{int} i):} Return mesh face with given ID if it
-  exists; return NULL otherwise. If faces are not explicitly
-  represented in the mesh, it will return NULL.
-
-\item[]\textbf{\textit{MRegion\_ptr}
-    MESH\_RegionFromID(\textit{Mesh\_ptr} mesh, \textit{int} id):}
-  Return mesh region with given ID if it exists; return NULL
-  otherwise.
-
-\item[]
-
-\item[]\textbf{\textit{int} MESH\_Num\_Attribs(\textit{Mesh\_ptr}
-    mesh):} Number of attributes associated with the mesh.
-
-\item[]\textbf{\textit{MAttrib\_ptr} MESH\_Attrib(\textit{Mesh\_ptr}
-    mesh, \textit{int} i):} Return the i'th attribute in the mesh.
-  Returns NULL if i $<$ 0 or i $>$ number of mesh attributes.
-
-\item[]\textbf{\textit{MAttrib\_ptr}
-    MESH\_Next\_Attrib(\textit{Mesh\_ptr} mesh, \textit{int} *index):}
-  Returns the next attribute while iterating through the attributes of
-  the mesh. See the routine List\_Next\_Entry above for an explanation
-  of how the iteration works.
-
-\item[]\textbf{\textit{MAttrib\_ptr}
-    MESH\_AttribByName(\textit{Mesh\_ptr} mesh, \textit{const char}
-    *name):} Return a mesh attribute with given name if it exisists in
-  the mesh. Returns NULL if mesh has no such attribute.
-
-\item[]
-
-\item[]\textbf{\textit{void} MESH\_Set\_GModel(\textit{Mesh\_ptr} mesh,
-GModel\_ptr geom):} Assign a geometric model handle to the mesh.
-
-\item[]\textbf{\textit{int} MESH\_Change\_RepType(\textit{Mesh\_ptr} mesh,
-\textit{int} nurep):} Change the representation type of the mesh. This
-routine can be used to modify the representation type dynamically to
-suit different algorithms.  However, the cost of making the change and
-reordering all adjacencies and creating or deleting entities has to be
-considered while invoking this routine. Also, once a conversion is
-made from a full representation to a reduced representation, not all
-information may be retrievable when switching back to a full
-representation.  (particularly classification information, i.e.,
-relationship of mesh entities to the geometric model).
-
-\item[]
-
-\item[]\textbf{\textit{void} MESH\_Add\_Vertex(\textit{Mesh\_ptr}
-    mesh, \textit{MVertex\_ptr} v):} Add a vertex to the mesh. It is
-  assumed that the vertex and its coordinates set are properly
-  defined. Normally, one need not call this since
-  \textbf{\textit{MV\_New}} will add the vertex to the mesh. Use this
-  only if you know exactly what you are doing.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Edge(\textit{Mesh\_ptr} mesh,
-    \textit{MEdge\_ptr} e):} Add an edge to the mesh. It is assumed
-  that the edge is and its topology is defined. Normally, one need not
-  call this since \textbf{\textit{ME\_New}} will add the edge to the
-  mesh. Use this only if you know exactly what you are doing.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Face(\textit{Mesh\_ptr} mesh,
-    \textit{MFace\_ptr} f):} Add a face to the mesh. It is assumed
-  that the face and its topology is properly defined. Normally, one
-  need not call this since \textbf{\textit{MF\_New}} will add the face
-  to the mesh. Use this only if you know exactly what you are doing.
-
-\item[]\textbf{\textit{void} MESH\_Add\_Region(\textit{Mesh\_ptr}
-    mesh, \textit{MRegion\_ptr} r):} Add a region to the mesh. It is
-  assumed that the region and its topology is properly defined.
-  Normally, one need not call this since \textbf{\textit{MR\_New}}
-  will add the region to the mesh. Use this only if you know exactly
-  what you are doing.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Vertex(\textit{Mesh\_ptr}
-    mesh, \textit{MVertex\_ptr} v):} Remove vertex from mesh. Vertex
-  is not deleted and must be deleted afterward separately. Normally,
-  one need not call this since \textbf{\textit{MV\_Delete}} will
-  remove the vertex from the mesh. Use this only if you know exactly
-  what you are doing.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Edge(\textit{Mesh\_ptr} mesh,
-    \textit{MEdge\_ptr} e):} Remove edge from mesh. Edge is not
-  deleted and must be deleted afterward separately. Normally, one need
-  not call this since \textbf{\textit{ME\_Delete}} will remove the
-  edge from the mesh. Use this only if you know exactly what you are
-  doing.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Face(\textit{Mesh\_ptr} mesh,
-    \textit{MFace\_ptr} f):} Remove face from mesh. Face is not
-  deleted and must be deleted afterward separately. Normally, one need
-  not call this since \textbf{\textit{MF\_Delete}} will remove the
-  face from the mesh. Use this only if you know exactly what you are
-  doing.
-
-\item[]\textbf{\textit{void} MESH\_Rem\_Region(\textit{Mesh\_ptr}
-    mesh, \textit{MRegion\_ptr} r):} Remove region from mesh. Region
-  is not deleted and must be deleted afterward separately. Normally,
-  one need not call this since \textbf{\textit{MR\_Delete}} will
-  remove the region from the mesh. Use this only if you know exactly
-  what you are doing.
-\end{description}
-
-\newpage
-\subsection{Mesh Vertex Object}
-
-\begin{description}
-\item[]\textbf{\textit{MVertex\_ptr} MV\_New(\textit{Mesh\_ptr}
-    mesh):} Create a new vertex object. No geometric or topological
-  information is embedded in the vertex when it is created. The vertex
-  only knows which mesh it belongs to. The ID of the vertex is set by
-  this function.
-
-\item[]\textbf{\textit{void} MV\_Delete(\textit{MVertex\_ptr} mvertex,
-    \textit{int} keep):} Delete the vertex. If \textbf{keep} is 0, the
-  vertex is removed from the mesh and all topological and geometric
-  information embedded in the vertex is destroyed. If \textbf{keep} is
-  1, the vertex is marked as type \textit{MDELVERTEX} and removed from
-  the mesh but the vertex is not destroyed.
-
-\item[]\textbf{\textit{void} MV\_Restore(\textit{MVertex\_ptr}
-    mvertex):} Restore a deleted vertex. The vertex type is restored
-  from \textit{MDELVERTEX} to \textit{MVERTEX} and the vertex is
-  added back to the mesh.
-
-\item[]
-
-\item[]\textbf{\textit{void} MV\_Set\_Coords(\textit{MVertex\_ptr}
-    mvertex, double *xyz):} Set the coordinates of the vertex.
-
-\item[]\textbf{\textit{void} V\_Set\_GEntity(\textit{MVertex\_ptr}
-    mvertex, GEntity\_ptr gent):} Set the geometric model entity on
-  which vertex is classified.
-
-\item[]\textbf{\textit{void} MV\_Set\_GEntDim(\textit{MVertex\_ptr} mvertex,
-\textit{int} gdim):} Set topological dimension of model entity on which
-vertex is classified.
-
-\item[]\textbf{\textit{void} MV\_Set\_GEntID(\textit{MVertex\_ptr} mvertex,
-\textit{int} gid):} Set ID of model entity on which vertex is
-classified.
-
-\item[]\textbf{\textit{void} MV\_Add\_AdjVertex(\textit{MVertex\_ptr}
-mvertex, \textit{MVertex\_ptr} adjvertex):} Add neighboring vertex,
-adjvertex, to ajdacent vertex list of vertex, mvertex.
-
-\item[]\textbf{\textit{void} MV\_Rem\_AdjVertex(\textit{MVertex\_ptr}
-mvertex, \textit{MVertex\_ptr} adjvertex):} Delete neighboring vertex
-of given vertex.
-
-\item[]\textbf{\textit{void} MV\_Set\_ID(\textit{MVertex\_ptr} mvertex,
-\textit{int} id):} Explicitly set ID of a vertex and overwrite the ID
-set by the MV\_New operator. Does not check for duplication of edge
-IDs.
-
-\item[]
-
-\item[]\textbf{\textit{Mesh\_ptr} MV\_Mesh(\textit{MVertex\_ptr} mv):}
-Returns the mesh that this vertex belongs to.
-
-\item[]\textbf{\textit{int} MV\_ID(\textit{MVertex\_ptr} mvertex):} Returns
-the ID of the vertex.
-
-\item[]\textbf{\textit{int} MV\_GEntDim(\textit{MVertex\_ptr} mvertex):}
-Returns the dimension of the geometric model entity that the vertex is
-classified on. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} MV\_GEntID(\textit{MVertex\_ptr} mvertex):}
-Returns the ID of the geometric model entity that the vertex is
-classified on. Returns 0 if this information is not known.
-
-\item[]\textbf{\textit{GEntity\_ptr} MV\_GEntity(\textit{MVertex\_ptr}
-  mvertex):} Returns a pointer or handle to the geometric model entity
-  that the vertex is classified on. Returns NULL if this information
-  is not known.
-
-\item[]\textbf{\textit{void} MV\_Coords(\textit{MVertex\_ptr} mvertex,
-double *xyz):} Returns the coordinates of the vertex.
-
-\item[]
-
-\item[]\textbf{\textit{int} MV\_Num\_AdjVertices(\textit{MVertex\_ptr}
-mvertex):} Returns the number of edge connected neighboring vertices of
-vertex. \textit{Not efficient for all representations}.
-
-\item[]\textbf{\textit{int} MV\_Num\_Edges(\textit{MVertex\_ptr} mvertex):}
-Returns the number of edges connected to the vertex.
-
-\item[]\textbf{\textit{int} MV\_Num\_Faces(\textit{MVertex\_ptr} mvertex):}
-Returns the number of faces connected to the vertex.
-
-\item[]\textbf{\textit{int} MV\_Num\_Regions(\textit{MVertex\_ptr} mvertex):}
-Returns the number of regions connected to the vertex
-
-\item[]\textbf{\textit{List\_ptr} MV\_AdjVertices(\textit{MVertex\_ptr}
-mvertex):} List of adjacent or edge connected neighboring vertices of vertex.
-
-\item[]\textbf{\textit{List\_ptr} MV\_Edges(\textit{MVertex\_ptr} mvertex):}
-List of edges connected to the vertex.
-
-\item[]\textbf{\textit{List\_ptr} MV\_Faces(\textit{MVertex\_ptr} mvertex):}
-List of faces connected to the vertex.
-
-\item[]\textbf{\textit{List\_ptr} MV\_Regions(\textit{MVertex\_ptr} mvertex):}
-List of regions connected to the vertex.
-\end{description}
-
-\newpage
-\subsection{Mesh Edge Object}
-
-\begin{description}
-\item[]\textbf{\textit{MEdge\_ptr} ME\_New(\textit{Mesh\_ptr} mesh):}
-  Create a new edge object. No topological information is embedded in
-  the edge when it is created. The edge only knows which mesh it
-  belongs to. The ID of the edge is set by this function.
-
-\item[]\textbf{\textit{void} ME\_Delete(\textit{MEdge\_ptr} medge,
-    \textit{int} keep):} Delete the edge. If \textbf{keep} is 0, the
-  edge is removed from the mesh and all topological and geometric
-  information embedded in the edge is destroyed. If \textbf{keep} is
-  1, the edge is marked as type \textit{MDELEDGE} and removed from the
-  mesh but the edge is not destroyed. Also, the vertices of this edge
-  no longer point to this edge.
-
-\item[]\textbf{\textit{void} ME\_Restore(\textit{MEdge\_ptr}
-    medge):} Restore a temporarily deleted edge. The edge type is
-  restored from \textit{MDELEDGE} to \textit{MEDGE} and the edge is
-  added back to the mesh. The vertices of the edge once again point
-  back to the edge.
-
-\item[]
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntity(\textit{MEdge\_ptr}
-    medge, GEntity\_ptr gent):} Set the geometric model entity on
-  which the edge is classified.
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntDim(\textit{MEdge\_ptr}
-    medge, \textit{int} gdim):} Set the topological dimension of model
-  entity on which edge is classified.
-
-\item[]\textbf{\textit{void} ME\_Set\_GEntID(\textit{MEdge\_ptr}
-    medge, \textit{int} gid):} Set ID of model entity on which edge is
-  classified.
-
-\item[]\textbf{\textit{void} ME\_Set\_ID(\textit{MEdge\_ptr} medge,
-    \textit{int} id):} Explicitly set ID of an edge and overwrite the
-  ID set by the ME\_New function. Does not check for duplication of
-  edge IDs.
-
-\item[]\textbf{\textit{void} ME\_Set\_Vertex(\textit{MEdge\_ptr}
-    medge, \textit{int} i, \textit{MVertex\_ptr} vertex):} Set the
-  i'th vertex of the edge. i can be 0 or 1.
-
-\item[]\textbf{\textit{void} ME\_Replace\_Vertex(\textit{MEdge\_ptr}
-    medge, \textit{MVertex\_ptr} vert, \textit{MVertex\_ptr} nuvert):}
-  Replace i'th vertex by new vertex.
-
-\item[]
-
-\item[]\textbf{\textit{Mesh\_ptr} ME\_Mesh(\textit{MEdge\_ptr}
-    medge):} Returns the mesh that this edge belongs to.
-
-\item[]\textbf{\textit{int} ME\_ID(\textit{MEdge\_ptr} medge):}
-  Returns the ID of the vertex. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} ME\_GEntDim(\textit{MEdge\_ptr} medge):}
-  Returns the dimension of the geometric model entity that the vertex
-  is classified on. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} ME\_GEntID(\textit{MEdge\_ptr} medge):}
-  Returns the ID of the geometric model entity that the vertex is
-  classified on.  Returns 0 if this information is not known.
-
-\item[]\textbf{\textit{GEntity\_ptr} ME\_GEntity(\textit{MEdge\_ptr}
-    medge):} Returns a pointer or handle to the geometric model entity
-  that the vertex is classified on. Returns NULL if this information
-  is not known.
-
-\item[]
-
-\item[]\textbf{\textit{int} ME\_Num\_Faces(\textit{MEdge\_ptr}
-    medge):} Returns the number of faces connected to the edge.
-
-\item[]\textbf{\textit{int} ME\_Num\_Regions(\textit{MEdge\_ptr}
-    medge):} Returns the number of regions connected to the edge.
-
-\item[]\textbf{\textit{MVertex\_ptr} ME\_Vertex(\textit{MEdge\_ptr}
-    medge, \textit{int} i):} Returns the i'th vertex of the edge. i=0
-  returns the first vertex and i=1 returns the second vertex.
-
-\item[]\textbf{\textit{MVertex\_ptr} ME\_OppVertex(\textit{MEdge\_ptr}
-    medge, \textit{MVertex\_ptr} ov):} Return the vertex opposite to
-  given vertex in edge.
-
-\item[]\textbf{\textit{int} ME\_UsesEntity(\textit{MEdge\_ptr} medge,
-    \textit{MEntity\_ptr} mentity, \textit{int} etype):} Check if edge
-  uses given lower dimension entity, \textit{mentity}. The dimension
-  of the entity is specified by the \textit{etype} variable. For an
-  edge, the only lower dimensional entity is a vertex. If the edge
-  uses the vertex, the function returns 1; otherwise it returns 0. If
-  any other type of entity is specified, the function returns 0.
-
-\item[]\textbf{\textit{List\_ptr} ME\_Faces(\textit{MEdge\_ptr}
-    medge):} Returns the set of faces using this edge.
-
-\item[]\textbf{\textit{List\_ptr} ME\_Regions(\textit{MEdge\_ptr}
-    medge):} Returns the set of regions using this edge.
-
-\item[]
-
-\item[]\textbf{\textit{MEdge\_ptr}
-    MVs\_CommonEdge(\textit{MVertex\_ptr} v1, \textit{MVertex\_ptr}
-    v2):} Return the edge connecting vertices v1 and v2, if it exists.
-  If such an edge does not exist, the function returns 0.
-
-\item[]\textbf{\textit{double} ME\_Len(\textit{MEdge\_ptr} e):} Return
-  the length of the straight line connecting the two vertices of the
-  edge.
-
-\item[]\textbf{\textit{double} ME\_LenSqr(\textit{MEdge\_ptr} e):}
-  Return the square of the length of the straight line connecting the
-  two vertices of the edge.
-
-\item[]\textbf{\textit{void} ME\_Vec(\textit{MEdge\_ptr} e, double
-    *evec):} Return the vector going from the first vertex of the edge
-  to the second vertex of the edge.
-\end{description}
-
-\newpage
-\subsection{Mesh Face Object}
-
-\begin{description}
-\item[]\textbf{\textit{MFace\_ptr} MF\_New(\textit{Mesh\_ptr} mesh):} Create
-a new face object. No topological information is embedded in the face
-when it is created. The face only knows which mesh it belongs
-to. The ID of the face is set by this function.
-
-\item[]\textbf{\textit{void} MF\_Delete(\textit{MFace\_ptr} mface):}
-  Delete the face. If \textbf{keep} is 0, the face is removed from the
-  mesh and all topological and geometric information embedded in the
-  face is destroyed. If \textbf{keep} is 1, the face is marked as type
-  \textit{MDELFACE} and removed from the mesh but the face is not
-  destroyed. Also, the vertices/edges of this face no longer point up
-  to this face.
-
-\item[]\textbf{\textit{void} MF\_Restore(\textit{MFace\_ptr}
-    mface):} Restore a temporarily deleted face. The face type is
-  restored from \textit{MDELFACE} to \textit{MFACE} and the face is
-  added back to the mesh. The vertices/edges of the face once again point
-  back to the face.
-
-\item[]
-
-\item[]\textbf{\textit{void} MF\_Set\_GEntity(\textit{MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the face is classified.
-
-\item[]\textbf{\textit{void} MF\_Set\_GEntDim(\textit{MFace\_ptr} mface,
-\textit{int} gdim):} Set the dimension of the geometric model entity on
-which the face is classified.
-
-\item[]\textbf{\textit{void} MF\_Set\_GEntID(\textit{MFace\_ptr} mface,
-\textit{int} gid):} Set the ID of the geometric model entity on which
-the face is classified.
-
-\item[]\textbf{\textit{void} MF\_Set\_ID(\textit{MFace\_ptr} mface,
-\textit{int} id):} Explicitly set ID of a face and overwrite the ID
-set by the \textbf{MF\_New} operator. Does not check for duplication of face
-IDs.
-
-\item[]\textbf{\textit{void} MF\_Set\_Edges(\textit{MFace\_ptr} mface,
-\textit{int} n, \textit{MEdge\_ptr} *edges, \textit{int} *dirs):} Set
-the edges of the face along with their directions. The ordered set of
-edge pointers and their directions are passed in through arrays along
-with the number of edges. The edges are assumed to be ordered
-clockwise around the face. If an edge direction is along the clockwise
-direction of the face then the entry in the 'dirs' array must be 1;
-otherwise it must be 0. This function is relevant only for full
-representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Set\_Vertices(\textit{MFace\_ptr}
-    mface, \textit{int} n, \textit{MVertex\_ptr} *verts):} Set the
-  vertices of the face. The ordered set of vertices (ccw around the
-  face) is passed in through an array along with the number of
-  vertices. This routine will collect/build all lower order
-  topological information (edges, edge directions in the face) as needed by the face.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Edge(\textit{MFace\_ptr}
-    mface, \textit{MEdge\_ptr} edge, \textit{MEdge\_ptr} nuedge,
-    \textit{int} dir):} \textit{THIS FUNCTION HAS BEEN SUPERSEDED BY
-    \textbf{MF\_Replace\_Edges}}.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Edge\_i(\textit{MFace\_ptr}
-    mface, \textit{int} i, \textit{MEdge\_ptr} nuedge, \textit{int}
-    dir):} \textit{THIS FUNCTION HAS BEEN SUPERSEDED BY
-    \textbf{MF\_Replace\_Edges\_i}}.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Edges(\textit{MFace\_ptr}
-    mface, \textit{int} nold, \textit{MEdge\_ptr} *oldedges,
-    \textit{int} nnu, \textit{MEdge\_ptr} *nuedges):} Replace a set of
-  edges in the face with another set of edges. The direction in which
-  the new edges are to be used in the face are automatically deduced.
-  This function is relevant only for full representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Edges\_i(\textit{MFace\_ptr}
-    mface, \textit{int} nold, \textit{int} i, \textit{int} nnu,
-    \textit{MEdge\_ptr} *nuedges):} Replace the \textit{nold} edges in
-  the face starting with the i'th edge and going ccw around the face
-  with a new set of edges. The directions in which the new edges are
-  used in the face are automatically deduced. This function is
-  relevant only for full representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Vertex(\textit{MFace\_ptr} mface,
-\textit{MVertex\_ptr} mvertex, \textit{MVertex\_ptr} nuvertex):}
-Replace a vertex in the face with another vertex. This function is
-relevant only for reduced representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
-mface, \textit{int} i, \textit{MVertex\_ptr} nuvertex):} Replace the
-i'th vertex in the face with a new vertex. This function is relevant
-only for reduced representations in MSTK.
-
-\item[]\textbf{\textit{void} MF\_Insert\_Vertex(\textit{MFace\_ptr} mface,
-\textit{MVertex\_ptr} nuvertex, \textit{MVertex\_ptr} b4vertex):}
-Insert a vertex in the face before \textbf{b4vertex} w.r.t. a ccw
-ordering of the vertex faces.
-
-\item[]\textbf{\textit{void} MF\_Replace\_Vertex\_i(\textit{MFace\_ptr}
-mface, \textit{MVertex\_ptr} nuvertex, \textit{int} i):} Insert a new
-vertex before vertex i in the face w.r.t. a ccw ordering of vertices.
-
-\item[]
-
-\item[]\textbf{\textit{Mesh\_ptr} MF\_Mesh(\textit{MFace\_ptr} mf):} Returns the mesh that this mesh belongs to.
-
-\item[]\textbf{\textit{int} MF\_ID(\textit{MFace\_ptr} mface):} Returns the ID of the face. Returns 0 if not known.
-
-\item[]\textbf{\textit{int} MF\_GEntDim(\textit{MFace\_ptr} mface):} Returns the dimensions of the geometric model entity that the vertex is classified on. Returns -1 if not known.
-
-\item[]\textbf{\textit{int} MF\_GEntID(\textit{MFace\_ptr} mface):} Returns the ID of the geometric model entity that the vertex is classified on. Returns 0 if this information is not known.
-
-\item[]\textbf{\textit{GEntity\_ptr} MF\_GEntity(\textit{MFace\_ptr} mface):} Returns a pointer or handle to the geometric model entity that the vertex is classified on.
-Returns NULL if this information is not known.
-
-\item[]
-
-\item[]\textbf{\textit{int} MF\_Num\_Vertices(\textit{MFace\_ptr} mface):} Returns the number of vertices of the face.
-
-\item[]\textbf{\textit{int} MF\_Num\_Edges(\textit{MFace\_ptr} mface):} Returns the number of edges of the face.
-
-\item[]\textbf{\textit{int} MF\_Num\_AdjFaces(\textit{MFace\_ptr}
-    mface):} Returns the number of adjacent faces of a face. This
-  operator is relevant only in planar or surface meshes, i.e., for
-  boundary faces not connected to any regions.
-
-\item[]\textbf{\textit{List\_ptr} MF\_Vertices(\textit{MFace\_ptr}
-    mface, \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
-  ordered set of the vertices of the face. The vertices are ordered in
-  ccw direction while looking down the face 'normal', if 'dir' is 1
-  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
-  vertex set is reordered so that it is the first vertex (This
-  argument will be added soon to the function. For now, omit this
-  argument). The behavior of this function can be illustrated using
-  Figure~\ref{fig:face_def}. For the face shown in the figure, a
-  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
-  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
-  A vertex set with ccw ordering starting with vertex $V_2$ is
-  ${V_2,V_3,V_0,V_1}$.
-
-\begin{figure}[!h]
-  \begin{center}
-  \includegraphics[scale=1.0]{figures/face_def} 
-  \caption{Face definition}
-  \label{fig:face_def}
-  \end{center}
-\end{figure}
-
-\item[]\textbf{\textit{List\_ptr} MF\_Edges(\textit{MFace\_ptr} mface,
-    \textit{int} dir, \textit{MVertex\_ptr} mvert):} Return the
-  ordered set of edges of the face. The edges are ordered in the ccw
-  while looking down the face 'normal' if dir is 1 and in the cw if
-  dir is 0. If 'mvert' is specified, the edge set is reordered so that
-  the first edge in the set contains this vertex. More precisely, if
-  'dir' is 1, and the first edge is 'e' used in the face in direction
-  'd', then \textbf{ME\_Vertex(e,!d)} = mvert. With reference to
-  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
-  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
-  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
-  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
-  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
-  ${E_1,E_0,E_3,E_2}$.
-
-\item[]\textbf{\textit{List\_ptr} MF\_AdjFaces(\textit{MFace\_ptr} mface):}
-List of adacent faces of a face. This operator is relevant only in
-planar or surface surface meshes, i.e., for boundary faces not
-connected to any regions.
-
-\item[]\textbf{\textit{int} MF\_EdgeDir(\textit{MFace\_ptr} mface,
-\textit{MEdge\_ptr} medge):} Returns the direction in which the face
-uses the given edge. If the faces use the edge in the positive direction,
-the function returns 1; otherwise it returns 0.
-
-\item[]\textbf{\textit{int} MF\_EdgeDir\_i(\textit{MFace\_ptr} mface,
-\textit{int} i):} Returns the direction in which the face uses its i'th
-edge. If the face uses the edge in the positive direction the function
-returns 1; otherwise it returns 0;
-
-\item[]\textbf{\textit{List\_ptr} MF\_Regions(\textit{MFace\_ptr} mface):}
-Return the set of regions connected to the face. If the face is not
-used by any regions, the function returns NULL to indicate that the
-set is empty. If not the set may contain one or two regions.
-
-\item[]\textbf{\textit{MRegion\_ptr} MF\_Region(\textit{MFace\_ptr} mface,
-\textit{int} side):} Returns the region on the specified side of the
-face. The positive side of the face (side = 1) is the side towards
-which the face normal points. The negative side of the face (side = 0)
-is the opposite side.
-
-\item[]
-
-\item[]\textbf{\textit{int} MF\_UsesEntity(\textit{MFace\_ptr} mface,
-\textit{MEntity\_ptr} mentity, \textit{int} type):} Check if the face
-uses the given lower dimension entity, 'mentity'. The type of the
-entity is specified by the 'etype' variable. For a face, a lower
-dimensional entity can be a vertex or an edge. If the face uses the
-vertex or the edge, the function returns 1; otherwise it returns 0. If
-any other type of entity is specified, the function returns 0.
-
-\item[]\textbf{\textit{MFace\_ptr} MVs\_CommonFace(\textit{int} nv,
-    \textit{MVertex\_ptr} *fverts):} Return a face connected to all
-  the given mesh vertices. Returns NULL if no such face exists.
-
-\item[]\textbf{\textit{MFace\_ptr} MEs\_CommonFace(\textit{int} ne,
-    \textit{MEdge\_ptr} *fedges):} Return a face connected to all
-  the given mesh edges. Returns NULL if no such face exists.
-
-\item[]
-
-\item[]\textbf{\textit{void} MF\_Coords(\textit{MFace\_ptr} mface,
-\textit{int} *n, double (*xyz)[3], \textit{int} dir):} Returns the
-coordinates of the face vertices in an array along with the number of
-vertices. If 'dir' is 1, the coordinates are returned with a  ccw
-ordering while looking down the face normal; if 'dir' is 0, they are
-returned with a cw ordering (See Figure~\ref{fig:face_def}).
-\end{description}
-
-\newpage
-\subsection{Mesh Region Object}
-
-\begin{description}
-\item[]\textbf{\textit{MRegion\_ptr} MR\_New(\textit{Mesh\_ptr}
-    mesh):} Create a new region object. No topological information is
-  embedded in the region when it is created. The region only knows
-  which mesh it belongs to. The ID of the region is set by this
-  function.
-
-\item[]\textbf{\textit{void} MR\_Delete(\textit{MRegion\_ptr} mregion,
-    \textit{int} keep):} Delete the region. Deletes all topological
-  information embedded in the region. If \textbf{keep} is 0, the
-  region is removed from the mesh and all topological and geometric
-  information embedded in the region is destroyed. If \textbf{keep} is
-  1, the region is marked as type \textit{MDELREGION} and removed from
-  the mesh but the region is not destroyed. Also, the vertices, edges
-  and faces of this region no longer point up to this region.
-
-\item[]\textbf{\textit{void} MR\_Restore(\textit{MRegion\_ptr}
-    mregion):} Restore a temporarily deleted region. The region type
-  is restored from \textit{MDELREGION} to \textit{MREGION} and the
-  region is added back to the mesh. The vertices, edges and faces of
-  the region once again point back to the region.
-
-\item[]
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntity(\textit{MRegion\_ptr}
-    mregion, GEntity\_ptr gent):} Set the geometric model entity on
-  which the region is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntDim(\textit{MRegion\_ptr}
-    mregion, \textit{int} gdim):} Set the dimension of the geometric
-  model entity on which the region is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_GEntID(\textit{MRegion\_ptr}
-    mregion, \textit{int} gid):} Set the ID of the geometric model
-  entity on which the region is classified.
-
-\item[]\textbf{\textit{void} MR\_Set\_ID(\textit{MRegion\_ptr}
-    mregion, \textit{int} id):} Explicitly set ID of a region and
-  overwrite the ID set by the \textbf{MR\_New} function. Does not
-  check for duplication of region IDs.
-
-\item[]\textbf{\textit{void} MR\_Set\_Faces(\textit{MRegion\_ptr}
-    mregion, \textit{int} nf, \textit{MFace\_ptr} *mfaces,
-    \textit{int} *dirs):} Set the faces of the region along with their
-  directions. The \textit{unordered} set of faces and their directions
-  are passed in through arrays along with the number of faces. If the
-  normal of the face points out of the region, the associated
-  direction to be passed in is 1; otherwise it is 0. This function is
-  only relevant for full representations in MSTK.
-
-\item[]\textbf{\textit{void} MR\_Set\_Vertices(\textit{MRegion\_ptr}
-    mregion, \textit{int} nv, \textit{MVertex\_ptr} *mvertices):} Set
-  the vertices of the region. This routine will collect/build all
-  lower order topological information (edges, edge directions in the
-  face) as needed by the face. For standard elements, the vertices
-  must be ordered as indicated in Appendix~\ref{app:reg_conventions},
-  while \textbf{nf} can be 0 and \textbf{rfv\_template} can be NULL.
-  For non-standard elements, \textbf{nf} is the number of faces of the
-  polyhedron and \textbf{rfv\_template} gives the vertices used in each
-  face. More precisely, \textbf{rfv\_template} has \textbf{nfv}
-  pointers to integer array $i$. The first entry of each
-  \textbf{rfv\_template$[i]$} represents the number of vertices in that
-  face and the remaining entries represent the list of vertices of the
-  face listed so that the face defined by them points outward.
-
-\item[]\textbf{\textit{void} MR\_Replace\_Face(\textit{MRegion\_ptr}
-    mregion, \textit{MFace\_ptr} mface, \textit{MFace\_ptr} nuface,
-    \textit{int} dir):} Replace a face of the region with another
-  face. The direction in which the new face is used in the region must
-  also be supplied. This function is only relevant for full
-  representations in MSTK.
-
-\item[]\textbf{\textit{void} MR\_Replace\_Vertex(\textit{MRegion\_ptr}
-    mregion, \textit{MVertex\_ptr} mvertex, \textit{MVertex\_ptr}
-    nuvertex):} Replace a vertex of a region with another vertex. This
-  function is relevant only for reduced representations in MSTK.
-
-\item[]\textbf{\textit{void}
-    MR\_Replace\_Face\_i(\textit{MRegion\_ptr} mregion, \textit{int}
-    i, \textit{MFace\_ptr} mface, \textit{int} dir):} Replace the i'th
-  face in the region with another face.The direction in which the new
-  face is used in the region must also be supplied. This function is
-  only relevant for full representations in MSTK.
-
-\item[]\textbf{\textit{void}
-    MR\_Replace\_Vertex\_i(\textit{MRegion\_ptr} mregion, \textit{int}
-    i, \textit{MVertex\_ptr} mvertex):} Replace the i'th vertex of the
-  region with another vertex. This function is only relevant for
-  reduced representations in MSTK.
-
-\item[]
-
-\item[]\textbf{\textit{Mesh\_ptr} MR\_Mesh(\textit{MRegion\_ptr}
-    mregion):} Returns the mesh that the region belongs to.
-
-\item[]\textbf{\textit{int} MR\_ID(\textit{MRegion\_ptr} mregion):}
-  Returns the ID of the region. Returns 0 if not known.
-
-\item[]\textbf{\textit{int} MR\_GEntDim(\textit{MRegion\_ptr}
-    mregion):} Returns the dimension of the geometric model entity the
-  region is classified on. Always returns 3 since a mesh region can be
-  classified only on a model region.
-
-\item[]\textbf{\textit{int} MR\_GEntID(\textit{MRegion\_ptr}
-    mregion):} Returns the ID of the geometric model entity that the
-  region is classified on. Returns 0 if not known.
-
-\item[]\textbf{\textit{GEntity\_ptr} MR\_GEntity(\textit{MRegion\_ptr}
-    mregion):} Returns a pointer or handle to the geometric model
-  entity that the vertex is classified on. Returns NULL if this
-  information is not known.
-
-\item[]
-
-\item[]\textbf{\textit{int} MR\_Num\_Vertices(\textit{MRegion\_ptr}
-mregion):} Returns the number of vertices of a region.
-
-\item[]\textbf{\textit{int} MR\_Num\_Edges(\textit{MRegion\_ptr} mregion):}
-Returns the number of edges of a region.
-
-\item[]\textbf{\textit{int} MR\_Num\_Faces(\textit{MRegion\_ptr} mregion):}
-Returns the number of faces of a region.
-
-\item[]\textbf{\textit{int} MR\_Num\_AdjRegions(\textit{MRegion\_ptr}
-mregion):} Returns the number of adjacent regions of a region, i.e.,
-regions sharing a face with this region.
-
-\item[]
-
-\item[]\textbf{\textit{List\_ptr} MR\_Vertices(\textit{MRegion\_ptr}
-mregion):} Returns the set of vertices of a region. For standard
-elements the vertices are ordered as indicated in
-Appendix~\ref{app:reg_conventions}. For non-standard elements the set
-is unordered.
-
-\item[]\textbf{\textit{List\_ptr} MR\_Edges(\textit{MRegion\_ptr} mregion):}
-Return the \underline{unordered} set of edges of a region.
-
-\item[]\textbf{\textit{List\_ptr} MR\_Faces(\textit{MRegion\_ptr} mregion):}
-Returns the set of faces of a region.
-
-\item[]\textbf{\textit{List\_ptr} MR\_AdjRegions(\textit{MRegion\_ptr}
-mregion):} Returns the set of adjacent regions of a region, i.e.,
-regions sharing a face with this region. The set is not ordered.
-
-\item[]\textbf{\textit{int} MR\_FaceDir(\textit{MRegion\_ptr} mregion,
-\textit{MFace\_ptr} mface):} Returns the direction in which the region
-uses the given face. Returns 1 if the face normal points out of the
-region and returns 0 if the face normal points into the region.
-
-\item[]\textbf{\textit{int} MR\_FaceDir\_i(\textit{MRegion\_ptr} mregion,
-\textit{int} i):} Returns the direction in which the region uses the
-i'th face. Returns 1 if the face normal points out of the region and
-returns 0 if the face normal points into the region.
-
-\item[]
-
-\item[]\textbf{\textit{int} MR\_UsesEntity(\textit{MRegion\_ptr} mregion,
-\textit{MEntity\_ptr} ment, \textit{int} type):} Check if the region
-uses the given lower dimension entity, 'mentity'. The type of the
-entity is
-
-\item[]
-
-\item[]\textbf{\textit{void} MR\_Coords(\textit{MRegion\_ptr} mregion,
-\textit{int} *n, double (*xyz)[3]):} Returns the coordinates of the
-region vertices in an array along with the number of vertices. For
-standard elements, the ordering is as given in
-Appendix~\ref{app:reg_conventions}. For non-standard elements, the
-ordering is arbitrary.
-\end{description}
-
-\newpage
-\subsection{Generic Entity Object}
-
-The following functions operate on generic mesh entities of type
-\textit{MEntity\_ptr}. This implies that variables of type
-\textit{MVertex\_ptr}, \textit{MEdge\_ptr}, \textit{MFace\_ptr},
-\textit{MRegion\_ptr} can all be passed in place of
-\textit{MEntity\_ptr} variables in the following functions.
-
-\begin{description}
-%%  \textit{void} MEnt\_Set\_GEntity(\textit{MEntity\_ptr} mentity, GEntity\_ptr gent);
-%%
-%%  \textit{void} MEnt\_Set\_GEntDim(\textit{MEntity\_ptr} mentity, \textit{int} gdim);
-%%
-%%  \textit{void} MEnt\_Set\_ID(\textit{MEntity\_ptr} mentity, \textit{int} id);
-
-\item[]\textbf{\textit{int} MEnt\_ID(\textit{MEntity\_ptr} mentity):} Returns
-the ID of a generic entity.
-
-\item[]\textbf{\textit{int} MEnt\_Dim(\textit{MEntity\_ptr} mentity):}
-Returns the topological dimension or type of generic entity.
-
-\item[]\textbf{\textit{int} MEnt\_OrigDim(\textit{MEntity\_ptr} mentity):}
-Returns the original topological dimension or type of generic entity before it
-was temporarily deleted.
-
-\item[]\textbf{\textit{Mesh\_ptr} MEnt\_Mesh(\textit{MEntity\_ptr} mentity):}
-Returns the mesh that the entity belongs to.
-
-\item[]\textbf{\textit{int} MEnt\_GEntDim(\textit{MEntity\_ptr}
-    mentity):} Returns the dimension of the geometric model entity
-  that the entity is classified on.
-
-\item[]\textbf{\textit{GEntity\_ptr}
-    MEnt\_GEntity(\textit{MEntity\_ptr} mentity):} Returns a pointer
-  or handle to geometric model entity that the entity is classified
-  on.
-
-\item[]
-
-\item[]\textbf{\textit{void} MEnt\_Set\_AttVal(\textit{MEntity\_ptr}
-    ent, \textit{MAttrib\_ptr} attrib, \textit{int} ival,
-    \textit{double} lval, \textit{void} *pval):} Set the value of an
-  attribute for given mesh entity. Depending on whether the attribute
-  is integer, double or pointer, the appropriate entry in the argument
-  list is specified.
-
-\item[]\textbf{\textit{void} MEnt\_Rem\_AttVal(\textit{MEntity\_ptr}
-    ent, \textit{MAttrib\_ptr mattrib}:} Clear the value of the given
-  attribute from the entity.
-
-\item[]\textbf{\textit{int} MEnt\_Get\_AttVal(\textit{MEntity\_ptr}
-    ent, \textit{MAttrib\_ptr} int *ival, double *rval, double
-    **pval):} Get the value of an attribute from an entity.  Depending
-  on the attribute type, either \textbf{ival}, \textbf{rval} or
-  \textbf{pval} is returned.
-
-\end{description}
-
-\newpage
-\subsection{Mesh Attributes}
-
-\begin{description}
-
-\item[]\textbf{\textit{MAttrib\_ptr} MAttrib\_New(\textit{Mesh\_ptr}
-    mesh, \textit{const char} *att\_name, \textit{MAttType} att\_type,
-    \textit{MType} entdim):} Define a new mesh attribute. Along with
-  the mesh to which this attribute is assigned (\textbf{mesh}) and the
-  name of the attribute (\textbf{att\_name}), the type of the
-  attribute (\textbf{att\_type}) must be specified as \textbf{INT},
-  \textbf{DOUBLE} or \textbf{POINTER}. Also, the dimension of the entity
-    for which the attribute value can be set must be specified as
-    \textbf{MVERTEX}, \textbf{MEDGE}, \textbf{MFACE}, \textbf{MREGION}
-    or \textbf{MALLTYPE}.
-
-\item[]\textbf{\textit{char} *MAttrib\_Get\_Name(\textit{MAttrib\_ptr}
-    attrib, \textit{char} *att\_name):} Get the name of the given attribute.
-
-\item[]\textbf{\textit{MAttType}
-    MAttrib\_Get\_Type(\textit{MAttrib\_ptr} attrib):} Get the type of
-  the attribute (Can return \textbf{INT}, \textbf{DOUBLE},
-  \textbf{PVAL}).
-
-\item[]\textbf{\textit{MType}
-    MAttrib\_Get\_EntDim(\textit{MAttrib\_ptr} attrib):} Get the
-  dimension (or type) of entity attribute can be asssigned to. Can be
-  \textbf{MVERTEX}, \textbf{MEDGE}, \textbf{MFACE}, \textbf{MREGION}
-  or \textbf{MALLTYPE}.
-
-\item[]\textbf{\textit{void} MAttrib\_Delete(\textit{MAttrib\_ptr}
-    attrib):} Delete an attribute.
-
-\end{description}
-
-\newpage
-\subsection{Entity Marks}
-
-Entity marks or markers are a way of tagging entities. Such
-functionality is useful in algorithms which must keep track of
-processed entities to avoid duplication of work. An example of such an
-operation is creating a union of entity sets while extracting upward
-adjacency information such as the regions connected to an edge.
-
-\begin{description}
-\item[]\textbf{\textit{int} MSTK\_GetMarker():} Returns a unique marker ID
-which may be used to tag entities.
-
-\item[]\textbf{\textit{void} MEnt\_Mark(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Mark an entity with the given marker 'mkr'.
-
-\item[]\textbf{\textit{int} MEnt\_IsMarked(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Check if an entity is marked with the given marker 'mkr'.
-
-\item[]\textbf{\textit{void} MEnt\_Unmark(\textit{MEntity\_ptr} ent, \textit{int} mkr):} Unmark an entity with respect to the given marker 'mkr'
-
-\item[]\textbf{\textit{void} List\_Mark(\textit{List\_ptr} list, \textit{int} mkr):} Mark a set of entities with given marker.
-
-\item[]\textbf{\textit{void} List\_Unmark(\textit{List\_ptr} list, \textit{int} mkr):}
-Unmark a set of entities with respect to the given marker.
-
-\item[]\textbf{\textit{void} MSTK\_FreeMarker(\textit{int} mkr):} Release the
-marker ID given by \textbf{MSTK\_GetMarker()} so that it can be
-reused. Care must be taken to unmark all entities marked with this
-marker ID before releasing it. If not, subsequent operations with
-reassigned marker will find a tag on some entities and mistake them
-for being processed.
-\end{description}
-
-\newpage
-\subsection{Mesh Modification}
-
-\begin{description}
-
-\item[]\textbf{\textit{int} ME\_Swap2D(\textit{MEdge\_ptr} e,
-\textit{MEdge\_ptr} *enew, \textit{MFace\_ptr} fnew[2]):} Swap an edge
-in a triangular mesh. No checks are performed for topological or
-geometric validity.
-
-\item[]\textbf{\textit{MFace\_ptr} MFs\_Join(\textit{MFace\_ptr} f1,
-\textit{MFace\_ptr} f2, \textit{MEdge\_ptr} e):} Join two faces along
-common edge and create new face by eliminating the common edge as
-shown in Figure~\ref{fig:MFs_Join}. If 'f1' has 'n1' edges and 'f2'
-has 'n2' edges, then the new face has ('n1'+'n2'-2) edges.
-
-\end{description}
-
-\begin{figure}[!h]
-  \begin{center}
-    \includegraphics[scale=1.0]{figures/MFs_Join}
-
-    \caption{Joining two faces (a) Two faces $F_1$ and $F_2$ sharing a
-      common edge (b) New pentagonal face $F_{new}$ created by
-      eliminating the common edge.  }
-    \label{fig:MFs_Join}
-  \end{center}
-\end{figure}
-
-\newpage
-\subsection{Utilities}
-
-\begin{description}
-
-\item[]\textbf{\textit{void} MSTK\_Report(\textit{char} *module,
-  \textit{char} *message, \textit{ErrType} severity):} Error handler
-  for MSTK. 'module' is the name of the function in which the error
-  occurs. 'message' is the error message and is recommended to be less
-  than 1024 characters in length. 'severity' is an error code and can
-  be MESSG, WARN, ERROR or FATAL. If the error code is FATAL, the
-  program will quit after printing the error. If the same message is
-  repeated successively, then the message is printed only the first
-  time.
-
-\item[]\textbf{\textit{void} List\_PrintID(\textit{List\_ptr} l):} Debugging
-utility to print the IDs of the entities in a set.
-
-\item[]\textbf{\textit{void} MV\_Print(\textit{MVertex\_ptr} v, int
-    lev):} Debugging utility to print information about a mesh vertex,
-  \textbf{v}. The argument \textbf{lev} controls the level of detail
-  of the information printed. \textbf{lev} = 0 prints the minimum
-  information, i.e., vertex pointer, its ID and its coordinates. If
-  \textbf{lev} = 1, the function prints classification information for
-  the vertex (if available), i.e., ID and dimension of the model
-  entity that the vertex is on. If \textbf{lev} $>$ 1, then upward
-  detailed adjacency information is also printed for the vertex, i.e.,
-  information is printed about the edges, faces and regions connected
-  to the vertex.
-
-\item[]\textbf{\textit{void} ME\_Print(\textit{MEdge\_ptr} e, int
-    lev):} Debugging utility to print information about a mesh edge,
-  \textbf{e}. The argument \textbf{lev} controls the level of detail
-  of the information printed. \textbf{lev} = 0 prints the minimum
-  information, i.e., edge pointer, its ID and the IDs of its two
-  vertices. If \textbf{lev} = 1, the function prints classification
-  information for the edge (if available), i.e., ID and dimension of
-  the model entity that the edge is on. Also, more detailed vertex
-  information printed in this case. If \textbf{lev} $>$ 1, the
-  function prints detailed upward adjacency information for the edge,
-  i.e., information is printed about the faces and regions connected
-  to the edge.
-
-  \item[]\textbf{\textit{void} MF\_Print(\textit{MFace\_ptr} f, int
-      lev):} Debugging utility to print information about a mesh face,
-    \textbf{f}. The argument \textbf{lev} controls the level of detail
-    of the information printed. \textbf{lev} = 0 prints the minimum
-    information, i.e., the face pointer and its ID. If \textbf{lev} =
-    1, the function prints classification information for the edge (if
-    available), i.e., ID and dimension of the model entity that the
-    face is on. Also, a signed list of the edges of the face is
-    printed.  If \textbf{lev} $>$ 1, the function prints detailed
-    downward and upward adjacency information for the face, i.e.,
-    information is printed about the edges and vertices of the face,
-    and about the regions connected to the face.
-
-  \item[]\textbf{\textit{void} MR\_Print(\textit{MRegion\_ptr} r, int
-      lev):} Debugging utility to print information about a mesh
-    region, \textbf{r}. The argument \textbf{lev} controls the level
-    of detail of the information printed. \textbf{lev} = 0 prints the
-    minimum information, i.e., region pointer and its ID. If
-    \textbf{lev} = 1, the function prints classification information
-    for the region (if available), i.e., ID of the model entity that
-    the region is on. Also, a signed list of the faces of the region
-    is printed. If \textbf{lev} $>$ 1, the function prints detailed
-    downward adjacency information for the region, i.e., information
-    is printed about the faces, edges and vertices forming the region.
-
-\end{description}
-
-\newpage
-\appendix
-
-\section{Mesh Representation Types  in MSTK}
-\label{app:reptypes}
+\subsection{Why should I use MSTK?}
+
+MSTK offers a flexible infrastructure for representing and
+manipulating meshes so that mesh-based application developers do not
+have to deal with the complexitiies of managing their own data
+structures. The functional interface of MSTK is designed to provide
+easy access to mesh data and allow for easy-to-read but efficient code
+to be written for mesh based applications. Accessing the mesh data
+through the functional interface allows the high level application to
+be unchanged even if the lower level data structures in MSTK
+change. MSTK supports a wide array of element types and several
+representations, all of which can take a new developer years to write
+and make robust. New users of MSTK can typically start writing their
+own codes to query and manipulate meshes within a few days of
+familiarizing themselves with MSTK. Finally, use of MSTK allows easy
+integration with other codes using MSTK as their mesh data framework.
+
+\section{MSTK Concepts}
+\subsection{Unstructured mesh representation}
+
+Meshes are made up of topological entities of different dimensions. In
+a ``traditional'' 3D finite element mesh, nodes are topological
+entities called vertices and are topologically 0-dimensional entities,
+while elements are topological entities called regions and are
+topologically 3-dimensional entities. In general, meshes can be
+described using vertices (0-dimensional entities), edges
+(1-dimensional entities), faces (2-dimensional entities) and regions
+(3-dimensional entities). MSTK has multiple ways of representing
+meshes as shown in the below. For example, representation F1 has
+entities of all dimensions up to the dimension of the mesh while
+representation R1 has entities of the lowest and highest dimension
+only. Currently, the application has to choose a particular
+representation type when a mesh is created although it may choose
+different representations for different meshes in the same program. In
+the future, the code will be allowed to switch between different
+representations depending on the needs of the particular algorithm
+being executed at that time.
+
+{\bf NOTE: CURRENTLY THE F1 REPRESENTATION IS THE MOST ROBUST
+  IMPLEMENTATION SINCE IT IS MOST COMMONLY USED. THIS REPRESENTATION
+  HAS GENERALLY BEEN STABLE FOR SEVERAL YEARS. THE REDUCED
+  REPRESENTATION HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT IS CURRENTLY
+  BROKEN. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
+  PLEASE REPORT ANY PROBLEMS IMMEDIATELY TO \texttt{rao@lanl.gov}.  }
+
+Even when using reduced representations, MSTK can provide the full set
+of topological adjacencies as if a full representation were being
+used. When necessary, entities not represented explicitly in the mesh
+are created on-the-fly and represented as volatile entities. Even
+though it is possible to take an algorithm that is written for a full
+representation and use it as is with a reduced representation,
+application developers are urged to be aware of the costs of each
+operation for different representations and use this information to
+design algorithms optimized for each representation.
 
@@
 
+\subsection{Mesh entity classification}
+
+The concept of mesh entities of various dimensions is actually derived
+from field of B-rep geometric modeling in which a model is comprised
+of a hierarchy of topological entities of different dimensions. {\em
+  Classification} is the relationship of each mesh entity to the model
+entity is represents the whole or a part of.
+
+To clarify further, every mesh is a discrete representation of a
+geometric model. This geometric model may exist in the analyst's mind,
+as schematic on paper or as a full model in a geometric modeling
+system. Encoding the relationship of the mesh to this geometric model
+to the extent possible provides enormous algorithmic benefits to the
+users of meshes.
+
+Recognizing that one may have varying degrees of information about the
+geometric model in different situations, MSTK allows mesh entities to
+store different levels of classification information or no
+classification information.
+
+Every mesh entity can store the dimension of the model entity it is on
+({\bf GEntDim}), the ID of the model entity it is on ({\bf GEntID})
+and a pointer to the model entity it is on ({\bf GEntity}). The
+following picture illustrates the classification of various entities
+in a simple mesh.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/classfn}
+  \end{center}
+  \caption{Example mesh and model showing the classification of mesh entities on geometric model entities}
+\end{figure}
+
+Classification information is very useful to have for simplifying
+algorithms in meshing and analysis applications. For example, using
+classification information, one can trivially extract all nodes on the
+boundary and apply a different smoothing algorithm than nodes in the
+interior of the domain. An FEA code can apply boundary conditions on
+nodes and faces more easily if it has classification information.
+
+If the application developer has nothing but the material IDs of the
+mesh elements, MSTK can build the classification information to the
+best of its ability using topological and geometric
+information. However, it must be noted that this is procedure is not
+perfect and some classification information cannot be derived
+unambiguously. So, wherever possible, users are encouraged to generate
+classification information during mesh generation and supply it to
+MSTK.
+
+\subsection{MSTK Application Programming Interface (API)}
+
+Even though MSTK is written in C, MSTK is designed in an object
+oriented manner. Mesh entities, mesh attributes and entity lists are
+considered to be objects which have their private data and a set of
+operators to query and manipulate this data. Even though methods may
+be devised to circumvent the data hiding in MSTK, this is highly
+discouraged. If you think that a particular operator is not efficient
+or you don'thave access to some mesh data, please contact the
+developer for assistance.
+
+An important feature of MSTK is its ability to provide the full set of
+mesh operators regardless of the representation used. When using
+reduced representations, entities not represented explicitly in the
+mesh are returned as volatile objects. However, it is possible to use
+most MSTK operators on such objects as well.
+
+Currently, MSTK offers only a C API but C++ and Fortran APIs are
+planned for the near future.
+
+\subsection{Getting meshes in and out of MSTK}
+
+There are a few ways of getting mesh data into MSTK. The first is to
+prepare an MSTK file with the full connectivity of an F1
+representation and to read it in using MESH\_InitFromFile. This is
+typically a difficult task for many applications which only have
+element-node data. In such cases, one can prepare an MSTK file in the
+R1 file format (in which the node listing is followed by the
+region-node connectivity) and read it in to any of the other
+representations, say F1. The other possibility is to write a file in
+the GMV format and use MESH\_ImportFromFile to import the mesh in.
+
+MSTK can write meshes out to different formats including GMV, STL and
+FLAG X3D. When compiled with the appropriate options and using a
+partitioner from an external library, MSTK can also write out a
+parallel X3D file.
+
 \newpage
+\section{MSTK Data Types}
+
+\begin{description}
+\item[]{\bf {\em List\_ptr}}: Handle to a List object.
+
+\item[]{\bf {\em Mesh\_ptr}}: Handle to a Mesh object.
+
+\item[]{\bf {\em MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
+Dimension 0).
+
+\item[]{\bf {\em MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1).
+
+\item[]{\bf {\em MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2).
+
+\item[]{\bf {\em MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3).
+
+\item[]{\bf {\em MEntity\_ptr}}: Handle to a generic Mesh Entity
+  object. Any of the above types of entities can be cast as
+  {\em MEntity\_ptr}.
+
+\item[]{\bf {\em GModel\_ptr}}: Handle to a Geometric Model object.
+\item[]{\bf {\em GEntity\_ptr}}: Handle to a Geometric Entity object.
+
+\item[]{\bf {\em MAttrib\_ptr}}: Handle to a mesh attribute.
+
+\item[]{\bf {\em RepType}}: Enumerated type describing the type
+  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
+  See Appendix~\ref{app:reptypes} for schematics of these
+  representations. {\em Currently only representation types F1 and
+    F4 are supported}.
+
+\item[]{\bf {\em MType}}: Enumerated type for mesh entity type. Can be MDELETED, MVERTEX, MEDGE, MFACE, MREGION, MUNKNOWNTYPE, MALLTYPE.
+
+\item[]{\bf {\em MFType}}: Enumerated type for mesh face type.
+Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON.
+
+\item[]{\bf {\em MRType}}: Enumerated type for mesh region type.
+Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED.
+
+\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
+\end{description}
+
+\newpage
+\section{MSTK Functional Interface}
+\subsection{List}
+
+Unordered sets or lists of entities in MSTK are returned as type
+{\em List\_ptr}. The following are the set operations available in
+MSTK:
+
+\begin{description}
+\item[]{\bf {\em List\_ptr} List\_New({\em int} inisize):} Create a
+new list with an initial size, {\em inisize}. If {\em inisize} is
+0, the initial size is set to be 10.
+
+\item[]{\bf {\em void} List\_Delete({\em List\_ptr} l):} Delete a list.
+
+\item[]{\bf {\em List\_ptr} List\_Compress({\em List\_ptr}
+    l):} Compress a list (delete all the null entries created when
+  entities are removed from the list).  Doing this while an algorithm
+  is iterating through the set can currently cause problems!! Calling
+  {\bf List\_Compress} could change the pointer for the list due to
+  reallocation.
+
+\item[]{\bf {\em List\_ptr} List\_Copy({\em List\_ptr} l):} Return a
+copy of a list.
+
+\item[]{\bf {\em List\_ptr} List\_Add({\em List\_ptr} l, {\em void
+  *}entry):} Add an entry to the list. The entry is strictly appended to the end
+of the list.
+
+\item[]{\bf {\em List\_ptr} List\_ChknAdd({\em List\_ptr} l,
+{\em void *}entry):} Add an entry to a list only if it is not already
+in the list.
+
+\item[]{\bf {\em List\_ptr} List\_Insert({\em List\_ptr} l,
+{\em void *}nuentry, {\em void *}b4entry):} Insert an entry into the list in the position before 'b4entry'.
+
+\item[]{\bf {\em List\_ptr} List\_Inserti({\em List\_ptr} l,
+    {\em void *}nuentry, {\em int} i):} Insert an entry into the
+  list at the i'th valid position and push the entries previously at
+  the i'th and later position back.
+
+\item[]{\bf {\em int} List\_Rem({\em List\_ptr} l, {\em void
+  *}entry):} Remove an entry from the list. Returns 1 if successful, 0
+otherwise.
+
+\item[]{\bf {\em int} List\_Remi({\em List\_ptr} l, {\em int} i):}
+Remove the i'th valid entry in the list. Returns 1 if successful, 0
+otherwise.
+
+\item[]{\bf {\em int} List\_RemSorted({\em List\_ptr} l, {\em void}
+    *entry, int (*entry2int)(void *)):} Remove an entry from a list
+  that is sorted according to some correspondence between an entry and
+  an integer value, e.g., {\bf MEnt\_ID(entry)} gives the ID of an
+  entity. The mapping between the entry and the integer is given by a
+  call to the routine {\bf entry2int}. The list could be one from which
+  entities have previously been removed. The routine makes use of the
+  sorted nature of the list to locate entries in $O(log(n))$ time
+  where $n$ is the size of the list. An example call to this routine
+  is
+
+  \par {\bf fnd = List\_RemSorted(mregionslist,region,\&(MR\_ID));}
+
+\item[]{\bf {\em int} List\_Replace({\em List\_ptr} l, {\em void
+  *}entry, {\em void *}nuentry):} Replace 'entry' with 'nuentry' in
+set. Returns 1 if successful, 0 otherwise.
+
+\item[]{\bf {\em int} List\_Replacei({\em List\_ptr} l, {\em int}
+i, void *nuentry):} Replace the i'th valid entry in the list with
+'nuentry'. Returns 1 if successful, 0 otherwise.
+
+\item[]{\bf {\em int} List\_Contains({\em List\_ptr} l, void *entry):}
+Returns 1 if list contains the entry, 0 otherwise.
+
+\item[]{\bf {\em int} List\_Locate({\em List\_ptr} l, void *entry):}
+Returns the positional index of the entry in the list. Returns -1 if
+the list does not contain the entry.
+
+\item[]{\bf {\em void *}List\_Entry({\em List\_ptr} l, {\em int}
+i):} Return the i'th valid entry in the list.  Returns a NULL pointer if
+the i'th valid entry could not be found.
+
+\item[]{\bf {\em void *}List\_Next\_Entry({\em List\_ptr} l,
+{\em int} *i):} Return the next valid entry in the list. This routine
+works like an iterator. To start iterating through the list, set the
+iteration index i=0 and call the routine to get the first entry in the
+set. Subsequent calls to the routine will iterate through the entries
+in the list. The routine will return a NULL to indicate that the end of
+the set is reached.
+
+The value of the iteration index i will be modified by the routine on
+each call to indicate where in the list it is. This value should not be
+modified externally while iterating through the list. Also, no specific
+meaning should be derived from the iteration index by other applications
+since the internal implementation and interpretation of the index may
+change at any time.
+
+Finally, there may be unexpected consequences if entries are removed
+from the list while an iterator is iterating through it using {\bf
+  List\_Next\_Entry}.
+
+\item[]{\bf {\em int} List\_Num\_Entries({\em List\_ptr} l):} Return
+the number of entries in a list.
+\end{description}
+
+\newpage
+\subsection{Mesh Object}
+
+A mesh object is a set of vertices (nodes) possibly connected by other
+entities such as edges, faces, regions. Depending on the
+representation chosen and type of mesh, some or all of the entities
+may be explicitly stored.  Full representations contain all types of
+entities up to the highest dimension of the mesh. For example, a full
+representation of a tetrahedral mesh contains vertices, edges, faces
+and regions. However, one type of reduced representation of this mesh
+may contain only vertices and regions. For a surface mesh, a full
+representation includes vertices, edges and faces while a reduced
+representation only has vertices and faces. Also, depending on the
+type of representation, some adjacencies (information about which
+entities are connected to which other entities) are stored and others
+are derived.
+
+\begin{description}
+\item[]{\bf {\em Mesh\_ptr} MESH\_New({\em RepType} type):}
+  Initialize a new mesh object with the given representation type
+  which can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Of these types
+  F1, F4, R1, R2 and R4 are implemented. If the representation type is
+  not known at the present time (e.g. before reading the mesh from a
+  file), the representation type of UNKNOWN\_REP can be specified.
+  Note that this only initializes a mesh object, it does not create or
+  generate a mesh which is the work of high level mesh generation
+  routines.
+
+\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr}
+    mesh, const char *filename):} Initialize or read a mesh from file
+  into the given mesh object. Returns 1 if successful, 0 otherwise. It
+  is possible to have a an MSTK file in the R1, R2 and R4 format into
+  a mesh initialized as type F1 or F4.
+
+\item[]{\bf {\em int} MESH\_InitFromGenDesc({\em Mesh\_ptr} mesh, {\em
+      int} nv, {\em double} (*xyz)[3], {\em int} *nfv, {\em int}
+    **fvids, {\em int} nr, {\em int} *nrv, {\em int} **rvids, {\em
+      int} *nrf, {\em int} ***rfvtemplate):} Initialize a mesh from a
+  minimal description of the mesh passed into the routine. In the
+  routine, {\bf nv} is the number of vertices or nodes and {\bf xyz}
+  is the array of node coordinates. If the mesh is a surface mesh,
+  {\bf nf} is the number of mesh faces, {\bf nfv} gives the number of
+  vertices for each face, {\bf fvids} gives the array indices of the
+  face vertices in ccw manner (starting from 0, not 1). If the mesh is
+  a solid mesh, {\bf nr} specifies the number of solid elements or
+  regions. If the mesh has only standard element types (tets,
+  pyramids, triangular prisms and hexes), then {\bf nrv} indicates the
+  number of vertices of each region and {\bf rvids} gives the array
+  indices of the region vertices. If, on the other hand, the mesh has
+  polyhedral elements, regions have to be described in terms of faces
+  which are in turn described in terms of vertices. Therefore, {\bf
+    nrf} indicates the number of faces for each region and {\bf
+    rfvtemplate} gives the array indices of the vertices for each face
+  of the region.
+
+\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr}
+    mesh, const char *filename, const char *format):} Import a mesh
+  data from an external file format and construct MSTK mesh. Currently
+  the only format supported is the
+  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
+  file format. The routine imports as many attributes as it can from
+  the input file. It also uses special attributes or keywords as data
+  about element and node classification.  For GMV files, the routine
+  uses the ``material'' data to indicate region or face classification
+  depending on the dimensionality of the mesh. If ``material'' data is
+  not specified, it uses the ``itetclr'' attribute to assign region or
+  face classification. It also uses the ``icr'' keyword describing the
+  number of constraints on a node to interpret if the node is
+  classified on a model region, model face, model edge or a model
+  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
+  meshes generated by
+  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.
+
+\item[]{\bf {\em int} MESH\_BuildClassfn({\em Mesh\_ptr} mesh):} Build
+  classification information for mesh entities if only partial
+  information is present. In other words, if the only data known is
+  the IDs of geometric model entities (``material regions'') of the
+  highest level mesh entities (faces or regions), then this procedure
+  will build information about the type of geometric model entities
+  that the lower dimension entities are on. Faces will be classified
+  as being on a model face (external or interior) or inside model
+  region. Edges will be classified as being on a model edge, model
+  face or in a model region. Vertices will be classified as being on a
+  model vertex, model edge, model face or in a model region. If no
+  classification data is associated with even the highest level
+  entities, the procedure will assume that all the highest level
+  entities are classified on one model entitiy of that dimension. If
+  partial information is available for lower order entities, this
+  routine will not destroy that information. The procedure also tries
+  to detect mesh edges that should be classified on model edges based
+  on the fact that the dihedral angle between the boundary faces
+  connected to the edge is smaller than some tolerance.
+
+\item[]{\bf {\em int} MESH\_DelInterior({\em Mesh\_ptr}
+    mesh):} Delete the interior of a mesh and retain only its
+    boundaries (including interior boundaries). For solid meshes, this
+    results in a surface mesh (i.e., all mesh regions, and  mesh
+    faces, edges and vertices classified on model regions are deleted). For
+    surface meshes, this results in curve mesh. For an edge mesh, only
+    the end vertices are retained (unlikely to be used this
+    way). Undefined for a vertex mesh.
+
+\item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr}
+    mesh, const char *filename):} Save a mesh to a filename. The file
+  is created if it does not exists. It is recommended that the .mstk
+  extension be used for MSTK mesh files.  However, there is no such
+  requirement.
+
+%% \item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr}
+%%     mesh, const char *filname, const char *format, int natt, char
+%%     **attnames):} Export a mesh to an external file format. Currently,
+%%   the formats supported are the GMV file (format string: ``gmv'') and
+%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
+%%   ``dx''). Only integer and double attributes of the mesh are exported
+%%   to the files.
+
+\item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr} mesh, const
+    char *filname, const char *format, int natt, char **attnames):}
+  Export a mesh to an external file format. Currently, the only
+  formats supported is the GMV file format (format string: ``gmv'')
+  and X3D (format string: ``flag''). Only integer and double
+  attributes of the mesh are exported to GMV files and no attributes
+  are exported to X3D files.
+
+\item[]{\bf {\em void} MESH\_ExportToGMV({\em Mesh\_ptr} mesh, const
+    char *filname, const char *format, int natt, char **attnames):} Export a mesh to a GMV file format.
+
+\item[]{\bf {\em void} MESH\_ExportToFLAGX3D({\em Mesh\_ptr} mesh, const
+    char *filname, const char *format, int natt, char **attnames):} Export a mesh to the FLAG X3D format.
+
+\item[]{\bf {\em void} MESH\_ExportToFLAGX3D\_Par({\em Mesh\_ptr} mesh,
+    const char *filname, const char *format, int natt, char
+    **attnames, int *procids):} Export a mesh to the FLAG X3D format
+  given a partitioned mesh. The owning processor for each element is
+  given in the array {\bf procids}.
+
+\item[]{\bf {\em void} MESH\_ExportToSTL({\em Mesh\_ptr} mesh, const
+    char *filname):} Export a mesh to the STL format (SURFACE MESHES ONLY!).
+
+\item[]{\bf {\em void} MESH\_Tet2Hex({\em Mesh\_ptr} tetmesh, {\em
+      Mesh\_ptr} *hexmesh):} Convert a tet mesh to a hex mesh by
+  splitting the tets - the quality of the resulting hex mesh is
+  usually not very good.
+
+\item[]{\bf {\em void} MESH\_Renumber({\em Mesh\_ptr} mesh):} Renumber
+  the entities of a mesh to avoid gaps in entity IDs. Note that in the
+  current implementation, renumbering mesh entities can make removal
+  of and searching for mesh entities much slower, so this must be
+  avoided as much as possible.
+
+\item[]{\bf {\em int} MESH\_PartitionWithMetis({\em Mesh\_ptr} mesh,
+    {\em int} nparts, {\em int} **part):} Partition a mesh with the
+  METIS libraries. 'part' is an array that contains the partition
+  number for each mesh face (surface meshes) or mesh region (volume
+  meshes).
+
+  \par {\bf NOTE THAT MSTK HAS TO HAVE BEEN COMPILED USING THE COMMAND
+    'make PAR=1' AND THE CALLING APPLICATION MUST LINK WITH THE METIS
+    LIBRARY.}
+
+\item[]
+
+\item[]{\bf {\em GModel\_ptr} MESH\_GModel({\em Mesh\_ptr} mesh):}
+Return a handle to the underlying geometric model. If there is no
+geometric model associated with the mesh, NULL pointer is returned.
+
+\item[]{\bf {\em RepType} MESH\_RepType({\em Mesh\_ptr} mesh):}
+Representation type currently being used by the mesh.
+
+\item[]{\bf {\em char } *MESH\_RepType({\em Mesh\_ptr} mesh):}
+  Representation type currently being used by the mesh returned as a
+  2-character string.
+
+\item[]
+
+\item[]{\bf {\em int} MESH\_Num\_Vertices({\em Mesh\_ptr} mesh):}
+Number of vertices in the mesh.
+
+\item[]{\bf {\em int} MESH\_Num\_Edges({\em Mesh\_ptr} mesh):}
+Number of edges in the mesh. For reduced representations, this routine
+returns 0 since it is impractically expensive to count the number of
+edges when they do not explicitly exist. Applications must find a way
+to avoid using this routine for reduced representations.
+
+\item[]{\bf {\em int} MESH\_Num\_Faces({\em Mesh\_ptr} mesh):}
+Number of faces in the mesh. For reduced representations R1 or R2,
+this routine counts only the faces that are explicitly represented
+i.e. faces not connected to any mesh region. Therefore, a value of 0
+will be returned for the number of faces of a tetrahedral mesh with
+representation R1 or R2 but the correct number will be reported for a
+tetrahedral mesh in other representations. Also, the correct number
+will be reported for the number of faces in a surface mesh in
+representation R1 or R2.  Therefore, this routine must be used
+carefully.
+
+\item[]{\bf {\em int} MESH\_Num\_Regions({\em Mesh\_ptr} mesh):}
+Number of regions in the mesh.
+
+\item[]
+
+\item[]{\bf {\em MVertex\_ptr} MESH\_Vertex({\em Mesh\_ptr} mesh,
+{\em int} i):} Return the i'th vertex in the mesh. Returns NULL if i
+$<$ 0 or i $>$ number of mesh vertices.
+
+\item[]{\bf {\em MEdge\_ptr} MESH\_Edge({\em Mesh\_ptr} mesh,
+{\em int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
+0 or i $>$ number of mesh edges. Returns NULL for reduced
+representations.
+
+\item[]{\bf {\em MFace\_ptr} MESH\_Face({\em Mesh\_ptr} mesh,
+{\em int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
+0 or i $>$ number of mesh faces. Only faces explicitly represented in
+the mesh are returned for reduced representation (See explanation for
+MESH\_Num\_Faces).
+
+\item[]{\bf {\em MRegion\_ptr} MESH\_Region({\em Mesh\_ptr} mesh, {\em
+      int} i):} Return the i'th region in the mesh. Returns NULL if i
+  $<$ 0 or i $>$ number of mesh region.
+
+\item[]
+
+\item[]{\bf {\em MVertex\_ptr} MESH\_Next\_Vertex({\em Mesh\_ptr}
+    mesh, {\em int} *idx):} Returns the next vertex while iterating
+  through the vertices of the mesh. See the routine {\bf
+    List\_Next\_Entry} above for an explanation of how the iteration
+  works. This routine is in general faster than using the routine {\bf
+    MESH\_Vertex}.
+
+\item[]{\bf {\em MEdge\_ptr} MESH\_Next\_Edge({\em Mesh\_ptr} mesh,
+{\em int} *idx):} Returns the next edge while iterating through the
+edges of the mesh. See the routine List\_Next\_Entry above for an
+explanation of how the iteration works.  The routine always returns
+NULL for reduced representations.
+
+\item[]{\bf {\em MFace\_ptr} MESH\_Next\_Face({\em Mesh\_ptr} mesh,
+{\em int} *idx):} Returns the next face while iterating through the
+faces of the mesh. See the routine {\bf List\_Next\_Entry} above for
+an explanation of how the iteration works.  Only faces explicitly
+represented in the mesh are returned for reduced representation (See
+explanation for {\bf MESH\_Num\_Faces}).
+
+\item[]{\bf {\em MRegion\_ptr}
+    MESH\_Next\_Region({\em Mesh\_ptr} mesh, {\em int} *idx):}
+  Returns the next region while iterating through the regions of the
+  mesh. See the routine {\bf List\_Next\_Entry} above for an
+  explanation of how the iteration works.
+
+\item[]
+
+\item[]{\bf {\em MVertex\_ptr}
+    MESH\_VertexFromID({\em Mesh\_ptr} mesh, {\em int} id):}
+  Return mesh vertex with given ID if it exists; return NULL
+  otherwise.
+
+\item[]{\bf {\em MEdge\_ptr} MESH\_EdgeFromID({\em Mesh\_ptr}
+    mesh, {\em int} id):} Return mesh edge with given ID if it
+  exists; return NULL otherwise. This routine will return NULL for all
+  reduced representations.
+
+\item[]{\bf {\em MFace\_ptr} MESH\_FaceFromID({\em Mesh\_ptr}
+    mesh, {\em int} i):} Return mesh face with given ID if it
+  exists; return NULL otherwise. If faces are not explicitly
+  represented in the mesh, it will return NULL.
+
+\item[]{\bf {\em MRegion\_ptr}
+    MESH\_RegionFromID({\em Mesh\_ptr} mesh, {\em int} id):}
+  Return mesh region with given ID if it exists; return NULL
+  otherwise.
+
+\item[]
+
+\item[]{\bf {\em int} MESH\_Num\_Attribs({\em Mesh\_ptr}
+    mesh):} Number of attributes associated with the mesh.
+
+\item[]{\bf {\em MAttrib\_ptr} MESH\_Attrib({\em Mesh\_ptr}
+    mesh, {\em int} i):} Return the i'th attribute in the mesh.
+  Returns NULL if i $<$ 0 or i $>$ number of mesh attributes.
+
+\item[]{\bf {\em MAttrib\_ptr}
+    MESH\_Next\_Attrib({\em Mesh\_ptr} mesh, {\em int} *index):}
+  Returns the next attribute while iterating through the attributes of
+  the mesh. See the routine List\_Next\_Entry above for an explanation
+  of how the iteration works.
+
+\item[]{\bf {\em MAttrib\_ptr}
+    MESH\_AttribByName({\em Mesh\_ptr} mesh, {\em const char}
+    *name):} Return a mesh attribute with given name if it exisists in
+  the mesh. Returns NULL if mesh has no such attribute.
+
+\item[]
+
+\item[]{\bf {\em void} MESH\_Set\_GModel({\em Mesh\_ptr} mesh,
+GModel\_ptr geom):} Assign a geometric model handle to the mesh.
+
+%\item[]{\bf {\em int} MESH\_Change\_RepType({\em Mesh\_ptr} mesh,
+%{\em int} nurep):} Change the representation type of the mesh. This
+%routine can be used to modify the representation type dynamically to
+%suit different algorithms.  However, the cost of making the change and
+%reordering all adjacencies and creating or deleting entities has to be
+%considered while invoking this routine. Also, once a conversion is
+%made from a full representation to a reduced representation, not all
+%information may be retrievable when switching back to a full
+%representation.  (particularly classification information, i.e.,
+%relationship of mesh entities to the geometric model).
+
+\item[]
+
+\item[]{\bf {\em void} MESH\_Add\_Vertex({\em Mesh\_ptr}
+    mesh, {\em MVertex\_ptr} v):} Add a vertex to the mesh. It is
+  assumed that the vertex and its coordinates set are properly
+  defined. Normally, one need not call this since
+  {\bf {\em MV\_New}} will add the vertex to the mesh. Use this
+  only if you know exactly what you are doing.
+
+\item[]{\bf {\em void} MESH\_Add\_Edge({\em Mesh\_ptr} mesh,
+    {\em MEdge\_ptr} e):} Add an edge to the mesh. It is assumed
+  that the edge is and its topology is defined. Normally, one need not
+  call this since {\bf {\em ME\_New}} will add the edge to the
+  mesh. Use this only if you know exactly what you are doing.
+
+\item[]{\bf {\em void} MESH\_Add\_Face({\em Mesh\_ptr} mesh,
+    {\em MFace\_ptr} f):} Add a face to the mesh. It is assumed
+  that the face and its topology is properly defined. Normally, one
+  need not call this since {\bf {\em MF\_New}} will add the face
+  to the mesh. Use this only if you know exactly what you are doing.
+
+\item[]{\bf {\em void} MESH\_Add\_Region({\em Mesh\_ptr}
+    mesh, {\em MRegion\_ptr} r):} Add a region to the mesh. It is
+  assumed that the region and its topology is properly defined.
+  Normally, one need not call this since {\bf {\em MR\_New}}
+  will add the region to the mesh. Use this only if you know exactly
+  what you are doing.
+
+\item[]{\bf {\em void} MESH\_Rem\_Vertex({\em Mesh\_ptr}
+    mesh, {\em MVertex\_ptr} v):} Remove vertex from mesh. Vertex
+  is not deleted and must be deleted afterward separately. Normally,
+  one need not call this since {\bf {\em MV\_Delete}} will
+  remove the vertex from the mesh. Use this only if you know exactly
+  what you are doing.
+
+\item[]{\bf {\em void} MESH\_Rem\_Edge({\em Mesh\_ptr} mesh,
+    {\em MEdge\_ptr} e):} Remove edge from mesh. Edge is not
+  deleted and must be deleted afterward separately. Normally, one need
+  not call this since {\bf {\em ME\_Delete}} will remove the
+  edge from the mesh. Use this only if you know exactly what you are
+  doing.
+
+\item[]{\bf {\em void} MESH\_Rem\_Face({\em Mesh\_ptr} mesh,
+    {\em MFace\_ptr} f):} Remove face from mesh. Face is not
+  deleted and must be deleted afterward separately. Normally, one need
+  not call this since {\bf {\em MF\_Delete}} will remove the
+  face from the mesh. Use this only if you know exactly what you are
+  doing.
+
+\item[]{\bf {\em void} MESH\_Rem\_Region({\em Mesh\_ptr}
+    mesh, {\em MRegion\_ptr} r):} Remove region from mesh. Region
+  is not deleted and must be deleted afterward separately. Normally,
+  one need not call this since {\bf {\em MR\_Delete}} will
+  remove the region from the mesh. Use this only if you know exactly
+  what you are doing.
+\end{description}
+
+\newpage
+\subsection{Mesh Vertex Object}
+
+\begin{description}
+\item[]{\bf {\em MVertex\_ptr} MV\_New({\em Mesh\_ptr}
+    mesh):} Create a new vertex object. No geometric or topological
+  information is embedded in the vertex when it is created. The vertex
+  only knows which mesh it belongs to. The ID of the vertex is set by
+  this function.
+
+\item[]{\bf {\em void} MV\_Delete({\em MVertex\_ptr} mvertex,
+    {\em int} keep):} Delete the vertex. If {\bf keep} is 0, the
+  vertex is removed from the mesh and all topological and geometric
+  information embedded in the vertex is destroyed. If {\bf keep} is
+  1, the vertex is marked as type {\em MDELVERTEX} and removed from
+  the mesh but the vertex is not destroyed.
+
+  \par {\bf NOTE: {\em MV\_Delete} AND ITS COUNTERPARTS WILL, IN GENERAL,
+    REMOVE AN ENTITY IN $O(log(N))$ TIME WHERE $N$ IS THE TOTAL NUMBER
+    OF ENTITIES ADDED TO THE MESH SINCE THE START OF THE
+    PROGRAM. HOWEVER, IF THE MESH ENTITIES HAVE NOT BEEN RENUUMBERED
+    (EITHER USING MESH\_RENUMBER OR EXPLICITLY USING MENT\_SET\_ID) OR
+    THE MESH ENTITY LISTS HAVE NOT BEEN COMPRESSED THEN THE REMOVAL
+    TIME IS $O(1)$ WHICH IS CLEARLY MUCH SUPERIOR.}
+
+\item[]{\bf {\em void} MV\_Restore({\em MVertex\_ptr}
+    mvertex):} Restore a deleted vertex. The vertex type is restored
+  from {\em MDELVERTEX} to {\em MVERTEX} and the vertex is
+  added back to the mesh.
+
+\item[]
+
+\item[]{\bf {\em void} MV\_Set\_Coords({\em MVertex\_ptr}
+    mvertex, double *xyz):} Set the coordinates of the vertex.
+
+\item[]{\bf {\em void} MV\_Set\_GEntity({\em MVertex\_ptr}
+    mvertex, GEntity\_ptr gent):} Set the geometric model entity on
+  which vertex is classified.
+
+\item[]{\bf {\em void} MV\_Set\_GEntDim({\em MVertex\_ptr} mvertex,
+{\em int} gdim):} Set topological dimension of model entity on which
+vertex is classified.
+
+\item[]{\bf {\em void} MV\_Set\_GEntID({\em MVertex\_ptr} mvertex,
+{\em int} gid):} Set ID of model entity on which vertex is
+classified.
+
+\item[]{\bf {\em void} MV\_Add\_AdjVertex({\em MVertex\_ptr}
+mvertex, {\em MVertex\_ptr} adjvertex):} Add neighboring vertex,
+adjvertex, to ajdacent vertex list of vertex, mvertex.
+
+\item[]{\bf {\em void} MV\_Rem\_AdjVertex({\em MVertex\_ptr}
+mvertex, {\em MVertex\_ptr} adjvertex):} Delete neighboring vertex
+of given vertex.
+
+\item[]{\bf {\em void} MV\_Set\_ID({\em MVertex\_ptr} mvertex,
+{\em int} id):} Explicitly set ID of a vertex and overwrite the ID
+set by the MV\_New operator. Does not check for duplication of edge
+IDs.
+
+\item[]
+
+\item[]{\bf {\em Mesh\_ptr} MV\_Mesh({\em MVertex\_ptr} mv):}
+Returns the mesh that this vertex belongs to.
+
+\item[]{\bf {\em int} MV\_ID({\em MVertex\_ptr} mvertex):} Returns
+the ID of the vertex.
+
+\item[]{\bf {\em int} MV\_GEntDim({\em MVertex\_ptr} mvertex):}
+Returns the dimension of the geometric model entity that the vertex is
+classified on. Returns -1 if not known.
+
+\item[]{\bf {\em int} MV\_GEntID({\em MVertex\_ptr} mvertex):}
+Returns the ID of the geometric model entity that the vertex is
+classified on. Returns 0 if this information is not known.
+
+\item[]{\bf {\em GEntity\_ptr} MV\_GEntity({\em MVertex\_ptr}
+  mvertex):} Returns a pointer or handle to the geometric model entity
+  that the vertex is classified on. Returns NULL if this information
+  is not known.
+
+\item[]{\bf {\em void} MV\_Coords({\em MVertex\_ptr} mvertex,
+double *xyz):} Returns the coordinates of the vertex.
+
+\item[]
+
+\item[]{\bf {\em int} MV\_Num\_AdjVertices({\em MVertex\_ptr}
+mvertex):} Returns the number of edge connected neighboring vertices of
+vertex. {\em Not efficient for all representations}.
+
+\item[]{\bf {\em int} MV\_Num\_Edges({\em MVertex\_ptr} mvertex):}
+Returns the number of edges connected to the vertex.
+
+\item[]{\bf {\em int} MV\_Num\_Faces({\em MVertex\_ptr} mvertex):}
+Returns the number of faces connected to the vertex.
+
+\item[]{\bf {\em int} MV\_Num\_Regions({\em MVertex\_ptr} mvertex):}
+Returns the number of regions connected to the vertex
+
+\item[]{\bf {\em List\_ptr} MV\_AdjVertices({\em MVertex\_ptr}
+mvertex):} List of adjacent or edge connected neighboring vertices of vertex.
+
+\item[]{\bf {\em List\_ptr} MV\_Edges({\em MVertex\_ptr} mvertex):}
+List of edges connected to the vertex.
+
+\item[]{\bf {\em List\_ptr} MV\_Faces({\em MVertex\_ptr} mvertex):}
+List of faces connected to the vertex.
+
+\item[]{\bf {\em List\_ptr} MV\_Regions({\em MVertex\_ptr} mvertex):}
+List of regions connected to the vertex.
+\end{description}
+
+\newpage
+\subsection{Mesh Edge Object}
+
+\begin{description}
+\item[]{\bf {\em MEdge\_ptr} ME\_New({\em Mesh\_ptr} mesh):}
+  Create a new edge object. No topological information is embedded in
+  the edge when it is created. The edge only knows which mesh it
+  belongs to. The ID of the edge is set by this function.
+
+\item[]{\bf {\em void} ME\_Delete({\em MEdge\_ptr} medge,
+    {\em int} keep):} Delete the edge. If {\bf keep} is 0, the
+  edge is removed from the mesh and all topological and geometric
+  information embedded in the edge is destroyed. If {\bf keep} is
+  1, the edge is marked as type {\em MDELEDGE} and removed from the
+  mesh but the edge is not destroyed. Also, the vertices of this edge
+  no longer point to this edge.
+
+\item[]{\bf {\em void} ME\_Restore({\em MEdge\_ptr}
+    medge):} Restore a temporarily deleted edge. The edge type is
+  restored from {\em MDELEDGE} to {\em MEDGE} and the edge is
+  added back to the mesh. The vertices of the edge once again point
+  back to the edge.
+
+\item[]
+
+\item[]{\bf {\em void} ME\_Set\_GEntity({\em MEdge\_ptr}
+    medge, GEntity\_ptr gent):} Set the geometric model entity on
+  which the edge is classified.
+
+\item[]{\bf {\em void} ME\_Set\_GEntDim({\em MEdge\_ptr}
+    medge, {\em int} gdim):} Set the topological dimension of model
+  entity on which edge is classified.
+
+\item[]{\bf {\em void} ME\_Set\_GEntID({\em MEdge\_ptr}
+    medge, {\em int} gid):} Set ID of model entity on which edge is
+  classified.
+
+\item[]{\bf {\em void} ME\_Set\_GInfo\_Auto({\em MEdge\_ptr} medge):}
+  Derive the classification (GEntDim, GEntID) of the mesh edge
+  automatically (if possible) from the classification of its
+  vertices. If it is not possible to unambiguously get this
+  information, the procedure will keep the default information. {\bf THIS
+  MAY GET WRONG OR INCORRECT INFORMATION WHEN THERE IS AMBIGUOUS
+  DATA. FOR EXAMPLE, IF THE VERTICES OF A MESH EDGE ARE CLASSIFIED ON
+  TWO DIFFERENT MODEL VERTICES, IT IS NOT POSSIBLE TO KNOW WHAT ENTITY
+  THE MESH EDGE IS CLASSIFIED USING JUST THE VERTEX INFORMATION.}
+
+\item[]{\bf {\em void} ME\_Set\_ID({\em MEdge\_ptr} medge,
+    {\em int} id):} Explicitly set ID of an edge and overwrite the
+  ID set by the ME\_New function. Does not check for duplication of
+  edge IDs.
+
+\item[]{\bf {\em void} ME\_Set\_Vertex({\em MEdge\_ptr}
+    medge, {\em int} i, {\em MVertex\_ptr} vertex):} Set the
+  i'th vertex of the edge. i can be 0 or 1.
+
+\item[]{\bf {\em void} ME\_Replace\_Vertex({\em MEdge\_ptr}
+    medge, {\em MVertex\_ptr} vert, {\em MVertex\_ptr} nuvert):}
+  Replace i'th vertex by new vertex.
+
+\item[]
+
+\item[]{\bf {\em Mesh\_ptr} ME\_Mesh({\em MEdge\_ptr}
+    medge):} Returns the mesh that this edge belongs to.
+
+\item[]{\bf {\em int} ME\_ID({\em MEdge\_ptr} medge):}
+  Returns the ID of the vertex. Returns -1 if not known.
+
+\item[]{\bf {\em int} ME\_GEntDim({\em MEdge\_ptr} medge):}
+  Returns the dimension of the geometric model entity that the vertex
+  is classified on. Returns -1 if not known.
+
+\item[]{\bf {\em int} ME\_GEntID({\em MEdge\_ptr} medge):}
+  Returns the ID of the geometric model entity that the vertex is
+  classified on.  Returns 0 if this information is not known.
+
+\item[]{\bf {\em GEntity\_ptr} ME\_GEntity({\em MEdge\_ptr}
+    medge):} Returns a pointer or handle to the geometric model entity
+  that the vertex is classified on. Returns NULL if this information
+  is not known.
+
+\item[]
+
+\item[]{\bf {\em int} ME\_Num\_Faces({\em MEdge\_ptr}
+    medge):} Returns the number of faces connected to the edge.
+
+\item[]{\bf {\em int} ME\_Num\_Regions({\em MEdge\_ptr}
+    medge):} Returns the number of regions connected to the edge.
+
+\item[]{\bf {\em MVertex\_ptr} ME\_Vertex({\em MEdge\_ptr}
+    medge, {\em int} i):} Returns the i'th vertex of the edge. i=0
+  returns the first vertex and i=1 returns the second vertex.
+
+\item[]{\bf {\em MVertex\_ptr} ME\_OppVertex({\em MEdge\_ptr}
+    medge, {\em MVertex\_ptr} ov):} Return the vertex opposite to
+  given vertex in edge.
+
+\item[]{\bf {\em int} ME\_UsesEntity({\em MEdge\_ptr} medge,
+    {\em MEntity\_ptr} mentity, {\em int} etype):} Check if edge
+  uses given lower dimension entity, {\em mentity}. The dimension
+  of the entity is specified by the {\em etype} variable. For an
+  edge, the only lower dimensional entity is a vertex. If the edge
+  uses the vertex, the function returns 1; otherwise it returns 0. If
+  any other type of entity is specified, the function returns 0.
+
+\item[]{\bf {\em List\_ptr} ME\_Faces({\em MEdge\_ptr}
+    medge):} Returns the set of faces using this edge.
+
+\item[]{\bf {\em List\_ptr} ME\_Regions({\em MEdge\_ptr}
+    medge):} Returns the set of regions using this edge.
+
+\item[]
+
+\item[]{\bf {\em MEdge\_ptr}
+    MVs\_CommonEdge({\em MVertex\_ptr} v1, {\em MVertex\_ptr}
+    v2):} Return the edge connecting vertices v1 and v2, if it exists.
+  If such an edge does not exist, the function returns 0.
+
+\item[]{\bf {\em double} ME\_Len({\em MEdge\_ptr} e):} Return
+  the length of the straight line connecting the two vertices of the
+  edge.
+
+\item[]{\bf {\em double} ME\_LenSqr({\em MEdge\_ptr} e):}
+  Return the square of the length of the straight line connecting the
+  two vertices of the edge.
+
+\item[]{\bf {\em void} ME\_Vec({\em MEdge\_ptr} e, double
+    *evec):} Return the vector going from the first vertex of the edge
+  to the second vertex of the edge.
+
+\item[]{\bf {\em int} MEs\_AreSame({\em MEdge\_ptr} e1, {\em
+      MEdge\_ptr} e2):} {\em Applicable only to reduced
+    representations.} Check if two edges described by only their
+  vertices are the same. In a reduced representation, edges are
+  temporary objects described by their end vertices. If two temporary
+  edges are described by the same vertices, the edges are considered
+  to be the same.
+
+\end{description}
+
+\newpage
+\subsection{Mesh Face Object}
+
+\begin{description}
+\item[]{\bf {\em MFace\_ptr} MF\_New({\em Mesh\_ptr} mesh):} Create
+a new face object. No topological information is embedded in the face
+when it is created. The face only knows which mesh it belongs
+to. The ID of the face is set by this function.
+
+\item[]{\bf {\em void} MF\_Delete({\em MFace\_ptr} mface):}
+  Delete the face. If {\bf keep} is 0, the face is removed from the
+  mesh and all topological and geometric information embedded in the
+  face is destroyed. If {\bf keep} is 1, the face is marked as type
+  {\em MDELFACE} and removed from the mesh but the face is not
+  destroyed. Also, the vertices/edges of this face no longer point up
+  to this face.
+
+\item[]{\bf {\em void} MF\_Restore({\em MFace\_ptr}
+    mface):} Restore a temporarily deleted face. The face type is
+  restored from {\em MDELFACE} to {\em MFACE} and the face is
+  added back to the mesh. The vertices/edges of the face once again point
+  back to the face.
+
+\item[]
+
+\item[]{\bf {\em void} MF\_Set\_GEntity({\em MFace\_ptr} mface, GEntity\_ptr gent):} Set the geometric model entity on which the face is classified.
+
+\item[]{\bf {\em void} MF\_Set\_GEntDim({\em MFace\_ptr} mface,
+{\em int} gdim):} Set the dimension of the geometric model entity on
+which the face is classified.
+
+\item[]{\bf {\em void} MF\_Set\_GEntID({\em MFace\_ptr} mface,
+{\em int} gid):} Set the ID of the geometric model entity on which
+the face is classified.
+
+\item[]{\bf {\em void} MF\_Set\_GInfo\_Auto({\em MFace\_ptr} mface):}
+  Derive the classification (GEntDim, GEntID) of the mesh face
+  automatically (if possible) from the classification of its
+  vertices or edges. If it is not possible to unambiguously get this
+  information, the procedure will keep the default information. {\bf THIS
+  MAY GET WRONG OR INCORRECT INFORMATION WHEN THERE IS AMBIGUOUS
+  DATA.}
+
+\item[]{\bf {\em void} MF\_Set\_ID({\em MFace\_ptr} mface,
+{\em int} id):} Explicitly set ID of a face and overwrite the ID
+set by the {\bf MF\_New} operator. Does not check for duplication of face
+IDs.
+
+\item[]{\bf {\em void} MF\_Set\_Edges({\em MFace\_ptr} mface,
+{\em int} n, {\em MEdge\_ptr} *edges, {\em int} *dirs):} Set
+the edges of the face along with their directions. The ordered set of
+edge pointers and their directions are passed in through arrays along
+with the number of edges. The edges are assumed to be ordered
+clockwise around the face. If an edge direction is along the clockwise
+direction of the face then the entry in the 'dirs' array must be 1;
+otherwise it must be 0. This function is relevant only for full
+representations in MSTK.
+
+\item[]{\bf {\em void} MF\_Set\_Vertices({\em MFace\_ptr}
+    mface, {\em int} n, {\em MVertex\_ptr} *verts):} Set the
+  vertices of the face. The ordered set of vertices (ccw around the
+  face) is passed in through an array along with the number of
+  vertices. This routine will collect/build all lower order
+  topological information (edges, edge directions in the face) as needed by the face.
+
+\item[]{\bf {\em void} MF\_Replace\_Edges({\em MFace\_ptr} mface, {\em
+      int} nold, {\em MEdge\_ptr} *oldedges, {\em int} nnu, {\em
+      MEdge\_ptr} *nuedges):} Replace a set of edges in the face with
+  another set of edges. The direction in which the new edges are to be
+  used in the face are automatically deduced.  This function is
+  relevant only for full representations in MSTK. The old and new edge
+  sets must be comprised of edges connected to each other and the end
+  vertices of the new set must match the end vertices of the new set.
+
+\item[]{\bf {\em void} MF\_Replace\_Edges\_i({\em MFace\_ptr}
+    mface, {\em int} nold, {\em int} i, {\em int} nnu,
+    {\em MEdge\_ptr} *nuedges):} Replace the {\em nold} edges in
+  the face starting with the i'th edge and going ccw around the face
+  with a new set of edges. The directions in which the new edges are
+  used in the face are automatically deduced. This function is
+  relevant only for full representations in MSTK.
+
+\item[]{\bf {\em void} MF\_Replace\_Vertex({\em MFace\_ptr} mface,
+{\em MVertex\_ptr} mvertex, {\em MVertex\_ptr} nuvertex):}
+Replace a vertex in the face with another vertex. This function is
+relevant only for reduced representations in MSTK.
+
+\item[]{\bf {\em void} MF\_Replace\_Vertex\_i({\em MFace\_ptr}
+mface, {\em int} i, {\em MVertex\_ptr} nuvertex):} Replace the
+i'th vertex in the face with a new vertex. This function is relevant
+only for reduced representations in MSTK.
+
+\item[]{\bf {\em void} MF\_Insert\_Vertex({\em MFace\_ptr} mface,
+{\em MVertex\_ptr} nuvertex, {\em MVertex\_ptr} b4vertex):}
+Insert a vertex in the face before {\bf b4vertex} w.r.t. a ccw
+ordering of the vertex faces.
+
+\item[]{\bf {\em void} MF\_Insert\_Vertex\_i({\em MFace\_ptr}
+mface, {\em MVertex\_ptr} nuvertex, {\em int} i):} Insert a new
+vertex before vertex i in the face w.r.t. a ccw ordering of vertices.
+
+\item[]
+
+\item[]{\bf {\em Mesh\_ptr} MF\_Mesh({\em MFace\_ptr} mf):} Returns the mesh that this mesh belongs to.
+
+\item[]{\bf {\em int} MF\_ID({\em MFace\_ptr} mface):} Returns the ID of the face. Returns 0 if not known.
+
+\item[]{\bf {\em int} MF\_GEntDim({\em MFace\_ptr} mface):} Returns the dimensions of the geometric model entity that the vertex is classified on. Returns -1 if not known.
+
+\item[]{\bf {\em int} MF\_GEntID({\em MFace\_ptr} mface):} Returns the ID of the geometric model entity that the vertex is classified on. Returns 0 if this information is not known.
+
+\item[]{\bf {\em GEntity\_ptr} MF\_GEntity({\em MFace\_ptr} mface):} Returns a pointer or handle to the geometric model entity that the vertex is classified on.
+Returns NULL if this information is not known.
+
+\item[]
+
+\item[]{\bf {\em int} MF\_Num\_Vertices({\em MFace\_ptr} mface):} Returns the number of vertices of the face.
+
+\item[]{\bf {\em int} MF\_Num\_Edges({\em MFace\_ptr} mface):} Returns the number of edges of the face.
+
+\item[]{\bf {\em int} MF\_Num\_AdjFaces({\em MFace\_ptr}
+    mface):} Returns the number of adjacent faces of a face. This
+  operator is relevant only in planar or surface meshes, i.e., for
+  boundary faces not connected to any regions.
+
+\item[]{\bf {\em List\_ptr} MF\_Vertices({\em MFace\_ptr}
+    mface, {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
+  ordered set of the vertices of the face. The vertices are ordered in
+  ccw direction while looking down the face 'normal', if 'dir' is 1
+  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
+  vertex set is reordered so that it is the first vertex (This
+  argument will be added soon to the function. For now, omit this
+  argument). The behavior of this function can be illustrated using
+  Figure~\ref{fig:face_def}. For the face shown in the figure, a
+  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
+  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
+  A vertex set with ccw ordering starting with vertex $V_2$ is
+  ${V_2,V_3,V_0,V_1}$.
+
+\begin{figure}[!h]
+  \begin{center}
+  \includegraphics[scale=1.0]{figures/face_def} 
+  \caption{Face definition}
+  \label{fig:face_def}
+  \end{center}
+\end{figure}
+
+\item[]{\bf {\em List\_ptr} MF\_Edges({\em MFace\_ptr} mface,
+    {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
+  ordered set of edges of the face. The edges are ordered in the ccw
+  while looking down the face 'normal' if dir is 1 and in the cw if
+  dir is 0. If 'mvert' is specified, the edge set is reordered so that
+  the first edge in the set contains this vertex. More precisely, if
+  'dir' is 1, and the first edge is 'e' used in the face in direction
+  'd', then {\bf ME\_Vertex(e,!d)} = mvert. With reference to
+  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
+  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
+  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
+  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
+  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
+  ${E_1,E_0,E_3,E_2}$.
+
+\item[]{\bf {\em List\_ptr} MF\_AdjFaces({\em MFace\_ptr} mface):}
+List of adacent faces of a face. This operator is relevant only in
+planar or surface surface meshes, i.e., for boundary faces not
+connected to any regions.
+
+\item[]{\bf {\em int} MF\_EdgeDir({\em MFace\_ptr} mface,
+{\em MEdge\_ptr} medge):} Returns the direction in which the face
+uses the given edge. If the faces use the edge in the positive direction,
+the function returns 1; otherwise it returns 0.
+
+\item[]{\bf {\em int} MF\_EdgeDir\_i({\em MFace\_ptr} mface,
+{\em int} i):} Returns the direction in which the face uses its i'th
+edge. If the face uses the edge in the positive direction the function
+returns 1; otherwise it returns 0;
+
+\item[]{\bf {\em List\_ptr} MF\_Regions({\em MFace\_ptr} mface):}
+Return the set of regions connected to the face. If the face is not
+used by any regions, the function returns NULL to indicate that the
+set is empty. If not the set may contain one or two regions.
+
+\item[]{\bf {\em MRegion\_ptr} MF\_Region({\em MFace\_ptr} mface,
+{\em int} side):} Returns the region on the specified side of the
+face. The positive side of the face (side = 1) is the side towards
+which the face normal points. The negative side of the face (side = 0)
+is the opposite side.
+
+\item[]
+
+\item[]{\bf {\em int} MF\_UsesEntity({\em MFace\_ptr} mface,
+{\em MEntity\_ptr} mentity, {\em int} type):} Check if the face
+uses the given lower dimension entity, 'mentity'. The type of the
+entity is specified by the 'etype' variable. For a face, a lower
+dimensional entity can be a vertex or an edge. If the face uses the
+vertex or the edge, the function returns 1; otherwise it returns 0. If
+any other type of entity is specified, the function returns 0.
+
+\item[]{\bf {\em MFace\_ptr} MVs\_CommonFace({\em int} nv,
+    {\em MVertex\_ptr} *fverts):} Return a face connected to all
+  the given mesh vertices. Returns NULL if no such face exists.
+
+\item[]{\bf {\em MFace\_ptr} MEs\_CommonFace({\em int} ne,
+    {\em MEdge\_ptr} *fedges):} Return a face connected to all
+  the given mesh edges. Returns NULL if no such face exists.
+
+\item[]{\bf {\em int} MFs\_AreSame({\em MFace\_ptr} f1, {\em
+      MFace\_ptr} f2):} {\em Applicable only to reduced
+    representations}. Check if two faces described by vertices are
+  equivalent. For example, a face described by vertices $\{1,2,3\}$ is
+  equivalent to faces described by vertices $\{1,3,2\}$ and
+  $\{3,1,2\}$.
+
+\item[]
+
+\item[]{\bf {\em void} MF\_Coords({\em MFace\_ptr} mface,
+{\em int} *n, double (*xyz)[3], {\em int} dir):} Returns the
+coordinates of the face vertices in an array along with the number of
+vertices. 
+% If 'dir' is 1, the coordinates are returned with a  ccw
+% ordering while looking down the face normal; if 'dir' is 0, they are
+% returned with a cw ordering (See Figure~\ref{fig:face_def}).
+\end{description}
+
+\newpage
+\subsection{Mesh Region Object}
+
+\begin{description}
+\item[]{\bf {\em MRegion\_ptr} MR\_New({\em Mesh\_ptr}
+    mesh):} Create a new region object. No topological information is
+  embedded in the region when it is created. The region only knows
+  which mesh it belongs to. The ID of the region is set by this
+  function.
+
+\item[]{\bf {\em void} MR\_Delete({\em MRegion\_ptr} mregion,
+    {\em int} keep):} Delete the region. Deletes all topological
+  information embedded in the region. If {\bf keep} is 0, the
+  region is removed from the mesh and all topological and geometric
+  information embedded in the region is destroyed. If {\bf keep} is
+  1, the region is marked as type {\em MDELREGION} and removed from
+  the mesh but the region is not destroyed. Also, the vertices, edges
+  and faces of this region no longer point up to this region.
+
+\item[]{\bf {\em void} MR\_Restore({\em MRegion\_ptr}
+    mregion):} Restore a temporarily deleted region. The region type
+  is restored from {\em MDELREGION} to {\em MREGION} and the
+  region is added back to the mesh. The vertices, edges and faces of
+  the region once again point back to the region.
+
+\item[]
+
+\item[]{\bf {\em void} MR\_Set\_GEntity({\em MRegion\_ptr}
+    mregion, GEntity\_ptr gent):} Set the geometric model entity on
+  which the region is classified.
+
+\item[]{\bf {\em void} MR\_Set\_GEntDim({\em MRegion\_ptr}
+    mregion, {\em int} gdim):} Set the dimension of the geometric
+  model entity on which the region is classified.
+
+\item[]{\bf {\em void} MR\_Set\_GEntID({\em MRegion\_ptr}
+    mregion, {\em int} gid):} Set the ID of the geometric model
+  entity on which the region is classified.
+
+\item[]{\bf {\em void} MR\_Set\_GInfo\_Auto({\em MRegion\_ptr} mregion):}
+  Derive the classification (GEntDim, GEntID) of the mesh region
+  automatically (if possible) from the classification of its
+  vertices or faces. If it is not possible to unambiguously get this
+  information, the procedure will keep the default information. {\bf THIS
+  MAY GET WRONG OR INCORRECT INFORMATION WHEN THERE IS AMBIGUOUS
+  DATA.}
+
+\item[]{\bf {\em void} MR\_Set\_ID({\em MRegion\_ptr}
+    mregion, {\em int} id):} Explicitly set ID of a region and
+  overwrite the ID set by the {\bf MR\_New} function. Does not
+  check for duplication of region IDs.
+
+\item[]{\bf {\em void} MR\_Set\_Faces({\em MRegion\_ptr}
+    mregion, {\em int} nf, {\em MFace\_ptr} *mfaces,
+    {\em int} *dirs):} Set the faces of the region along with their
+  directions. The {\em unordered} set of faces and their directions
+  are passed in through arrays along with the number of faces. If the
+  normal of the face points out of the region, the associated
+  direction to be passed in is 1; otherwise it is 0. This function is
+  only relevant for full representations in MSTK.
+
+\item[]{\bf {\em void} MR\_Set\_Vertices({\em MRegion\_ptr}
+    mregion, {\em int} nv, {\em MVertex\_ptr} *mvertices, 
+    {\em int} nf, {\em int} **rfv\_template):} Set
+  the vertices of the region. This routine will collect/build all
+  lower order topological information (edges, edge directions in the
+  face) as needed by the face. For standard elements, the vertices
+  must be ordered as indicated in Appendix~\ref{app:reg_conventions},
+  while {\bf nf} can be 0 and {\bf rfv\_template} can be NULL.
+  For non-standard elements, {\bf nf} is the number of faces of the
+  polyhedron and {\bf rfv\_template} gives the vertices used in each
+  face. More precisely, {\bf rfv\_template} has {\bf nfv}
+  pointers to integer array $i$. The first entry of each
+  {\bf rfv\_template$[i]$} represents the number of vertices in that
+  face and the remaining entries represent the list of vertices of the
+  face listed so that the face defined by them points out of the region.
+
+\item[]{\bf {\em void} MR\_Replace\_Face({\em MRegion\_ptr}
+    mregion, {\em MFace\_ptr} mface, {\em MFace\_ptr} nuface,
+    {\em int} dir):} Replace a face of the region with another
+  face. The direction in which the new face is used in the region must
+  also be supplied. This function is only relevant for full
+  representations in MSTK.
+
+\item[]{\bf {\em void} MR\_Replace\_Vertex({\em MRegion\_ptr}
+    mregion, {\em MVertex\_ptr} mvertex, {\em MVertex\_ptr}
+    nuvertex):} Replace a vertex of a region with another vertex. This
+  function is relevant only for reduced representations in MSTK.
+
+\item[]{\bf {\em void}
+    MR\_Replace\_Face\_i({\em MRegion\_ptr} mregion, {\em int}
+    i, {\em MFace\_ptr} mface, {\em int} dir):} Replace the i'th
+  face in the region with another face.The direction in which the new
+  face is used in the region must also be supplied. This function is
+  only relevant for full representations in MSTK.
+
+\item[]{\bf {\em void}
+    MR\_Replace\_Vertex\_i({\em MRegion\_ptr} mregion, {\em int}
+    i, {\em MVertex\_ptr} mvertex):} Replace the i'th vertex of the
+  region with another vertex. This function is only relevant for
+  reduced representations in MSTK.
+
+\item[]
+
+\item[]{\bf {\em Mesh\_ptr} MR\_Mesh({\em MRegion\_ptr}
+    mregion):} Returns the mesh that the region belongs to.
+
+\item[]{\bf {\em int} MR\_ID({\em MRegion\_ptr} mregion):}
+  Returns the ID of the region. Returns 0 if not known.
+
+\item[]{\bf {\em int} MR\_GEntDim({\em MRegion\_ptr}
+    mregion):} Returns the dimension of the geometric model entity the
+  region is classified on. Always returns 3 since a mesh region can be
+  classified only on a model region.
+
+\item[]{\bf {\em int} MR\_GEntID({\em MRegion\_ptr}
+    mregion):} Returns the ID of the geometric model entity that the
+  region is classified on. Returns 0 if not known.
+
+\item[]{\bf {\em GEntity\_ptr} MR\_GEntity({\em MRegion\_ptr}
+    mregion):} Returns a pointer or handle to the geometric model
+  entity that the vertex is classified on. Returns NULL if this
+  information is not known.
+
+\item[]
+
+\item[]{\bf {\em int} MR\_Num\_Vertices({\em MRegion\_ptr}
+mregion):} Returns the number of vertices of a region.
+
+\item[]{\bf {\em int} MR\_Num\_Edges({\em MRegion\_ptr} mregion):}
+Returns the number of edges of a region.
+
+\item[]{\bf {\em int} MR\_Num\_Faces({\em MRegion\_ptr} mregion):}
+Returns the number of faces of a region.
+
+\item[]{\bf {\em int} MR\_Num\_AdjRegions({\em MRegion\_ptr}
+mregion):} Returns the number of adjacent regions of a region, i.e.,
+regions sharing a face with this region.
+
+\item[]
+
+\item[]{\bf {\em List\_ptr} MR\_Vertices({\em MRegion\_ptr}
+mregion):} Returns the set of vertices of a region. For standard
+elements the vertices are ordered as indicated in
+Appendix~\ref{app:reg_conventions}. For non-standard elements the set
+is unordered.
+
+\item[]{\bf {\em List\_ptr} MR\_Edges({\em MRegion\_ptr} mregion):}
+Return the \underline{unordered} set of edges of a region.
+
+\item[]{\bf {\em List\_ptr} MR\_Faces({\em MRegion\_ptr} mregion):}
+Returns the set of faces of a region.
+
+\item[]{\bf {\em List\_ptr} MR\_AdjRegions({\em MRegion\_ptr}
+mregion):} Returns the set of adjacent regions of a region, i.e.,
+regions sharing a face with this region. The set is not ordered.
+
+\item[]{\bf {\em int} MR\_FaceDir({\em MRegion\_ptr} mregion,
+{\em MFace\_ptr} mface):} Returns the direction in which the region
+uses the given face. Returns 1 if the face normal points out of the
+region and returns 0 if the face normal points into the region.
+
+\item[]{\bf {\em int} MR\_FaceDir\_i({\em MRegion\_ptr} mregion,
+{\em int} i):} Returns the direction in which the region uses the
+i'th face. Returns 1 if the face normal points out of the region and
+returns 0 if the face normal points into the region.
+
+\item[]
+
+\item[]{\bf {\em int} MR\_UsesEntity({\em MRegion\_ptr} mregion,
+{\em MEntity\_ptr} ment, {\em int} type):} Check if the region
+uses the given lower dimension entity, 'mentity'. The type of the
+entity is
+
+\item[]
+
+\item[]{\bf {\em void} MR\_Coords({\em MRegion\_ptr} mregion,
+{\em int} *n, double (*xyz)[3]):} Returns the coordinates of the
+region vertices in an array along with the number of vertices. For
+standard elements, the ordering is as given in
+Appendix~\ref{app:reg_conventions}. For non-standard elements, the
+ordering is arbitrary.
+\end{description}
+
+\newpage
+\subsection{Generic Entity Object}
+
+The following functions operate on generic mesh entities of type
+{\em MEntity\_ptr}. This implies that variables of type
+{\em MVertex\_ptr}, {\em MEdge\_ptr}, {\em MFace\_ptr},
+{\em MRegion\_ptr} can all be passed in place of
+{\em MEntity\_ptr} variables in the following functions.
+
+\begin{description}
+%%  {\em void} MEnt\_Set\_GEntity({\em MEntity\_ptr} mentity, GEntity\_ptr gent);
+%%
+%%  {\em void} MEnt\_Set\_GEntDim({\em MEntity\_ptr} mentity, {\em int} gdim);
+%%
+%%  {\em void} MEnt\_Set\_ID({\em MEntity\_ptr} mentity, {\em int} id);
+
+\item[]{\bf {\em int} MEnt\_ID({\em MEntity\_ptr} mentity):} Returns
+the ID of a generic entity.
+
+\item[]{\bf {\em int} MEnt\_Dim({\em MEntity\_ptr} mentity):}
+Returns the topological dimension or type of generic entity.
+
+\item[]{\bf {\em int} MEnt\_OrigDim({\em MEntity\_ptr} mentity):}
+Returns the original topological dimension or type of generic entity before it
+was temporarily deleted.
+
+\item[]{\bf {\em int} MEnt\_IsVolatile({\em MEntity\_ptr}) mentity):}
+  Is the entity a temporary (volatile) entity in a reduced
+  representation or an explicitly represented persistent entity. For
+  example, this query will return 1 on any edge in a reduced
+  representation.
+
+\item[]{\bf {\em Mesh\_ptr} MEnt\_Mesh({\em MEntity\_ptr} mentity):}
+Returns the mesh that the entity belongs to.
+
+\item[]{\bf {\em int} MEnt\_GEntDim({\em MEntity\_ptr}
+    mentity):} Returns the dimension of the geometric model entity
+  that the entity is classified on.
+
+\item[]{\bf {\em GEntity\_ptr}
+    MEnt\_GEntity({\em MEntity\_ptr} mentity):} Returns a pointer
+  or handle to geometric model entity that the entity is classified
+  on.
+
+\item[]{\bf {\em void} MEnt\_Delete({\em MEntity\_ptr} mentity, {\em
+      int} keep):} Delete a generic mesh entity. If {\bf keep} is 0,
+  the entity is removed from the mesh and all topological and
+  geometric information embedded in the entity is destroyed. If {\bf
+    keep} is 1, the region is marked as deleted and removed from the
+  mesh but the entity is not destroyed. Also, lower order entities in
+  the mesh no longer point up to this entity.
+
+\item[]
+
+\item[]{\bf {\em void} MEnt\_Set\_AttVal({\em MEntity\_ptr}
+    ent, {\em MAttrib\_ptr} attrib, {\em int} ival,
+    {\em double} lval, {\em void} *pval):} Set the value of an
+  attribute for given mesh entity. Depending on whether the attribute
+  is integer, double or pointer, the appropriate entry in the argument
+  list is specified.
+
+\item[]{\bf {\em void} MEnt\_Rem\_AttVal({\em MEntity\_ptr}
+    ent, {\em MAttrib\_ptr mattrib}:} Clear the value of the given
+  attribute from the entity.
+
+\item[]{\bf {\em int} MEnt\_Get\_AttVal({\em MEntity\_ptr}
+    ent, {\em MAttrib\_ptr} int *ival, double *rval, double
+    **pval):} Get the value of an attribute from an entity.  Depending
+  on the attribute type, either {\bf ival}, {\bf rval} or
+  {\bf pval} is returned.
+
+\item[]{\bf {\em void} MEnt\_Rem\_AllAttVals({\em MEntity\_ptr}):}
+  Remove all attribute values attached to an entity.
+
+\end{description}
+
+\newpage
+\subsection{Mesh Attributes}
+
+\begin{description}
+
+\item[]{\bf {\em MAttrib\_ptr} MAttrib\_New({\em Mesh\_ptr}
+    mesh, {\em const char} *att\_name, {\em MAttType} att\_type,
+    {\em MType} entdim):} Define a new mesh attribute. Along with
+  the mesh to which this attribute is assigned ({\bf mesh}) and the
+  name of the attribute ({\bf att\_name}), the type of the
+  attribute ({\bf att\_type}) must be specified as {\bf INT},
+  {\bf DOUBLE} or {\bf POINTER}. Also, the dimension of the entity
+    for which the attribute value can be set must be specified as
+    {\bf MVERTEX}, {\bf MEDGE}, {\bf MFACE}, {\bf MREGION}
+    or {\bf MALLTYPE}.
+
+\item[]{\bf {\em char} *MAttrib\_Get\_Name({\em MAttrib\_ptr}
+    attrib, {\em char} *att\_name):} Get the name of the given attribute.
+
+\item[]{\bf {\em MAttType}
+    MAttrib\_Get\_Type({\em MAttrib\_ptr} attrib):} Get the type of
+  the attribute (Can return {\bf INT}, {\bf DOUBLE},
+  {\bf PVAL}).
+
+\item[]{\bf {\em MType}
+    MAttrib\_Get\_EntDim({\em MAttrib\_ptr} attrib):} Get the
+  dimension (or type) of entity attribute can be asssigned to. Can be
+  {\bf MVERTEX}, {\bf MEDGE}, {\bf MFACE}, {\bf MREGION}
+  or {\bf MALLTYPE}.
+
+\item[]{\bf {\em void} MAttrib\_Delete({\em MAttrib\_ptr}
+    attrib):} Delete an attribute.
+
+\end{description}
+
+\newpage
+\subsection{Entity Marks}
+
+Entity marks or markers are a way of tagging entities. Such
+functionality is useful in algorithms which must keep track of
+processed entities to avoid duplication of work. An example of such an
+operation is creating a union of entity sets while extracting upward
+adjacency information such as the regions connected to an edge. Use of
+entity marks avoids calling {\bf List\_ChknAdd} (search through the
+list and add item if it is not there) or using attributes to tag
+entities with 0 or 1. As a result it is much more efficient than
+either option.
+
+\begin{description}
+\item[]{\bf {\em int} MSTK\_GetMarker():} Returns a unique marker ID
+which may be used to tag entities.
+
+\item[]{\bf {\em void} MEnt\_Mark({\em MEntity\_ptr} ent, {\em int} mkr):} Mark an entity with the given marker 'mkr'.
+
+\item[]{\bf {\em int} MEnt\_IsMarked({\em MEntity\_ptr} ent, {\em int} mkr):} Check if an entity is marked with the given marker 'mkr'.
+
+\item[]{\bf {\em void} MEnt\_Unmark({\em MEntity\_ptr} ent, {\em int} mkr):} Unmark an entity with respect to the given marker 'mkr'
+
+\item[]{\bf {\em void} List\_Mark({\em List\_ptr} list, {\em int} mkr):} Mark a set of entities with given marker.
+
+\item[]{\bf {\em void} List\_Unmark({\em List\_ptr} list, {\em int} mkr):}
+Unmark a set of entities with respect to the given marker.
+
+\item[]{\bf {\em void} MSTK\_FreeMarker({\em int} mkr):} Release the
+marker ID given by {\bf MSTK\_GetMarker()} so that it can be
+reused. Care must be taken to unmark all entities marked with this
+marker ID before releasing it. If not, subsequent operations with
+reassigned marker will find a tag on some entities and mistake them
+for being processed.
+\end{description}
+
+An example use of entity marks is given below:
+
+\newpage
+\begin{verbatim}
+
+/* Code to get the faces of a vertex */
+/* NOTE: This code is for illustration of entity marks only. */
+/*       This functionality is already available through MV_Faces */
+
+vedges = MV_Edges(v);      /* Edges of vertex */
+
+mkid = MSTK_GetMarker();   /* Get a new marker */
+vfaces = List_New(10);     /* Initialize list of faces cncted to v */
+
+idx = 0;
+while ((ve = List_Next_Entry(vedges,&idx))) {   /* For each edge */
+
+  efaces = ME_Faces(ve);                /* Get faces of edge */
+  idx1 = 0;
+  while ((ef = List_Next_Entry(efaces,&idx1))) {
+    if (!MEnt_IsMarked(ef,mkid)) {         /* Is the face already in list? */
+      List_Add(vfaces,ef);                 /* No? Add it to the list and   */
+      MEnt_Mark(ef,mkid);                  /* mark it as being in the list */
+    }
+  }
+  List_Delete(efaces);
+
+}
+
+List_Delete(vedges);
+
+List_Unmark(vfaces,mkid);                 /* Unmark all the marked faces */
+MSTK_FreeMarker(mkid);                    /* Very important to free the marker!! */
+                                          /* otherwise, we'll run out of markers */
+\end{verbatim}
+
+\newpage
+\subsection{Mesh Modification}
+
+\begin{description}
+
+\item[]{\bf {\em int} ME\_Swap2D({\em MEdge\_ptr} e,
+{\em MEdge\_ptr} *enew, {\em MFace\_ptr} fnew[2]):} Swap an edge
+in a triangular mesh. No checks are performed for topological or
+geometric validity.
+
+\item[]{\bf {\em MVertex\_ptr} MVs\_Merge({\em MVertex\_ptr} v1, {\em
+      MVertex\_ptr} v2):} Merge two vertices, v1 and v2, and return
+  the retained vertex. By default, v1 is retained.
+
+\item[]{\bf {\em MFace\_ptr} MFs\_Join({\em MFace\_ptr} f1,
+{\em MFace\_ptr} f2, {\em MEdge\_ptr} e):} Join two faces along
+common edge and create new face by eliminating the common edge as
+shown in Figure~\ref{fig:MFs_Join}. If 'f1' has 'n1' edges and 'f2'
+has 'n2' edges, then the new face has ('n1'+'n2'-2) edges.
+
+\end{description}
+
+\begin{figure}[!h]
+  \begin{center}
+    \includegraphics[scale=1.0]{figures/MFs_Join}
+
+    \caption{Joining two faces (a) Two faces $F_1$ and $F_2$ sharing a
+      common edge (b) New pentagonal face $F_{new}$ created by
+      eliminating the common edge.  }
+    \label{fig:MFs_Join}
+  \end{center}
+\end{figure}
+
+\newpage
+\subsection{Utilities}
+
+\begin{description}
+
+\item[]{\bf {\em void} MSTK\_Report({\em char} *module,
+  {\em char} *message, {\em ErrType} severity):} Error handler
+  for MSTK. 'module' is the name of the function in which the error
+  occurs. 'message' is the error message and is recommended to be less
+  than 1024 characters in length. 'severity' is an error code and can
+  be MESSG, WARN, ERROR or FATAL. If the error code is FATAL, the
+  program will quit after printing the error. If the same message is
+  repeated successively, then the message is printed only the first
+  time.
+
+\item[]{\bf {\em void} List\_PrintID({\em List\_ptr} l):} Debugging
+utility to print the IDs of the entities in a set.
+
+\item[]{\bf {\em void} MV\_Print({\em MVertex\_ptr} v, int
+    lev):} Debugging utility to print information about a mesh vertex,
+  {\bf v}. The argument {\bf lev} controls the level of detail
+  of the information printed. {\bf lev} = 0 prints the minimum
+  information, i.e., vertex pointer, its ID and its coordinates. If
+  {\bf lev} = 1, the function prints classification information for
+  the vertex (if available), i.e., ID and dimension of the model
+  entity that the vertex is on. If {\bf lev} $>$ 1, then upward
+  detailed adjacency information is also printed for the vertex, i.e.,
+  information is printed about the edges, faces and regions connected
+  to the vertex.
+
+\item[]{\bf {\em void} ME\_Print({\em MEdge\_ptr} e, int
+    lev):} Debugging utility to print information about a mesh edge,
+  {\bf e}. The argument {\bf lev} controls the level of detail
+  of the information printed. {\bf lev} = 0 prints the minimum
+  information, i.e., edge pointer, its ID and the IDs of its two
+  vertices. If {\bf lev} = 1, the function prints classification
+  information for the edge (if available), i.e., ID and dimension of
+  the model entity that the edge is on. Also, more detailed vertex
+  information printed in this case. If {\bf lev} $>$ 1, the
+  function prints detailed upward adjacency information for the edge,
+  i.e., information is printed about the faces and regions connected
+  to the edge.
+
+  \item[]{\bf {\em void} MF\_Print({\em MFace\_ptr} f, int
+      lev):} Debugging utility to print information about a mesh face,
+    {\bf f}. The argument {\bf lev} controls the level of detail
+    of the information printed. {\bf lev} = 0 prints the minimum
+    information, i.e., the face pointer and its ID. If {\bf lev} =
+    1, the function prints classification information for the edge (if
+    available), i.e., ID and dimension of the model entity that the
+    face is on. Also, a signed list of the edges of the face is
+    printed.  If {\bf lev} $>$ 1, the function prints detailed
+    downward and upward adjacency information for the face, i.e.,
+    information is printed about the edges and vertices of the face,
+    and about the regions connected to the face.
+
+  \item[]{\bf {\em void} MR\_Print({\em MRegion\_ptr} r, int
+      lev):} Debugging utility to print information about a mesh
+    region, {\bf r}. The argument {\bf lev} controls the level
+    of detail of the information printed. {\bf lev} = 0 prints the
+    minimum information, i.e., region pointer and its ID. If
+    {\bf lev} = 1, the function prints classification information
+    for the region (if available), i.e., ID of the model entity that
+    the region is on. Also, a signed list of the faces of the region
+    is printed. If {\bf lev} $>$ 1, the function prints detailed
+    downward adjacency information for the region, i.e., information
+    is printed about the faces, edges and vertices forming the region.
+
+\end{description}
+
+\newpage
+\appendix
+
 \section{Conventions for Vertex, Edge Numbering in Standard Region Types}
@@
 \begin{verse}
-\# \textit{This is a comment} \\
-\# \textit{The string ``MSTK'' and File version number (1.0)} \\
+\# {\em This is a comment} \\
+\# {\em The string ``MSTK'' and File version number (1.0)} \\
 \vspace{1ex}
@@
 
-\# \textit{char *reptype - Type of representation} \\  
-\# \textit{\textit{int} NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
+\# {\em char *reptype - Type of representation} \\  
+\# {\em {\em int} NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
 \vspace{1ex}
@@
 
-\# \textit{\textbf{VERTEX INFO}} \\
-\# \textit{Each record has} \\
-\# \textit{double X,Y,Z Coordinates} \\
-\# \textit{\textit{int} Mdim - Topological type or dimension of model entity that} \\
-\# \textit{\hspace{4.7em} the vertex is on} \\
-\# \textit{\textit{int} Mid - ID of model entity the vertex is on} \\
-\# \textit{Mdim and Mid can be -1 and 0 resp. if model info. is absent} 
+\# {\em {\bf VERTEX INFO}} \\
+\# {\em Each record has} \\
+\# {\em double X,Y,Z Coordinates} \\
+\# {\em {\em int} Mdim - Topological type or dimension of model entity that} \\
+\# {\em \hspace{4.7em} the vertex is on} \\
+\# {\em {\em int} Mid - ID of model entity the vertex is on} \\
+\# {\em Mdim and Mid can be -1 and 0 resp. if model info. is absent} 
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NV times}
+\# {\em Repeated NV times}
 \vspace{2ex}
 
-\# \textit{\textbf{EDGE INFO} - present only if NE $\ne$ 0} \\
-\# \textit{Keyword 'edges' followed by edge records}
-\# \textit{Each edge record has} \\
-\# \textit{\textit{int} Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
-\# \textit{\textit{int} Mdim, Mid}
+\# {\em {\bf EDGE INFO} - present only if NE $\ne$ 0} \\
+\# {\em Keyword 'edges' followed by edge records}
+\# {\em Each edge record has} \\
+\# {\em {\em int} Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
+\# {\em {\em int} Mdim, Mid}
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NEdges times}
+\# {\em Repeated NEdges times}
 \vspace{2ex}
@@
 \newpage
-\# \textit{\textbf{FACE INFO} - present only if NF $\ne$ 0} \\
-\# \textit{Keyword 'faces'}
-\# \textit{char *FLtype: Keyword for lower order entity describing faces} \\
-\# \textit{Values: Vertex, Edge (case insensitive), e.g. VeRteX or EDGE} \\
+\# {\em {\bf FACE INFO} - present only if NF $\ne$ 0} \\
+\# {\em Keyword 'faces'}
+\# {\em char *FLtype: Keyword for lower order entity describing faces} \\
+\# {\em Values: Vertex, Edge (case insensitive), e.g. VeRteX or EDGE} \\
 \vspace{1ex}
@@
 
-\# \textit{If face described by vertices, then each face record has} \\
-\# \textit{\textit{int} NFV - Number of face vertices} \\
-\# \textit{\textit{int} Vid\_1 - ID of first vertex of face} \\
-\# \textit{\textit{int} Vid\_2 - ID of second vertex of face} \\
+\# {\em If face described by vertices, then each face record has} \\
+\# {\em {\em int} NFV - Number of face vertices} \\
+\# {\em {\em int} Vid\_1 - ID of first vertex of face} \\
+\# {\em {\em int} Vid\_2 - ID of second vertex of face} \\
 \# . . . \\
-\# \textit{\textit{int} Vid\_1 - ID of NFV'th vertex of face} \\
-\# \textit{\textit{int} Mdim, Mid} 
+\# {\em {\em int} Vid\_1 - ID of NFV'th vertex of face} \\
+\# {\em {\em int} Mdim, Mid} 
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NFaces times}
+\# {\em Repeated NFaces times}
 \vspace{2ex}\vspace{1ex}
 
-\# \textit{If face described by edges, then each face record has} \\
-\# \textit{\textit{int} NFE - Number of face edges} \\
-\# \textit{\textit{int} $\pm$Eid\_1 - signed ID of first edge of face} \\
-\# \textit{\textit{int} $\pm$Eid\_2 - signed ID of second edge of face} \\
+\# {\em If face described by edges, then each face record has} \\
+\# {\em {\em int} NFE - Number of face edges} \\
+\# {\em {\em int} $\pm$Eid\_1 - signed ID of first edge of face} \\
+\# {\em {\em int} $\pm$Eid\_2 - signed ID of second edge of face} \\
 \# . . . \\
-\# \textit{\textit{int} $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
-\# \textit{\textit{int} Mdim, Mid} \\
+\# {\em {\em int} $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
+\# {\em {\em int} Mdim, Mid} \\
 \# \\
-\# \textit{if sign of edge is +, face uses edge in direction it was defined} \\
-\# \textit{if sign of edge is -, face uses edge in opposite direction} \\
+\# {\em if sign of edge is +, face uses edge in direction it was defined} \\
+\# {\em if sign of edge is -, face uses edge in opposite direction} \\
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NFaces times}
+\# {\em Repeated NFaces times}
 \vspace{2ex}\vspace{1ex}
 
-\# \textit{\textbf{REGION INFO} - present only if NR $\ne$ 0} \\
-\# \textit{Keyword 'regions'}
-\# \textit{char *RLtype - keyword for lower order entity describing region} \\
-\# \textit{Values: Vertex, Face (case insensitive), e.g. VERtex or faCE}
+\# {\em {\bf REGION INFO} - present only if NR $\ne$ 0} \\
+\# {\em Keyword 'regions'}
+\# {\em char *RLtype - keyword for lower order entity describing region} \\
+\# {\em Values: Vertex, Face (case insensitive), e.g. VERtex or faCE}
 \vspace{1ex}
@@
 
-\# \textit{if region described by vertices, then each region record has} \\
-\# \textit{\textit{int} NRV - Number of region vertices} \\
-\# \textit{\textit{int} Vid\_1 - ID of first vertex of region} \\
-\# \textit{\textit{int} Vid\_2 - ID of second vertex of region} \\
+\# {\em if region described by vertices, then each region record has} \\
+\# {\em {\em int} NRV - Number of region vertices} \\
+\# {\em {\em int} Vid\_1 - ID of first vertex of region} \\
+\# {\em {\em int} Vid\_2 - ID of second vertex of region} \\
 \# . . . \\
-\# \textit{\textit{int} Vid\_NFE - ID of NRV'th vertex of region} \\
-\# \textit{\textit{int} Mid, (NOTE: Mdim is not specified, since it has to be 3)}
+\# {\em {\em int} Vid\_NFE - ID of NRV'th vertex of region} \\
+\# {\em {\em int} Mid, (NOTE: Mdim is not specified, since it has to be 3)}
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeat NR times} 
+\# {\em Repeat NR times} 
 \vspace{3ex}
 
-\# \textit{if region described by faces, then each region record has} \\
-\# \textit{\textit{int} NRF - Number of region faces} \\
-\# \textit{\textit{int} Fid\_1 - signed ID of first face of region} \\
-\# \textit{\textit{int} Fid\_2 - signed ID of second face of region} \\
+\# {\em if region described by faces, then each region record has} \\
+\# {\em {\em int} NRF - Number of region faces} \\
+\# {\em {\em int} Fid\_1 - signed ID of first face of region} \\
+\# {\em {\em int} Fid\_2 - signed ID of second face of region} \\
 \# . . . \\
-\# \textit{\textit{int} Fid\_NRF - signed ID of NRF'th face of region} \\
-\# \textit{\textit{int} Mdim, Mid} \\
+\# {\em {\em int} Fid\_NRF - signed ID of NRF'th face of region} \\
+\# {\em {\em int} Mdim, Mid} \\
 \#  \\
-\# \textit{if sign of face is +, face normal points out of region} \\
-\# \textit{if sign of edge is -, face normal points into region}
+\# {\em if sign of face is +, face normal points out of region} \\
+\# {\em if sign of edge is -, face normal points into region}
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NR times} 
+\# {\em Repeated NR times} 
 \vspace{2ex}
 
-\# \textit{\textbf{NOT IMPLEMENTED}}
-
-\# \textit{\textbf{VERTEX ATTRIBUTES}} \\
-\# \textit{\textit{int} NVA - Number of Vertex attributes} \\
+\# {\em {\bf NOT IMPLEMENTED}}
+
+\# {\em {\bf VERTEX ATTRIBUTES}} \\
+\# {\em {\em int} NVA - Number of Vertex attributes} \\
 \# \\
-\# \textit{char *VA\_name\_1 - Name of first vertex attribute} \\
-\# \textit{\textit{int} VA\_type\_1 - Type of first vertex attribute} \\
-\# \textit{\textit{int} VA\_dim\_1 - Dimension of first vertex attribute} \\
+\# {\em char *VA\_name\_1 - Name of first vertex attribute} \\
+\# {\em {\em int} VA\_type\_1 - Type of first vertex attribute} \\
+\# {\em {\em int} VA\_dim\_1 - Dimension of first vertex attribute} \\
 \# \\
-\# \textit{char *VA\_name\_2 - Name of second vertex attribute} \\
-\# \textit{\textit{int} VA\_type\_2 - Type of second vertex attribute} \\
-\# \textit{\textit{int} VA\_dim\_2 - Dimension of first vertex attribute} \\
+\# {\em char *VA\_name\_2 - Name of second vertex attribute} \\
+\# {\em {\em int} VA\_type\_2 - Type of second vertex attribute} \\
+\# {\em {\em int} VA\_dim\_2 - Dimension of first vertex attribute} \\
 \# \\
@@
 \# \\
-\# \textit{char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
-\# \textit{\textit{int} VA\_type\_NVA - Type of NVA'th vertex attribute} \\
-\# \textit{\textit{int} VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
+\# {\em char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
+\# {\em {\em int} VA\_type\_NVA - Type of NVA'th vertex attribute} \\
+\# {\em {\em int} VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
 \# \\
-\# \textit{VA\_type can be 1 (int), 2 (double), 3 (string)} \\
-\# \textit{VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
-\# \textit{VA\_dim can only be 1 when VA\_type is string}
+\# {\em VA\_type can be 1 (int), 2 (double), 3 (string)} \\
+\# {\em VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
+\# {\em VA\_dim can only be 1 when VA\_type is string}
 \vspace{1ex}
@@
 
-\# \textit{For each vertex attribute record, set of attribute values} \\
-\# \textit{E.G., there are 3 attributes for each vertex:} \\
-\# \textit{a scalar int, a vector of 3 doubles and a string}
-
-VA\_\textit{int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
-VA\_\textit{int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
+\# {\em For each vertex attribute record, set of attribute values} \\
+\# {\em E.G., there are 3 attributes for each vertex:} \\
+\# {\em a scalar int, a vector of 3 doubles and a string}
+
+VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
+VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
 . \\
@@
 . \\
-\# \textit{Repeated NV times}
+\# {\em Repeated NV times}
 \vspace{2ex}
@@
 \newpage
-\# \textit{\textbf{EDGE ATTRIBUTES}} \\
-\# \textit{Similar to Vertex attribute description}
+\# {\em {\bf EDGE ATTRIBUTES}} \\
+\# {\em Similar to Vertex attribute description}
 \vspace{1ex}
@@
 
-\# \textit{For each edge attribute record, set of attribute values} \\
-\# \textit{E.G., a scalar int, a scalar double and a string}
-\vspace{1ex}
-
-EA\_\textit{int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
-EA\_\textit{int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
+\# {\em For each edge attribute record, set of attribute values} \\
+\# {\em E.G., a scalar int, a scalar double and a string}
+\vspace{1ex}
+
+EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
+EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
 . . . \\
-\# \textit{Repeated NE times}
+\# {\em Repeated NE times}
 \vspace{2ex}
 
-\# \textit{\textbf{FACE ATTRIBUTES}} \\
-\# \textit{Similar to Vertex attribute description}
+\# {\em {\bf FACE ATTRIBUTES}} \\
+\# {\em Similar to Vertex attribute description}
 \vspace{1ex}
@@
 
-\# \textit{For each face attribute record, set of attribute values} \\
-\# \textit{E.G., a vector of 2 doubles and a string}
+\# {\em For each face attribute record, set of attribute values} \\
+\# {\em E.G., a vector of 2 doubles and a string}
 \vspace{1ex}
@@
 . . . \\
-\# \textit{Repeated NF times}
+\# {\em Repeated NF times}
 \vspace{2ex}
@@
 \newpage
-\# \textit{\textbf{REGION ATTRIBUTES}} \\
-\# \textit{Similar to Vertex attribute description}
+\# {\em {\bf REGION ATTRIBUTES}} \\
+\# {\em Similar to Vertex attribute description}
 \vspace{1ex}
@@
 
-\# \textit{For each Region attribute record, set of attribute values} \\
-\# \textit{E.G., a vector of 3 ints}
+\# {\em For each Region attribute record, set of attribute values} \\
+\# {\em E.G., a vector of 3 ints}
 \vspace{1ex}
@@
 . . .\\
-\# \textit{Repeated NR times}
+\# {\em Repeated NR times}
 \vspace{2ex}
@@
  strcat(meshname,"-copy.mstk");
- MESH_WriteToFile(mesh,meshname);
+ MESH_WriteToFile(mesh,meshname,F1);
 

### 2010-02-25 docs/MSTK.tex
Updated!
Added section on volatile entities and description for related routines

@@
   HAS GENERALLY BEEN STABLE FOR SEVERAL YEARS. THE REDUCED
-  REPRESENTATION HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT IS CURRENTLY
-  BROKEN. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
+  REPRESENTATIONS HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT MAY HAVE
+  SOME BUGS. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
   PLEASE REPORT ANY PROBLEMS IMMEDIATELY TO \texttt{rao@lanl.gov}.  }
@@
 
+\subsection{Volatile mesh entities in reduced representations}
+
+As discussed above reduced representations, do not explicitly
+represent all types of mesh entities. For example, R1 representations
+do not explicitly represent edges and faces while R4 representations
+do explicitly represent edges only. However, whenever an implicit
+entity is requested from MSTK, the software creates a temporary entity
+so that it appears that the entity actually exists in the
+database. Since these entities are created-on-the-fly they are called
+{\em volatile} mesh entities. For example, if an application asked for
+the faces of a region in an R1 representation, MSTK will create as
+many volatile faces as necessary, put them in a list and return them
+to the calling application. Thus the application can pretend to work
+with a full representation (although it is not always efficient to do so).
+
+Volatile entities are stored for a period of time in internal data
+structures in MSTK. Whenever there is a request for a new volatile
+mesh entity, the code first checks if this entity has already been
+created. If the volatile already exists in the database, then the
+entity is returned as is and if it does not, it is created. This
+ensures that many copies of the same entity are not created in
+MSTK. However, to ensure that MSTK also does not store every volatile
+entity forever (thereby, recreating a full representation), there is a
+mechanism to perform garbage cleaning on volatile
+entities. Periodically through the execution of the code, volatile
+entities that have not been used for a long time are deleted from the
+database freeing up valuable memory (which was the main point of using
+a reduced representation after all).
+
+To ensure that garbage cleaning does not delete a volatile mesh entity
+that is being processed or stored by a calling application, MSTK also
+provides a mechanism for locking entities. One can specify an autolock
+mechanism for the entire mesh which means that no volatile entity will
+ever be deleted once it is created. On the other hand, applications
+may lock specific entities and unlock them when they cease to be
+useful.
+
 \subsection{Mesh entity classification}
@@
   what you are doing.
+
+\item[]{\bf {\em void} MESH\_Set\_AutoLock({\em Mesh\_ptr} mesh, {\em
+      int} autolock):} If autolock is 1, all volatile mesh entities
+  created in reduced representations will be automatically locked and
+  cannot be removed by garbage cleaning procedures (They can still be
+  removed by explicitly calling a delete on them). If it 0, volatile
+  mesh entities can be deleted internally after a period of disuse.
+
+\item[]{\bf {\em int} MESH\_AutoLock({\em Mesh\_ptr} mesh):} Return
+the autolock status for volatile entities (See {\bf
+  MESH\_Set\_AutoLock} for details).
+
 \end{description}
@@
   to be the same.
+
+\item[]
+
+\item[]{\bf {\em void} ME\_Lock({\em MEdge\_ptr} e):} Lock a volatile
+  edge so that it cannot be deleted by garbage cleaning procedures (It
+  can still be deleted by {\bf ME\_Delete}.
+
+\item[]{\bf {\em void} ME\_UnLock({\em MEdge\_ptr} e):} Unlock a volatile
+  edge so that it may be deleted freely by garbage cleaning procedures.
 
@@
 % returned with a cw ordering (See Figure~\ref{fig:face_def}).
+
+\item[]
+
+\item[]{\bf {\em void} MF\_Lock({\em MFace\_ptr} f):} Lock a volatile
+  face so that it cannot be deleted by garbage cleaning procedures (It
+  can still be deleted by {\bf MF\_Delete}.
+
+\item[]{\bf {\em void} MF\_UnLock({\em MFace\_ptr} f):} Unlock a volatile
+  face so that it may be deleted freely by garbage cleaning procedures.
+
 \end{description}

### 2010-03-01 docs/MSTK.tex
Added section about handling physics variables and boundary conditions

@@
 oriented manner. Mesh entities, mesh attributes and entity lists are
-considered to be objects which have their private data and a set of
-operators to query and manipulate this data. Even though methods may
-be devised to circumvent the data hiding in MSTK, this is highly
-discouraged. If you think that a particular operator is not efficient
-or you don'thave access to some mesh data, please contact the
-developer for assistance.
+considered to be like objects which have their private data and a set
+of operators to query and manipulate this data. Even though methods
+may be devised to circumvent the data hiding in MSTK, since it is
+written in C, this is highly discouraged. If you think that a
+particular operator is not efficient or you don'thave access to some
+mesh data, please contact the MSTK developer for assistance.
 
@@
 planned for the near future.
-
-\subsection{Getting meshes in and out of MSTK}
-
-There are a few ways of getting mesh data into MSTK. The first is to
-prepare an MSTK file with the full connectivity of an F1
-representation and to read it in using MESH\_InitFromFile. This is
-typically a difficult task for many applications which only have
-element-node data. In such cases, one can prepare an MSTK file in the
-R1 file format (in which the node listing is followed by the
-region-node connectivity) and read it in to any of the other
-representations, say F1. The other possibility is to write a file in
-the GMV format and use MESH\_ImportFromFile to import the mesh in.
-
-MSTK can write meshes out to different formats including GMV, STL and
-FLAG X3D. When compiled with the appropriate options and using a
-partitioner from an external library, MSTK can also write out a
-parallel X3D file.
 
@@
 \appendix
+
+\section{Getting meshes in and out of MSTK}
+
+There are a few ways of getting mesh data into MSTK. The first is to
+prepare an MSTK file with the full connectivity of an F1
+representation and to read it in using MESH\_InitFromFile. This is
+typically a difficult task for many applications which only have
+element-node data. In such cases, one can prepare an MSTK file in the
+R1 file format (in which the node listing is followed by the
+region-node connectivity) and read it in to any of the other
+representations, say F1. The other possibility is to write a file in
+the GMV format and use MESH\_ImportFromFile to import the mesh in.
+
+MSTK can write meshes out to different formats including GMV, STL and
+FLAG X3D. When compiled with the appropriate options and using a
+partitioner from an external library, MSTK can also write out a
+parallel X3D file.
+
+\section{Using MSTK in a numerical simulation code}
+
+In addition to mesh generation and mesh manipulation algorithms, MSTK
+can be used to write mesh-based numerical simulation codes such as
+Finite element, finite volume and finite difference methods. MSTK is
+particularly suited to developing methods which operate on general
+unstructured meshes. There is some overhead to using operators to
+access mesh data instead of directly accessing arrays; however, we
+believe that the ease of managing complex mesh data through an API
+outweighs this cost. Without a framework like MSTK, applications have
+to devise their own unstructured mesh data structures and the code is
+littered with direct accesses to these data structures (often
+complicated array-based data structures). This makes even small
+changes to the data structure very onerous because thousands to tens
+of thousands of lines of code are affected. In the end, individual
+developers create adhoc patches (often duplicated in different ways)
+to overcome any deficiencies in the data structures and eventually the
+code becomes unmanageable. We believe this is poor software
+engineering.
+
+\subsection{Simulation Variables}
+
+When using MSTK in a physics based code, the question of storing
+physics variables (node velocities, element
+pressures) is an important one.
+
+There are two ways of storing physics in combination with MSTK. The
+first is to use the attributes feature in MSTK. Scalar variables can
+be stored with entities as double precision values of mesh attributes;
+vector and tensor variables can be stored as pointer variables. When
+storing vectors as pointers, it is more efficient to allocate a large
+array to store the vectors for all entities and store pointers into
+that large array with the individual entities. This way the
+application does not allocate a huge number of small arrays. This
+method is very convenient as the variables are readily available using
+the handle of the entity, i.e., by calling {\bf MEnt\_Get\_AttVal}.
+However, this method could be somewhat inefficient and also requires a
+paradigm shift in how codes are written.
+
+The other way of managing physics variables is to directly allocate
+linear arrays for each variable and using the entity ID to reference
+entries in this array. For example, one might have a linear array {\bf
+  rho} for the densities of all elements and we use {\bf
+  rho[MR\_ID(r)-1]} to find the density of a particular element {\bf
+  r}. The ID of elements is obtained quite easily and stays static
+unless {\bf MESH\_Renumber} is called explicitly. If there is mesh
+modification, there will be intermediate parts of this array which
+will no longer be relevant but this is easily managed. This method is
+closer to traditional approaches and can be adopted quite easily. It
+could be more efficient than using attributes but is less
+``object-oriented.''
+
+\subsection{Boundary Conditions, Material Properties}
+
+Applying boundary conditions in an MSTK-based simulation code can be
+much easier than in a ``traditional'' simulation code, {\em if} the
+mesh carries the right classification information. In a
+``traditional'' simulation code, boundary conditions are specified
+using one of two ways:
+
+\begin{itemize}
+\item An implicit geometric entity is defined and the boundary
+  condition is specified on that geometry, e.g., normal velocity is 1.0
+  for all nodes on the plane [1.35 2.7695 1]. This requires one to check
+  the coordinates of all nodes to see if they lie on this plane and
+  apply the boundary condition on those nodes.
+
+\item Entity sets like node sets or face sets are defined as a
+  collection of entity IDs on which a particular boundary conditions
+  are applied. The user has pre-selected a set of nodes and applied
+  the boundary condition one them.
+\end{itemize}
+
+MSTK, on the other hand, provides for an easier way to specify
+boundary conditions, initial conditions and material properties
+provided the mesh has classification information. For example, if we
+knew that model face 51 has a pressure boundary condition on it, we
+only need to find all faces that are classified on model face 51
+(MF\_GEntDim(f)=2, MF\_GEntID(f)=51) and specify the pressure on that
+face in the code. The input file for the boundary conditions does not
+have to explicitly list any face sets nor do we have to use error
+prone geometric checks on the entire mesh to apply boundary
+conditions. In fact, if the mesh is directly tied to a geometric model
+(using a GEntity\_ptr), it is possible to specify a boundary condition
+on a complex surface such as a Trimmed NURBS surface and still apply
+the condition without error-prone geometric checks or large input
+files. Similarly, material properties can be specified per geometric
+model region and applied automatically to every mesh region classified
+in that model region without the need for element sets in input files.
 

### 2010-03-20 docs/MSTK.tex
Updated version of MSTK manual with new operators, new examples and new sections

--- 2010-02-25 docs/MSTK.tex
@@
   \end{center}
+  \caption{Mesh representations in MSTK}
+  \label{fig:reptypes}
 \end{figure}
@@
   of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
-  See Appendix~\ref{app:reptypes} for schematics of these
+  See Appendix~\ref{fig:reptypes} for schematics of these
   representations. {\em Currently only representation types F1 and
@@
 
-\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
+\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE, POINTER, VECTOR or TENSOR.
 \end{description}
@@
 
-\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr}
-    mesh, const char *filename):} Initialize or read a mesh from file
-  into the given mesh object. Returns 1 if successful, 0 otherwise. It
-  is possible to have a an MSTK file in the R1, R2 and R4 format into
-  a mesh initialized as type F1 or F4.
+\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr} mesh, const
+    char *filename):} Initialize or read a mesh from a file in the
+  MSTK format into the given mesh object. Returns 1 if successful, 0
+  otherwise. It is possible to have a an MSTK file in the R1, R2 and
+  R4 format into a mesh initialized as type F1 or F4. This routine
+  imports any attributes that are into the mesh.
 
@@
   meshes generated by
-  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.
+  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov}}.
 
@@
 
-\item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr}
-    mesh, const char *filename):} Save a mesh to a filename. The file
-  is created if it does not exists. It is recommended that the .mstk
-  extension be used for MSTK mesh files.  However, there is no such
-  requirement.
+  \item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr} mesh, const
+      char *filename):} Save a mesh to a filename. The file is created
+    if it does not exists. It is recommended that the .mstk extension
+    be used for MSTK mesh files.  However, there is no such
+    requirement. The routine will write out the mesh and any
+    attributes attached to the mesh except attributes of the type
+    POINTER.
 
@@
 \item[]{\bf {\em List\_ptr} MV\_Edges({\em MVertex\_ptr} mvertex):}
-List of edges connected to the vertex.
+List of edges connected to the vertex. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
 \item[]{\bf {\em List\_ptr} MV\_Faces({\em MVertex\_ptr} mvertex):}
-List of faces connected to the vertex.
+List of faces connected to the vertex. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
 \item[]{\bf {\em List\_ptr} MV\_Regions({\em MVertex\_ptr} mvertex):}
-List of regions connected to the vertex.
+List of regions connected to the vertex. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 \end{description}
@@
 \item[]{\bf {\em List\_ptr} ME\_Faces({\em MEdge\_ptr}
-    medge):} Returns the set of faces using this edge.
+    medge):} Returns the set of faces using this edge. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
 \item[]{\bf {\em List\_ptr} ME\_Regions({\em MEdge\_ptr}
-    medge):} Returns the set of regions using this edge.
+    medge):} Returns the set of regions using this edge. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
   A vertex set with ccw ordering starting with vertex $V_2$ is
-  ${V_2,V_3,V_0,V_1}$.
+  ${V_2,V_3,V_0,V_1}$.  {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
   ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
-  ${E_1,E_0,E_3,E_2}$.
+  ${E_1,E_0,E_3,E_2}$.  {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
 planar or surface surface meshes, i.e., for boundary faces not
-connected to any regions.
+connected to any regions.  {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
 used by any regions, the function returns NULL to indicate that the
-set is empty. If not the set may contain one or two regions.
+set is empty. If not the set may contain one or two regions.  {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
 Appendix~\ref{app:reg_conventions}. For non-standard elements the set
-is unordered.
+is unordered. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
 \item[]{\bf {\em List\_ptr} MR\_Edges({\em MRegion\_ptr} mregion):}
-Return the \underline{unordered} set of edges of a region.
+Return the \underline{unordered} set of edges of a region.  {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
 \item[]{\bf {\em List\_ptr} MR\_Faces({\em MRegion\_ptr} mregion):}
-Returns the set of faces of a region.
+Returns the set of faces of a region. {\bf The list returned by this
+  operator must be deleted by the calling application using {\em List\_Delete}}.
 
@@
 
-\item[]{\bf {\em void} MEnt\_Set\_AttVal({\em MEntity\_ptr}
-    ent, {\em MAttrib\_ptr} attrib, {\em int} ival,
-    {\em double} lval, {\em void} *pval):} Set the value of an
-  attribute for given mesh entity. Depending on whether the attribute
-  is integer, double or pointer, the appropriate entry in the argument
-  list is specified.
+\item[]{\bf {\em void} MEnt\_Set\_AttVal({\em MEntity\_ptr} ent, {\em
+      MAttrib\_ptr} attrib, {\em int} ival, {\em double} lval, {\em
+      void} *pval):} Set the value of an attribute for given mesh
+  entity. Depending on the attribute type, different arguments of this
+  function are relevant. If the attribute type is:
+
+\begin{itemize}
+  \item {\bf INT} -- specify the integer value in {\bf ival},
+  \item {\bf DOUBLE} -- specify the the real value in {\bf rval},
+  \item {\bf VECTOR} -- specify a pointer to an array of
+    doubles in {\bf pval}\footnote{Remember to not free the array specified in {\bf pval} without telling the attribute system because MSTK stores the pointer to the array directly and does not make a copy of the array.} - the number of doubles should equal the number components specified in the definition of the VECTOR attribute,
+  \item {\bf TENSOR} -- specify a pointer
+    to an array of doubles in {\bf pval} - the number of doubles should equal the number of components specified in the definition of the TENSOR attribute,
+  \item {\bf POINTER} -- specify the pointer value in {\bf pval}.
+  \end{itemize}
 
@@
 
-\item[]{\bf {\em int} MEnt\_Get\_AttVal({\em MEntity\_ptr}
-    ent, {\em MAttrib\_ptr} int *ival, double *rval, double
-    **pval):} Get the value of an attribute from an entity.  Depending
-  on the attribute type, either {\bf ival}, {\bf rval} or
-  {\bf pval} is returned.
+\item[]{\bf {\em int} MEnt\_Get\_AttVal({\em MEntity\_ptr} ent, {\em
+      MAttrib\_ptr} int *ival, double *rval, double **pval):} Get the
+  value of an attribute from an entity.  Depending on the attribute
+  type, different arguments of this function are relevant. If the
+  attribute type is:
+
+  \begin{itemize}
+  \item {\bf INT} -- {\bf ival} contains the integer value,
+  \item {\bf DOUBLE} -- {\bf rval} contains the real value,
+  \item {\bf VECTOR} -- {\bf pval} contains a pointer to an array of doubles,
+  \item {\bf TENSOR} -- {\bf pval} contains a pointer
+    to an array of doubles,
+  \item {\bf POINTER} -- {\bf pval} conntains the pointer value.
+  \end{itemize}
 
@@
 
-\item[]{\bf {\em MAttrib\_ptr} MAttrib\_New({\em Mesh\_ptr}
-    mesh, {\em const char} *att\_name, {\em MAttType} att\_type,
-    {\em MType} entdim):} Define a new mesh attribute. Along with
-  the mesh to which this attribute is assigned ({\bf mesh}) and the
-  name of the attribute ({\bf att\_name}), the type of the
-  attribute ({\bf att\_type}) must be specified as {\bf INT},
-  {\bf DOUBLE} or {\bf POINTER}. Also, the dimension of the entity
-    for which the attribute value can be set must be specified as
-    {\bf MVERTEX}, {\bf MEDGE}, {\bf MFACE}, {\bf MREGION}
-    or {\bf MALLTYPE}.
+\item[]{\bf {\em MAttrib\_ptr} MAttrib\_New({\em Mesh\_ptr} mesh, {\em
+      const char} *att\_name, {\em MAttType} att\_type, {\em MType}
+    entdim, $<${\em int ncomponents}$>$):} Define a new mesh
+  attribute. Along with the mesh to which this attribute is assigned
+  ({\bf mesh}) and the name of the attribute ({\bf att\_name}), the
+  type of the attribute ({\bf att\_type}) must be specified as {\bf
+    INT}, {\bf DOUBLE}, {\bf VECTOR}, {\bf TENSOR} or {\bf
+    POINTER}. Also, the dimension of the entity for which the
+  attribute value can be set must be specified as {\bf MVERTEX}, {\bf
+    MEDGE}, {\bf MFACE}, {\bf MREGION} or {\bf MALLTYPE}. The last
+  argument {\em {\bf ncomponents}} is an optional argument for attributes of
+  type {\bf INT}, {\bf DOUBLE} or {\bf POINTER} but is required for
+  attributes of type {\bf VECTOR} and {\bf TENSOR}.
 
@@
 
-\# {\em {\bf NOT IMPLEMENTED}}
-
-\# {\em {\bf VERTEX ATTRIBUTES}} \\
-\# {\em {\em int} NVA - Number of Vertex attributes} \\
+\# {\em {\bf ATTRIBUTES}} \\
 \# \\
-\# {\em char *VA\_name\_1 - Name of first vertex attribute} \\
-\# {\em {\em int} VA\_type\_1 - Type of first vertex attribute} \\
-\# {\em {\em int} VA\_dim\_1 - Dimension of first vertex attribute} \\
-\# \\
-\# {\em char *VA\_name\_2 - Name of second vertex attribute} \\
-\# {\em {\em int} VA\_type\_2 - Type of second vertex attribute} \\
-\# {\em {\em int} VA\_dim\_2 - Dimension of first vertex attribute} \\
-\# \\
-\# . . .\\
-\# \\
-\# {\em char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
-\# {\em {\em int} VA\_type\_NVA - Type of NVA'th vertex attribute} \\
-\# {\em {\em int} VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
-\# \\
-\# {\em VA\_type can be 1 (int), 2 (double), 3 (string)} \\
-\# {\em VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
-\# {\em VA\_dim can only be 1 when VA\_type is string}
+\# {\em char *Attrib\_name\_1 - Name of first attribute} \\
+\# {\em int Attrib\_type\_1 - Type of  attribute (INT, DOUBLE, VECTOR, TENSOR} \\
+\# {\em int Attrib\_len\_1 - 0 (scalars), number of components (vectors), number of components (tensors)} \\
+\# {\em char *Attrib\_ent\_1 - Type of mesh entity attribute is applicable to (VERTEX, EDGE, FACE, REGION, ALLTYPES)} \\
 \vspace{1ex}
-
-NVA \hspace{0.5ex} \\
-VA\_name\_1 \hspace{0.5ex} VA\_type\_1 \hspace{0.5ex} VA\_dim\_1 \\ 
-VA\_name\_2 \hspace{0.5ex} VA\_type\_2 \hspace{0.5ex} VA\_dim\_2 \\
-. \\
-VA\_dim\_NVA \hspace{0.5ex} VA\_type\_NVA \hspace{0.5ex} VA\_dim\_2  
+\# {\em int Nent - Number entities for which this attribute is specified - Useful if an attribute is to be specified for only a small subset of mesh entities, e.g. boundary conditions} \\
+\# {\em Entity\_dim\_1 \hspace{0.5ex} Entity\_ID\_1 \hspace{0.5ex} Attrib\_components} \\ 
+\# {\em Entity\_dim\_2 \hspace{0.5ex} Entity\_ID\_2 \hspace{0.5ex} Attrib\_components} \\ 
+\# {\em Entity\_dim\_3 \hspace{0.5ex} Entity\_ID\_3 \hspace{0.5ex} Attrib\_components} \\ 
+.\\
+.\\
+.\\
+\# {\em Entity\_dim\_n \hspace{0.5ex} Entity\_ID\_n \hspace{0.5ex} Attrib\_components}
 \vspace{1ex}
 
-\# {\em For each vertex attribute record, set of attribute values} \\
-\# {\em E.G., there are 3 attributes for each vertex:} \\
-\# {\em a scalar int, a vector of 3 doubles and a string}
-
-VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
-VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
 . \\
@@
 . \\
-\# {\em Repeated NV times}
+
+\# {\em char *Attrib\_name\_N - Name of first attribute} \\
+\# {\em int Attrib\_type\_N - Type of  attribute (INT, DOUBLE, VECTOR, TENSOR)} \\
+\# {\em int Attrib\_len\_N - 0 (scalars), number of components (vectors), number of components (tensors)} \\
+\# {\em char *Attrib\_ent\_N - Type of mesh entity attribute is applicable to (VERTEX, EDGE, FACE, REGION, ALLTYPES)} \\
+\vspace{1ex}
+\# {\em int Nent - Number entities for which this attribute is specified} \\
+\# {\em Entity\_dim\_1 \hspace{0.5ex} Entity\_ID\_1 \hspace{0.5ex} Attrib\_components} \\ 
+\# {\em Entity\_dim\_2 \hspace{0.5ex} Entity\_ID\_2 \hspace{0.5ex} Attrib\_components} \\ 
+\# {\em Entity\_dim\_3 \hspace{0.5ex} Entity\_ID\_3 \hspace{0.5ex} Attrib\_components} \\ 
+.\\
+.\\
+.\\
+\# {\em Entity\_dim\_n \hspace{0.5ex} Entity\_ID\_n \hspace{0.5ex} Attrib\_components}
+\vspace{1ex}
 \vspace{2ex}
 
+\end{verse}
+
 \newpage
-\# {\em {\bf EDGE ATTRIBUTES}} \\
-\# {\em Similar to Vertex attribute description}
-\vspace{1ex}
-
-NEA \hspace{0.5ex} \\
-EA\_name\_1 \hspace{0.5ex} EA\_type\_1 \hspace{0.5ex} EA\_dim\_1 \\ 
-EA\_name\_2 \hspace{0.5ex} EA\_type\_2 \hspace{0.5ex} EA\_dim\_2 \\
-. \\
-EA\_dim\_NEA \hspace{0.5ex} EA\_type\_NEA \hspace{0.5ex} EA\_dim\_2  
-\vspace{1ex}
-
-\# {\em For each edge attribute record, set of attribute values} \\
-\# {\em E.G., a scalar int, a scalar double and a string}
-\vspace{1ex}
-
-EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
-EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
-. . . \\
-\# {\em Repeated NE times}
-\vspace{2ex}
-
-\# {\em {\bf FACE ATTRIBUTES}} \\
-\# {\em Similar to Vertex attribute description}
-\vspace{1ex}
-
-NFA \hspace{0.5ex} \\
-FA\_name\_1 \hspace{0.5ex} FA\_type\_1 \hspace{0.5ex} FA\_dim\_1 \\ 
-FA\_name\_2 \hspace{0.5ex} FA\_type\_2 \hspace{0.5ex} FA\_dim\_2 \\
-. \\
-FA\_dim\_NFA \hspace{0.5ex} FA\_type\_NEA \hspace{0.5ex} FA\_dim\_2  
-\vspace{1ex}
-
-\# {\em For each face attribute record, set of attribute values} \\
-\# {\em E.G., a vector of 2 doubles and a string}
-\vspace{1ex}
-
-FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
-FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
-. . . \\
-\# {\em Repeated NF times}
-\vspace{2ex}
+
+\section{Example Mesh and Files}
+
+In this section we show a simple square geometric model and an example
+2D mesh of that model comprised of triangles, quads (convex and
+non-convex) and polygons. Assume that the mesh vertices have a
+velocity associated with them and mesh faces have a symmetric
+second-order diffusivity tensor associated with them. We then
+present the MSTK files describing the mesh in the F1 and R1 formats.
+
+\begin{figure}[!h]
+  \begin{center}
+    \begin{minipage}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[scale=0.85]{figures/examplemodel}
+      \end{center}
+    \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[scale=0.85]{figures/examplemesh}
+      \end{center}
+    \end{minipage}
+  \end{center}
+  \caption{Geometric model and a two-dimensional mesh of mixed elements of the model.}
+\end{figure}
 
 \newpage
-\# {\em {\bf REGION ATTRIBUTES}} \\
-\# {\em Similar to Vertex attribute description}
-\vspace{1ex}
-
-NRA \hspace{0.5ex} \\
-RA\_name\_1 \hspace{0.5ex} RA\_type\_1 \hspace{0.5ex} RA\_dim\_1 \\ 
-RA\_name\_2 \hspace{0.5ex} RA\_type\_2 \hspace{0.5ex} RA\_dim\_2 \\
-. \\
-RA\_dim\_NRA \hspace{0.5ex} RA\_type\_NRA \hspace{0.5ex} RA\_dim\_NRA 
-\vspace{1ex}
-
-\# {\em For each Region attribute record, set of attribute values} \\
-\# {\em E.G., a vector of 3 ints}
-\vspace{1ex}
-
-RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
-RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
-. . .\\
-\# {\em Repeated NR times}
-\vspace{2ex}
-
-\end{verse}
+
+File in F1 format:
+\begin{verbatim}
+MSTK 1.0
+F1 18 31 14 0
+vertices
+0.00 0.00 0.00  0 1
+0.33 0.00 0.00  1 1
+0.66 0.00 0.00  1 1
+1.00 0.00 0.00  0 2
+0.00 0.33 0.00  1 4
+0.66 0.66 0.00  2 1
+0.66 1.00 0.00  1 3
+0.16 0.84 0.00  2 1
+0.00 0.66 0.00  1 4
+0.16 0.33 0.00  2 1
+0.50 0.33 0.00  2 1
+0.40 0.84 0.00  2 1
+1.00 0.66 0.00  1 2
+0.00 1.00 0.00  0 4
+1.00 1.00 0.00  0 3
+0.66 0.50 0.00  2 1
+0.33 1.00 0.00  1 3
+1.00 0.33 0.00  1 2
+edges 
+2 1    1 1
+2 3    1 1
+3 4    1 1
+4 18   1 2
+2 11   2 1
+15 7   1 3
+14 8   2 1
+18 16  2 1
+12 16  2 1
+10 5   2 1
+11 16  2 1
+15 13  1 2
+14 17  1 3
+17 7   1 3
+8  9   2 1
+18 13  1 2
+14 9   1 4
+10 11  2 1
+10 2   2 1
+9 10   2 1
+16 6   2 1
+6 13   2 1
+7 6    2 1
+17 12  2 1
+8 17   2 1
+3 11   2 1
+11 18  2 1
+1 5    1 4
+5 9    1 4
+7 12   2 1
+8 12   2 1
+faces edge
+4  3 4 -27 -26            2 1
+6  18 11 -9 -31 15 20     2 1
+3  25 -13 7               2 1
+3  31 -24 -25             2 1
+3  2 26 -5                2 1
+4  -28 -1 -19 10          2 1
+4  16 -22 -21 -8          2 1
+4  22 -12 6 23            2 1
+3  -30 -14 24             2 1
+3  -10 -20 -29            2 1
+3  19 5 -18               2 1
+4  9 21 -23 30            2 1
+3  -15 -7 17              2 1
+3  27 8 -11               2 1
+attributes
+velocity
+VECTOR
+3
+MVERTEX
+18
+0 1    0.0  0.0  0.0
+0 2    3.3  0.0  0.0
+0 3    6.6  0.0  0.0
+0 4   10.0  0.0  0.0
+0 5    0.0  3.3  0.0
+0 6    6.6  6.6  0.0
+0 7    6.6 10.0  0.0
+0 8    1.6  8.4  0.0
+0 9    0.0  6.6  0.0
+0 10   1.6  3.3  0.0
+0 11   5.0  3.3  0.0
+0 12   4.0  6.6  0.0
+0 13  10.0  6.6  0.0
+0 14   0.0 10.0  0.0
+0 15  10.0 10.0  0.0
+0 16   6.6  5.0  0.0
+0 17   3.3 10.0  0.0
+0 18  10.0  3.3  0.0
+diffusity
+TENSOR
+3
+MFACE
+14
+2 1    1.0 10.0 2.0
+2 2    1.0 10.0 2.0
+2 3    1.0 10.0 2.0
+2 4    1.0 10.0 2.0
+2 5    1.0 10.0 2.0
+2 6    1.0 10.0 2.0
+2 7    1.0 10.0 2.0
+2 8    1.0 10.0 2.0
+2 9    1.0 10.0 2.0
+2 10   1.0 10.0 2.0
+2 11   1.0 10.0 2.0
+2 12   1.0 10.0 2.0
+2 13   1.0 10.0 2.0
+2 14   1.0 10.0 2.0
+\end{verbatim}
 
 \newpage
+File in R1 format:
+\begin{verbatim}
+MSTK 1.0
+R1 18 0 14 0
+vertices
+0.00 0.00 0.00  0 1
+0.33 0.00 0.00  1 1
+0.66 0.00 0.00  1 1
+1.00 0.00 0.00  0 2
+0.00 0.33 0.00  1 4
+0.66 0.66 0.00  2 1
+0.66 1.00 0.00  1 3
+0.16 0.84 0.00  2 1
+0.00 0.66 0.00  1 4
+0.16 0.33 0.00  2 1
+0.50 0.33 0.00  2 1
+0.40 0.84 0.00  2 1
+1.00 0.66 0.00  1 2
+0.00 1.00 0.00  0 4
+1.00 1.00 0.00  0 3
+0.66 0.50 0.00  2 1
+0.33 1.00 0.00  1 3
+1.00 0.33 0.00  1 2
+faces vertex
+4  3 4 18 11              2 1
+6  9 10 11 16 12 8        2 1
+3  14 8 17                2 1
+3  8 12 17                2 1
+3  2 3 11                 2 1
+4  1 2 10 5               2 1
+4  6 16 18 13             2 1
+4  6 13 15 7              2 1
+3  7 17 12                2 1
+3  5 10 9                 2 1
+3  2 11 10                2 1
+4  12 16 6 7              2 1
+3  9 8 14                 2 1
+3  11 18 16               2 1
+attributes
+velocity
+VECTOR
+3
+MVERTEX
+18
+0 1    0.0  0.0  0.0
+0 2    3.3  0.0  0.0
+0 3    6.6  0.0  0.0
+0 4   10.0  0.0  0.0
+0 5    0.0  3.3  0.0
+0 6    6.6  6.6  0.0
+0 7    6.6 10.0  0.0
+0 8    1.6  8.4  0.0
+0 9    0.0  6.6  0.0
+0 10   1.6  3.3  0.0
+0 11   5.0  3.3  0.0
+0 12   4.0  6.6  0.0
+0 13  10.0  6.6  0.0
+0 14   0.0 10.0  0.0
+0 15  10.0 10.0  0.0
+0 16   6.6  5.0  0.0
+0 17   3.3 10.0  0.0
+0 18  10.0  3.3  0.0
+diffusity
+TENSOR
+3
+MFACE
+14
+2 1    1.0 10.0 2.0
+2 2    1.0 10.0 2.0
+2 3    1.0 10.0 2.0
+2 4    1.0 10.0 2.0
+2 5    1.0 10.0 2.0
+2 6    1.0 10.0 2.0
+2 7    1.0 10.0 2.0
+2 8    1.0 10.0 2.0
+2 9    1.0 10.0 2.0
+2 10   1.0 10.0 2.0
+2 11   1.0 10.0 2.0
+2 12   1.0 10.0 2.0
+2 13   1.0 10.0 2.0
+2 14   1.0 10.0 2.0
+\end{verbatim}
 
@@
 
+ if (argc == 1) {
+   fprintf(stderr,"Usage: %s meshfilename (without .mstk extension)\n",argv[0]);
+   exit(-1);
+ }
+
  /* Initialize MSTK - Always do this even if it does
@@
  if (!ok) {
-   fprintf(stderr,"Cannot file input file %s\n\n\n",meshname);
+   fprintf(stderr,"Cannot find input file %s\n\n\n",meshname);
    exit(-1);
@@
 
+ if ((natt = MESH_Num_Attribs(mesh))) {
+
+   fprintf(stderr,"Attributes on the mesh:\n\n");
+
+   for (i = 0; i < natt; i++) {
+
+     attrib = MESH_Attrib(mesh,i);
+
+     attentdim = MAttrib_Get_EntDim(attrib);
+
+     /* Won't print out edge, face and region based attributes
+        but the code should be quite similar */
+
+     if (attentdim != MVERTEX) continue;
+
+     fprintf(stderr,"Attribute Number %-d\n",i+1);
+
+     MAttrib_Get_Name(attrib,attname);
+     fprintf(stderr,"Name: %-s\n",attname);
+
+     atttype = MAttrib_Get_Type(attrib);
+     switch(atttype) {
+     case INT:
+       fprintf(stderr,"Type: Integer\n");
+       break;
+     case DOUBLE:
+       fprintf(stderr,"Type: Double\n");
+       break;
+     case VECTOR:
+       fprintf(stderr,"Type: Vector\n");
+       break;
+     case TENSOR:
+       fprintf(stderr,"Type: Tensor\n");
+       break;
+     default:
+       fprintf(stderr,"Unrecognizable or unprintable attribute type\n");
+       continue;	
+     }
+
+     switch(attentdim) {
+     case MVERTEX:
+       fprintf(stderr,"Applicable to vertices only\n");
+       break;
+     case MEDGE:
+     case MFACE:
+     case MREGION:
+     case MALLTYPE:
+       /* Won't print out edge, face and region based attributes
+          but the code should be quite similar */
+       continue; 
+     default:
+       fprintf(stderr,"Unrecognized entity type\n");
+       continue;
+     }
+
+     ncomp = MAttrib_Get_NumComps(attrib);
+     fprintf(stderr,"Number of components: %-d\n",ncomp);
+
+     switch(attentdim) {
+     case MVERTEX:
+       idx = 0;
+       while ((v = MESH_Next_Vertex(mesh,&idx))) {
+         if (MEnt_Get_AttVal(v,attrib,&ival,&rval,&pval)) {
+           fprintf(stderr,"V %-d: ",MV_ID(v));
+           switch (atttype) {
+           case INT:
+             fprintf(stderr," %-d\n",ival);
+             break;
+           case DOUBLE: 
+             fprintf(stderr," %-lf ",rval);
+             break;
+           case VECTOR: case TENSOR:
+             rval_arr = (double *) pval;
+             for (k = 0; k < ncomp; k++)
+       	fprintf(stderr," %-lf ",rval_arr[k]);
+             break;
+           default:
+             break;
+           }
+           fprintf(stderr,"\n");
+         }
+       }
+       break;
+     case MEDGE: case MFACE: case MREGION: case MALLTYPE:
+       /* Skipped code to print out attributes for all other
+          entity types but it is almost identical */
+       break;	
+     default:
+       break;
+     } /* switch (attentdim) */
+
+   } /* for (i = 0; i < natt) */
+
+ } /* if (Mesh_Num_Attribs(mesh)) */
+
  /* Write out a copy of the mesh */

### 2010-03-23 docs/MSTK.tex
Added LAUR numbers and acknowledgement

@@
 
-\title{MSTK: Mesh Toolkit, v 1.7 - DRAFT}
+\title{MSTK: Mesh Toolkit, v 1.7}
 
@@
   Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
-  rao@lanl.gov}
+  rao@lanl.gov \\
+  \vspace{5em}
+  LA-UR-04-0878}
 
@@
 
+\section{Acknowledgements}
+
+This work was performed under the auspices of the National Nuclear
+Security Administration of the US Department of Energy at Los Alamos
+National Laboratory under contract No. DE-AC52-06NA25396 with partial support
+from the Advanced Simulation Capability (ASC) program.
+
+This manual has been reviewed by LANL and deemed suitable for unlimited
+release (LA-UR-04-0878). MSTK has been reviewed by LANL and deemed suitable
+for unlimited distribution (LA-CC-04-010). MSTK is distributed under
+an LGPL open-source license.
+
 \end{document}

### 2011-02-18 docs/MSTK.tex
Updated documentation for version 1.8

@@
 
-\par MSTK will eventually support distributed meshes for parallel
-computing. However, this is still not in place.
+\par MSTK has preliminary support of distributed meshes for parallel
+computing.
 
@@
 MSTK offers a flexible infrastructure for representing and
-manipulating meshes so that mesh-based application developers do not
-have to deal with the complexitiies of managing their own data
-structures. The functional interface of MSTK is designed to provide
-easy access to mesh data and allow for easy-to-read but efficient code
-to be written for mesh based applications. Accessing the mesh data
-through the functional interface allows the high level application to
-be unchanged even if the lower level data structures in MSTK
-change. MSTK supports a wide array of element types and several
-representations, all of which can take a new developer years to write
-and make robust. New users of MSTK can typically start writing their
-own codes to query and manipulate meshes within a few days of
-familiarizing themselves with MSTK. Finally, use of MSTK allows easy
-integration with other codes using MSTK as their mesh data framework.
+manipulating meshes in a distributed environment so that mesh-based
+application developers do not have to deal with the complexitiies of
+managing their own data structures. The functional interface of MSTK
+is designed to provide easy access to mesh data and allow for
+easy-to-read but efficient code to be written for mesh based
+applications. Accessing the mesh data through the functional interface
+allows the high level application to be unchanged even if the lower
+level data structures in MSTK change. MSTK supports a wide array of
+element types and several representations, all of which can take a new
+developer years to write and make robust. New users of MSTK can
+typically start writing their own codes to query and manipulate meshes
+within a few days of familiarizing themselves with MSTK. Finally, use
+of MSTK allows easy integration with other codes using MSTK as their
+mesh data framework.
 
@@
 
+\subsection{Parallel Mesh Management}
+
+MSTK now has preliminary support for representation of distributed
+surface and volume meshes for supporting parallel
+simulations. \textit{This is an alpha functionality and the API or
+  behavior of some functions is subject to revision.}
+
+Currently, the only way to distribute meshes on multiple processors is
+to perform a serial read on processor 0 and then distribute it to
+however many processors are necessary. Mesh attributes and mesh entity
+sets are appropriately partitioned and distributed to the different
+processors. Functionality to read a pre-partitioned mesh is planned
+for the future.
+
+MSTK currently supports one (or no) layer of overlapped or ghost
+entities between processors. If a surface mesh requests a layer of
+ghost faces, then all lower order entities (edges and vertices)
+forming the ghost faces are also ghosted. Similarly, if a volume mesh
+requests a layer of ghost regions, all lower order entities (faces,
+edges and vertices) forming the ghost regions are reproduced. Entities
+that are owned by a processor but duplicated on another processor are
+called {\em Overlap} entities. Entities that are duplicates of overlap
+entities are called {\em Ghost} entities. The master processor ID of a
+ghost entity is the processor ID that its corresponding overlap entity
+lives on.
+
+MSTK supports automatic updating of scalars, vectors and tensors on
+mesh entities of all dimensions. The application can update the
+attribute values on its owned entities and call MSTK to update the
+attribute values on ghosts on other processors.
+
+\textit{MSTK does not yet support parallel modification of meshes}.
+
 \subsection{MSTK Application Programming Interface (API)}
@@
 region-node connectivity) and read it in to any of the other
-representations, say F1. The other possibility is to write a file in
-the GMV format and use MESH\_ImportFromFile to import the mesh in.
+representations, say F1.
+
+It is also possible to import meshes in other formats into MSTK. Some
+formats that MSTK can read from are the GMV format and the ExodusII
+format. Applications can use MESH\_ImportFromFile to import the mesh
+into MSTK.
 
 MSTK can write meshes out to different formats including GMV, STL and
-FLAG X3D. When compiled with the appropriate options and using a
-partitioner from an external library, MSTK can also write out a
-parallel X3D file.
+ExodusII and FLAG X3D. When compiled with the appropriate options and
+using a partitioner from an external library, MSTK can also write out
+a parallel X3D file.
 
@@
 
-Unordered sets or lists of entities in MSTK are returned as type
-{\em List\_ptr}. The following are the set operations available in
-MSTK:
+Unordered lists of entities in MSTK are returned as type {\em
+  List\_ptr} (See MSets for named entity sets). The following are the
+list operations available in MSTK:
 
@@
 
-\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr}
-    mesh, const char *filename, const char *format):} Import a mesh
-  data from an external file format and construct MSTK mesh. Currently
-  the only format supported is the
+\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr} mesh, const
+    char *filename, const char *format):} Import a mesh data from an
+  external file format and construct MSTK mesh. Currently the formats
+  supported are the
   GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
-  file format. The routine imports as many attributes as it can from
-  the input file. It also uses special attributes or keywords as data
-  about element and node classification.  For GMV files, the routine
-  uses the ``material'' data to indicate region or face classification
-  depending on the dimensionality of the mesh. If ``material'' data is
-  not specified, it uses the ``itetclr'' attribute to assign region or
-  face classification. It also uses the ``icr'' keyword describing the
+  file format and the Exodus II file format. The routine imports as
+  many attributes as it can from the input file. It also uses special
+  attributes or keywords as data about element and node
+  classification.  For GMV files, the routine uses the ``material''
+  data to indicate region or face classification depending on the
+  dimensionality of the mesh. If ``material'' data is not specified,
+  it uses the ``itetclr'' attribute to assign region or face
+  classification. It also uses the ``icr'' keyword describing the
   number of constraints on a node to interpret if the node is
@@
   meshes generated by
-  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov}}.
+  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov}}. For Exodus II file
+  formats, element blocks are used to create sets of mesh regions with
+  the name 'matset\_N' (where N is the element block ID), sidesets are
+  used to create sets of faces named 'sideset\_N' and nodesets are used
+  to create sets of vertices named 'nodeset\_N'.
 
@@
 
-%% \item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr}
-%%     mesh, const char *filname, const char *format, int natt, char
-%%     **attnames):} Export a mesh to an external file format. Currently,
-%%   the formats supported are the GMV file (format string: ``gmv'') and
-%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
-%%   ``dx''). Only integer and double attributes of the mesh are exported
-%%   to the files.
-
-\item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr} mesh, const
-    char *filname, const char *format, int natt, char **attnames):}
-  Export a mesh to an external file format. Currently, the only
-  formats supported is the GMV file format (format string: ``gmv'')
-  and X3D (format string: ``flag''). Only integer and double
-  attributes of the mesh are exported to GMV files and no attributes
-  are exported to X3D files.
+  \item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr} mesh,
+      const char *filname, const char *format, int natt, char
+      **attnames):} Export a mesh to an external file
+    format. Currently, the formats supported are the GMV file format
+    (format: ``gmv'') and X3D (format: ``flag''), Exodus II file
+    format (format: ``exodusii''). In addition surface meshes may be
+    exported to STL (format: ``stl'') and IBM Data Explorer (format:
+    ``dx'') format. Only integer and double attributes of the mesh are
+    exported to GMV files and no attributes are exported to X3D files.
 
   \item[]{\bf {\em void} MESH\_ExportToGMV({\em Mesh\_ptr} mesh, const
-    char *filname, const char *format, int natt, char **attnames):} Export a mesh to a GMV file format.
-
-\item[]{\bf {\em void} MESH\_ExportToFLAGX3D({\em Mesh\_ptr} mesh, const
-    char *filname, const char *format, int natt, char **attnames):} Export a mesh to the FLAG X3D format.
-
-\item[]{\bf {\em void} MESH\_ExportToFLAGX3D\_Par({\em Mesh\_ptr} mesh,
-    const char *filname, const char *format, int natt, char
-    **attnames, int *procids):} Export a mesh to the FLAG X3D format
-  given a partitioned mesh. The owning processor for each element is
-  given in the array {\bf procids}.
-
-\item[]{\bf {\em void} MESH\_ExportToSTL({\em Mesh\_ptr} mesh, const
-    char *filname):} Export a mesh to the STL format (SURFACE MESHES ONLY!).
+      char *filname, int natt, char **attnames):} Export a mesh to a
+    GMV file format.
+
+\item[]{\bf {\em void} MESH\_ExportToFLAGX3D({\em Mesh\_ptr} mesh,
+    const char *filname, int natt, char **attnames):} Export a mesh to
+  the FLAG X3D format.
+
+\item[]{\bf {\em void} MESH\_ExportToFLAGX3D\_Par({\em Mesh\_ptr}
+    mesh, const char *filname, int natt, char **attnames, int
+    *procids):} Export a mesh to the FLAG X3D format given a
+  partitioned mesh. The owning processor for each element is given in
+  the array {\bf procids}.
+
+\item[]{\bf {\em void} MESH\_ExportToExodusII({\em Mesh\_ptr} mesh,
+    const char *filname, int natt, char **attnames):} Export a mesh to
+  a GMV file format.
+
+\item[]{\bf {\em void} MESH\_ExportToDX({\em Mesh\_ptr} mesh, const
+    char *filname):} Export a mesh to the IBM Data Explorer format
+  (SURFACE MESHES ONLY!).
 
@@
 
+\item[]{\bf {\em MEntity\_ptr} MESH\_EntityFromID({\em Mesh\_ptr}
+    mesh, {\em MType mtype}, {\em int} id):} Return mesh entity of
+  given type and given ID if it exists; return NULL otherwise.
+
 \item[]
@@
     MESH\_AttribByName({\em Mesh\_ptr} mesh, {\em const char}
-    *name):} Return a mesh attribute with given name if it exisists in
+    *name):} Return a mesh attribute with given name if it exists in
   the mesh. Returns NULL if mesh has no such attribute.
+
+\item[]
+
+\item[]{\bf {\em int} MESH\_Num\_MSets({\em Mesh\_ptr}
+    mesh):} Number of entity sets associated with the mesh.
+
+\item[]{\bf {\em MSet\_ptr} MESH\_MSet({\em Mesh\_ptr}
+    mesh, {\em int} i):} Return the i'th entity set in the mesh.
+  Returns NULL if i $<$ 0 or i $>$ number of entity sets.
+
+\item[]{\bf {\em MSet\_ptr}
+    MESH\_Next\_MSet({\em Mesh\_ptr} mesh, {\em int} *index):}
+  Returns the next entity set while iterating through the sets of
+  the mesh. See the routine List\_Next\_Entry above for an explanation
+  of how the iteration works.
+
+\item[]{\bf {\em MSet\_ptr}
+    MESH\_MSetByName({\em Mesh\_ptr} mesh, {\em const char}
+    *name):} Return a mesh entity set with given name if it exists in
+  the mesh. Returns NULL if mesh has no such set.
 
@@
   operator must be deleted by the calling application using {\em List\_Delete}}.
+
+\item[]
+
+\item[]  {\bf {\em PType} MV\_PType({\em MVertex\_ptr} v)}: Parallel entity type for vertex - can be PINTERIOR, POVERLAP or PGHOST.
+\item[]  {\bf {\em int}   MV\_MasterParID({\em MVertex\_ptr} v)}:
+\item[]  {\bf {\em int}   MV\_GlobalID({\em MVertex\_ptr} v)}:
+
+\item[]
+\item[]  The following three functions affect how mesh entities communicate and how their attributes are updated.
+IF YOU DON'T KNOW WHAT YOU ARE DOING WHEN YOU CALL THIS FUNCTION, YOU WILL GET WHAT YOU DESERVE.
+\item[]  {\bf {\em void}  MV\_Set\_PType({\em MVertex\_ptr} v, {\em PType} ptype)}: 
+\item[]  {\bf {\em void}  MV\_Set\_MasterParID({\em MVertex\_ptr} v, {\em int} masterparid)}
+\item[]  {\bf {\em void}  MV\_Set\_GlobalID({\em MVertex\_ptr} v, {\em int} globalid)}:
+
 \end{description}
@@
   edge so that it may be deleted freely by garbage cleaning procedures.
+
+\item[]
+
+\item[]  {\bf {\em PType} ME\_PType({\em MEdge\_ptr} v)}: Parallel entity type for edge - can be PINTERIOR, POVERLAP or PGHOST.
+\item[]  {\bf {\em int}   ME\_MasterParID({\em MEdge\_ptr} v)}:
+\item[]  {\bf {\em int}   ME\_GlobalID({\em MEdge\_ptr} v)}:
+
+\item[]
+\item[]  The following three functions affect how mesh entities communicate and how their attributes are updated.
+IF YOU DON'T KNOW WHAT YOU ARE DOING WHEN YOU CALL THIS FUNCTION, YOU WILL GET WHAT YOU DESERVE.
+\item[]  {\bf {\em void}  ME\_Set\_PType({\em MEdge\_ptr} v, {\em PType} ptype)}: 
+\item[]  {\bf {\em void}  ME\_Set\_MasterParID({\em MEdge\_ptr} v, {\em int} masterparid)}
+\item[]  {\bf {\em void}  ME\_Set\_GlobalID({\em MEdge\_ptr} v, {\em int} globalid)}:
 
@@
 
+\item[]  {\bf {\em PType} MF\_PType({\em MFace\_ptr} v)}: Parallel entity type for face - can be PINTERIOR, POVERLAP or PGHOST.
+\item[]  {\bf {\em int}   MF\_MasterParID({\em MFace\_ptr} v)}:
+\item[]  {\bf {\em int}   MF\_GlobalID({\em MFace\_ptr} v)}:
+
+\item[]
+\item[]  The following three functions affect how mesh entities communicate and how their attributes are updated.
+IF YOU DON'T KNOW WHAT YOU ARE DOING WHEN YOU CALL THIS FUNCTION, YOU WILL GET WHAT YOU DESERVE.
+\item[]  {\bf {\em void}  MF\_Set\_PType({\em MFace\_ptr} v, {\em PType} ptype)}: 
+\item[]  {\bf {\em void}  MF\_Set\_MasterParID({\em MFace\_ptr} v, {\em int} masterparid)}
+\item[]  {\bf {\em void}  MF\_Set\_GlobalID({\em MFace\_ptr} v, {\em int} globalid)}:
+
+\item[]
+
 \item[]{\bf {\em void} MF\_Coords({\em MFace\_ptr} mface,
@@
 
+\item[]  {\bf {\em PType} MR\_PType({\em MRegion\_ptr} v)}: Parallel entity type for edge - can be PINTERIOR, POVERLAP or PGHOST.
+\item[]  {\bf {\em int}   MR\_MasterParID({\em MRegion\_ptr} v)}:
+\item[]  {\bf {\em int}   MR\_GlobalID({\em MRegion\_ptr} v)}:
+
+\item[]
+\item[]  The following three functions affect how mesh entities communicate and how their attributes are updated.
+IF YOU DON'T KNOW WHAT YOU ARE DOING WHEN YOU CALL THIS FUNCTION, YOU WILL GET WHAT YOU DESERVE.
+\item[]  {\bf {\em void}  MR\_Set\_PType({\em MRegion\_ptr} v, {\em PType} ptype)}: 
+\item[]  {\bf {\em void}  MR\_Set\_MasterParID({\em MRegion\_ptr} v, {\em int} masterparid)}
+\item[]  {\bf {\em void}  MR\_Set\_GlobalID({\em MRegion\_ptr} v, {\em int} globalid)}:
+
+\item[]
+
 \item[]{\bf {\em void} MR\_Coords({\em MRegion\_ptr} mregion,
@@
 \newpage
+\subsection{Mesh Sets}
+
+Mesh sets are unordered but named sets of entities. MSets can be
+queried and modified just like Lists. However, unlike Lists, MSets are
+associated with a mesh, can be constrained to contain only particular
+types of entities and have boolean operations performed on them.
+
+\begin{description}
+
+\item[]{\bf {\em MSet\_ptr} MSet\_New({\em Mesh\_ptr} mesh, {\em const
+      char} *set\_name, {\em MType} entdim):} Define a new mesh entity
+  set. Along with the mesh to which this set is assigned ({\bf mesh})
+  and the name of the set ({\bf att\_name}), the the dimension of the
+  entities that the set can contain must be specified as {\bf
+    MVERTEX}, {\bf MEDGE}, {\bf MFACE}, {\bf MREGION} or {\bf
+    MALLTYPE}.
+
+\item[]{\bf {\em char} *MSet\_Name({\em MSet\_ptr} set, {\em char}
+    *att\_name):} Get the name of the given set.
+
+\item[]{\bf {\em MType} MSet\_EntDim({\em MSet\_ptr} set):} Get the
+  dimension of entities the set can contain. Can be {\bf MVERTEX},
+  {\bf MEDGE}, {\bf MFACE}, {\bf MREGION} or {\bf MALLTYPE}.
+
+\item[]{\bf {\em Mesh\_ptr} MSet\_Mesh({\em MSet\_ptr} set):}
+Returns the mesh that the entity set belongs to.
+
+\item[]{\bf {\em void} MSet\_Delete({\em MSet\_ptr} set):} Delete an
+  set.
+
+\item[]{\bf {\em MSet\_ptr} MSet\_Add({\em MSet\_ptr} l, {\em void
+      *}entry):} Add an entry to the set. The entry is strictly
+  appended to the end of the set.
+
+\item[]{\bf {\em MSet\_ptr} MSet\_ChknAdd({\em MSet\_ptr} l, {\em void
+      *}entry):} Add an entry to a set only if it is not already in
+  the set.
+
+\item[]{\bf {\em MSet\_ptr} MSet\_Insert({\em MSet\_ptr} l, {\em void
+      *}nuentry, {\em void *}b4entry):} Insert an entry into the set
+  in the position before 'b4entry'.
+
+\item[]{\bf {\em MSet\_ptr} MSet\_Inserti({\em MSet\_ptr} l, {\em void
+      *}nuentry, {\em int} i):} Insert an entry into the set at the
+  i'th valid position and push the entries previously at the i'th and
+  later position back.
+
+\item[]{\bf {\em int} MSet\_Rem({\em MSet\_ptr} l, {\em void
+      *}entry):} Remove an entry from the set. Returns 1 if
+  successful, 0 otherwise.
+
+\item[]{\bf {\em int} MSet\_Remi({\em MSet\_ptr} l, {\em int} i):}
+  Remove the i'th valid entry in the set. Returns 1 if successful, 0
+  otherwise.
+
+\item[]{\bf {\em int} MSet\_Replace({\em MSet\_ptr} l, {\em void
+      *}entry, {\em void *}nuentry):} Replace 'entry' with 'nuentry'
+  in set. Returns 1 if successful, 0 otherwise.
+
+\item[]{\bf {\em int} MSet\_Replacei({\em MSet\_ptr} l, {\em int} i,
+    void *nuentry):} Replace the i'th valid entry in the set with
+  'nuentry'. Returns 1 if successful, 0 otherwise.
+
+\item[]{\bf {\em int} MSet\_Contains({\em MSet\_ptr} l, void *entry):}
+  Returns 1 if set contains the entry, 0 otherwise.
+
+\item[]{\bf {\em int} MSet\_Locate({\em MSet\_ptr} l, void *entry):}
+  Returns the positional index of the entry in the set. Returns -1 if
+  the set does not contain the entry.
+
+\item[]{\bf {\em void *}MSet\_Entry({\em MSet\_ptr} l, {\em int} i):}
+  Return the i'th valid entry in the set.  Returns a NULL pointer if
+  the i'th valid entry could not be found.
+
+\item[]{\bf {\em void *}MSet\_Next\_Entry({\em MSet\_ptr} l, {\em int}
+    *i):} Return the next valid entry in the set. This routine works
+  just like List\_Next\_Entry.
+
+\item[]{\bf {\em int} MSet\_Num\_Entries({\em MSet\_ptr} l):} Return
+  the number of entries in a set.
+
+\item[]{\bf {\em MSet\_ptr} MSets\_Union({\em MSet\_ptr} s1, {\em
+      MSet\_ptr} s2):} Create a new set from the union of two sets.
+
+\item[]{\bf {\em MSet\_ptr} MSets\_Intersect({\em MSet\_ptr} s1, {\em
+      MSet\_ptr} s2):} Create a new set from the intersection of two
+  sets. Routine returns null if the intersection is a null set.
+
+\item[]{\bf {\em MSet\_ptr} MSets\_Subtract({\em MSet\_ptr} s1, {\em
+      MSet\_ptr} s2):} Create a new set by subtracting set 2 from set
+  1. Routine returns null if the result is a null set.
+
+\end{description}
+
+\newpage
 \subsection{Entity Marks}
@@
 
-\item[]{\bf {\em void} List\_Mark({\em List\_ptr} list, {\em int} mkr):} Mark a set of entities with given marker.
+\item[]{\bf {\em void} List\_Mark({\em List\_ptr} list, {\em int} mkr):} Mark a list of entities with given marker.
 
 \item[]{\bf {\em void} List\_Unmark({\em List\_ptr} list, {\em int} mkr):}
+Unmark a list of entities with respect to the given marker.
+
+\item[]{\bf {\em void} MSet\_Mark({\em MSet\_ptr} list, {\em int} mkr):} Mark a set of entities with given marker.
+
+\item[]{\bf {\em void} MSet\_Unmark({\em MSet\_ptr} list, {\em int} mkr):}
 Unmark a set of entities with respect to the given marker.
@@
 \end{verbatim}
+
+\newpage 
+\subsection{Parallel Mesh Management}
+
+The following are the highest level of parallel mesh management functions.
+
+\begin{description}
+
+\item[] {\bf {\em int} MSTK\_Mesh\_Read\_Distribute({\em Mesh\_ptr
+      *}recv\_mesh, {\em const char*} global\_mesh\_name, {\em int} dim,
+    {\em int} ring, {\em int} with\_attr, {\em int} rank, {\em int}
+    num, {\em MPI\_Comm} comm)}: Routine reads a mesh of dimension
+  'dim' from file 'global\_mesh\_name' on processor 'rank' and
+  distributes it 'num' processors. 'comm' is the communicator to be
+  used by MPI. If 'ring' is 0, there is no overlap of the highest
+  level entities between processors - only entities on processor
+  boundaries will be duplicated on other processors. If 'ring' is N,
+  then 'N' layers of the highest level entities will overlap between
+  processors ({\em Currently, only 1 layer of overlap elements are
+    supported}). If 'with\_attr' is 1, then MSTK sets up the links
+  necessary to synchronize attribute data across processors. This routine must be called from only one processor.
+
+\item[] {\bf {\em int} MSTK\_Mesh\_Distribute({\em Mesh\_ptr
+      *}recv\_mesh, {\em int} ring, {\em int} with\_attr, {\em int}
+    rank, {\em int} num, {\em MPI\_Comm} comm)}: Similar to the above
+  Read\_Distribute routine except that it does do any read. On the
+  processor, that is doing the distributing, recv\_mesh holds the
+  input and out meshes.
+
+\item[] {\bf {\em int} MSTK\_UpdateAttr({\em Mesh\_ptr} mesh, {\em
+      int} rank, {\em int} num, {\em MPI\_Comm} comm)}: Update the
+  attributes of duplicated entities on remote processors.
+
+\end{description}
+
+For users requiring more fine grain control of their application, the
+following a list of lower level parallel mesh functions. Descriptions
+of these functions will be added soon. {\bf NOTE THAT THESE ROUTINES
+  HAVE NOT IN ANYWAY BEEN MADE IDIOT-PROOF AND REQUIRE GOOD KNOWLEDGE
+  OF THE PARALLEL MODEL IN MSTK.}
+
+\begin{description}
+
+  \item[] {\bf {\em int} MSTK\_SendMesh({\em Mesh\_ptr} mesh, {\em int} rank, {\em int} with\_attr, {\em MPI\_Comm} comm)}:
+
+  \item[] {\bf {\em int} MSTK\_RecvMesh({\em Mesh\_ptr} mesh, {\em int} dim, {\em int} send\_rank, {\em int} rank, {\em int} with\_attr, {\em MPI\_Comm} comm)}:
+
+  \item[] {\bf {\em int} MESH\_PartitionWithMetis({\em Mesh\_ptr} mesh, {\em int} nparts, {\em int} **part)}:
+  \item[] {\bf {\em int}        MSTK\_Mesh\_Partition({\em Mesh\_ptr} mesh, {\em Mesh\_ptr} *submeshes, {\em int} num, 				 {\em int} ring, {\em int} with\_attr)}
+  \item[] {\bf {\em int}        MESH\_Partition({\em Mesh\_ptr} mesh, {\em Mesh\_ptr} *submeshes, {\em int} num)}
+  \item[] {\bf {\em int}        MESH\_CopyAttr({\em Mesh\_ptr} mesh, {\em Mesh\_ptr} submesh, {\em const char} *attr\_name)}
+
+  \item[] {\bf {\em int}        MESH\_BuildPBoundary({\em Mesh\_ptr} mesh, {\em Mesh\_ptr} submesh)}
+
+  \item[] {\bf {\em int}        MESH\_AddGhost({\em Mesh\_ptr} mesh, {\em Mesh\_ptr} submesh, {\em int} part\_no, {\em int} ring)}
+
+  \item[] {\bf {\em int}        MESH\_SendMesh({\em Mesh\_ptr} mesh, {\em int} rank, {\em MPI\_Comm} comm)}
+  \item[] {\bf {\em int}        MESH\_SendAttr({\em Mesh\_ptr} mesh, {\em const char} *attr\_name, {\em int} rank, {\em MPI\_Comm} comm)}
+  \item[] {\bf {\em int}        MESH\_RecvMesh({\em Mesh\_ptr} mesh, {\em int} dim, {\em int} send\_rank, {\em int} rank, {\em MPI\_Comm} comm)}
+  \item[] {\bf {\em int}        MESH\_RecvAttr({\em Mesh\_ptr} mesh, {\em const char} *attr\_name, {\em int} rank, {\em int} recv\_rank, {\em MPI\_Comm} comm)}
+
+  \item[]{\bf {\em int} MESH\_Num\_GhostVertices({\em Mesh\_ptr} mesh):}
+    Number of ghost vertices in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_GhostEdges({\em Mesh\_ptr} mesh):}
+    Number of ghost edges in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_GhostFaces({\em Mesh\_ptr} mesh):}
+    Number of ghost faces in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_GhostRegions({\em Mesh\_ptr} mesh):}
+    Number of ghost regions in the mesh.
+
+  \item[]
+
+  \item[]{\bf {\em MVertex\_ptr} MESH\_GhostVertex({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th ghost vertex in the mesh. Returns NULL if i
+    $<$ 0 or i $>$ number of mesh vertices.
+
+  \item[]{\bf {\em MEdge\_ptr} MESH\_GhostEdge({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th ghost edge in the mesh. Returns NULL if i $<$
+    0 or i $>$ number of mesh edges. Returns NULL for reduced
+    representations.
+
+  \item[]{\bf {\em MFace\_ptr} MESH\_GhostFace({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
+    0 or i $>$ number of mesh faces. Only faces explicitly represented in
+    the mesh are returned for reduced representation.
+
+  \item[]{\bf {\em MRegion\_ptr} MESH\_GhostRegion({\em Mesh\_ptr} mesh, {\em
+        int} i):} Return the i'th region in the mesh. Returns NULL if i
+    $<$ 0 or i $>$ number of mesh region.
+
+  \item[]
+
+  \item[]{\bf {\em MVertex\_ptr} MESH\_Next\_GhostVertex({\em Mesh\_ptr}
+      mesh, {\em int} *idx):} Returns the next ghost vertex while iterating
+    through the vertices of the mesh. See the routine {\bf
+      List\_Next\_Entry} above for an explanation of how the iteration
+    works. This routine is in general faster than using the routine {\bf
+      MESH\_GhostVertex}.
+
+  \item[]{\bf {\em MEdge\_ptr} MESH\_Next\_GhostEdge({\em Mesh\_ptr} mesh,
+      {\em int} *idx):} Returns the next ghost edge while iterating through the
+    edges of the mesh. See the routine List\_Next\_Entry above for an
+    explanation of how the iteration works.  The routine always returns
+    NULL for reduced representations.
+
+  \item[]{\bf {\em MFace\_ptr} MESH\_Next\_GhostFace({\em Mesh\_ptr} mesh,
+      {\em int} *idx):} Returns the next ghost face while iterating through the
+    faces of the mesh. See the routine {\bf List\_Next\_Entry} above for
+    an explanation of how the iteration works.  Only faces explicitly
+    represented in the mesh are returned for reduced representation.
+
+  \item[]{\bf {\em MRegion\_ptr}
+      MESH\_Next\_GhostRegion({\em Mesh\_ptr} mesh, {\em int} *idx):}
+    Returns the next region while iterating through the regions of the
+    mesh. See the routine {\bf List\_Next\_Entry} above for an
+    explanation of how the iteration works.
+
+  \item[]
+
+  \item[]{\bf {\em int} MESH\_Num\_OverlapVertices({\em Mesh\_ptr} mesh):}
+    Number of overlap vertices in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_OverlapEdges({\em Mesh\_ptr} mesh):}
+    Number of overlap edges in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_OverlapFaces({\em Mesh\_ptr} mesh):}
+    Number of overlap faces in the mesh.
+
+  \item[]{\bf {\em int} MESH\_Num\_OverlapRegions({\em Mesh\_ptr} mesh):}
+    Number of overlap regions in the mesh.
+
+  \item[]
+
+  \item[]{\bf {\em MVertex\_ptr} MESH\_OverlapVertex({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th overlap vertex in the mesh. Returns NULL if i
+    $<$ 0 or i $>$ number of mesh vertices.
+
+  \item[]{\bf {\em MEdge\_ptr} MESH\_OverlapEdge({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th overlap edge in the mesh. Returns NULL if i $<$
+    0 or i $>$ number of mesh edges. Returns NULL for reduced
+    representations.
+
+  \item[]{\bf {\em MFace\_ptr} MESH\_OverlapFace({\em Mesh\_ptr} mesh,
+      {\em int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
+    0 or i $>$ number of mesh faces. Only faces explicitly represented in
+    the mesh are returned for reduced representation.
+
+  \item[]{\bf {\em MRegion\_ptr} MESH\_OverlapRegion({\em Mesh\_ptr} mesh, {\em
+        int} i):} Return the i'th region in the mesh. Returns NULL if i
+    $<$ 0 or i $>$ number of mesh region.
+
+  \item[]
+
+  \item[]{\bf {\em MVertex\_ptr} MESH\_Next\_OverlapVertex({\em Mesh\_ptr}
+      mesh, {\em int} *idx):} Returns the next overlap vertex while iterating
+    through the vertices of the mesh. See the routine {\bf
+      List\_Next\_Entry} above for an explanation of how the iteration
+    works. This routine is in general faster than using the routine {\bf
+      MESH\_OverlapVertex}.
+
+  \item[]{\bf {\em MEdge\_ptr} MESH\_Next\_OverlapEdge({\em Mesh\_ptr} mesh,
+      {\em int} *idx):} Returns the next overlap edge while iterating through the
+    edges of the mesh. See the routine List\_Next\_Entry above for an
+    explanation of how the iteration works.  The routine always returns
+    NULL for reduced representations.
+
+  \item[]{\bf {\em MFace\_ptr} MESH\_Next\_OverlapFace({\em Mesh\_ptr} mesh,
+      {\em int} *idx):} Returns the next overlap face while iterating through the
+    faces of the mesh. See the routine {\bf List\_Next\_Entry} above for
+    an explanation of how the iteration works.  Only faces explicitly
+    represented in the mesh are returned for reduced representation.
+
+  \item[]{\bf {\em MRegion\_ptr}
+      MESH\_Next\_OverlapRegion({\em Mesh\_ptr} mesh, {\em int} *idx):}
+    Returns the next region while iterating through the regions of the
+    mesh. See the routine {\bf List\_Next\_Entry} above for an
+    explanation of how the iteration works.
+
+\end{description}
 

### 2011-11-21 docs/MSTK.tex
Forgot to update version number to 1.8
Modified the arguments for MSTK_Mesh_Distribute and MSTK_Mesh_Read_Distribute

@@
 
-\title{MSTK: Mesh Toolkit, v 1.7}
+\title{MSTK: Mesh Toolkit, v 1.83}
 
@@
 \item[] {\bf {\em int} MSTK\_Mesh\_Read\_Distribute({\em Mesh\_ptr
-      *}recv\_mesh, {\em const char*} global\_mesh\_name, {\em int} dim,
+      *}recv\_mesh, {\em const char*} global\_mesh\_name, {\em int} *dim,
     {\em int} ring, {\em int} with\_attr, {\em int} rank, {\em int}
@@
 \item[] {\bf {\em int} MSTK\_Mesh\_Distribute({\em Mesh\_ptr
-      *}recv\_mesh, {\em int} ring, {\em int} with\_attr, {\em int}
-    rank, {\em int} num, {\em MPI\_Comm} comm)}: Similar to the above
-  Read\_Distribute routine except that it does do any read. On the
-  processor, that is doing the distributing, recv\_mesh holds the
-  input and out meshes.
+      *}recv\_mesh, {\em int} *dim, {\em int} ring, {\em int}
+    with\_attr, {\em int} rank, {\em int} num, {\em MPI\_Comm} comm)}:
+  Similar to the above Read\_Distribute routine except that it does
+  not do any read. On the root processor, 'recv\_mesh' is an
+  input-output parameter - the application uses it to send in the full
+  mesh to be distributed and the mesh partition belonging to the root
+  processor is returned in it. On other processors, this is an output
+  parameter only. Also, 'dim', indicating the topological dimension of
+  the input mesh, is an input parameter on the root processor while it
+  is an output parameter on the other processors.
 

### 2012-08-20 docs/MSTK.tex
Fixed error in declaration of MSTK_InitFromGenDesc

@@
 \item[]{\bf {\em int} MESH\_InitFromGenDesc({\em Mesh\_ptr} mesh, {\em
-      int} nv, {\em double} (*xyz)[3], {\em int} *nfv, {\em int}
-    **fvids, {\em int} nr, {\em int} *nrv, {\em int} **rvids, {\em
-      int} *nrf, {\em int} ***rfvtemplate):} Initialize a mesh from a
+      int} nv, {\em double} (*xyz)[3], {\em int} nf, {\em int} *nfv, 
+    {\em int} **fvids, {\em int} nr, {\em int} *nrv, {\em int}
+    **rvids, {\em int} *nrf, {\em int} ***rfvtemplate):} Initialize a
+  mesh from a
   minimal description of the mesh passed into the routine. In the

### 2015-06-02 docs/MSTK-present.tex
short summary of MSTK

\documentclass{beamer}

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{algorithmic}
\usepackage[boxed]{algorithm}
\usepackage{rotating}
\usepackage{fancybox}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{color}
\usepackage{multicol}
\usepackage{lmodern} % to allow arbitrary font sizes

\usepackage{listings}
\lstset{basicstyle=\color{blue}\small\normalfont,
        commentstyle=\itshape\color{red},
        language=c++}

%------------------------------------------------------------------%
%---BEGIN GRAPHICS BOX DEFN FOR LOGOS------------------------------%
%------------------------------------------------------------------%
\newsavebox{\lanllogo}
\sbox{\lanllogo}{\includegraphics[scale=0.4]{/home/rao/docs/logos/70th-lanllogo-color}}
\newsavebox{\lanllogofooter}
\sbox{\lanllogofooter}{\includegraphics[scale=0.2]{/home/rao/docs/logos/70th-lanllogo-color}}
%------------------------------------------------------------------%
%---END GRAPHICS BOX DEFN FOR LOGOS--------------------------------%
%------------------------------------------------------------------%

%------------------------------------------------------------------%
%--- BEGIN HEADERS AND FOOTERS ------------------------------------%
%------------------------------------------------------------------%

%
% Empty header
%

\setbeamertemplate{headline}{}

%
% If you want just a logo in the right footer
%

%\logo{\usebox{\lanllogofooter}}

%
% If you want a logo on the left, short title and short author in the 
% center and page numbers on the right
%

\setbeamertemplate{footline}
{%
  \raisebox{0.5em}{
    \usebox{\lanllogofooter}\hspace{0.2\textwidth}%
    \raisebox{0.75em}{
      \insertshorttitle---\insertshortauthor%
      \hspace{0.35\textwidth}%
      \insertpagenumber
    }
  }
}

%
% Suppress beamer's naviagation icons
%

\setbeamertemplate{navigation symbols}{}

%------------------------------------------------------------------%
%--- END HEADERS AND FOOTERS --------------------------------------%
%------------------------------------------------------------------%

%------------------------------------------------------------------%
%---END PREAMBLE---------------------------------------------------%
%------------------------------------------------------------------%

\newenvironment{list1}
{\begin{list}{$\bullet$}{\setlength{\leftmargin}{0.0em}\setlength{\topsep}{0.0em}\setlength{\itemsep}{0.3em}}
}
{\end{list}}
\newenvironment{list2}
{\begin{list}{-}{\setlength{\leftmargin}{1.0em}\setlength{\topsep}{0.0em}\setlength{\itemsep}{0.3em}}
}
{\end{list}}

\begin{document}

%------------------------------------------------------------------%
%
%%%% You can use this to change the font
 \renewcommand{\rmdefault}{ptm} %% Roman family to Postscript times roman
 \renewcommand{\sfdefault}{phv} %% San Serif family to Postscript helvetica

 \renewcommand{\ttdefault}{ptc} %% Typewriter family to Postscript courier
 \sffamily                      %% Use the helvetica font
%
%%%% Or this
% \newfont{\"fnt_name''}{``name'' at ``size''}
%
%%%% Or a few other ways - See book
%
%
% Use roman bold font
%\rm\bf

\bfseries  %% Make all letters bold
\small

%------------------------------------------------------------------%
%---END CHANGE FONTS-----------------------------------------------%
%------------------------------------------------------------------%

\raggedright

%------------------------------------------------------------------%
%---END PRELIMS----------------------------------------------------%
%------------------------------------------------------------------%

\title{MSTK: A Flexible Parallel Unstructured Mesh Framework for Multi-physics Simulations}

\author[Garimella]{Rao V. Garimella (\texttt{rao@lanl.gov})}
\institute{T-5 Group, Los Alamos National Laboratory, Los Alamos, NM, USA}
\date{Sep 19, 2014}

\begin{frame}
\thispagestyle{empty}
\titlepage
  \vfill
  \begin{center}
    \usebox{\lanllogo}
  \end{center}
\end{frame}

\begin{frame}
\frametitle{\underline{What is MSTK?}}

\begin{itemize}
\item Parallel unstructured mesh framework for large scale numerical simulations.
\item Allows users to represent, manipulate and query unstructured 3D meshes with arbitrary topology elements.
\item  Applications do not need to code their own data structures. 
\item MSTK is NOT a mesh generator but can be used to write one more easily. 
\item MSTK cannot answer computational
questions related to the mesh (for example, is this point in the given
mesh element?).
\item Open source under LGPL.
\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{\underline{Salient Features of MSTK}}

\begin{itemize} 
\item Allows different mesh representations under common interfaces.
\item Developers can choose represention with all or some intermediate
  entities (edges, faces) but vertices and regions are always represented explicitly.
\item Can query connectivity (adjacency) between any two types of entities
\item Multiple meshes supported simultaneously
\item Meshes can be distributed across large numbers of processors
\item Easy-to-use application programming interface (API) to
  query and modify information about mesh
\item Object-oriented like API - applications do not modify data structures directly
\item Supports named entity attributes and named entity sets
\end{itemize}

%Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
%FEA Applications,'' {\em International Journal of Numerical Methods
%  in Engineering}, v55 n4, pp. 441-478, 2002.
\end{frame}

\begin{frame}
\frametitle{Why you should MSTK?}

\begin{itemize}
\item Offers flexible infrastructure for representing and
manipulating meshes in a distributed environment 
\item ``Object-oriented'' like functional interface for safe, easy access to mesh data
\item Application developers do not have to manage their own mesh data 
\item Use of API cause minimum disruption even if lower level data structures change
\item Easy integration with other codes using MSTK
\item Startup time for new application developers is days not months.
\item Smaller footprint and faster than comparable toolkits$^\dag$
\end{itemize}

$\dag$ ``Mesh Infrastructure for Coupled Multiprocess Geophysical
Simulations'', in Proceedings of the 23rd International Meshing Roundtable,
London, UK, Oct 2014.
\end{frame}

\begin{frame}
\frametitle{MSTK Concepts}

{\underline Mesh Entities:}

\begin{itemize}
\item Nodes/Vertices: topologically 0-dimensional entities, lowest order entity
\item Edges: topologically 1-dimensional entities, bounded by two vertices
\item Faces: topologically 2-dimensional entities (tri, quad, polygon)
\item Regions/Elements/Cells: 3-dimensional entities (tet, hex, prism, polyhedra)
\item[]
\item No higher order elements 
\end{itemize}

{\underline Attributes}: Named application data (integer, real or
pointer) attached to mesh entities

{\underline Sets}: Named groups of mesh entities (may be heterogeneous)

{\underline Lists}: Smart expandable arrays of mesh entities, with low
cost addition and removal of entities

{\underline Markers}: Lightweight binary tags for entities (useful for
searching and for boolean operations)

\end{frame}

\begin{frame}
\frametitle{MSTK Mesh Representations}

\begin{itemize}
\item MSTK has multiple ways of representing meshes
\item Full representations explicitly store all types of mesh entities
\item Reduced representations do not store some intermediate types of entities
\item MSTK can create implicit entities on the fly and automatically
  delete them if they are not used for a while
\item The F1 representation is most tested followed by R1
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Example Mesh Representations}

\begin{figure}[!ht]
  \begin{center}
    \begin{minipage}{2.525in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repF1} \\
        Representation F1
      \end{center}
    \end{minipage}
    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR1} \\
        Representation R1
      \end{center}
    \end{minipage}
  \end{center}
  \label{fig:reptypes}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Mesh entity classification}

{\em Classification} - relationship of each mesh entity to the model
entity it represents the whole or a part of.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/classfn}
  \end{center}
  \caption{Example mesh and model showing the classification of mesh entities on geometric model entities}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Mesh entity classification - continued}

\begin{itemize}
\item Classification useful for simplifying algorithms in meshing and
  analysis applications.
  \begin{itemize}
  \item Specialization of mesh operations such as smoothing on
    boundary of domain
  \item Checking for topological conformity and guarding against
    dimensional reduction in mesh modification
  \item Assignment of material properties
  \item Application of boundary conditions
  \end{itemize}
\item Where possible, applications should supply classification data
\item If unavailable, classification of entities can be derived based
  on ``material IDs'' - but its not foolproof
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Parallel Mesh Management}

\begin{itemize}
\item Support for surface and volume meshes distributed across arbitrary number of processors
\item Currently allows one (or no) layer of overlapped or ghost entities around each partition
\item Ghost layers support all entities and connectivities
\item Mesh can be imported/created on processor 0 and distributed to other processors
\item Mesh can be partitioned using Metis or Zoltan
\item Meshes that are very thin in one direction can be partitioned in lateral direction only
\item Pre-partitioned mesh can be read on each processor and ``woven'' together to establish interprocessor connectivity
\item Support for updating vertex coordinates or attributes across processors
\item No support yet for parallel mesh modification
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Getting meshes in and out of MSTK}

Import:
\begin{itemize}
\item MSTK (Can convert to different representation in import)
\item Exodus II (serial)
\item Exodus II/Nemesis I (parallel)
\item GMV (General Mesh Viewer)
\end{itemize}

Export:
\begin{itemize}
\item MSTK (Can convert to different representation in export)
\item Exodus II(serial)
\item Exodus II/Nemesis I (parallel)
\item GMV
\item STL
\item FLAG X3D
\end{itemize}

Standalone tool 'meshconvert' can also be used for partitioning meshes
and converting between formats.
\end{frame}

\begin{frame}
\frametitle{MSTK based Applications and Software}

\begin{itemize}
  \item Subsurface flow and reactive transport (Amanzi)
  \item Arctic Terrestrial Simulator to study effects of warming on
  arctic permafrost (ATS)
  \item Unstructured polyhedral mesh optimization and untangling (optmesh)
  \item Adaptive Mesh Refinement in 2D
  \item Polyhedral Mesh Generation
  \item Interface Reconstruction
  \item Mimetic Finite Difference based solver for elliptic problems
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{MSTK Computational Cost}

Amanzi software for subsurface flow and transport incorporates three
different mesh frameworks - MSTK (LANL), STKmesh (SNL) and MOAB
(ANL\/U.of.Wisc).

Computational cost of Amanzi's components with the three frameworks
for a subsurface flow problem is shown below:

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/AmanziPerf-MSTK-STK-MOAB.pdf}
\end{center}
\end{figure}

Graph shows performance without (left 3 columns) and with (right 3
columns) caching of frequently used mesh data within Amanzi.

\end{frame}

\begin{frame}
\frametitle{MSTK Memory Usage}

The memory usage of MSTK (using the full F1 representation) compared
to STKmesh (SNL) and MOAB (ANL\/U.of.Wisc) for a simple subsurface flow
problem in Amanzi is shown below (statistics collected by Massif):

\begin{table}
  \begin{center}
  \begin{tabular}{|c|c|c|}
    \hline%
    MSTK & STK-Mesh & MOAB \\
    \hline
    165 MB  & 195 MB  &  240 MB \\
  \hline
  \end{tabular}
  \end{center}
  \label{tab:memtable}
\end{table}

MSTK's memory usage for a million elements on a 64-bit machine (no
extra application data) is as shown:

\begin{table}
  \begin{center}
  \begin{tabular}{|c||c|c|}
    \hline%
     {} & F1 representation & R1 representation \\
    \hline%
    1 Million Tetrahedra  &  650 MB  &  210 MB \\
    1 Million Hexahedra  & 1400 MB  &  430 MB \\
    \hline
  \end{tabular}
  \end{center}
  \label{tab:memtable}
\end{table}

\end{frame}

\begin{frame}
\frametitle{Parallel Scaling}

\end{frame}

\begin{frame}[fragile]
\frametitle{Example program - Initialize mesh}

\begin{verbatim}
#include "MSTK.h"

int main(int argc, char *argv[]) {
 int i, idx, idx2, ok, edir;
 double xyz[3];
 char meshname[256] = "testmesh.mstk";
 Mesh_ptr mesh;
 MVertex_ptr v;
 MEdge_ptr e;
 MFace_ptr f;

 MSTK_Init(); /* Initialize MSTK */

 mesh = MESH_New(UNKNOWN_REP);          /* Create a mesh object */
 if (!(ok = MESH_InitFromFile(mesh,meshname)) { /* Load the mesh */
   fprintf(stderr,"Cannot load input file %s\n\n\n",meshname);
   exit(-1);
 }

\end{verbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example (contd.) - Print vertex information}

\begin{verbatim}
 int nv = MESH_Num_Vertices(mesh);
 for (i = 0; i < nv; i++) \{
  v = MESH_Vertex(mesh,i);

  /* Basic info */
  printf("\nVertex: 0x%-x   ID: %-d   ", v, MV_ID(v));

  /* Classification w.r.t. geometric model */

  int gdim = MV_GEntDim(v);
  int gid = MV_GEntID(v);
  printf("GEntID: %-d   GEntDim: %-d\n", gid, gdim);

  MV_Coords(v,xyz);   /* Coordinates */
  printf("Coords: %16.8lf %16.8lf %16.8lf\n",xyz[0],xyz[1],xyz[2]);
 \}
\end{verbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example (contd.) - Print face information}

\begin{verbatim}
 idx = 0;
 while ((f = MESH_Next_Face(mesh,&idx))) {

  printf("\nFace: 0x%-x   ID: %-d   ", f, MF_ID(f));

  printf("Vertex IDs:\n");
  List_ptr fvertices = MF_Vertices(f,1,0);

  idx2 = 0;
  while ((v = List_Next_Entry(fvertices,&idx2)))
    printf("%-d\n",MV_ID(v));
  printf("\n");

  List_Delete(fvertices);
 }

 MESH_Delete(mesh);
 return 1;
}
\end{verbatim}
\end{frame}

\begin{frame}
\frametitle{Acknowledgements}

This work was performed under the auspices of the National Nuclear
Security Administration of the US Department of Energy at Los Alamos
National Laboratory under contract No. DE-AC52-06NA25396 as part of
projects funded by the DOE Office of Science SciDAC program, the DOE
Advanced Simulation Capability (ASC) program, the DOE Advanced
Simulation Capability for Environmental Management (ASCEM) program and
Los Alamos National Laboratory LDRD program.

\end{frame}

\end{document}

### 2016-01-19 docs/MSTK.tex
Added enum for gather operations on entity attributes

@@
 \item[]{\bf {\em int} MESH\_Num\_Vertices({\em Mesh\_ptr} mesh):}
-Number of vertices in the mesh.
-
-\item[]{\bf {\em int} MESH\_Num\_Edges({\em Mesh\_ptr} mesh):}
-Number of edges in the mesh. For reduced representations, this routine
-returns 0 since it is impractically expensive to count the number of
-edges when they do not explicitly exist. Applications must find a way
-to avoid using this routine for reduced representations.
-
-\item[]{\bf {\em int} MESH\_Num\_Faces({\em Mesh\_ptr} mesh):}
-Number of faces in the mesh. For reduced representations R1 or R2,
-this routine counts only the faces that are explicitly represented
-i.e. faces not connected to any mesh region. Therefore, a value of 0
-will be returned for the number of faces of a tetrahedral mesh with
-representation R1 or R2 but the correct number will be reported for a
-tetrahedral mesh in other representations. Also, the correct number
-will be reported for the number of faces in a surface mesh in
-representation R1 or R2.  Therefore, this routine must be used
-carefully.
+Number of vertices in the mesh. This includes owned and ghost
+vertices.
+
+\item[]{\bf {\em int} MESH\_Num\_Edges({\em Mesh\_ptr} mesh):} Number
+  of edges in the mesh (owned and ghost). For reduced representations,
+  this routine returns 0 since it is impractically expensive to count
+  the number of edges when they do not explicitly exist. Applications
+  must find a way to avoid using this routine for reduced
+  representations.
+
+\item[]{\bf {\em int} MESH\_Num\_Faces({\em Mesh\_ptr} mesh):} Number
+  of faces in the mesh (owned and ghost). For reduced representations
+  R1 or R2, this routine counts only the faces that are explicitly
+  represented i.e. faces not connected to any mesh region. Therefore,
+  a value of 0 will be returned for the number of faces of a
+  tetrahedral mesh with representation R1 or R2 but the correct number
+  will be reported for a tetrahedral mesh in other
+  representations. Also, the correct number will be reported for the
+  number of faces in a surface mesh in representation R1 or R2.
+  Therefore, this routine must be used carefully.
 
 \item[]{\bf {\em int} MESH\_Num\_Regions({\em Mesh\_ptr} mesh):}
-Number of regions in the mesh.
+Number of regions in the mesh (owned and ghost).
 
@@
 {\em int} i):} Return the i'th vertex in the mesh. Returns NULL if i
-$<$ 0 or i $>$ number of mesh vertices.
-
-\item[]{\bf {\em MEdge\_ptr} MESH\_Edge({\em Mesh\_ptr} mesh,
-{\em int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
-0 or i $>$ number of mesh edges. Returns NULL for reduced
-representations.
-
-\item[]{\bf {\em MFace\_ptr} MESH\_Face({\em Mesh\_ptr} mesh,
-{\em int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
-0 or i $>$ number of mesh faces. Only faces explicitly represented in
-the mesh are returned for reduced representation (See explanation for
-MESH\_Num\_Faces).
+$<$ 0 or i $>$ number of mesh vertices. The vertex may be owned or ghost.
+
+\item[]{\bf {\em MEdge\_ptr} MESH\_Edge({\em Mesh\_ptr} mesh, {\em
+    int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
+  0 or i $>$ number of mesh edges. The edge may be owned or
+  ghost. Returns NULL for reduced representations.
+
+\item[]{\bf {\em MFace\_ptr} MESH\_Face({\em Mesh\_ptr} mesh, {\em
+    int} i):} Return the i'th face in the mesh. Returns NULL if i $<$
+  0 or i $>$ number of mesh faces. The face may be owned or
+  ghost. Only faces explicitly represented in the mesh are returned
+  for reduced representation (See explanation for MESH\_Num\_Faces).
 
@@
     int} i):} Return the i'th region in the mesh. Returns NULL if i
-  $<$ 0 or i $>$ number of mesh region.
+  $<$ 0 or i $>$ number of mesh region. The region may be owned or
+  ghost.
 
@@
   mesh, {\em int} *idx):} Returns the next vertex while iterating
-  through the vertices of the mesh. See the routine {\bf
-    List\_Next\_Entry} above for an explanation of how the iteration
-  works. This routine is in general faster than using the routine {\bf
-    MESH\_Vertex}.
+  through the vertices of the mesh. The vertex may be owned or ghost
+  and the type must be checked using {\bf MV\_PType}. See the routine
+  {\bf List\_Next\_Entry} above for an explanation of how the
+  iteration works. This routine is in general faster than using the
+  routine {\bf MESH\_Vertex}.
 
@@
   {\em int} *idx):} Returns the next edge while iterating through the
-edges of the mesh. See the routine List\_Next\_Entry above for an
-explanation of how the iteration works.  The routine always returns
-NULL for reduced representations.
+  edges of the mesh. The edge may be owned or ghost
+  and the type must be checked using {\bf ME\_PType}. See the routine
+  List\_Next\_Entry above for an explanation of how the iteration
+  works.  The routine always returns NULL for reduced representations.
 
@@
   {\em int} *idx):} Returns the next face while iterating through the
-faces of the mesh. See the routine {\bf List\_Next\_Entry} above for
-an explanation of how the iteration works.  Only faces explicitly
-represented in the mesh are returned for reduced representation (See
-explanation for {\bf MESH\_Num\_Faces}).
-
-\item[]{\bf {\em MRegion\_ptr}
-    MESH\_Next\_Region({\em Mesh\_ptr} mesh, {\em int} *idx):}
-  Returns the next region while iterating through the regions of the
-  mesh. See the routine {\bf List\_Next\_Entry} above for an
-  explanation of how the iteration works.
+  faces of the mesh. The face may be owned or ghost
+  and the type must be checked using {\bf MF\_PType}. See the routine
+  {\bf List\_Next\_Entry} above for an explanation of how the
+  iteration works.  Only faces explicitly represented in the mesh are
+  returned for reduced representation (See explanation for {\bf
+    MESH\_Num\_Faces}).
+
+\item[]{\bf {\em MRegion\_ptr} MESH\_Next\_Region({\em Mesh\_ptr}
+  mesh, {\em int} *idx):} Returns the next region while iterating
+  through the regions of the mesh. The region may be owned or ghost
+  and the type must be checked using {\bf MR\_PType}. See the routine
+  {\bf List\_Next\_Entry} above for an explanation of how the
+  iteration works.
 

### 2016-01-19 include/MSTK_types.h

@@ -54,16 +54,19 @@ extern "C" {
   typedef void     *Hash_ptr;
 #endif
 
-  typedef enum RepType {F1=0, F4, R1, R2, R4, UNKNOWN_REP} RepType;
+  typedef enum {F1=0, F4, R1, R2, R4, UNKNOWN_REP} RepType;
 
-typedef enum MFType {FDELETED=-1, FUNKNOWN=0, TRI=3, QUAD, POLYGON} MFType;
-typedef enum MRType {RDELETED=-1, RUNKNOWN=0, TET, PYRAMID, PRISM, HEX, POLYHED} MRType;
+typedef enum {FDELETED=-1, FUNKNOWN=0, TRI=3, QUAD, POLYGON} MFType;
+typedef enum {RDELETED=-1, RUNKNOWN=0, TET, PYRAMID, PRISM, HEX, POLYHED} MRType;
 
-  typedef enum MType {MDELETED=-1, MVERTEX=0, MEDGE=1, MFACE=2, MREGION=3, MUNKNOWNTYPE=4, MALLTYPE=5, MANYTYPE=6} MType;
+  typedef enum {MDELETED=-1, MVERTEX=0, MEDGE=1, MFACE=2, MREGION=3, MUNKNOWNTYPE=4, MALLTYPE=5, MANYTYPE=6} MType;
 
 /* typedefs needed for attributes */
-typedef enum MAttType {INT=0, DOUBLE, POINTER, VECTOR, TENSOR} MAttType;
+typedef enum {INT=0, DOUBLE, POINTER, VECTOR, TENSOR} MAttType;
 
+/* typedefs to indicate how to aggregate attribute values from different processors */
+
+typedef enum {ATTOP_UNDEF, ATTOP_MAX, ATTOP_MIN, ATTOP_SUM, ATTOP_AVG} MAttOpType;
 
 #ifdef MSTK_HAVE_MPI  
 

### 2016-01-19 src/io/MESH_ExportToFLAGX3D.c

@@ -4,7 +4,7 @@
 #include <time.h>
 
 #include "MSTK.h"
-
+#include "MSTK_private.h"
 
 #ifdef __cplusplus
 extern "C" {
@@ -33,8 +33,8 @@ int MESH_ExportToFLAGX3D(Mesh_ptr mesh, const char *filename, const int natt,
   MEdge_ptr             edge;
   MFace_ptr	        face;
   MRegion_ptr           region;
-  MAttrib_ptr           attrib, *nodatts=NULL, *cellatts=NULL, oppatt;
-  MAttrib_ptr           vidatt, eidatt, fidatt, ridatt;
+  MAttrib_ptr           attrib, *nodatts=NULL, *cellatts=NULL;
+  MAttrib_ptr           vidatt, eidatt, fidatt, ridatt, oppatt, opppidatt;
   MAttType              atttype;
   char                  attname[256], matname[256], tmpstr[256];
   int                   jv, je, jr, jf;
@@ -51,33 +51,34 @@ int MESH_ExportToFLAGX3D(Mesh_ptr mesh, const char *filename, const int natt,
   FILE		        *fp;
 
 #ifdef MSTK_HAVE_MPI
-  int rank, numprocs;
+  int pid, numprocs;
   MPI_Comm_size(comm,&numprocs);
-  MPI_Comm_rank(comm,&rank);
+  MPI_Comm_rank(comm,&pid);
+  pid += 1;  /* FLAG X3D counts processor IDs from 1 */
 #endif
 
   vidatt = MAttrib_New(mesh,"vidatt",INT,MVERTEX);
   eidatt = MAttrib_New(mesh,"eidatt",INT,MEDGE);
   fidatt = MAttrib_New(mesh,"fidatt",INT,MFACE);
   ridatt = MAttrib_New(mesh,"ridatt",INT,MREGION);
-
+  
   idx = 0; i = 0;
   while ((vertex = MESH_Next_Vertex(mesh,&idx)))
     MEnt_Set_AttVal(vertex,vidatt,++i,0.0,NULL);
-
+  
   idx = 0; i = 0;
   while ((edge = MESH_Next_Edge(mesh,&idx)))
     MEnt_Set_AttVal(edge,eidatt,++i,0.0,NULL);
-
+  
   idx = 0; i = 0;
   while ((face = MESH_Next_Face(mesh,&idx)))
     MEnt_Set_AttVal(face,fidatt,++i,0.0,NULL);
-
[…]
